### 2011-01-10 text/src/article.tex
Initial project setup

%% BioMed_Central_Tex_Template_v1.05
%%                                      %
%  bmc_article.tex            ver: 1.05 %
%                                       %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%         <27 January 2006>           %%
%%                                     %%
%%                                     %%
%% Uses:                               %%
%% cite.sty, url.sty, bmc_article.cls  %%
%% ifthen.sty. multicol.sty		       %%
%%									   %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%	
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.pdf and the instructions for    %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%   
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %% 
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %% 
%%  submitted article.                         %% 
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}

\newboolean{publ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                              %%
%% You may change the following style settings  %%
%% Should you wish to format your article       %%
%% in a publication style for printing out and  %%
%% sharing with colleagues, but ensure that     %%
%% before submitting to BMC that the style is   %%
%% returned to the Review style setting.        %%
%%                                              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

% Begin ...
\begin{document}
\begin{bmcformat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A sample article title}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Ensure \and is entered between all but   %%
%% the last two authors. This will be       %%
%% replaced by a comma in the final article %%
%%                                          %%
%% Ensure there are no trailing spaces at   %% 
%% the ends of the lines                    %%     	
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Charles A Darwin\correspondingauthor$^{1,2}$%
       \email{Charles A Darwin\correspondingauthor - charles@londonzoo.co.uk}%
      \and
         Jane E Doe\correspondingauthor$^2$%
         \email{Jane E Doe\correspondingauthor - jane.e.doe@cambridge.co.uk}
       and 
         John RS Smith$^3$%
         \email{John RS Smith - john.RS.Smith@cambridge.co.uk}%
      }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address{%
    \iid(1)Life Sciences Department, Kings College London, Cornwall House,%
        Waterloo Road, London, UK\\
    \iid(2)Department of Zoology, Cambridge, Waterloo Road, London, UK\\
    \iid(3)Marine Ecology Department, Institute of Marine Sciences Kiel, %
        D\"{u}sternbrooker Weg 20, 24105 Kiel, Germany
}%

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the Instructions for       %%
%% authors on http://www.biomedcentral.com  %%
%% and change the section headings          %%
%% accordingly.                             %%   
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
        % Do not use inserted blank lines (ie \\) until main body of text.
        \paragraph*{Background:} Text for this section of the abstract.

        \paragraph*{Results:} Text for this section of the abstract \ldots

        \paragraph*{Conclusions:} Text for this section of the abstract \ldots
\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the instructions for       %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and change the section headings          %%
%% accordingly.                             %% 
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
 Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results and Discussion %%
%%
\section*{Results and Discussion}
  \subsection*{Results sub-heading}
    \subsubsection*{This is a sub-sub-heading}
      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
      pb at end of lines ensures correct paragraph spacing.\pb
	  Text for this sub-sub-section \ldots
    \subsubsection*{Another sub-sub-sub-heading}
      Text for this sub-sub-section \ldots

  \subsection*{Another results sub-heading}
    Text for this sub-section \ldots

  \subsection*{Yet another results sub-heading}
    Text for this sub-section.  More results \ldots

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
  Text for this section \ldots

%%%%%%%%%%%%%%%%%%
\section*{Methods}
  \subsection*{Methods sub-heading for this section}
    Text for this sub-section \ldots

  \subsection*{Another methods sub-heading for this section}
    Text for this sub-section \ldots

  \subsection*{Yet another sub-heading for this section}
    Text for this sub-section \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
    Text for this section \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
  \ifthenelse{\boolean{publ}}{\small}{}
  Text for this section \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%              
%%  Bmc_article.bst  will be used to                       %%
%%  create a .BBL file for submission, which includes      %%
%%  XML structured for BMC.                                %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %% 
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %% 
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article} }     % Bibliography file (usually '*.bib' )

%%%%%%%%%%%

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}
  \subsection*{Figure 1 - Sample figure title}
      A short description of the figure content
      should go here.

  \subsection*{Figure 2 - Sample figure title}
      Figure legend text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Sample table title}
    Here is an example of a \emph{small} table in \LaTeX\ using  
    \verb|\tabular{...}|. This is where the description of the table 
    should go. \par \mbox{}
    \par
    \mbox{
      \begin{tabular}{|c|c|c|}
        \hline \multicolumn{3}{|c|}{My Table}\\ \hline
        A1 & B2  & C3 \\ \hline
        A2 & ... & .. \\ \hline
        A3 & ..  & .  \\ \hline
      \end{tabular}
      }
  \subsection*{Table 2 - Sample table title}
    Large tables are attached as separate files but should
    still be described here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{bmcformat}
\end{document}

### 2011-01-10 text/build.sh

@@ -1,0 +1,36 @@
+#!/bin/sh
+
+# Targets
+SRC=src
+LIB=lib
+OUT=out
+TEMP=tmp
+
+ARTICLE=article
+
+# Clean directory
+if [ -d $TEMP ] ; then
+        rm $TEMP/*
+        rmdir $TEMP
+fi	
+mkdir $TEMP
+
+if [ -d $OUT ] ; then
+        rm $OUT/*
+        rmdir $OUT
+fi
+mkdir $OUT
+
+# Copy required files
+cp $SRC/* $TEMP
+cp $LIB/* $TEMP
+
+# Create the pdf and ps files
+make -C $TEMP
+
+# Copy produced files to out
+cp $TEMP/$ARTICLE.pdf $OUT
+
+# Remove the temporary directory
+rm $TEMP/*
+rmdir $TEMP

### 2011-01-11 text/src/article.tex
Add vim folding for manuscript
Correct permissions for LaTeX files

@@
+% {{{ Preamble
+
 %% BioMed_Central_Tex_Template_v1.05
@@
 \begin{bmcformat}
-
+% }}}
+% {{{ Title
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \maketitle
-
+%}}}
+% {{{ Abstract
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% The Section headings here are those for  %%
-%% a Research article submitted to a        %%
-%% BMC-Series journal.                      %%  
-%%                                          %%
-%% If your article is not of this type,     %%
-%% then refer to the instructions for       %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and change the section headings          %%
-%% accordingly.                             %% 
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%
-%% Background %%
-%%
+% }}}
+% {{{ Background
 \section*{Background}
  Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
+% }}}
+% {{{ Implementation
+%%%%%%%%%%%%%%%%%%%%%%
+\section*{Implementation}
+  Text for this section \ldots
+% }}}
+% {{{ Results and Discussion
 
@@
     Text for this sub-section.  More results \ldots
-
+% }}}
+% {{{ Conclusions
 %%%%%%%%%%%%%%%%%%%%%%
@@
   Text for this section \ldots
-
-%%%%%%%%%%%%%%%%%%
-\section*{Methods}
-  \subsection*{Methods sub-heading for this section}
-    Text for this sub-section \ldots
-
-  \subsection*{Another methods sub-heading for this section}
-    Text for this sub-section \ldots
-
-  \subsection*{Yet another sub-heading for this section}
-    Text for this sub-section \ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% }}}
+% {{{ Availability and Requirements
+%%%%%%%%%%%%%%%%%%%%%%
+\section*{Availability and Requirements}
+  Text for this section \ldots
+% }}}
+% {{{ Competing Interests
 \section*{Authors contributions}
     Text for this section \ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% }}}
+% {{{ Author Contributions
+\section*{Authors contributions}
+    Text for this section \ldots
+% }}}
+% {{{ Author's Information
+\section*{Authors contributions}
+    Text for this section \ldots
+% }}}
+% {{{ Acknowledgements
 \section*{Acknowledgements}
@@
   Text for this section \ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%              
-%%  Bmc_article.bst  will be used to                       %%
-%%  create a .BBL file for submission, which includes      %%
-%%  XML structured for BMC.                                %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %% 
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %% 
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%}}}
+% {{{ Bibliography
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
@@
   \bibliography{article} }     % Bibliography file (usually '*.bib' ) 
-
-%%%%%%%%%%%
-
+%}}}
+% {{{ Figures
 \ifthenelse{\boolean{publ}}{\end{multicols}}{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
 
@@
       Figure legend text.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
+% }}}
+% {{{ Tables
 \section*{Tables}
@@
     still be described here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% }}}
+% {{{ Additional Files
 \section*{Additional Files}
@@
 \end{document}
+% }}}

### 2011-01-11 text/src/article.tex
Add working title
Update authors

@@
 % {{{ Title
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\title{A sample article title}
+\title{Scaffolder - Simple and reproducable genome scaffolding using
+text configuration files.}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Ensure \and is entered between all but   %%
-%% the last two authors. This will be       %%
-%% replaced by a comma in the final article %%
-%%                                          %%
-%% Ensure there are no trailing spaces at   %% 
-%% the ends of the lines                    %%     	
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author{Charles A Darwin\correspondingauthor$^{1,2}$%
-       \email{Charles A Darwin\correspondingauthor - charles@londonzoo.co.uk}%
+\author{
+  Michael D Barton$^{1}$%
+  \email{Michael D Barton - mail@michaelbarton.me.uk}%
 \and
-         Jane E Doe\correspondingauthor$^2$%
-         \email{Jane E Doe\correspondingauthor - jane.e.doe@cambridge.co.uk}
-       and 
-         John RS Smith$^3$%
-         \email{John RS Smith - john.RS.Smith@cambridge.co.uk}%
+  Hazel A Barton\correspondingauthor$^1$%
+  \email{Hazel A Barton\correspondingauthor - bartonh@nku.edu}%
       }
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address{%
-    \iid(1)Life Sciences Department, Kings College London, Cornwall House,%
-        Waterloo Road, London, UK\\
-    \iid(2)Department of Zoology, Cambridge, Waterloo Road, London, UK\\
-    \iid(3)Marine Ecology Department, Institute of Marine Sciences Kiel, %
-        D\"{u}sternbrooker Weg 20, 24105 Kiel, Germany
-}%
+\address{\iid(1) Department of Biological Sciences, Northern Kentucky%
+University, SC 204D Nunn Drive, Highland Heights, KY 41076 }%
 

### 2011-01-11 text/src/article.tex
Set abstract sections
Fix address

@@
 \address{\iid(1) Department of Biological Sciences, Northern Kentucky%
-University, SC 204D Nunn Drive, Highland Heights, KY 41076 }%
+University, Nunn Drive, Highland Heights, KY 41076 }%
 
@@
 % {{{ Abstract
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% The Section headings here are those for  %%
-%% a Research article submitted to a        %%
-%% BMC-Series journal.                      %%  
-%%                                          %%
-%% If your article is not of this type,     %%
-%% then refer to the Instructions for       %%
-%% authors on http://www.biomedcentral.com  %%
-%% and change the section headings          %%
-%% accordingly.                             %%   
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{abstract}
-        % Do not use inserted blank lines (ie \\) until main body of text.
-        \paragraph*{Background:} Text for this section of the abstract.
-
-        \paragraph*{Results:} Text for this section of the abstract \ldots
-
-        \paragraph*{Conclusions:} Text for this section of the abstract \ldots
+  \paragraph*{Background:} Background \dots
+  \paragraph*{Methods:} Methods \dots
+  \paragraph*{Results:} Results \ldots
+  \paragraph*{Conclusions:} Conclusions \ldots
 \end{abstract}

### 2011-01-12 text/src/article.tex
Initial draft of abstract

@@
 \begin{abstract}
-  \paragraph*{Background:} Background \dots
-  \paragraph*{Methods:} Methods \dots
-  \paragraph*{Results:} Results \ldots
-  \paragraph*{Conclusions:} Conclusions \ldots
+
+  \paragraph*{Background:} Assembly of short read sequencing data can result in
+  a fragmented and non-contiguous final sequence. A common step in a genome
+  sequencing project is finish the sequence be joining contigs together and
+  filling gaps generated by PCR. This scaffolding step however is non-trivial
+  as manipulating large blocks of fasta sequence is error-prone. Further more
+  join sequences together in single fasta file hides the source of each region
+  of sequence in the final genome build making reproducing the steps or
+  editting an existing genome build difficult.
+
+  \paragraph*{Methods:} Scaffolder is tool designed for scaffolding contigs
+  togther to build a draft genome sequence. Scaffolder is implemented in the
+  ruby programming language and is installed via the rubygem package management
+  system. Genome scaffolds are written in YAML a markup language designed to be
+  both human and machine readable. Command line binaries and extensive
+  documentation are also available.
+
+  \paragraph*{Results:} Scaffolder allows a genome build to be defined in in
+  a scaffold file using a simple human readable syntax. The scaffold file
+  highlights where each contig is used to build the draft genome sequence
+  making the genome scaffolding reproducible. Changes and updates to the genome
+  sequence are therefore comparatively easier to make compared to editing
+  fasta. Scaffolder supports adding inserts to the scaffold to bridge gaps or
+  replace regions in the scaffold. Scaffolder can be installed through the ruby
+  gems.
+
+  \paragraph*{Conclusions:} Conclusions \ldots Scaffolder represents an easy to
+  use genome scaffolding tool. This tool promotes reproducibility and
+  maintenence in building a genome. Scaffolder can be found at
+  http://michaelbarton.me.uk/sfdr.
+
 \end{abstract}

### 2011-01-12 text/src/article.tex
Revise abstract and title

--- 2011-01-11 text/src/article.tex
@@
 
-\title{Scaffolder - Simple and reproducable genome scaffolding using
+\title{Scaffolder - Simple and reproducible genome scaffolding using
 text configuration files.}
@@
 \begin{abstract}
-  \paragraph*{Background:} Background \dots
-  \paragraph*{Methods:} Methods \dots
-  \paragraph*{Results:} Results \ldots
-  \paragraph*{Conclusions:} Conclusions \ldots
+
+  \paragraph*{Background:} Assembly of short read sequencing data can result in
+  a fragmented series of contigs. Therefore a common step in a genome project
+  is piece separate contigs together and fill gaps generated by PCR. This
+  scaffolding step however is non-trivial requiring manipulation of large
+  blocks of nucleotide sequence. Manually joining contigs also hides the source
+  of each region in the final genome fasta file. Taken together these
+  considerations make reproducing or editing an existing genome build
+  difficult.
+
+  \paragraph*{Methods:} The software outlined "Scaffolder" is implemented in
+  the ruby programming language and installed via the RubyGems software
+  management system. Genome scaffolds are defined using YAML - a markup
+  language which is both human and machine readable. Command line binaries and
+  extensive documentation are also available for the package.
+
+  \paragraph*{Results:} This software allows a genome build to be defined in
+  terms of the constituent contigs using a relatively simple to write and edit
+  scaffold file syntax. This scaffold file defines where each contig is used
+  making the genome build reproducible and comparatively easier to edit than
+  fasta sequence. This syntax also supports adding inserts to the scaffold for
+  the purpose of bridging gaps or replace regions.
+
+  \paragraph*{Conclusions:} Scaffolder is easy to use genome scaffolding
+  software. This tool promotes reproducibility and maintenance in building
+  a genome. Scaffolder can be found at http://michaelbarton.me.uk/sfdr.
+
 \end{abstract}

### 2011-01-13 text/src/article.tex
Revise abstract

@@
   is piece separate contigs together and fill gaps generated by PCR. This
-  scaffolding step however is non-trivial requiring manipulation of large
+  scaffolding step, however, is non-trivial requiring manipulation of large
   blocks of nucleotide sequence. Manually joining contigs also hides the source
@@
   management system. Genome scaffolds are defined using YAML - a markup
-  language which is both human and machine readable. Command line binaries and
+  language, which is both human and machine-readable. Command line binaries and
   extensive documentation are also available for the package.

### 2011-01-13 text/src/article.tex
Further revise abstract

@@
   a fragmented series of contigs. Therefore a common step in a genome project
-  is piece separate contigs together and fill gaps generated by PCR. This
-  scaffolding step, however, is non-trivial requiring manipulation of large
-  blocks of nucleotide sequence. Manually joining contigs also hides the source
-  of each region in the final genome fasta file. Taken together these
-  considerations make reproducing or editing an existing genome build
-  difficult.
+  is to join neighbouring sequence regions together and fill gaps using insert
+  data generated through PCR. This scaffolding step, however, is non-trivial
+  and requires manually editing large blocks of nucleotide sequence. Joining
+  sequence contigs together also hides the source of each region in the final
+  genome sequence. Taken together, these considerations may make reproducing or
+  editing an existing genome build difficult.
 
   \paragraph*{Methods:} The software outlined "Scaffolder" is implemented in
-  the ruby programming language and installed via the RubyGems software
+  the Ruby programming language and can be installed via the RubyGems software
   management system. Genome scaffolds are defined using YAML - a markup
   language, which is both human and machine-readable. Command line binaries and
-  extensive documentation are also available for the package.
+  extensive documentation are provided.
 
@@
   terms of the constituent contigs using a relatively simple to write and edit
-  scaffold file syntax. This scaffold file defines where each contig is used
-  making the genome build reproducible and comparatively easier to edit than
-  fasta sequence. This syntax also supports adding inserts to the scaffold for
-  the purpose of bridging gaps or replace regions.
+  scaffold file syntax. The Scaffolder syntax further allows inserts to be
+  added to the scaffold for the purpose of filling unknown sequence regions.
+  Defining the genome construction in a file makes the scaffolding process
+  reproducible and comparatively easier to edit than raw fasta sequence.
 

### 2011-01-18 text/src/article.tex

@@
 
+% Scaffolder url
+\urldef{\scaffolder}\url{http://www.michaelbarton.me.uk/projects/scaffolder/}
+
 % Begin ...
@@
 
-\title{Scaffolder - Simple and reproducible genome scaffolding using
-text configuration files.}
+\title{Scaffolder - Software for Microbial Genome Scaffolding.}
 
@@
   Defining the genome construction in a file makes the scaffolding process
-  reproducible and comparatively easier to edit than raw fasta sequence.
+  reproducible and comparatively easier to edit than raw fasta.
 
@@
   software. This tool promotes reproducibility and maintenance in building
-  a genome. Scaffolder can be found at http://michaelbarton.me.uk/sfdr.
+  a genome. Scaffolder can be found at \scaffolder.
 

### 2011-01-18 text/src/article.tex
Add license and requirements
No competing interests
Author contributions

@@
 % {{{ Availability and Requirements
-%%%%%%%%%%%%%%%%%%%%%%
 \section*{Availability and Requirements}
-  Text for this section \ldots
+
+  \begin{description}
+    \item[Project name:] Scaffolder
+    \item[Project home page:] \scaffolder
+    \item[Operating system:] Platform Independent. Tested on Mac OS X and
+    Debian.
+    \item[Programming language:] Ruby
+    \item[Other requirements:] RubyGems
+    \item[License:] MIT
+    \item[Any restrictions to use by non-academics:] None
+  \end{description}
+
 % }}}
@@
 \section*{Authors contributions}
-    Text for this section \ldots
+  The authors declare no competing interests.
 % }}}
@@
 \section*{Authors contributions}
-    Text for this section \ldots
+
+MDB developed and maintains the scaffolder tool. MDB and HAB wrote the
+manuscript.
+
 % }}}

### 2011-01-18 text/src/article.tex
Initial draft of implementation section

@@
 \section*{Background}
- Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
+ Text for this section.
 % }}}
 % {{{ Implementation
-%%%%%%%%%%%%%%%%%%%%%%
 \section*{Implementation}
-  Text for this section \ldots
+
+Scaffolder is written in Ruby \cite{ruby-lang} using version 1.8.7. Unit and
+ntegration tests use RSpec and Cucumber\cite{rspec}. The source code is
+documented in-line using Yard\cite{yard} and the Unix manual pages are
+generated using Ronn\cite{ronn}.
+
+The sequence scaffold is defined in YAML\cite{yaml}. YAML is a machine
+readable data format written using indentation of white space. YAML is designed
+to be human readable which is ideal for manual editing of a genome scaffold.
+Each stretch of sequence in the genome scaffold is defined as a list entry in
+the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
+each scaffold entry to create an instance of the Scaffolder::Region class.
+
+The Region class is inherited by all types of sequence in the scaffold. The
+Region provides helper methods common to the tasks in manipulating scaffold
+sequence, subsequencing and reversing. In particular the region class has
+a class annotation method\cite{class-annotations} which allows attributes in the
+scaffold method to be defined in the class. This negates having to write
+a custom method to deal with each scaffold variable.  Furthermore this makes
+the process of transferring attributes from the scaffold file to each region
+instance much simpler - each yaml attribute corresponds to a method call on the
+instantiated class. The scaffold file effectively becomes a DSL for populating
+the scaffold region instances while still maintaining the machine readable data
+qualities of YAML.
+
+The region class provides a sequence hook method. This method returns by
+default the sequence corresponding to the fasta sequence defined by the source
+tag which is then sublclassed or reversed if necessary. Subclasses of the region
+class should override this method to provide custom functionality in
+manipulating the source sequence. For example, the Scaffolder::Region::Sequence
+class overrides the sequence method add any sequence inserts specified in the
+scaffold file.
+
 % }}}
@@
   The authors declare no competing interests.
+% }}}
+% {{{ Availability and Requirements
+\section*{List of Abbreviations Used}
+
+  \begin{description}
+    \item[YAML:] YAML Ain't Markup Language\cite{yaml}
+  \end{description}
+
 % }}}

### 2011-01-19 text/src/article.tex
Insert paragraph breaks

@@
 documented in-line using Yard\cite{yard} and the Unix manual pages are
-generated using Ronn\cite{ronn}.
+generated using Ronn\cite{ronn}.\pb
 
@@
 the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
-each scaffold entry to create an instance of the Scaffolder::Region class.
+each scaffold entry to create an instance of the Scaffolder::Region class.\pb
 
@@
 sequence, subsequencing and reversing. In particular the region class has
-a class annotation method\cite{class-annotations} which allows attributes in the
-scaffold method to be defined in the class. This negates having to write
+a class annotation method\cite{class-annotations} which allows attributes in
+the scaffold method to be defined in the class. This negates having to write
 a custom method to deal with each scaffold variable.  Furthermore this makes
 the process of transferring attributes from the scaffold file to each region
-instance much simpler - each yaml attribute corresponds to a method call on the
-instantiated class. The scaffold file effectively becomes a DSL for populating
-the scaffold region instances while still maintaining the machine readable data
-qualities of YAML.
+instance much simpler - each yaml attribute corresponds to a method call on
+the instantiated class. The scaffold file effectively becomes a DSL for
+populating the scaffold region instances while still maintaining the machine
+readable data qualities of YAML.\pb
 
@@
 default the sequence corresponding to the fasta sequence defined by the source
-tag which is then sublclassed or reversed if necessary. Subclasses of the region
-class should override this method to provide custom functionality in
-manipulating the source sequence. For example, the Scaffolder::Region::Sequence
-class overrides the sequence method add any sequence inserts specified in the
-scaffold file.
+tag which is then sublclassed or reversed if necessary. Subclasses of the
+region class should override this method to provide custom functionality in
+manipulating the source sequence. For example, the
+Scaffolder::Region::Sequence class overrides the sequence method add any
+sequence inserts specified in the scaffold file.\pb
 

### 2011-01-19 text/src/article.tex
Draft outline for results section

--- 2011-01-18 text/src/article.tex
@@
 % }}}
-% {{{ Results and Discussion
+\section*{Results and Discussion} %{{{1
+\subsection*{Scaffolder Description} %{{{2
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results and Discussion %%
-%%
-\section*{Results and Discussion}
-  \subsection*{Results sub-heading}
-    \subsubsection*{This is a sub-sub-heading}
-      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
-      pb at end of lines ensures correct paragraph spacing.\pb
-	  Text for this sub-sub-section \ldots
-    \subsubsection*{Another sub-sub-sub-heading}
-      Text for this sub-sub-section \ldots
+  * Join contigs together
+  * Add inserts
+  * Add unknown gap regions
+  * Specified in a YAML file format
 
-  \subsection*{Another results sub-heading}
-    Text for this sub-section \ldots
+\subsection*{Scaffold file format} %{{{2
 
-  \subsection*{Yet another results sub-heading}
-    Text for this sub-section.  More results \ldots
-% }}}
-% {{{ Conclusions
-%%%%%%%%%%%%%%%%%%%%%%
-\section*{Conclusions}
+  * Screen shot with example file
+  * Illustrate three types of region in the scaffold file
+  * Explanation of what each region does
+  * Expected output from each region
+
+\subsection*{Command line usage} %{{{2
+
+  * Illustrate running command line to generate output
+  * Demonstrate scaffold validate to test inserts overlap
+
+\subsection*{Programmatic usage of API} %{{{2
+
+  * Demonstrate how scaffolder can be manipulated programmatically
+  * Examine scaffolder data using a ruby program
+  * Example of extending sequence class with a new method \%GC content
+  * Create a new class to do something
+
+\subsection*{Installation} %{{{2
+
+  * Install through rubygems
+  * Make sure the bin is in the class path
+
+\section*{Conclusions} %{{{1
   Text for this section \ldots

### 2011-01-21 text/src/article.tex
Revise first two sections of implementation

--- 2011-01-19 text/src/article.tex
@@
 % }}}
-% {{{ Implementation
-\section*{Implementation}
+\section*{Implementation} %{{{1
+
+\subsection*{Code and Dependencies} %{{{2
 
 Scaffolder is written in Ruby \cite{ruby-lang} using version 1.8.7. Unit and
-ntegration tests use RSpec and Cucumber\cite{rspec}. The source code is
-documented in-line using Yard\cite{yard} and the Unix manual pages are
-generated using Ronn\cite{ronn}.\pb
+integration tests are implemented using the RSpec and Cucumber libraries
+respectively\cite{rspec}. The Scaffolder source code is documented using the
+Yard\cite{yard} library and the Unix manual pages are generated for the command
+line tools using the Ronn library\cite{ronn}. Programmatic manipulation of
+biological sequences is performed using BioRuby \cite{bioruby}. A full list of
+the scaffolder dependencies can be found in the Gemfile in the source code root
+directory.\pb
 
-The sequence scaffold is defined in YAML\cite{yaml}. YAML is a machine
-readable data format written using indentation of white space. YAML is designed
-to be human readable which is ideal for manual editing of a genome scaffold.
+\subsection*{Scaffold file syntax} %{{{2
+
+The orientation of contigs and inserts in a Scaffolder file is specified using
+YAML syntax\cite{yaml}. YAML was selected as a machine readable markup with
+parsers available for most programming languages. Since YAML is a common data
+interchange format this should make it easier for developers to computationally
+generate a genome scaffold and output to the Scaffolder format. Furthermore the
+YAML syntax can be validated using third party tools such as
+Kwalify\cite{kwalify}.\pb
+
+In addition to being machine readable the YAML syntax is designed to be easier
+to manually edit when compared with XML. YAML uses whitespace to delimit node
+in a directed acyclic graph compared with XML which uses angular braces (`$<$'
+and `$>$') with open and closing tags. By choosing a human readable syntax it is
+aimed that the scaffold should be easier to edit and change manually in simple
+text edit. A scenario which we anticipate is common in the on-going process of
+scaffolding a genome sequence. \pb
+
+The entries in the YAML scaffold are defined as a list. Each entry in the list
+corresponds to a region sequence in the final output sequence which follows
+the previous entry in the scaffold and precedes the following entry. Each
+attribute defined for each entry in the list allows specifications for the
+entry such as if any starting sequence should be clipped from the beginning or
+end, or if the sequence should be reversed. \pb
+
+INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.
+
+\subsection*{Parsing the YAML scaffold} %{{{2
+
 Each stretch of sequence in the genome scaffold is defined as a list entry in
@@
 the process of transferring attributes from the scaffold file to each region
-instance much simpler - each yaml attribute corresponds to a method call on
-the instantiated class. The scaffold file effectively becomes a DSL for
-populating the scaffold region instances while still maintaining the machine
-readable data qualities of YAML.\pb
+instance much simpler - each yaml attribute corresponds to a method call on the
+instantiated class. The scaffold file effectively becomes a DSL for populating
+the scaffold region instances while still maintaining the machine readable data
+qualities of YAML.\pb
 
@@
 region class should override this method to provide custom functionality in
-manipulating the source sequence. For example, the
-Scaffolder::Region::Sequence class overrides the sequence method add any
-sequence inserts specified in the scaffold file.\pb
+manipulating the source sequence. For example, the Scaffolder::Region::Sequence
+class overrides the sequence method add any sequence inserts specified in the
+scaffold file.\pb
 
-% }}}
-% {{{ Results and Discussion
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results and Discussion %%
-%%
-\section*{Results and Discussion}
+\section*{Results and Discussion} %{{{1
   \subsection*{Results sub-heading}

### 2011-01-21 text/src/article.tex
Write first section in results

--- 2011-01-19 text/src/article.tex
@@
 \section*{Results and Discussion} %{{{1
+
 \subsection*{Scaffolder Description} %{{{2
 
-  * Join contigs together
-  * Add inserts
-  * Add unknown gap regions
-  * Specified in a YAML file format
+Scaffolder is a tool aimed at build genome sequence scaffolds through simple to
+edit configuration files. Scaffolder allows contigs to be joined together
+through defintition in the scaffold file. Each contigs is stored in a fasta
+file, as provided by the standard output from an assembler such as
+newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder allows inserts to be
+added to scaffold so that gaps in the assembled data may be bridged and filled
+- a common process in finishing a genome sequence. Scaffolder also allows the
+specification of unknown regions in the scaffold. These may be regions where
+the sequence is unknown but the approximate size can be estimated from the
+sequencing of insert pairs or from estimating distances from contigs aligned to
+reference genomes.
 
-\subsection*{Scaffold file format} %{{{2
+Scaffolder provides a ruby API that allows the scaffold and designated
+sequences to be manipulated through a ruby programme. This allows scaffolder to
+be integrated into exiting genomics works flows such as using Rake. Scaffolder
+all provides a command line binary for the manipulation of the scaffold. This
+binary behaves as a standard unix command line interface returning appropriate
+exit statuses and providing manual pages for each of the commands.
+
+\subsection*{Defining a scaffold in a text file} %{{{2
+
+The genome scaffold is defined YAML format. YAML is an easy to edit data format
+that is both machine readable and easy to edit manually.
+
+INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD
 

### 2011-01-24 text/src/article.tex
Revise first section of results

--- 2011-01-19 text/src/article.tex
@@
 \section*{Results and Discussion} %{{{1
+
 \subsection*{Scaffolder Description} %{{{2
 
-  * Join contigs together
-  * Add inserts
-  * Add unknown gap regions
-  * Specified in a YAML file format
+Individual reads produced from nucleotide sequencing can be assembled into
+contiguous regions (``contigs'') of many overlapping individual reads. These
+reads may however not represent a complete genome sequence where all contigs do
+not result in a single complete sequence. The software outlined here
+``Scaffolder'' allows these contigs to be further joined together into
+a continuous single sequence. This selection, order and direction of contigs
+into a final draft genome is often described as the `scaffold'. Scaffolder
+allows the manual or computational specification of a genome scaffold by
+specifying which contig sequences are used in simple to edit configuration
+files.
 
-\subsection*{Scaffold file format} %{{{2
+The scaffold file specifies how contigs should be joined together in the draft
+genome sequence. The raw nucleotide sequence for each contig is stored in
+a fasta file, as might be produced from assembly of sequencing reads using
+programs such as newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder then
+the contig to be selected by specifying the name of the contig in the scaffold
+file. The scaffold file also allows a sub-sequence or the reverse complement of
+a contig to used. This therefore preserves the original sequence data but also
+allowing sequences to manipulated in the final genome build.
+
+The majority of a genome can be estimated from scaffolding assembled contigs
+into the most parsimonious build. There may however still be internal gaps in
+the contigs from mate pair assembly based on estimated mate pair insert size.
+Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
+and filled - a common process in finishing a genome sequence. Inserts may be
+treated the same way as a contig where the raw sequence, as may be produced by
+PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
+in the contig.
+
+Scaffolder allows the specification of unknown regions in the scaffold. These
+are areas of the genome where the nucleotide sequence is unknown but the
+approximate size region may be estimated from mate pair insert size or from
+mapping contigs to reference genome and estimating the size of the gap. These
+unknown regions allow a continuous draft sequence to be scaffolded but whilst
+still highlighting regions waiting to be resolved.
+
+\subsection*{Defining a scaffold in a text file} %{{{2
+
+The genome scaffold is defined YAML format. YAML is an easy to edit data format
+that is both machine readable and easy to edit manually.
+
+INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD
 
@@
   * Expected output from each region
+
+Scaffolder provides a ruby API that allows the scaffold and designated
+sequences to be manipulated through a ruby programme. This allows scaffolder to
+be integrated into exiting genomics works flows such as using Rake. Scaffolder
+all provides a command line binary for the manipulation of the scaffold. This
+binary behaves as a standard Unix command line interface returning appropriate
+exit statuses and providing manual pages for each of the commands. These two
+methods are outlined in the following two sections.
 
@@
 
-  * Install through rubygems
+  * Install through RubyGems
   * Make sure the bin is in the class path

### 2011-01-24 text/src/article.tex
Update references for first results section

@@
 a fasta file, as might be produced from assembly of sequencing reads using
-programs such as newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder then
-the contig to be selected by specifying the name of the contig in the scaffold
-file. The scaffold file also allows a sub-sequence or the reverse complement of
-a contig to used. This therefore preserves the original sequence data but also
-allowing sequences to manipulated in the final genome build.
+programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
+Scaffolder then the contig to be selected by specifying the name of the contig
+in the scaffold file. The scaffold file also allows a sub-sequence or the
+reverse complement of a contig to used. This therefore preserves the original
+sequence data but also allowing sequences to manipulated in the final genome
+build.
 

### 2011-01-25 text/src/article.tex
Fix broken header
Results section for scaffolder file format

@@
 specifying which contig sequences are used in simple to edit configuration
-files.
+files. \pb
 
@@
 sequence data but also allowing sequences to manipulated in the final genome
-build.
+build. \pb
 
@@
 PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
-in the contig.
+in the contig. \pb
 
@@
 unknown regions allow a continuous draft sequence to be scaffolded but whilst
-still highlighting regions waiting to be resolved.
+still highlighting regions waiting to be resolved. \pb
 
@@
 
-The genome scaffold is defined YAML format. YAML is an easy to edit data format
-that is both machine readable and easy to edit manually.
+The genome scaffold is wriiten in YAML format. YAML is a data format designed
+to be both machine readable and easy to manually edit. An example of this
+format illustrating all the features of scaffolder is shown in Figure 1. \pb
 
-INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD
+\subsubsection*{Simple sequence region} %{{{3
 
-  * Screen shot with example file
-  * Illustrate three types of region in the scaffold file
-  * Explanation of what each region does
-  * Expected output from each region
+The first line of the file begins with three dashes. This is required to
+signify the start of a YAML document. Lines 2 begins with a dash character `-'
+which denotes the start of an entry in the sequence scaffold, in this case the
+first entry. Line 3 is indented by two spaces which therefore means this line
+contains an entry type. The ``sequence'' tag is used indicate a sequence is
+inserted at this region in the scaffold. The following line contain is
+indented by two spaces and indicates that this an attribute of the sequence
+entry. In this case the ``source'' keyword is used to indicate used which
+sequence should be inserted into the scaffold. This sequence is indentified by
+the first word of the fasta header of the selected sequence. This indicates
+a simple sequence region added to the scaffold. \pb
+
+\subsubsection*{Unresolved seqeunce region} %{{{3
+
+The following line has no indentation and begins with dash indicating the next
+entry in the scaffold. The following entry is indentified by the ``unresolved''
+tag. This identifies a region of unknown nucleotide sequence but is included
+because the approximate size of the region is known. The following line defines
+that the size of this unknown region in the genome scaffold should be 20
+nucleotides. This therefore results in the region being filled with 20 `N'
+characters in draft sequence. \pb
+
+\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3
+
+The following entry is another sequence using the sequence from 'sequence2' in
+the scaffold file. This sequence entry has three further attributes in addition
+to defining the source sequence. The first two are start and stop. These are
+used to trim the final sequence inserted into the scaffold. In this case the
+first two nucleotides of the sequence are removed so that that sequence starts
+at position 3 of the original raw sequence. Similarly the sequence after
+position is trimmed away after position 100. These sequence is also reverse
+complemented using the ``reverse'' tag. Reverse complementation is performed
+after the sequence has been trimmed to the specified coordinates. \pb
+
+The final attribute for this sequence entry defines inserts to be added to the
+sequence. These are additional regions of sequence that can be added to
+a sequence entry to alter the existing sequence such to fill gaps or correct
+sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
+is then listed similarly to the entries in the scaffold: beginning with a dash
+on one line then indented by an additional two spaces on the following lines.
+\pb
+
+Inserts are listed similarly to sequence entries where reverse, start and stop
+attributes can be used to trim and reverse the sequence. The `source' tag is
+used to indentify which fasta sequence entry should be used as the insert.  The
+insert has two additional tags ``open'' and ``close'' which determine where the
+insert is added to the hosting sequence. These coordinates are inclusive
+replaceing the coordinate nucleotides. The open and close positions need not be
+the same size as the insert size. Any difference in size to the replaced
+nucleotides will result in the length on the encoding sequence being adjusted.
+\pb
+
+The next insert entry shows an insert with only the open position specified.
+This is to illustrate that only the one of the `open' or `close' keywords is
+sufficient to add an insert. If only one of these words is used then the
+opposing insert position is calculated from the length on the insert sequence
+in the fasta file. This feature is designed for using inserts to close gaps
+where the length of the insert may not completely fill the gap but can still be
+used to fill a portion of the gap sequence. \pb
+
+\subsubsection*{Scaffolder file attribute processing} %{{{3
+
+Each scaffolder sequence entry can have multiple attributes which applied in
+different order can result in a different sequence. For instance reverse
+complementing the sequence then adding inserts will have a different effect
+compared with performing these actions in the reverse order. Scaffolder
+therefore follows the following protocol for processing each entry in the
+scaffold file.
+
+\paragraph{Process inserts}
+
+\subsection*{Scaffolder usage} %{{{2
 
@@
 exit statuses and providing manual pages for each of the commands. These two
-methods are outlined in the following two sections.
+methods are outlined in the following two sections. \pb
 
-\subsection*{Command line usage} %{{{2
+\subsubsection*{Installation} %{{{3
+
+  * Install through RubyGems
+
+  * Make sure the bin is in the class path
+
+\subsubsection*{Command line interface} %{{{3
 
@@
 
-\subsection*{Programmatic usage of API} %{{{2
+\subsubsection*{Accessible programmatic interface} %{{{3
 
@@
 
-\subsection*{Installation} %{{{2
-
-  * Install through RubyGems
-  * Make sure the bin is in the class path
+\subsection*{Discussion of scaffolder approach} %{{{2
 
 \section*{Conclusions} %{{{1
-  Text for this section \ldots
-% }}}
-% {{{ Availability and Requirements
+
+Text for this section \ldots }}} {{{ Availability and Requirements
+
 \section*{Availability and Requirements}
@@
   Text for this section \ldots
-%}}}
-% {{{ Bibliography
+%}}} {{{1 Bibliography
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
@@
   \bibliography{article} }     % Bibliography file (usually '*.bib' )
-%}}}
-% {{{ Figures
+
 \ifthenelse{\boolean{publ}}{\end{multicols}}{}
 
-\section*{Figures}
+\section*{Figures} %{{{1
+
   \subsection*{Figure 1 - Sample figure title}
-      A short description of the figure content
-      should go here.
 
-  \subsection*{Figure 2 - Sample figure title}
-      Figure legend text.
-% }}}
-% {{{ Tables
-\section*{Tables}
+  A short description of the figure content should go here.
+
+\section*{Tables} %{{{1
+
   \subsection*{Table 1 - Sample table title}
@@
     still be described here.
-% }}}
-% {{{ Additional Files
-\section*{Additional Files}
+
+\section*{Additional Files} %{{{1
+
   \subsection*{Additional file 1 --- Sample additional file title}
@@
 \end{bmcformat}
+
 \end{document}
-% }}}

### 2011-01-26 text/src/article.tex
Revision of first half of implementation section

--- 2011-01-24 text/src/article.tex
@@
 \section*{Background}
- Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
+ Text for this section.
 % }}}
-% {{{ Implementation
+\section*{Implementation} %{{{1
+
+\subsection*{Code and Dependencies} %{{{2
+
+Scaffolder is written in the Ruby programming language, version 1.8.7
+\cite{ruby-lang}. Unit tests were implemented during development to maintain
+functionality of individual elements of source code. Integration tests were
+also written to test the scaffold software interface as a whole. These unit and
+integration tests were written using the RSpec and Cucumber libraries
+respectively\cite{rspec}.\pb
+
+The Scaffolder source code is documented using the Yard library\cite{yard}.
+Unix manual pages for the command line interface were generated using the Ronn
+library\cite{ronn}. The manipulation of biological sequences in Scaffolder is
+performed using the BioRuby library\cite{bioruby}. A full list of the software
+dependencies in scaffolder can be found in the Gemfile in the root of source
+code directory.\pb
+
+\subsection*{Scaffold file syntax} %{{{2
+
+The specification of the nucleotide sequences comprising the genome scaffold is
+written using YAML \cite{yaml}. YAML is a data format using whitespace and
+identation to produce a standardised machine readable structure. As a commonly
+accepted data format third-party developers have the option to generate
+a genome scaffold using their programming language of choice. As a standardised
+format YAML formatted scaffold files can checked for syntax validity using
+third party tools such as Kwalify\cite{kwalify}. \pb
+
+In addition to being machine readable it also important that a genome scaffold
+file also be easy to edit manually. As such this was the second reason the YAML
+syntax was selected for the scaffold file data format. YAML has a concise and
+compact syntax and when compared with formats such as extensible markup
+language (XML) which require angular braces (`$<$' and `$>$') and open and
+closing tags to delimite each data field.
+
+By choosing a human readable syntax it is aimed that the scaffold should be
+easier to edit and change manually in simple text edit.  A scenario which we
+anticipate is common in the on-going process of scaffolding a genome sequence.
+\pb
+
+The entries in the YAML scaffold are defined as a list. Each entry in the list
+corresponds to a region sequence in the final output sequence which follows the
+previous entry in the scaffold and precedes the following entry.  Each
+attribute defined for each entry in the list allows specifications for the
+entry such as if any starting sequence should be clipped from the beginning or
+end, or if the sequence should be reversed. \pb
+
+INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.
+
+\subsection*{Parsing the YAML scaffold} %{{{2
+
+Each stretch of sequence in the genome scaffold is defined as a list entry in
+the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
+each scaffold entry to create an instance of the Scaffolder::Region class.\pb
+
+The Region class is inherited by all types of sequence in the scaffold. The
+Region provides helper methods common to the tasks in manipulating scaffold
+sequence, subsequencing and reversing. In particular the region class has
+a class annotation method\cite{class-annotations} which allows attributes in
+the scaffold method to be defined in the class. This negates having to write
+a custom method to deal with each scaffold variable.  Furthermore this makes
+the process of transferring attributes from the scaffold file to each region
+instance much simpler - each yaml attribute corresponds to a method call on the
+instantiated class. The scaffold file effectively becomes a DSL for populating
+the scaffold region instances while still maintaining the machine readable data
+qualities of YAML.\pb
+
+The region class provides a sequence hook method. This method returns by
+default the sequence corresponding to the fasta sequence defined by the source
+tag which is then sublclassed or reversed if necessary. Subclasses of the
+region class should override this method to provide custom functionality in
+manipulating the source sequence. For example, the Scaffolder::Region::Sequence
+class overrides the sequence method add any sequence inserts specified in the
+scaffold file.\pb
+
+\section*{Results and Discussion} %{{{1
+  \subsection*{Results sub-heading}
+    \subsubsection*{This is a sub-sub-heading}
+      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
+      pb at end of lines ensures correct paragraph spacing.\pb
+	  Text for this sub-sub-section \ldots
+    \subsubsection*{Another sub-sub-sub-heading}
+      Text for this sub-sub-section \ldots
+
+  \subsection*{Another results sub-heading}
+    Text for this sub-section \ldots
+
+  \subsection*{Yet another results sub-heading}
+    Text for this sub-section.  More results \ldots
+% }}}
+% {{{ Conclusions
 %%%%%%%%%%%%%%%%%%%%%%
-\section*{Implementation}
-  Text for this section \ldots
-% }}}
-\section*{Results and Discussion} %{{{1
-
-\subsection*{Scaffolder Description} %{{{2
-
-Individual reads produced from nucleotide sequencing can be assembled into
-contiguous regions (``contigs'') of many overlapping individual reads. These
-reads may however not represent a complete genome sequence where all contigs do
-not result in a single complete sequence. The software outlined here
-``Scaffolder'' allows these contigs to be further joined together into
-a continuous single sequence. This selection, order and direction of contigs
-into a final draft genome is often described as the `scaffold'. Scaffolder
-allows the manual or computational specification of a genome scaffold by
-specifying which contig sequences are used in simple to edit configuration
-files.
-
-The scaffold file specifies how contigs should be joined together in the draft
-genome sequence. The raw nucleotide sequence for each contig is stored in
-a fasta file, as might be produced from assembly of sequencing reads using
-programs such as newbler\cite{newbler} or abyss\cite{abyss}. Scaffolder then
-the contig to be selected by specifying the name of the contig in the scaffold
-file. The scaffold file also allows a sub-sequence or the reverse complement of
-a contig to used. This therefore preserves the original sequence data but also
-allowing sequences to manipulated in the final genome build.
-
-The majority of a genome can be estimated from scaffolding assembled contigs
-into the most parsimonious build. There may however still be internal gaps in
-the contigs from mate pair assembly based on estimated mate pair insert size.
-Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
-and filled - a common process in finishing a genome sequence. Inserts may be
-treated the same way as a contig where the raw sequence, as may be produced by
-PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
-in the contig.
-
-Scaffolder allows the specification of unknown regions in the scaffold. These
-are areas of the genome where the nucleotide sequence is unknown but the
-approximate size region may be estimated from mate pair insert size or from
-mapping contigs to reference genome and estimating the size of the gap. These
-unknown regions allow a continuous draft sequence to be scaffolded but whilst
-still highlighting regions waiting to be resolved.
-
-\subsection*{Defining a scaffold in a text file} %{{{2
-
-The genome scaffold is defined YAML format. YAML is an easy to edit data format
-that is both machine readable and easy to edit manually.
-
-INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD
-
-  * Screen shot with example file
-  * Illustrate three types of region in the scaffold file
-  * Explanation of what each region does
-  * Expected output from each region
-
-Scaffolder provides a ruby API that allows the scaffold and designated
-sequences to be manipulated through a ruby programme. This allows scaffolder to
-be integrated into exiting genomics works flows such as using Rake. Scaffolder
-all provides a command line binary for the manipulation of the scaffold. This
-binary behaves as a standard Unix command line interface returning appropriate
-exit statuses and providing manual pages for each of the commands. These two
-methods are outlined in the following two sections.
-
-\subsection*{Command line usage} %{{{2
-
-  * Illustrate running command line to generate output
-  * Demonstrate scaffold validate to test inserts overlap
-
-\subsection*{Programmatic usage of API} %{{{2
-
-  * Demonstrate how scaffolder can be manipulated programmatically
-  * Examine scaffolder data using a ruby program
-  * Example of extending sequence class with a new method \%GC content
-  * Create a new class to do something
-
-\subsection*{Installation} %{{{2
-
-  * Install through RubyGems
-  * Make sure the bin is in the class path
-
-\section*{Conclusions} %{{{1
+\section*{Conclusions}
   Text for this section \ldots

### 2011-01-27 text/src/article.tex
Revise second section of implementation

@@
 
-The specification of the nucleotide sequences comprising the genome scaffold is
-written using YAML \cite{yaml}. YAML is a data format using whitespace and
-identation to produce a standardised machine readable structure. As a commonly
-accepted data format third-party developers have the option to generate
-a genome scaffold using their programming language of choice. As a standardised
-format YAML formatted scaffold files can checked for syntax validity using
-third party tools such as Kwalify\cite{kwalify}. \pb
+The choice of nucleotide sequences comprising the genome scaffold is specified
+in YAML \cite{yaml}. YAML is a language format for data using whitespace and
+identation to produce a machine readable structure. As a standardised data
+format third-party developers have the option to generate a genome scaffold
+using any programming language for which a YAML parser exists. Furthermore YAML
+formatted scaffold files can be validated using third party tools such as
+Kwalify\cite{kwalify}.\pb
 
-In addition to being machine readable it also important that a genome scaffold
-file also be easy to edit manually. As such this was the second reason the YAML
-syntax was selected for the scaffold file data format. YAML has a concise and
-compact syntax and when compared with formats such as extensible markup
-language (XML) which require angular braces (`$<$' and `$>$') and open and
-closing tags to delimite each data field.
+The initial assembly of sequencing reads may result in an incomplete genome
+build. Comparison of contigs to reference genomes and the generation of
+additional sequence from PCR translates to genome finishing as an ongoing
+process. Therefore the scaffold file should be simple to edit manually in
+additon to being computationally tractable. This requirement for the scaffold
+files to be modifable in a simple simple text editor is the second reason YAML
+was selected as the format for scaffold files.\pb
 
-By choosing a human readable syntax it is aimed that the scaffold should be
-easier to edit and change manually in simple text edit.  A scenario which we
-anticipate is common in the on-going process of scaffolding a genome sequence.
-\pb
+\subsection*{Parsing the YAML scaffold} %{{{2
 
@@
 entry such as if any starting sequence should be clipped from the beginning or
-end, or if the sequence should be reversed. \pb
+end, or if the sequence should be reversed.\pb
 
 INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.
-
-\subsection*{Parsing the YAML scaffold} %{{{2
 

### 2011-01-27 text/src/article.tex
Revise last section of implementation

@@
 The choice of nucleotide sequences comprising the genome scaffold is specified
-in YAML \cite{yaml}. YAML is a language format for data using whitespace and
-identation to produce a machine readable structure. As a standardised data
+in YAML \cite{yaml}. YAML is a language format for data using white-space and
+indentation to produce a machine readable structure. As a standardised data
 format third-party developers have the option to generate a genome scaffold
@@
 process. Therefore the scaffold file should be simple to edit manually in
-additon to being computationally tractable. This requirement for the scaffold
-files to be modifable in a simple simple text editor is the second reason YAML
+addition to being computationally tractable. This requirement for the scaffold
+files to be modifiable in a simple simple text editor is the second reason YAML
 was selected as the format for scaffold files.\pb
 
-\subsection*{Parsing the YAML scaffold} %{{{2
+\subsection*{Processing the scaffold file} %{{{2
 
-The entries in the YAML scaffold are defined as a list. Each entry in the list
-corresponds to a region sequence in the final output sequence which follows the
-previous entry in the scaffold and precedes the following entry.  Each
-attribute defined for each entry in the list allows specifications for the
-entry such as if any starting sequence should be clipped from the beginning or
-end, or if the sequence should be reversed.\pb
+The form of the scaffold file is a list of entries. Each entry corresponds to
+a length of sequence in the the draft genome sequence. When generating the
+genome sequence from the scaffold file each entry is processed in order. Each
+of the nucleotide sequence specified by the scaffold file entries are
+concatenated together to produce the draft genome sequence. Each entry in the
+scaffold file may have attributes which define whether a sub-sequence or the
+reverse complement of the sequence should be used is the scaffold. The types of
+attributes are outlined in the Results and Discussion section below.\pb
 
-INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.
+As the scaffold file is processed a corresponding ruby object instance is
+created to correspond to an entry in the scaffold file. as an instance of
+a subclass of Scaffolder::Region.  Scaffolder::Region is the super class for
+all region entries in the scaffold.  This Region class provides the helper
+methods for sub-sequencing and reversing complementing a sequence.  The Region
+class also provides a class annotation method\cite{class-annotations} which
+simplifies defining the attributes for each scaffold entry. Possible attributes
+for scaffold entries can be added by using this helper method. \pb
 
-Each stretch of sequence in the genome scaffold is defined as a list entry in
-the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
-each scaffold entry to create an instance of the Scaffolder::Region class.\pb
-
-The Region class is inherited by all types of sequence in the scaffold. The
-Region provides helper methods common to the tasks in manipulating scaffold
-sequence, subsequencing and reversing. In particular the region class has
-a class annotation method\cite{class-annotations} which allows attributes in
-the scaffold method to be defined in the class. This negates having to write
-a custom method to deal with each scaffold variable.  Furthermore this makes
-the process of transferring attributes from the scaffold file to each region
-instance much simpler - each yaml attribute corresponds to a method call on the
-instantiated class. The scaffold file effectively becomes a DSL for populating
-the scaffold region instances while still maintaining the machine readable data
-qualities of YAML.\pb
-
-The region class provides a sequence hook method. This method returns by
-default the sequence corresponding to the fasta sequence defined by the source
-tag which is then sublclassed or reversed if necessary. Subclasses of the
-region class should override this method to provide custom functionality in
-manipulating the source sequence. For example, the Scaffolder::Region::Sequence
-class overrides the sequence method add any sequence inserts specified in the
-scaffold file.\pb
+An noteworthy method of the Scaffolder::Region class is the ``raw_sequence''
+method. In the Scaffolder::Region class method returns by the sequence defined
+by the ``source'' tag in the scaffold file. This sequence is then sub-sequenced
+or reversed complemented if necessary. Subclasses of the Scaffolder::Region
+class override this method to additional manipulate the nucleotide sequence.
+For example, the Scaffolder::Region::Sequence class overrides the sequence
+method to insert any specified inserts specified in the scaffold file before
+this sequence is returned up the method chain.\pb
 

### 2011-01-27 text/src/article.tex
Remove two paragraphs from implementation section.

@@
 
-\subsection*{Processing the scaffold file} %{{{2
-
 The form of the scaffold file is a list of entries. Each entry corresponds to
@@
 
-As the scaffold file is processed a corresponding ruby object instance is
-created to correspond to an entry in the scaffold file. as an instance of
-a subclass of Scaffolder::Region.  Scaffolder::Region is the super class for
-all region entries in the scaffold.  This Region class provides the helper
-methods for sub-sequencing and reversing complementing a sequence.  The Region
-class also provides a class annotation method\cite{class-annotations} which
-simplifies defining the attributes for each scaffold entry. Possible attributes
-for scaffold entries can be added by using this helper method. \pb
-
-An noteworthy method of the Scaffolder::Region class is the ``raw_sequence''
-method. In the Scaffolder::Region class method returns by the sequence defined
-by the ``source'' tag in the scaffold file. This sequence is then sub-sequenced
-or reversed complemented if necessary. Subclasses of the Scaffolder::Region
-class override this method to additional manipulate the nucleotide sequence.
-For example, the Scaffolder::Region::Sequence class overrides the sequence
-method to insert any specified inserts specified in the scaffold file before
-this sequence is returned up the method chain.\pb
-
-\section*{Results and Discussion} %{{{1
-  \subsection*{Results sub-heading}
-    \subsubsection*{This is a sub-sub-heading}
-      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
-      pb at end of lines ensures correct paragraph spacing.\pb
-	  Text for this sub-sub-section \ldots
-    \subsubsection*{Another sub-sub-sub-heading}
-      Text for this sub-sub-section \ldots
+\section*{Results and Discussion} %{{{1 \subsection*{Results sub-heading}
+\subsubsection*{This is a sub-sub-heading} Sub-sub-sub-headings are made with
+the \textsl{\\subsubsection} command. \pb pb at end of lines ensures correct
+paragraph spacing.\pb Text for this sub-sub-section \ldots
+\subsubsection*{Another sub-sub-sub-heading} Text for this sub-sub-section
+\ldots
 

### 2011-01-27 text/src/article.tex
Results section for installation

@@
 \section*{Background}
- Text for this section.
+ Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
 % }}}
-\section*{Implementation} %{{{1
-
-\subsection*{Code and Dependencies} %{{{2
-
-Scaffolder is written in the Ruby programming language, version 1.8.7
-\cite{ruby-lang}. Unit tests were implemented during development to maintain
-functionality of individual elements of source code. Integration tests were
-also written to test the scaffold software interface as a whole. These unit and
-integration tests were written using the RSpec and Cucumber libraries
-respectively\cite{rspec}.\pb
-
-The Scaffolder source code is documented using the Yard library\cite{yard}.
-Unix manual pages for the command line interface were generated using the Ronn
-library\cite{ronn}. The manipulation of biological sequences in Scaffolder is
-performed using the BioRuby library\cite{bioruby}. A full list of the software
-dependencies in scaffolder can be found in the Gemfile in the root of source
-code directory.\pb
-
-\subsection*{Scaffold file syntax} %{{{2
-
-The choice of nucleotide sequences comprising the genome scaffold is specified
-in YAML \cite{yaml}. YAML is a language format for data using white-space and
-indentation to produce a machine readable structure. As a standardised data
-format third-party developers have the option to generate a genome scaffold
-using any programming language for which a YAML parser exists. Furthermore YAML
-formatted scaffold files can be validated using third party tools such as
-Kwalify\cite{kwalify}.\pb
-
-The initial assembly of sequencing reads may result in an incomplete genome
-build. Comparison of contigs to reference genomes and the generation of
-additional sequence from PCR translates to genome finishing as an ongoing
-process. Therefore the scaffold file should be simple to edit manually in
-addition to being computationally tractable. This requirement for the scaffold
-files to be modifiable in a simple simple text editor is the second reason YAML
-was selected as the format for scaffold files.\pb
-
-The form of the scaffold file is a list of entries. Each entry corresponds to
-a length of sequence in the the draft genome sequence. When generating the
-genome sequence from the scaffold file each entry is processed in order. Each
-of the nucleotide sequence specified by the scaffold file entries are
-concatenated together to produce the draft genome sequence. Each entry in the
-scaffold file may have attributes which define whether a sub-sequence or the
-reverse complement of the sequence should be used is the scaffold. The types of
-attributes are outlined in the Results and Discussion section below.\pb
-
-\section*{Results and Discussion} %{{{1 \subsection*{Results sub-heading}
-\subsubsection*{This is a sub-sub-heading} Sub-sub-sub-headings are made with
-the \textsl{\\subsubsection} command. \pb pb at end of lines ensures correct
-paragraph spacing.\pb Text for this sub-sub-section \ldots
-\subsubsection*{Another sub-sub-sub-heading} Text for this sub-sub-section
-\ldots
-
-  \subsection*{Another results sub-heading}
-    Text for this sub-section \ldots
-
-  \subsection*{Yet another results sub-heading}
-    Text for this sub-section.  More results \ldots
-% }}}
-% {{{ Conclusions
+% {{{ Implementation
 %%%%%%%%%%%%%%%%%%%%%%
-\section*{Conclusions}
+\section*{Implementation}
   Text for this section \ldots
 % }}}
-% {{{ Availability and Requirements
+\section*{Results and Discussion} %{{{1
+
+\subsection*{Scaffolder Description} %{{{2
+
+Individual reads produced from nucleotide sequencing can be assembled into
+contiguous regions (``contigs'') of many overlapping individual reads. These
+reads may however not represent a complete genome sequence where all contigs do
+not result in a single complete sequence. The software outlined here
+``Scaffolder'' allows these contigs to be further joined together into
+a continuous single sequence. This selection, order and direction of contigs
+into a final draft genome is often described as the `scaffold'. Scaffolder
+allows the manual or computational specification of a genome scaffold by
+specifying which contig sequences are used in simple to edit configuration
+files. \pb
+
+The scaffold file specifies how contigs should be joined together in the draft
+genome sequence. The raw nucleotide sequence for each contig is stored in
+a fasta file, as might be produced from assembly of sequencing reads using
+programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
+Scaffolder then the contig to be selected by specifying the name of the contig
+in the scaffold file. The scaffold file also allows a sub-sequence or the
+reverse complement of a contig to used. This therefore preserves the original
+sequence data but also allowing sequences to manipulated in the final genome
+build. \pb
+
+The majority of a genome can be estimated from scaffolding assembled contigs
+into the most parsimonious build. There may however still be internal gaps in
+the contigs from mate pair assembly based on estimated mate pair insert size.
+Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
+and filled - a common process in finishing a genome sequence. Inserts may be
+treated the same way as a contig where the raw sequence, as may be produced by
+PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
+in the contig. \pb
+
+Scaffolder allows the specification of unknown regions in the scaffold. These
+are areas of the genome where the nucleotide sequence is unknown but the
+approximate size region may be estimated from mate pair insert size or from
+mapping contigs to reference genome and estimating the size of the gap. These
+unknown regions allow a continuous draft sequence to be scaffolded but whilst
+still highlighting regions waiting to be resolved. \pb
+
+\subsection*{Defining a scaffold in a text file} %{{{2
+
+The genome scaffold is wriiten in YAML format. YAML is a data format designed
+to be both machine readable and easy to manually edit. An example of this
+format illustrating all the features of scaffolder is shown in Figure 1. \pb
+
+\subsubsection*{Simple sequence region} %{{{3
+
+The first line of the file begins with three dashes. This is required to
+signify the start of a YAML document. Lines 2 begins with a dash character `-'
+which denotes the start of an entry in the sequence scaffold, in this case the
+first entry. Line 3 is indented by two spaces which therefore means this line
+contains an entry type. The ``sequence'' tag is used indicate a sequence is
+inserted at this region in the scaffold. The following line contain is
+indented by two spaces and indicates that this an attribute of the sequence
+entry. In this case the ``source'' keyword is used to indicate used which
+sequence should be inserted into the scaffold. This sequence is indentified by
+the first word of the fasta header of the selected sequence. This indicates
+a simple sequence region added to the scaffold. \pb
+
+\subsubsection*{Unresolved seqeunce region} %{{{3
+
+The following line has no indentation and begins with dash indicating the next
+entry in the scaffold. The following entry is indentified by the ``unresolved''
+tag. This identifies a region of unknown nucleotide sequence but is included
+because the approximate size of the region is known. The following line defines
+that the size of this unknown region in the genome scaffold should be 20
+nucleotides. This therefore results in the region being filled with 20 `N'
+characters in draft sequence. \pb
+
+\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3
+
+The following entry is another sequence using the sequence from 'sequence2' in
+the scaffold file. This sequence entry has three further attributes in addition
+to defining the source sequence. The first two are start and stop. These are
+used to trim the final sequence inserted into the scaffold. In this case the
+first two nucleotides of the sequence are removed so that that sequence starts
+at position 3 of the original raw sequence. Similarly the sequence after
+position is trimmed away after position 100. These sequence is also reverse
+complemented using the ``reverse'' tag. Reverse complementation is performed
+after the sequence has been trimmed to the specified coordinates. \pb
+
+The final attribute for this sequence entry defines inserts to be added to the
+sequence. These are additional regions of sequence that can be added to
+a sequence entry to alter the existing sequence such to fill gaps or correct
+sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
+is then listed similarly to the entries in the scaffold: beginning with a dash
+on one line then indented by an additional two spaces on the following lines.
+\pb
+
+Inserts are listed similarly to sequence entries where reverse, start and stop
+attributes can be used to trim and reverse the sequence. The `source' tag is
+used to indentify which fasta sequence entry should be used as the insert.  The
+insert has two additional tags ``open'' and ``close'' which determine where the
+insert is added to the hosting sequence. These coordinates are inclusive
+replaceing the coordinate nucleotides. The open and close positions need not be
+the same size as the insert size. Any difference in size to the replaced
+nucleotides will result in the length on the encoding sequence being adjusted.
+\pb
+
+The next insert entry shows an insert with only the open position specified.
+This is to illustrate that only the one of the `open' or `close' keywords is
+sufficient to add an insert. If only one of these words is used then the
+opposing insert position is calculated from the length on the insert sequence
+in the fasta file. This feature is designed for using inserts to close gaps
+where the length of the insert may not completely fill the gap but can still be
+used to fill a portion of the gap sequence. \pb
+
+\subsubsection*{Scaffolder file attribute processing} %{{{3
+
+Each scaffolder sequence entry can have multiple attributes which applied in
+different order can result in a different sequence. For instance reverse
+complementing the sequence then adding inserts will have a different effect
+compared with performing these actions in the reverse order. Scaffolder
+therefore follows the following protocol for processing each entry in the
+scaffold file. \pb
+
+\paragraph{Reverse sort inserts:} If the entry is a sequence and has sequence
+inserts these are reverse sorted by insert close position. The insert with the
+largest close position is first and and the smallest close position is last.
+This is done to preserve the the insert open/close coordinates as they are
+added to the scaffold. \pb
+
+\paragraph{Update sequence with insert:} Each insert is added to the sequence
+replacing the region specified by the open an close position with the specified
+source sequence.\pb
+
+\paragraph{Update sequence start/stop co-ordinates:} If the insert added to the
+sequence is a different length than the region replaced the start and stop
+co-ordinates of the sequence are correspondingly updated if necessary. E.g. if
+a 7 bp insert filled a bp region this would result in a 2bp difference in the
+length of the sequence. The stop coordinated would be updated to reflect this.
+\pb
+
+\paragraph{Sub-sequence selected:} If the sequence is has start or stop
+sequences specified these are used to trip the sequence to these co-ordinates.
+\pb
+
+\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyworkd the sequence is reverse complemented.
+\pb
+
+\subsection*{Scaffolder usage} %{{{2
+
+Scaffolder provides a ruby API that allows the scaffold and designated
+sequences to be manipulated through a ruby programme. This allows scaffolder to
+be integrated into exiting genomics works flows such as using Rake.  Scaffolder
+all provides a command line binary for the manipulation of the scaffold. This
+binary behaves as a standard Unix command line interface returning appropriate
+exit statuses and providing manual pages for each of the commands. These two
+methods are outlined in the following three sections. \pb
+
+\subsubsection*{Installation} %{{{3
+
+Scaffolder relies on both the ruby programming language and the RubyGems
+package management system. These are commonly installed on recent versions of
+Mac OSX and linux distrubtions. The installation of these can be tested on the
+command line by typing the following commands. \pb
+
+\begin{verbatim}
+ruby -v
+gem -v
+\end{verbatim}
+
+If both these commands return without error this indicates the necessary
+requirements to install scaffolder. The scaffolder API and command line tools
+can then be installed with the follow command.
+
+\begin{verbatim}
+gem install scaffolder scaffolder-tools
+\end{verbatim}
+
+Finally to ensure that the command line tools are accessible these the
+RubyGems binary directory should be in the command line path. This can be done
+by adding the following to the ~/.bashrc file or other corresponding shell
+configure file. Assuming that ~/.gem is the install path for RubyGems.
+
+\begin{verbatim}
+export PATH=~/.gem/bin:$PATH
+\end{verbatim}
+
+\subsubsection*{Command line interface} %{{{3
+
+  * Illustrate running command line to generate output
+  * Demonstrate scaffold validate to test inserts overlap
+
+\subsubsection*{Accessible programmatic interface} %{{{3
+
+  * Demonstrate how scaffolder can be manipulated programmatically
+  * Examine scaffolder data using a ruby program
+  * Example of extending sequence class with a new method \%GC content
+  * Create a new class to do something
+
+\subsection*{Discussion of scaffolder approach} %{{{2
+
+\section*{Conclusions} %{{{1
+
+Text for this section \ldots
+
 \section*{Availability and Requirements}
@@
 % }}}
-% {{{ Availability and Requirements
-\section*{List of Abbreviations Used}
-
-  \begin{description}
-    \item[YAML:] YAML Ain't Markup Language\cite{yaml}
-  \end{description}
-
-% }}}
 % {{{ Author Contributions

### 2011-01-28 text/src/article.tex
Fix vim fold markder
Minor corrections

--- 2011-01-27 text/src/article.tex
@@
 concatenated together to produce the draft genome sequence. Each entry in the
-scaffold file may have attributes which define whether a sub-sequence or the
-reverse complement of the sequence should be used is the scaffold. The types of
-attributes are outlined in the Results and Discussion section below.\pb
+scaffold file may have attributes that define whether a sub-sequence or the
+reverse complement of the sequence should be used. The types of attributes are
+outlined in the Results and Discussion section below.\pb
 

### 2011-01-28 text/src/article.tex
Command line interface section of results

--- 2011-01-27 text/src/article.tex
@@
 
-  * Illustrate running command line to generate output
-  * Demonstrate scaffold validate to test inserts overlap
+Scaffolder provides a command line interface to validate and build a genome
+sequence from a scaffold. Assuming scaffolder has been installed and the
+binary is available in the command line path the following command can be used
+to invoke scaffolder. This will list the available commands for running
+scaffolder.
 
-\subsubsection*{Accessible programmatic interface} %{{{3
+\begin{verbatim}
+$ scaffolder
+\end{verbatim}
+
+Help on each command can be found by type `help' followed by the name of the
+command. This will return the Unix manual page for the command of interest.
+For example the following command will return the manual page for sequence
+listing the available options for running the sequence command.
+
+\begin{verbatim}
+$ scaffolder help sequence
+\end{verbatim}
+
+To the generate the fasta sequence from a scaffold the sequence command should
+be used. This should be passed the name of the file with containing the YAML
+formatted scaffold and the corresponding fasta file of sequences. This will
+then return the complete drafted sequence to the standard out.
+
+\begin{verbatim}
+$ scaffolder sequence scaffold.yml sequences.fna
+\end{verbatim}
+
+Scaffolder also provides a function to test the scaffold for overlapping
+insert sequence co-ordinates. As described in the section above overlapping
+inserts will result in unexpected behaviour as the coordinates for the first
+insert will change the expected insert location for the second insert. The
+validate command will check the scaffold for any overlapping inserts and
+return them as a YAML formatted list describing each one.
+
+\begin{verbatim}
+$ scaffolder validate scaffold.yml sequences.fna
+\end{verbatim}
+
+\subsubsection*{Application programming interface} %{{{3
 
@@
 
-\section*{Availability and Requirements}
+\section*{Availability and Requirements} %{{{1
 
@@
   Text for this section \ldots
-%}}} {{{1 Bibliography
+%{{{1 Bibliography
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}

### 2011-01-28 text/src/article.tex
Outline for scaffolder discussion section

@@
 \section*{Results and Discussion} %{{{1
+
+Pros
+  * Reproducable steps in building a scaffold
+  * Extensive documentation. Should be easy to use as possible.
+  * Scaffold file allows meaningful construction of a scaffold
+  * See of each sequence in building final draft genome
+  * Keeping genome sequence in a scaffold file means it is simpler to edit after the fact.
+  * Versionable plain text. Easy to store and see diffs
+  * Common format for scaffold file means it can be easily manipulated programmatically.
+  * Command line interface for using as is
+  * Ruby API allowing more complex manipulation of the scaffold
+  * Simple to install.  Automatically installed. No manual installation.
 

### 2011-01-28 text/src/article.tex
Subsection on using the scaffolder API

@@
 \section*{Results and Discussion} %{{{1
-
-Pros
-  * Reproducable steps in building a scaffold
-  * Extensive documentation. Should be easy to use as possible.
-  * Scaffold file allows meaningful construction of a scaffold
-  * See of each sequence in building final draft genome
-  * Keeping genome sequence in a scaffold file means it is simpler to edit after the fact.
-  * Versionable plain text. Easy to store and see diffs
-  * Common format for scaffold file means it can be easily manipulated programmatically.
-  * Command line interface for using as is
-  * Ruby API allowing more complex manipulation of the scaffold
-  * Simple to install.  Automatically installed. No manual installation.
 
@@
 to invoke scaffolder. This will list the available commands for running
-scaffolder.
+scaffolder. \pb
 
@@
 For example the following command will return the manual page for sequence
-listing the available options for running the sequence command.
+listing the available options for running the sequence command. \pb
 
@@
 formatted scaffold and the corresponding fasta file of sequences. This will
-then return the complete drafted sequence to the standard out.
+then return the complete drafted sequence to the standard out. \pb
 
@@
 validate command will check the scaffold for any overlapping inserts and
-return them as a YAML formatted list describing each one.
+return them as a YAML formatted list describing each one. \pb
 
@@
 
-  * Demonstrate how scaffolder can be manipulated programmatically
-  * Examine scaffolder data using a ruby program
-  * Example of extending sequence class with a new method \%GC content
-  * Create a new class to do something
+The scaffolder tool has a defined application programming interface (API) to
+allow ruby programs to manipulate the scaffold. Detailing the complete API
+would be overwhelming to this document. The scaffolder API is fully documented
+online using at the scaffolder website. The following code snippet gives and
+outline of how the API may be used to create a script to replicate the
+function of the `scaffolder sequence' command described above. \pb
+
+\begin{verbatim}
+  #!/usr/bin/env ruby
+
+  require 'rubygems'
+  require 'scaffolder'
+
+  scaffold_file = ARGV.shift
+  sequence_file = ARGV.shift
+
+  scaffold = Scaffolder.new(
+    YAML.load(File.read(scaffold_file)),
+    sequence_file)
+
+  sequence = scaffold.map{|entry| entry.sequence }.join
+
+  puts sequence
+\end{verbatim}
+
+This program illustrates that the scaffold object is initialised with two arguments: the YAML loaded scaffold file and the location of the fasta formatted sequence file.
+The scaffold is then enumerable as the next line illustrates the trivial code required to generate the scaffold.
+Further details of the attributes for each entry in the scaffold can be found in the documentation for the Scaffolder::Region class.
+\pb
 
 \subsection*{Discussion of scaffolder approach} %{{{2
+
+Pros
+  * Do one thing and do it well
+  * If your data structures are good enough, the algorithm to manipulate them should be trivial."
+
+  * Reproducable steps in building a scaffold
+  * Extensive documentation. Should be easy to use as possible.
+  * Scaffold file allows meaningful construction of a scaffold
+  * See of each sequence in building final draft genome
+  * Keeping genome sequence in a scaffold file means it is simpler to edit after the fact.
+  * Versionable plain text. Easy to store and see diffs
+  * Common format for scaffold file means it can be easily manipulated programmatically.
+  * Command line interface for using as is
+  * Ruby API allowing more complex manipulation of the scaffold
+  * Simple to install.  Automatically installed. No manual installation.
 

### 2011-01-28 text/src/article.tex
Final results section discussing scaffolder

@@
 
-Pros
-  * Do one thing and do it well
-  * If your data structures are good enough, the algorithm to manipulate them should be trivial."
+The main aims of scaffolder are two fold. The first is usuability and the second is reproducibility.
 
-  * Reproducable steps in building a scaffold
-  * Extensive documentation. Should be easy to use as possible.
-  * Scaffold file allows meaningful construction of a scaffold
-  * See of each sequence in building final draft genome
-  * Keeping genome sequence in a scaffold file means it is simpler to edit after the fact.
-  * Versionable plain text. Easy to store and see diffs
-  * Common format for scaffold file means it can be easily manipulated programmatically.
-  * Command line interface for using as is
-  * Ruby API allowing more complex manipulation of the scaffold
-  * Simple to install.  Automatically installed. No manual installation.
+\subsubsection{Simple to use}
+
+Scaffolder is designed to be as simple to use as possible. Assuming the
+existence of Ruby and RubyGems, scaffolder can be installed with a single
+command. Scaffolder uses a minimal and compact syntax to described the
+scaffold. This syntax is simple to construct and edit whilst being succint and
+readable in describing how the scaffold was constructed. Scaffolder provides
+extensive documentation both for the command line tools and for the Scaffolder
+Ruby API.
+
+\subsubsection{Reproducible genome construction}
+
+Constructing a genome sequence using a plain text scaffold file means the same
+sequence is always output for the same file. The sequence produced from
+manually joining separate contigs in a text editor can not be reliably
+reproduced. Constructing the scaffold in a plain text file means that the
+sequence is reproducible. Furthermore the source of each region in final
+sequence can be identified. The scaffold file provides a readable record of
+how the sequence was constructed. The scaffold file also means the scaffold is
+easer to edit once already constructed. Much more difficult to do when working
+on the sequence alone. The YAML text format is also much easier to compare
+differences using standard diff commands. This makes scaffold files much more
+simple to store in version control and to see the stages in the construction
+of the genome sequence.
 
@@
 
-Text for this section \ldots
+  * Do one thing and do it well
+  * If your data structures are good enough, the algorithm to manipulate them should be trivial."
 

### 2011-01-28 text/src/article.tex
Draft section for conclusions

@@
 
-  * Do one thing and do it well
-  * If your data structures are good enough, the algorithm to manipulate them should be trivial."
+Scaffolder follows the Unix philospophy ``Do one thing and do it well''.
+Scaffolder focuses on creating a simple to use and install tool for generating
+a genome scaffold. The YAML data format for creating a scaffold file is
+standardised and easily manipulated programmatically. This thereby follows
+a second Unix tenet: ``If your data structures are good enough, the algorithm
+to manipulate them should be trivial.''
 

### 2011-01-31 text/src/article.tex
Add initial text for background

--- 2011-01-28 text/src/article.tex
@@
 % }}}
-% {{{ Background
-\section*{Background}
- Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
-% }}}
-% {{{ Implementation
-%%%%%%%%%%%%%%%%%%%%%%
-\section*{Implementation}
+\section*{Background} %{{{1
+
+Large-scale parallel genome sequencing can produces hundreds of thousands to
+millions of nucleotide reads. These reads describe the nucleotide sequence in
+a region of the genome. The current status of parallel nucleotide sequencing
+limits the largest read length to $>$1000 nucleotides and to obtain
+a contiguous complete genome sequence these sequence reads must be joined
+together. \pb
+
+The process of joining these reads together is described as the `assembly'
+stage in a genome sequencing project. Assembly software joins these numerous
+sequence reads together, usually from the overlap between reads, into
+contiguous regions (`contigs') composing many of the individual reads. The goal
+of the assembly stage is produce a contiuous genome sequence describing the
+complete genome from start to finish. \pb
+
+These reads may however not represent a complete genome sequence where all
+contigs do not result in a single complete sequence. The software outlined here
+``Scaffolder'' allows these contigs to be further joined together into
+a continuous single sequence. This selection, order and direction of contigs
+into a final draft genome is often described as the `scaffold'. \pb
+
+\section*{Implementation} %{{{1
+
   Text for this section \ldots
 % }}}
-\section*{Results and Discussion} %{{{1
+% {{{ Results and Discussion
 
-\subsection*{Scaffolder Description} %{{{2
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Results and Discussion %%
+%%
+\section*{Results and Discussion}
+  \subsection*{Results sub-heading}
+    \subsubsection*{This is a sub-sub-heading}
+      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
+      pb at end of lines ensures correct paragraph spacing.\pb
+	  Text for this sub-sub-section \ldots
+    \subsubsection*{Another sub-sub-sub-heading}
+      Text for this sub-sub-section \ldots
 
-Individual reads produced from nucleotide sequencing can be assembled into
-contiguous regions (``contigs'') of many overlapping individual reads. These
-reads may however not represent a complete genome sequence where all contigs do
-not result in a single complete sequence. The software outlined here
-``Scaffolder'' allows these contigs to be further joined together into
-a continuous single sequence. This selection, order and direction of contigs
-into a final draft genome is often described as the `scaffold'. Scaffolder
-allows the manual or computational specification of a genome scaffold by
-specifying which contig sequences are used in simple to edit configuration
-files. \pb
+  \subsection*{Another results sub-heading}
+    Text for this sub-section \ldots
 
-The scaffold file specifies how contigs should be joined together in the draft
-genome sequence. The raw nucleotide sequence for each contig is stored in
-a fasta file, as might be produced from assembly of sequencing reads using
-programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
-Scaffolder then the contig to be selected by specifying the name of the contig
-in the scaffold file. The scaffold file also allows a sub-sequence or the
-reverse complement of a contig to used. This therefore preserves the original
-sequence data but also allowing sequences to manipulated in the final genome
-build. \pb
-
-The majority of a genome can be estimated from scaffolding assembled contigs
-into the most parsimonious build. There may however still be internal gaps in
-the contigs from mate pair assembly based on estimated mate pair insert size.
-Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
-and filled - a common process in finishing a genome sequence. Inserts may be
-treated the same way as a contig where the raw sequence, as may be produced by
-PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
-in the contig. \pb
-
-Scaffolder allows the specification of unknown regions in the scaffold. These
-are areas of the genome where the nucleotide sequence is unknown but the
-approximate size region may be estimated from mate pair insert size or from
-mapping contigs to reference genome and estimating the size of the gap. These
-unknown regions allow a continuous draft sequence to be scaffolded but whilst
-still highlighting regions waiting to be resolved. \pb
-
-\subsection*{Defining a scaffold in a text file} %{{{2
-
-The genome scaffold is wriiten in YAML format. YAML is a data format designed
-to be both machine readable and easy to manually edit. An example of this
-format illustrating all the features of scaffolder is shown in Figure 1. \pb
-
-\subsubsection*{Simple sequence region} %{{{3
-
-The first line of the file begins with three dashes. This is required to
-signify the start of a YAML document. Lines 2 begins with a dash character `-'
-which denotes the start of an entry in the sequence scaffold, in this case the
-first entry. Line 3 is indented by two spaces which therefore means this line
-contains an entry type. The ``sequence'' tag is used indicate a sequence is
-inserted at this region in the scaffold. The following line contain is
-indented by two spaces and indicates that this an attribute of the sequence
-entry. In this case the ``source'' keyword is used to indicate used which
-sequence should be inserted into the scaffold. This sequence is indentified by
-the first word of the fasta header of the selected sequence. This indicates
-a simple sequence region added to the scaffold. \pb
-
-\subsubsection*{Unresolved seqeunce region} %{{{3
-
-The following line has no indentation and begins with dash indicating the next
-entry in the scaffold. The following entry is indentified by the ``unresolved''
-tag. This identifies a region of unknown nucleotide sequence but is included
-because the approximate size of the region is known. The following line defines
-that the size of this unknown region in the genome scaffold should be 20
-nucleotides. This therefore results in the region being filled with 20 `N'
-characters in draft sequence. \pb
-
-\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3
-
-The following entry is another sequence using the sequence from 'sequence2' in
-the scaffold file. This sequence entry has three further attributes in addition
-to defining the source sequence. The first two are start and stop. These are
-used to trim the final sequence inserted into the scaffold. In this case the
-first two nucleotides of the sequence are removed so that that sequence starts
-at position 3 of the original raw sequence. Similarly the sequence after
-position is trimmed away after position 100. These sequence is also reverse
-complemented using the ``reverse'' tag. Reverse complementation is performed
-after the sequence has been trimmed to the specified coordinates. \pb
-
-The final attribute for this sequence entry defines inserts to be added to the
-sequence. These are additional regions of sequence that can be added to
-a sequence entry to alter the existing sequence such to fill gaps or correct
-sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
-is then listed similarly to the entries in the scaffold: beginning with a dash
-on one line then indented by an additional two spaces on the following lines.
-\pb
-
-Inserts are listed similarly to sequence entries where reverse, start and stop
-attributes can be used to trim and reverse the sequence. The `source' tag is
-used to indentify which fasta sequence entry should be used as the insert.  The
-insert has two additional tags ``open'' and ``close'' which determine where the
-insert is added to the hosting sequence. These coordinates are inclusive
-replaceing the coordinate nucleotides. The open and close positions need not be
-the same size as the insert size. Any difference in size to the replaced
-nucleotides will result in the length on the encoding sequence being adjusted.
-\pb
-
-The next insert entry shows an insert with only the open position specified.
-This is to illustrate that only the one of the `open' or `close' keywords is
-sufficient to add an insert. If only one of these words is used then the
-opposing insert position is calculated from the length on the insert sequence
-in the fasta file. This feature is designed for using inserts to close gaps
-where the length of the insert may not completely fill the gap but can still be
-used to fill a portion of the gap sequence. \pb
-
-\subsubsection*{Scaffolder file attribute processing} %{{{3
-
-Each scaffolder sequence entry can have multiple attributes which applied in
-different order can result in a different sequence. For instance reverse
-complementing the sequence then adding inserts will have a different effect
-compared with performing these actions in the reverse order. Scaffolder
-therefore follows the following protocol for processing each entry in the
-scaffold file. \pb
-
-\paragraph{Reverse sort inserts:} If the entry is a sequence and has sequence
-inserts these are reverse sorted by insert close position. The insert with the
-largest close position is first and and the smallest close position is last.
-This is done to preserve the the insert open/close coordinates as they are
-added to the scaffold. \pb
-
-\paragraph{Update sequence with insert:} Each insert is added to the sequence
-replacing the region specified by the open an close position with the specified
-source sequence.\pb
-
-\paragraph{Update sequence start/stop co-ordinates:} If the insert added to the
-sequence is a different length than the region replaced the start and stop
-co-ordinates of the sequence are correspondingly updated if necessary. E.g. if
-a 7 bp insert filled a bp region this would result in a 2bp difference in the
-length of the sequence. The stop coordinated would be updated to reflect this.
-\pb
-
-\paragraph{Sub-sequence selected:} If the sequence is has start or stop
-sequences specified these are used to trip the sequence to these co-ordinates.
-\pb
-
-\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyworkd the sequence is reverse complemented.
-\pb
-
-\subsection*{Scaffolder usage} %{{{2
-
-Scaffolder provides a ruby API that allows the scaffold and designated
-sequences to be manipulated through a ruby programme. This allows scaffolder to
-be integrated into exiting genomics works flows such as using Rake.  Scaffolder
-all provides a command line binary for the manipulation of the scaffold. This
-binary behaves as a standard Unix command line interface returning appropriate
-exit statuses and providing manual pages for each of the commands. These two
-methods are outlined in the following three sections. \pb
-
-\subsubsection*{Installation} %{{{3
-
-Scaffolder relies on both the ruby programming language and the RubyGems
-package management system. These are commonly installed on recent versions of
-Mac OSX and linux distrubtions. The installation of these can be tested on the
-command line by typing the following commands. \pb
-
-\begin{verbatim}
-ruby -v
-gem -v
-\end{verbatim}
-
-If both these commands return without error this indicates the necessary
-requirements to install scaffolder. The scaffolder API and command line tools
-can then be installed with the follow command.
-
-\begin{verbatim}
-gem install scaffolder scaffolder-tools
-\end{verbatim}
-
-Finally to ensure that the command line tools are accessible these the
-RubyGems binary directory should be in the command line path. This can be done
-by adding the following to the ~/.bashrc file or other corresponding shell
-configure file. Assuming that ~/.gem is the install path for RubyGems.
-
-\begin{verbatim}
-export PATH=~/.gem/bin:$PATH
-\end{verbatim}
-
-\subsubsection*{Command line interface} %{{{3
-
-Scaffolder provides a command line interface to validate and build a genome
-sequence from a scaffold. Assuming scaffolder has been installed and the
-binary is available in the command line path the following command can be used
-to invoke scaffolder. This will list the available commands for running
-scaffolder. \pb
-
-\begin{verbatim}
-$ scaffolder
-\end{verbatim}
-
-Help on each command can be found by type `help' followed by the name of the
-command. This will return the Unix manual page for the command of interest.
-For example the following command will return the manual page for sequence
-listing the available options for running the sequence command. \pb
-
-\begin{verbatim}
-$ scaffolder help sequence
-\end{verbatim}
-
-To the generate the fasta sequence from a scaffold the sequence command should
-be used. This should be passed the name of the file with containing the YAML
-formatted scaffold and the corresponding fasta file of sequences. This will
-then return the complete drafted sequence to the standard out. \pb
-
-\begin{verbatim}
-$ scaffolder sequence scaffold.yml sequences.fna
-\end{verbatim}
-
-Scaffolder also provides a function to test the scaffold for overlapping
-insert sequence co-ordinates. As described in the section above overlapping
-inserts will result in unexpected behaviour as the coordinates for the first
-insert will change the expected insert location for the second insert. The
-validate command will check the scaffold for any overlapping inserts and
-return them as a YAML formatted list describing each one. \pb
-
-\begin{verbatim}
-$ scaffolder validate scaffold.yml sequences.fna
-\end{verbatim}
-
-\subsubsection*{Application programming interface} %{{{3
-
-The scaffolder tool has a defined application programming interface (API) to
-allow ruby programs to manipulate the scaffold. Detailing the complete API
-would be overwhelming to this document. The scaffolder API is fully documented
-online using at the scaffolder website. The following code snippet gives and
-outline of how the API may be used to create a script to replicate the
-function of the `scaffolder sequence' command described above. \pb
-
-\begin{verbatim}
-  #!/usr/bin/env ruby
-
-  require 'rubygems'
-  require 'scaffolder'
-
-  scaffold_file = ARGV.shift
-  sequence_file = ARGV.shift
-
-  scaffold = Scaffolder.new(
-    YAML.load(File.read(scaffold_file)),
-    sequence_file)
-
-  sequence = scaffold.map{|entry| entry.sequence }.join
-
-  puts sequence
-\end{verbatim}
-
-This program illustrates that the scaffold object is initialised with two arguments: the YAML loaded scaffold file and the location of the fasta formatted sequence file.
-The scaffold is then enumerable as the next line illustrates the trivial code required to generate the scaffold.
-Further details of the attributes for each entry in the scaffold can be found in the documentation for the Scaffolder::Region class.
-\pb
-
-\subsection*{Discussion of scaffolder approach} %{{{2
-
-Pros
-  * Do one thing and do it well
-  * If your data structures are good enough, the algorithm to manipulate them should be trivial."
-
-  * Reproducable steps in building a scaffold
-  * Extensive documentation. Should be easy to use as possible.
-  * Scaffold file allows meaningful construction of a scaffold
-  * See of each sequence in building final draft genome
-  * Keeping genome sequence in a scaffold file means it is simpler to edit after the fact.
-  * Versionable plain text. Easy to store and see diffs
-  * Common format for scaffold file means it can be easily manipulated programmatically.
-  * Command line interface for using as is
-  * Ruby API allowing more complex manipulation of the scaffold
-  * Simple to install.  Automatically installed. No manual installation.
-
-\section*{Conclusions} %{{{1
-
+  \subsection*{Yet another results sub-heading}
+    Text for this sub-section.  More results \ldots
+% }}}
+% {{{ Conclusions
+%%%%%%%%%%%%%%%%%%%%%%
+\section*{Conclusions}
   Text for this section \ldots
-
-\section*{Availability and Requirements} %{{{1
+% }}}
+% {{{ Availability and Requirements
+\section*{Availability and Requirements}
 
@@
   Text for this section \ldots
-%{{{1 Bibliography
+%}}}
+% {{{ Bibliography
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
@@
   \bibliography{article} }     % Bibliography file (usually '*.bib' ) 
-
+%}}}
+% {{{ Figures
 \ifthenelse{\boolean{publ}}{\end{multicols}}{}
 
-\section*{Figures} %{{{1
+\section*{Figures}
+  \subsection*{Figure 1 - Sample figure title}
+      A short description of the figure content
+      should go here.
 
-  \subsection*{Figure 1 - Sample figure title}
-
-  A short description of the figure content should go here.
-
-\section*{Tables} %{{{1
-
+  \subsection*{Figure 2 - Sample figure title}
+      Figure legend text.
+% }}}
+% {{{ Tables
+\section*{Tables}
   \subsection*{Table 1 - Sample table title}
@@
     still be described here.
-
-\section*{Additional Files} %{{{1
-
+% }}}
+% {{{ Additional Files
+\section*{Additional Files}
   \subsection*{Additional file 1 --- Sample additional file title}
@@
 \end{bmcformat}
-
 \end{document}
+% }}}

### 2011-01-31 text/src/article.tex
Add initial text for background
Revise first section of the results

--- 2011-01-28 text/src/article.tex
@@
 
-\subsection*{Scaffolder Description} %{{{2
+\subsection*{Scaffolder Simplifies Genome Finishing} %{{{2
 
-Individual reads produced from nucleotide sequencing can be assembled into
-contiguous regions (``contigs'') of many overlapping individual reads. These
-reads may however not represent a complete genome sequence where all contigs do
-not result in a single complete sequence. The software outlined here
-``Scaffolder'' allows these contigs to be further joined together into
-a continuous single sequence. This selection, order and direction of contigs
-into a final draft genome is often described as the `scaffold'. Scaffolder
-allows the manual or computational specification of a genome scaffold by
-specifying which contig sequences are used in simple to edit configuration
-files. \pb
+The Scaffolder software allows manual or computational joining of mutliple
+discontiguous nucleotide sequences together into a contiguous super sequence.
+Simple plain-text configuration files are used to select the nucleotide
+sequences in the scaffold and how they should be joined together. This is
+called the ``scaffold file''. \pb
 
-The scaffold file specifies how contigs should be joined together in the draft
-genome sequence. The raw nucleotide sequence for each contig is stored in
-a fasta file, as might be produced from assembly of sequencing reads using
-programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
-Scaffolder then the contig to be selected by specifying the name of the contig
-in the scaffold file. The scaffold file also allows a sub-sequence or the
-reverse complement of a contig to used. This therefore preserves the original
-sequence data but also allowing sequences to manipulated in the final genome
-build. \pb
+The scaffold file is read by the scaffolder software to determine how the
+scaffold super sequence is generated. In addition to specifying which
+nucleotides should be used to generate the draft super sequence, the scaffold
+file allows the sequences to trimmed down to smaller sub-sequences and to
+reverse complement the sequence if necessary. \pb
 
-The majority of a genome can be estimated from scaffolding assembled contigs
-into the most parsimonious build. There may however still be internal gaps in
-the contigs from mate pair assembly based on estimated mate pair insert size.
-Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
-and filled - a common process in finishing a genome sequence. Inserts may be
-treated the same way as a contig where the raw sequence, as may be produced by
-PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
-in the contig. \pb
+There may be unknown regions in the sequences used in the scaffold. Region of
+nucleotide sequence the assembler was unable to determine the nucleotide
+sequence. The process of genome finishing may involve determining the correct
+nucleotide sequence for these regions then filling the gaps with the additional
+correct sequence. The scaffold file specifically allows these inserts to be
+added to scaffold. These inserts can be treated the same way as a larger
+contig: trimmed and/or reversed to match the required gap region. \pb
 
-Scaffolder allows the specification of unknown regions in the scaffold. These
-are areas of the genome where the nucleotide sequence is unknown but the
-approximate size region may be estimated from mate pair insert size or from
-mapping contigs to reference genome and estimating the size of the gap. These
-unknown regions allow a continuous draft sequence to be scaffolded but whilst
-still highlighting regions waiting to be resolved. \pb
+When scaffolding a genome there may be regions of sequence approximate size
+region may be estimated from mate-pair insert size or from mapping the contigs
+to reference genome. These unknown regions are useful to join non-contiguous
+sequences together by the estimated size of the region whilst still
+highlighting areas in the genome to be resolved. These unknown regions can be
+specified in the scaffold file for the above reasons. \pb
+
+The original raw nucleotide sequence for each assembled contig or insert
+sequence is maintained in a fasta file. The fasta file of contigs is a common
+output for assemblers and as such should be readily available after assembly
+without additional processing. The choice of the nucleotide sequence to use in
+the scaffold is specified using the first space-delimited word from the fasta
+header of the sequence. By maintaining the nucleotide sequences in a separate
+file this preserves the original raw sequence in the fasta file. Furthermore
+this separates the actual nucleotide sequence from the determination of how it
+is selected and editted in the final genome sequence. \pb
 

### 2011-01-31 text/src/article.tex
Spell check

@@
 
-The Scaffolder software allows manual or computational joining of mutliple
+The Scaffolder software allows manual or computational joining of multiple
 discontiguous nucleotide sequences together into a contiguous super sequence.
@@
 this separates the actual nucleotide sequence from the determination of how it
-is selected and editted in the final genome sequence. \pb
+is selected and edited in the final genome sequence. \pb
 
@@
 
-The genome scaffold is wriiten in YAML format. YAML is a data format designed
+The genome scaffold is written in YAML format. YAML is a data format designed
 to be both machine readable and easy to manually edit. An example of this
@@
 entry. In this case the ``source'' keyword is used to indicate used which
-sequence should be inserted into the scaffold. This sequence is indentified by
+sequence should be inserted into the scaffold. This sequence is identified by
 the first word of the fasta header of the selected sequence. This indicates
@@
 
-\subsubsection*{Unresolved seqeunce region} %{{{3
+\subsubsection*{Unresolved sequence region} %{{{3
 
 The following line has no indentation and begins with dash indicating the next
-entry in the scaffold. The following entry is indentified by the ``unresolved''
+entry in the scaffold. The following entry is identified by the ``unresolved''
 tag. This identifies a region of unknown nucleotide sequence but is included
@@
 attributes can be used to trim and reverse the sequence. The `source' tag is
-used to indentify which fasta sequence entry should be used as the insert.  The
+used to identify which fasta sequence entry should be used as the insert.  The
 insert has two additional tags ``open'' and ``close'' which determine where the
 insert is added to the hosting sequence. These coordinates are inclusive
-replaceing the coordinate nucleotides. The open and close positions need not be
+replacing the coordinate nucleotides. The open and close positions need not be
 the same size as the insert size. Any difference in size to the replaced
@@
 
-\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyworkd the sequence is reverse complemented.
+\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyword the sequence is reverse complemented.
 \pb
@@
 package management system. These are commonly installed on recent versions of
-Mac OSX and linux distrubtions. The installation of these can be tested on the
+Mac OSX and Linux distributions. The installation of these can be tested on the
 command line by typing the following commands. \pb
@@
 
-The main aims of scaffolder are two fold. The first is usuability and the second is reproducibility.
+The main aims of scaffolder are two fold. The first is usability and the second is reproducibility.
 
@@
 command. Scaffolder uses a minimal and compact syntax to described the
-scaffold. This syntax is simple to construct and edit whilst being succint and
+scaffold. This syntax is simple to construct and edit whilst being succinct and
 readable in describing how the scaffold was constructed. Scaffolder provides
@@
 how the sequence was constructed. The scaffold file also means the scaffold is
-easer to edit once already constructed. Much more difficult to do when working
+easier to edit once already constructed. Much more difficult to do when working
 on the sequence alone. The YAML text format is also much easier to compare
@@
 
-Scaffolder follows the Unix philospophy ``Do one thing and do it well''.
+Scaffolder follows the Unix philosophy ``Do one thing and do it well''.
 Scaffolder focuses on creating a simple to use and install tool for generating

### 2011-02-01 text/src/article.tex
Revise scaffold file section of results

@@
 
-The genome scaffold is written in YAML format. YAML is a data format designed
-to be both machine readable and easy to manually edit. An example of this
-format illustrating all the features of scaffolder is shown in Figure 1. \pb
+The scaffold file is written using `YAML', a data format designed to be both
+machine readable and manually editable. An example of a scaffold file in YAML
+format is shown in Figure 1. This example scaffold file illustrates the
+available features in scaffolder. The basic layout of the scaffold file is as
+a list - a series of entries where each entry defines a region of sequence to
+be generated for the final super sequence. In the final super sequence each of
+the entries are joined together to generate a single continuous sequences. \pb
 
@@
 
-The first line of the file begins with three dashes. This is required to
-signify the start of a YAML document. Lines 2 begins with a dash character `-'
-which denotes the start of an entry in the sequence scaffold, in this case the
-first entry. Line 3 is indented by two spaces which therefore means this line
-contains an entry type. The ``sequence'' tag is used indicate a sequence is
-inserted at this region in the scaffold. The following line contain is
-indented by two spaces and indicates that this an attribute of the sequence
-entry. In this case the ``source'' keyword is used to indicate used which
-sequence should be inserted into the scaffold. This sequence is identified by
-the first word of the fasta header of the selected sequence. This indicates
-a simple sequence region added to the scaffold. \pb
+The first line of the scaffold file begins with three dashes. This is
+a requirement to indicate the start of a YAML formatted document. Line 2 begins
+with a dash character `-' denoting the first entry in the sequence scaffold.
+Line 3 is indented by two spaces. Whitespace is used to delimite attributes in
+the file. Lines with an incremented level of white space are considered linked
+to the first preceeding non-indented line. \pb
+
+The ``sequence'' tag used on line 3 indicates that this region is in the
+scaffold will contain nucleotide sequence from the fasta file. The following
+line indented by two spaces and indicating an attribute of the sequence entry
+defined on line 3. In this case the ``source'' keyword identifies the name of
+the sequence to be inserted into the scaffold. As described above the first
+space-delimited word of the fasta header should be used to identify the
+sequence. Taken together lines 2-4 describe the first entry in the scaffold is
+a simple sequence which the identifier `sequence'. \pb
 
@@
 
-The following line has no indentation and begins with dash indicating the next
-entry in the scaffold. The following entry is identified by the ``unresolved''
-tag. This identifies a region of unknown nucleotide sequence but is included
-because the approximate size of the region is known. The following line defines
-that the size of this unknown region in the genome scaffold should be 20
-nucleotides. This therefore results in the region being filled with 20 `N'
-characters in draft sequence. \pb
+Line 5 has no whitespace indentation and begins with a dash and therefore
+describes the next entry in the scaffold. This entry is identified by the
+``unresolved'' tag, and specifies a region of unknown nucleotides but known
+length. The following line specifies the size of this unknown region. Lines 5-6
+therefore insert a region of 20 `N' characters in the super sequence directly
+following the region defined by the first entry on lines 2-5. \pb
 
@@
 
-The following entry is another sequence using the sequence from 'sequence2' in
-the scaffold file. This sequence entry has three further attributes in addition
-to defining the source sequence. The first two are start and stop. These are
-used to trim the final sequence inserted into the scaffold. In this case the
-first two nucleotides of the sequence are removed so that that sequence starts
-at position 3 of the original raw sequence. Similarly the sequence after
-position is trimmed away after position 100. These sequence is also reverse
-complemented using the ``reverse'' tag. Reverse complementation is performed
-after the sequence has been trimmed to the specified coordinates. \pb
+The last entry adds the sequence 'sequence2' in the scaffold file. This entry
+demonstrates three further attributes for a sequence entry.The `start' and
+`stop' tags are used to trim the sequence used in the scaffold to these
+coordinates inclusively. The ``reverse'' tag is used to reverse complement this
+region in the final super sequence. \pb
 
-The final attribute for this sequence entry defines inserts to be added to the
-sequence. These are additional regions of sequence that can be added to
-a sequence entry to alter the existing sequence such to fill gaps or correct
-sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
-is then listed similarly to the entries in the scaffold: beginning with a dash
-on one line then indented by an additional two spaces on the following lines.
-\pb
+This region further includes the ``inserts'' tag to update the inserted
+sequence with additional regions of sequence. These inserts are also added as
+a YAML list with each insert listed with a dash on one line followed by the
+remainder of the insert errata. \pb
 
-Inserts are listed similarly to sequence entries where reverse, start and stop
-attributes can be used to trim and reverse the sequence. The `source' tag is
-used to identify which fasta sequence entry should be used as the insert.  The
-insert has two additional tags ``open'' and ``close'' which determine where the
-insert is added to the hosting sequence. These coordinates are inclusive
-replacing the coordinate nucleotides. The open and close positions need not be
-the same size as the insert size. Any difference in size to the replaced
-nucleotides will result in the length on the encoding sequence being adjusted.
-\pb
+Lines 16-19 illustrate an insert using similar attributes to that of a sequence
+entry. The reverse, start and stop tags trim the inserted sequence and the
+`source' tag identifies the corresponding fasta sequence. The ``open'' and
+``close'' tags on lines 20-21 determine where the insert is added to the
+enclosing sequence.  These coordinates inclusively replace the region within
+these coordinates with the specified fasta sequence. The open and close
+positions need not be the same size as the inserted sequence. \pb
 
-The next insert entry shows an insert with only the open position specified.
-This is to illustrate that only the one of the `open' or `close' keywords is
-sufficient to add an insert. If only one of these words is used then the
-opposing insert position is calculated from the length on the insert sequence
-in the fasta file. This feature is designed for using inserts to close gaps
-where the length of the insert may not completely fill the gap but can still be
-used to fill a portion of the gap sequence. \pb
+The next insert on lines 23-24 specifies only the open position. This
+illustrate that only the one of the `open' or `close' tags is required.  When
+only open or close tag is used then the opposing position is calculated from
+the length on the insert fasta sequence. This allows inserts to partially fill
+gap regions in sequences. \pb
 

### 2011-02-01 text/src/article.tex
Revise scaffolder file processing section

@@
 
-Each scaffolder sequence entry can have multiple attributes which applied in
-different order can result in a different sequence. For instance reverse
-complementing the sequence then adding inserts will have a different effect
-compared with performing these actions in the reverse order. Scaffolder
-therefore follows the following protocol for processing each entry in the
-scaffold file. \pb
+The sequence entry in the scaffold file allows multiple sequence-editting
+attributes which when applied in different combinations will result in
+different output sequence. An example is reverse complementing the sequence
+then follwing adding sequence inserts. Performing these actions in the reverse
+order will produce a difference output sequence. This section therefore
+outlines the scaffolder protocol for processing sequence entries. \pb
 
-\paragraph{Reverse sort inserts:} If the entry is a sequence and has sequence
-inserts these are reverse sorted by insert close position. The insert with the
-largest close position is first and and the smallest close position is last.
-This is done to preserve the the insert open/close coordinates as they are
-added to the scaffold. \pb
+\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
+inserts these are reverse sorted by close position. The insert with the
+largest close position is first in the list and inserts with the smallest
+close position is last. This preserves the original indented coordinates for
+insert during the process of adding inserts to the sequence. \pb
 
-\paragraph{Update sequence with insert:} Each insert is added to the sequence
-replacing the region specified by the open an close position with the specified
-source sequence.\pb
+\paragraph{Update sequence with insert:} Each insert sequence is processed
+according to any specified attributes. The insert is added to the enclosing
+sequence in the region specified by the open and close coordinates. \pb
 
-\paragraph{Update sequence start/stop co-ordinates:} If the insert added to the
-sequence is a different length than the region replaced the start and stop
-co-ordinates of the sequence are correspondingly updated if necessary. E.g. if
-a 7 bp insert filled a bp region this would result in a 2bp difference in the
-length of the sequence. The stop coordinated would be updated to reflect this.
-\pb
+If the insert added to the sequence is a different length than the region
+replaced the sequence start and stop co-ordinates are updated. For example if
+a 7bp insert replaced a 5bp region the stop coordinate of the enclosing
+sequence would be incremented by the 2bp difference. \pb
 
-\paragraph{Sub-sequence selected:} If the sequence is has start or stop
-sequences specified these are used to trip the sequence to these co-ordinates.
-\pb
+\paragraph{Triming sequence:} If the sequence is has start or stop sequences
+specified these are used to trip the sequence to these co-ordinates
+inclusively. \pb
 
-\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyword the sequence is reverse complemented.
-\pb
+\paragraph{Reverse complement sequence:} If the ``reverse'' keyword is
+specified the sequence is reverse complemented. This is final step in
+processing the sequence with results in the sequence being returned. \pb
 

### 2011-02-01 text/src/article.tex
Revise third section of the results

@@
 
-Scaffolder provides a ruby API that allows the scaffold and designated
-sequences to be manipulated through a ruby programme. This allows scaffolder to
-be integrated into exiting genomics works flows such as using Rake.  Scaffolder
-all provides a command line binary for the manipulation of the scaffold. This
-binary behaves as a standard Unix command line interface returning appropriate
-exit statuses and providing manual pages for each of the commands. These two
-methods are outlined in the following three sections. \pb
+Scaffolder provides a standardised set of Ruby classes and methods (the
+application programming interface or API) allowing the scaffold to be
+manipulated programmatically. This allows scaffolder to be integrated into
+existing genomics workflows or used with Ruby build tools such as using Rake.
+\pb
+
+Scaffolder also provides a command line interface (CLI) through a scaffolder
+binary. This the scaffold sequence to be produced and validated without
+requiring any Ruby knowledge. The scaffolder CLI behaves as a standard Unix
+tool - returning appropriate exit statuses and providing appropriate manual
+pages. The use of both these scaffolder interfaces is outlined in the following
+sections. \pb
 
@@
 Scaffolder relies on both the ruby programming language and the RubyGems
-package management system. These are commonly installed on recent versions of
-Mac OSX and Linux distributions. The installation of these can be tested on the
-command line by typing the following commands. \pb
+package management system to work. These are either already present easily
+installed on recent distributions of Mac OSX or Linux. The presences of these
+tools can be tested by typing the following on the command line. \pb
 
@@
 If both these commands return without error this indicates the necessary
-requirements to install scaffolder. The scaffolder API and command line tools
-can then be installed with the follow command.
+system requirements to install and run scaffolder. The scaffolder API and CLI
+can then be installed through RubyGems with the following command.
 
@@
 
-Finally to ensure that the command line tools are accessible these the
-RubyGems binary directory should be in the command line path. This can be done
-by adding the following to the ~/.bashrc file or other corresponding shell
-configure file. Assuming that ~/.gem is the install path for RubyGems.
+To ensure that the scaffolder CLI is accessible the RubyGems `bin' directory
+should be in the command line path. Assuming that RubyGems are installed in
+~/.gem this can be done by adding the following to the ~/.bashrc file. This
+should be adjusted for other shells such as zsh etc..
 
@@
 
-\subsubsection*{Command line interface} %{{{3
+\subsubsection*{Command Line Interface} %{{{3
 
-Scaffolder provides a command line interface to validate and build a genome
-sequence from a scaffold. Assuming scaffolder has been installed and the
-binary is available in the command line path the following command can be used
-to invoke scaffolder. This will list the available commands for running
-scaffolder. \pb
+The Scaffolder CLI provides functions to validate the scaffold file and build the draft scaffold sequence. Assuming scaffolder has been installed and the
+binary is available in the command line path the following command should run without error and list the available functions. \pb
 
 \begin{verbatim}
-$ scaffolder
+  scaffolder
 \end{verbatim}
 
-Help on each command can be found by type `help' followed by the name of the
-command. This will return the Unix manual page for the command of interest.
-For example the following command will return the manual page for sequence
-listing the available options for running the sequence command. \pb
+Help on a specific command can be found by typing `help' followed by the name
+of the command. This will return the corresponding Unix manual page. The
+following example will return the manual page for the sequence command. \pb
 
 \begin{verbatim}
-$ scaffolder help sequence
+  scaffolder help sequence
 \end{verbatim}
 
-To the generate the fasta sequence from a scaffold the sequence command should
-be used. This should be passed the name of the file with containing the YAML
-formatted scaffold and the corresponding fasta file of sequences. This will
-then return the complete drafted sequence to the standard out. \pb
+The ``sequence'' command builds the fasta sequence from the scaffold and fasta
+file. Running this command requires the locations of both the scaffold file and
+the fasta file of sequences. This will then return the complete scaffold
+sequence. \pb
 
 \begin{verbatim}
-$ scaffolder sequence scaffold.yml sequences.fna
+  scaffolder sequence scaffold.yml sequences.fna
 \end{verbatim}
 
-Scaffolder also provides a function to test the scaffold for overlapping
-insert sequence co-ordinates. As described in the section above overlapping
-inserts will result in unexpected behaviour as the coordinates for the first
-insert will change the expected insert location for the second insert. The
-validate command will check the scaffold for any overlapping inserts and
-return them as a YAML formatted list describing each one. \pb
+Scaffolder provides a function to test the scaffold for overlapping insert
+coordinates. Overlapping inserts will result in unexpected behaviour as the
+coordinates for the first insert may change the insert location for subsequent
+inserts. The ``validate'' command will check the scaffold for any such inserts
+and return them as a YAML list. \pb
 
 \begin{verbatim}
-$ scaffolder validate scaffold.yml sequences.fna
+  scaffolder validate scaffold.yml sequences.fna
 \end{verbatim}
@@
 
-The scaffolder tool has a defined application programming interface (API) to
-allow ruby programs to manipulate the scaffold. Detailing the complete API
-would be overwhelming to this document. The scaffolder API is fully documented
-online using at the scaffolder website. The following code snippet gives and
-outline of how the API may be used to create a script to replicate the
+Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
+This API is detailed in full on the scaffolder website. The following code
+snippet outlines how the API may be used to create a script to replicate the
 function of the `scaffolder sequence' command described above. \pb
@@
 
-This program illustrates that the scaffold object is initialised with two arguments: the YAML loaded scaffold file and the location of the fasta formatted sequence file.
-The scaffold is then enumerable as the next line illustrates the trivial code required to generate the scaffold.
-Further details of the attributes for each entry in the scaffold can be found in the documentation for the Scaffolder::Region class.
-\pb
+This code snippet illustrates that the scaffold object is initialised with two
+arguments: the YAML loaded scaffold and the location of the fasta sequence
+file. The scaffold is an enumerable object as the next line illustrates the
+trivial code required to generate the scaffold sequence. Further details of
+the classes and method related to each scaffold entry can be found in the
+documentation for the Scaffolder::Region class. \pb
 

### 2011-02-01 text/src/article.tex
Revise last section of results

@@
 
-\subsection*{Discussion of scaffolder approach} %{{{2
+\subsection*{Discussion of Scaffolder Approach} %{{{2
 
-The main aims of scaffolder are two fold. The first is usability and the second is reproducibility.
+The main aims of scaffolder are two foldfold. The first is produce easy to
+install and simplifies the task of genome finishing. The second aim is to
+allow to create a tool to facilitate reproducibility in the task of joining
+large nucleotide sequences together. \pb
 
@@
 
-Scaffolder is designed to be as simple to use as possible. Assuming the
-existence of Ruby and RubyGems, scaffolder can be installed with a single
-command. Scaffolder uses a minimal and compact syntax to described the
-scaffold. This syntax is simple to construct and edit whilst being succinct and
-readable in describing how the scaffold was constructed. Scaffolder provides
+Scaffolder is designed to be simple to use as possible. Assuming the existence
+of Ruby and RubyGems, scaffolder can be installed in a single command line
+step. Scaffolder uses a minimal and compact syntax to described the genome
+scaffold. This syntax is simple to construct and edit whilst being succinct
+and readable. This aims to allow creating a genome scaffold to relatively
+simple to do in a common text editor. Furthermore scaffolder provides
 extensive documentation both for the command line tools and for the Scaffolder
-Ruby API.
+Ruby API. Therefore should any part of the scaffolder tool be unclear the
+documentation should make an solution easy to find. \pb
 
@@
 Constructing a genome sequence using a plain text scaffold file means the same
-sequence is always output for the same file. The sequence produced from
-manually joining separate contigs in a text editor can not be reliably
-reproduced. Constructing the scaffold in a plain text file means that the
-sequence is reproducible. Furthermore the source of each region in final
-sequence can be identified. The scaffold file provides a readable record of
-how the sequence was constructed. The scaffold file also means the scaffold is
-easier to edit once already constructed. Much more difficult to do when working
-on the sequence alone. The YAML text format is also much easier to compare
-differences using standard diff commands. This makes scaffold files much more
-simple to store in version control and to see the stages in the construction
-of the genome sequence.
+sequence is always output for the same file. Joining large nucleotide
+sequences manually cannot however be reliably reproduced. The layout of the
+scaffold file furthermore provides a human readable record of how the sequence
+was constructed. \pb
+
+Configuring the final sequence in the scaffold file means the build is easier
+to edit once already constructed compared with working on the sequence alone.
+The YAML text format is also much easier to compare differences using standard
+text comparison tools such as diff. This makes scaffold files easier to store
+in version control systems. This affords genome scaffolding the same benefits
+enjoyed by software developers. \pb
 

### 2011-02-07 text/src/article.tex

@@
 % {{{ Background
-\section*{Background}
- Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
-% }}}
-% {{{ Implementation
-%%%%%%%%%%%%%%%%%%%%%%
-\section*{Implementation}
+\section*{Background} %{{{1
+
+Large-scale parallel genome sequencing can produces hundreds of thousands to
+millions of nucleotide reads. These reads describe the nucleotide sequence in
+a region of the genome. The current status of parallel nucleotide sequencing
+limits the largest read length to $>$1000 nucleotides and to obtain
+a contiguous complete genome sequence these sequence reads must be joined
+together. \pb
+
+The process of joining these reads together is described as the `assembly'
+stage in a genome sequencing project. Assembly software joins these numerous
+sequence reads together, usually from the overlap between reads, into
+contiguous regions (`contigs') composing many of the individual reads. The goal
+of the assembly stage is produce a contiuous genome sequence describing the
+complete genome from start to finish. \pb
+
+These reads may however not represent a complete genome sequence where all
+contigs do not result in a single complete sequence. The software outlined here
+``Scaffolder'' allows these contigs to be further joined together into
+a continuous single sequence. This selection, order and direction of contigs
+into a final draft genome is often described as the `scaffold'. \pb
+
+\section*{Implementation} %{{{1
+
   Text for this section \ldots

### 2011-02-08 text/src/article.tex

--- 2011-02-01 text/src/article.tex
@@
 % }}}
-% {{{ Background
-\section*{Background}
- Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
-% }}}
-% {{{ Implementation
-%%%%%%%%%%%%%%%%%%%%%%
-\section*{Implementation}
+\section*{Background} %{{{1
+%{{{2
+
+High throughput sequencing can be reads short regions of a genome sequence to
+produce hundreds of thousands to millions of lines of nucleotide sequence. The
+current status of parallel nucleotide sequencing is limited to producing reads
+less than $>$1000 nucleotides long. Therefore to obtain a contiguous complete
+genome sequence these sequence reads must therefore be joined together. \pb
+
+The process of joining sequencing reads together is described as the
+`assembly' stage in a genome project. Assembly software joins these numerous
+sequence reads together into contiguous regions (`contigs') composed of many
+overlapping individual reads. The goal of the assembly stage is produce
+a continuous genome sequence describing the complete genome from start to
+finish. \pb
+
+Often repetitive nucleotide `repeat' regions in the genome or biased and
+incomplete sequencing data may prevent the genome being assembled into
+a continuous sequence. This sequence data can however describe the majority of
+the genetic information in the genome and in the case of microbes and moving
+forward with the current sequence data at this stage may be the best
+investment of time and money \cite{branscomb2002}. \pb
+
+A continuous high-quality `finished' genome sequence provides a much greater
+depth of information as compared to a draft sequence
+\cite{parkhill2002,fraser2002}. Examples of depth of information are complete
+resolution of repeat regions and the exact distances between genomic elements.
+The process of moving from a draft sequence to a finished continuous sequence
+involves the process of `scaffolding' the contigs produced by the assembly
+software together. \pb
+
+\subsection*{Scaffolding} %{{{2
+
+Scaffolding can be described as the process of moving from a series of
+disconnected contigs to complete continuous sequence of all the contigs joined
+together. The scaffolding process may not be completed but may still resolve
+and join some contigs together or add additional regions of sequence to fill
+gaps. In the case of a microbial genome, the process of scaffolding may
+include the following steps:
+
+\subsubsection*{Contig Orientation} %{{{3
+
+Sequence reads may be produced from either DNA strand during the sequencing
+process. Therefore the resulting contigs produced from these sequences may are
+likely to be representative from either of the genome strands. Contig
+orientation involves correcting the sequence data so that all contigs
+represent sequence data from the same DNA strand. In the case of microbes this
+orientation is often the 5`-3` strand from the origin of DNA replication.
+
+\subsubsection*{Contig Ordering} %{{{3
+
+The source location of each contig from the genome is unknown. Contig ordering
+represents determining the order of contigs which most accurately represent
+the actual genome sequence. This process can also determine if any contigs
+represent extra-genomic DNA such as plasmids and therefore should be separated
+from the genomic contigs. Contigs are usually ordered in reference to the
+origin of replication where the contig enclosing this region is the first
+contigs and subsequent contigs are labelled in the direction of DNA
+replication.
+
+\subsubsection*{Contig Distancing} %{{{3
+
+Once the order of contigs in the genome sequence is known it is useful to
+determine their approximate distance from each other. This allows the size of
+the gaps between contigs to be determined. Filling in these regions in with
+unknown nucleotide characters allows a draft continuous representation of the
+genome to determined.
+
+\subsubsection*{Gap Closing and Filling} %{{{3
+
+The final stage of producing a genome sequence is to close and fill gaps in
+the sequence. This may require returning to the laboratory to perform
+additional sequencing to close gaps or \emph{in silico} methods may be able to
+close gaps using existing data. Once all gaps in a scaffold have been closed
+the genome may be considered finished. Though additional sequencing may be
+performed to `proof' any low coverage regions for erroneous nucleotide
+insertions, deletions or substitutions.
+
+\subsection*{Computational Methods for Scaffolding} %{{{2
+
+Scaffolding may be considered a combination of \emph{in vitro} and \emph{in
+silico} methods. Several steps may be performed computationally by using
+available data to generate a scaffold. Examples may be the use of pair reads
+sequencing, where reads a known distance apart, or from comparing contigs to
+a reference genome. In the case where \emph{in silico} methods are unable to
+complete the scaffold in may be necessary to use PCR to amplify then use
+traditional Sanger sequencing to find the sequence and close gaps. Obviously
+\emph{in silico} methods are preferable as they less costly in laboratory time
+and materials \cite{nagaryan2010}. \pb
+
+For these reason may examples exist for software to generate genome scaffolds
+from assembly data. These software use a variety of approaches. Synteny
+approaches use a reference genome to order and orient contigs
+\cite{richter2007,zhao2008}. Repeat regions which may be responsible for many
+of the gaps between contigs may be reassembled using algorithms specifically
+for repeat resolution \cite{mulkov2002,koren2010}. Uniquely matching reads may
+be used to try and bridge gaps between scaffolds \cite{tsai2010}. Whole genome
+may use heuristic graph algorithms to build a path between contigs using
+multiple input data \cite{pop2004,dayarian2010}. Finally when the scaffold
+cannot be completely resolved \emph{in silico} software packages exist to
+suggest the likely primers necessary for determining the sequence in gap
+regions \cite{gordon2001,nagaryan2010}. Taken together these tools provide
+a wide variety of possible options for scaffolding a genome should an
+assembler be unable to produce a continuous sequence from the sequencing
+reads. \pb
+
+Scaffolding however will inevitably require manual editing if computational
+methods do not completely resolve the final genome sequence. This involves
+moving an joining larges blocks of nucleotide text together in what the genome
+finisher perceives to be the correct scaffold. This process may be expected to
+be error-prone. Further manual editing cannot be reliably reproduced which
+precludes reproducibility in a genome scaffolding process. \pb
+
+\subsection*{The Scaffolder Software} %{{{2
+
+The software outlined here ``Scaffolder'' is aimed at joining contigs together
+to build a genome scaffold. Scaffolder uses a specific data syntax to define
+how contigs are joined together. Furthermore scaffolder allows inserts into
+contigs and unknown regions in a scaffold to be defined. This syntax allows
+a scaffold to be edited in much easier way than editing large blocks of
+nucleotides. Further more this makes genome scaffolds reproducible.
+
+\section*{Implementation} %{{{1
+
   Text for this section \ldots
 % }}}
-\section*{Results and Discussion} %{{{1
+% {{{ Results and Discussion
 
-\subsection*{Scaffolder Simplifies Genome Finishing} %{{{2
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Results and Discussion %%
+%%
+\section*{Results and Discussion}
+  \subsection*{Results sub-heading}
+    \subsubsection*{This is a sub-sub-heading}
+      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
+      pb at end of lines ensures correct paragraph spacing.\pb
+	  Text for this sub-sub-section \ldots
+    \subsubsection*{Another sub-sub-sub-heading}
+      Text for this sub-sub-section \ldots
 
-The Scaffolder software allows manual or computational joining of multiple
-discontiguous nucleotide sequences together into a contiguous super sequence.
-Simple plain-text configuration files are used to select the nucleotide
-sequences in the scaffold and how they should be joined together. This is
-called the ``scaffold file''. \pb
+  \subsection*{Another results sub-heading}
+    Text for this sub-section \ldots
 
-The scaffold file is read by the scaffolder software to determine how the
-scaffold super sequence is generated. In addition to specifying which
-nucleotides should be used to generate the draft super sequence, the scaffold
-file allows the sequences to trimmed down to smaller sub-sequences and to
-reverse complement the sequence if necessary. \pb
-
-There may be unknown regions in the sequences used in the scaffold. Region of
-nucleotide sequence the assembler was unable to determine the nucleotide
-sequence. The process of genome finishing may involve determining the correct
-nucleotide sequence for these regions then filling the gaps with the additional
-correct sequence. The scaffold file specifically allows these inserts to be
-added to scaffold. These inserts can be treated the same way as a larger
-contig: trimmed and/or reversed to match the required gap region. \pb
-
-When scaffolding a genome there may be regions of sequence approximate size
-region may be estimated from mate-pair insert size or from mapping the contigs
-to reference genome. These unknown regions are useful to join non-contiguous
-sequences together by the estimated size of the region whilst still
-highlighting areas in the genome to be resolved. These unknown regions can be
-specified in the scaffold file for the above reasons. \pb
-
-The original raw nucleotide sequence for each assembled contig or insert
-sequence is maintained in a fasta file. The fasta file of contigs is a common
-output for assemblers and as such should be readily available after assembly
-without additional processing. The choice of the nucleotide sequence to use in
-the scaffold is specified using the first space-delimited word from the fasta
-header of the sequence. By maintaining the nucleotide sequences in a separate
-file this preserves the original raw sequence in the fasta file. Furthermore
-this separates the actual nucleotide sequence from the determination of how it
-is selected and edited in the final genome sequence. \pb
-
-\subsection*{Defining a scaffold in a text file} %{{{2
-
-The scaffold file is written using `YAML', a data format designed to be both
-machine readable and manually editable. An example of a scaffold file in YAML
-format is shown in Figure 1. This example scaffold file illustrates the
-available features in scaffolder. The basic layout of the scaffold file is as
-a list - a series of entries where each entry defines a region of sequence to
-be generated for the final super sequence. In the final super sequence each of
-the entries are joined together to generate a single continuous sequences. \pb
-
-\subsubsection*{Simple sequence region} %{{{3
-
-The first line of the scaffold file begins with three dashes. This is
-a requirement to indicate the start of a YAML formatted document. Line 2 begins
-with a dash character `-' denoting the first entry in the sequence scaffold.
-Line 3 is indented by two spaces. Whitespace is used to delimite attributes in
-the file. Lines with an incremented level of white space are considered linked
-to the first preceeding non-indented line. \pb
-
-The ``sequence'' tag used on line 3 indicates that this region is in the
-scaffold will contain nucleotide sequence from the fasta file. The following
-line indented by two spaces and indicating an attribute of the sequence entry
-defined on line 3. In this case the ``source'' keyword identifies the name of
-the sequence to be inserted into the scaffold. As described above the first
-space-delimited word of the fasta header should be used to identify the
-sequence. Taken together lines 2-4 describe the first entry in the scaffold is
-a simple sequence which the identifier `sequence'. \pb
-
-\subsubsection*{Unresolved sequence region} %{{{3
-
-Line 5 has no whitespace indentation and begins with a dash and therefore
-describes the next entry in the scaffold. This entry is identified by the
-``unresolved'' tag, and specifies a region of unknown nucleotides but known
-length. The following line specifies the size of this unknown region. Lines 5-6
-therefore insert a region of 20 `N' characters in the super sequence directly
-following the region defined by the first entry on lines 2-5. \pb
-
-\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3
-
-The last entry adds the sequence 'sequence2' in the scaffold file. This entry
-demonstrates three further attributes for a sequence entry.The `start' and
-`stop' tags are used to trim the sequence used in the scaffold to these
-coordinates inclusively. The ``reverse'' tag is used to reverse complement this
-region in the final super sequence. \pb
-
-This region further includes the ``inserts'' tag to update the inserted
-sequence with additional regions of sequence. These inserts are also added as
-a YAML list with each insert listed with a dash on one line followed by the
-remainder of the insert errata. \pb
-
-Lines 16-19 illustrate an insert using similar attributes to that of a sequence
-entry. The reverse, start and stop tags trim the inserted sequence and the
-`source' tag identifies the corresponding fasta sequence. The ``open'' and
-``close'' tags on lines 20-21 determine where the insert is added to the
-enclosing sequence.  These coordinates inclusively replace the region within
-these coordinates with the specified fasta sequence. The open and close
-positions need not be the same size as the inserted sequence. \pb
-
-The next insert on lines 23-24 specifies only the open position. This
-illustrate that only the one of the `open' or `close' tags is required.  When
-only open or close tag is used then the opposing position is calculated from
-the length on the insert fasta sequence. This allows inserts to partially fill
-gap regions in sequences. \pb
-
-\subsubsection*{Scaffolder file attribute processing} %{{{3
-
-The sequence entry in the scaffold file allows multiple sequence-editting
-attributes which when applied in different combinations will result in
-different output sequence. An example is reverse complementing the sequence
-then follwing adding sequence inserts. Performing these actions in the reverse
-order will produce a difference output sequence. This section therefore
-outlines the scaffolder protocol for processing sequence entries. \pb
-
-\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts these are reverse sorted by close position. The insert with the
-largest close position is first in the list and inserts with the smallest
-close position is last. This preserves the original indented coordinates for
-insert during the process of adding inserts to the sequence. \pb
-
-\paragraph{Update sequence with insert:} Each insert sequence is processed
-according to any specified attributes. The insert is added to the enclosing
-sequence in the region specified by the open and close coordinates. \pb
-
-If the insert added to the sequence is a different length than the region
-replaced the sequence start and stop co-ordinates are updated. For example if
-a 7bp insert replaced a 5bp region the stop coordinate of the enclosing
-sequence would be incremented by the 2bp difference. \pb
-
-\paragraph{Triming sequence:} If the sequence is has start or stop sequences
-specified these are used to trip the sequence to these co-ordinates
-inclusively. \pb
-
-\paragraph{Reverse complement sequence:} If the ``reverse'' keyword is
-specified the sequence is reverse complemented. This is final step in
-processing the sequence with results in the sequence being returned. \pb
-
-\subsection*{Scaffolder usage} %{{{2
-
-Scaffolder provides a standardised set of Ruby classes and methods (the
-application programming interface or API) allowing the scaffold to be
-manipulated programmatically. This allows scaffolder to be integrated into
-existing genomics workflows or used with Ruby build tools such as using Rake.
-\pb
-
-Scaffolder also provides a command line interface (CLI) through a scaffolder
-binary. This the scaffold sequence to be produced and validated without
-requiring any Ruby knowledge. The scaffolder CLI behaves as a standard Unix
-tool - returning appropriate exit statuses and providing appropriate manual
-pages. The use of both these scaffolder interfaces is outlined in the following
-sections. \pb
-
-\subsubsection*{Installation} %{{{3
-
-Scaffolder relies on both the ruby programming language and the RubyGems
-package management system to work. These are either already present easily
-installed on recent distributions of Mac OSX or Linux. The presences of these
-tools can be tested by typing the following on the command line. \pb
-
-\begin{verbatim}
-  ruby -v
-  gem -v
-\end{verbatim}
-
-If both these commands return without error this indicates the necessary
-system requirements to install and run scaffolder. The scaffolder API and CLI
-can then be installed through RubyGems with the following command.
-
-\begin{verbatim}
-  gem install scaffolder scaffolder-tools
-\end{verbatim}
-
-To ensure that the scaffolder CLI is accessible the RubyGems `bin' directory
-should be in the command line path. Assuming that RubyGems are installed in
-~/.gem this can be done by adding the following to the ~/.bashrc file. This
-should be adjusted for other shells such as zsh etc..
-
-\begin{verbatim}
-  export PATH=~/.gem/bin:$PATH
-\end{verbatim}
-
-\subsubsection*{Command Line Interface} %{{{3
-
-The Scaffolder CLI provides functions to validate the scaffold file and build the draft scaffold sequence. Assuming scaffolder has been installed and the
-binary is available in the command line path the following command should run without error and list the available functions. \pb
-
-\begin{verbatim}
-  scaffolder
-\end{verbatim}
-
-Help on a specific command can be found by typing `help' followed by the name
-of the command. This will return the corresponding Unix manual page. The
-following example will return the manual page for the sequence command. \pb
-
-\begin{verbatim}
-  scaffolder help sequence
-\end{verbatim}
-
-The ``sequence'' command builds the fasta sequence from the scaffold and fasta
-file. Running this command requires the locations of both the scaffold file and
-the fasta file of sequences. This will then return the complete scaffold
-sequence. \pb
-
-\begin{verbatim}
-  scaffolder sequence scaffold.yml sequences.fna
-\end{verbatim}
-
-Scaffolder provides a function to test the scaffold for overlapping insert
-coordinates. Overlapping inserts will result in unexpected behaviour as the
-coordinates for the first insert may change the insert location for subsequent
-inserts. The ``validate'' command will check the scaffold for any such inserts
-and return them as a YAML list. \pb
-
-\begin{verbatim}
-  scaffolder validate scaffold.yml sequences.fna
-\end{verbatim}
-
-\subsubsection*{Application programming interface} %{{{3
-
-Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
-This API is detailed in full on the scaffolder website. The following code
-snippet outlines how the API may be used to create a script to replicate the
-function of the `scaffolder sequence' command described above. \pb
-
-\begin{verbatim}
-  #!/usr/bin/env ruby
-
-  require 'rubygems'
-  require 'scaffolder'
-
-  scaffold_file = ARGV.shift
-  sequence_file = ARGV.shift
-
-  scaffold = Scaffolder.new(
-    YAML.load(File.read(scaffold_file)),
-    sequence_file)
-
-  sequence = scaffold.map{|entry| entry.sequence }.join
-
-  puts sequence
-\end{verbatim}
-
-This code snippet illustrates that the scaffold object is initialised with two
-arguments: the YAML loaded scaffold and the location of the fasta sequence
-file. The scaffold is an enumerable object as the next line illustrates the
-trivial code required to generate the scaffold sequence. Further details of
-the classes and method related to each scaffold entry can be found in the
-documentation for the Scaffolder::Region class. \pb
-
-\subsection*{Discussion of scaffolder approach} %{{{2
-
-The main aims of scaffolder are two fold. The first is usability and the second is reproducibility.
-
-\subsubsection{Simple to use}
-
-Scaffolder is designed to be as simple to use as possible. Assuming the
-existence of Ruby and RubyGems, scaffolder can be installed with a single
-command. Scaffolder uses a minimal and compact syntax to described the
-scaffold. This syntax is simple to construct and edit whilst being succinct and
-readable in describing how the scaffold was constructed. Scaffolder provides
-extensive documentation both for the command line tools and for the Scaffolder
-Ruby API.
-
-\subsubsection{Reproducible genome construction}
-
-Constructing a genome sequence using a plain text scaffold file means the same
-sequence is always output for the same file. The sequence produced from
-manually joining separate contigs in a text editor can not be reliably
-reproduced. Constructing the scaffold in a plain text file means that the
-sequence is reproducible. Furthermore the source of each region in final
-sequence can be identified. The scaffold file provides a readable record of
-how the sequence was constructed. The scaffold file also means the scaffold is
-easier to edit once already constructed. Much more difficult to do when working
-on the sequence alone. The YAML text format is also much easier to compare
-differences using standard diff commands. This makes scaffold files much more
-simple to store in version control and to see the stages in the construction
-of the genome sequence.
-
-\section*{Conclusions} %{{{1
-
-Scaffolder follows the Unix philosophy ``Do one thing and do it well''.
-Scaffolder focuses on creating a simple to use and install tool for generating
-a genome scaffold. The YAML data format for creating a scaffold file is
-standardised and easily manipulated programmatically. This thereby follows
-a second Unix tenet: ``If your data structures are good enough, the algorithm
-to manipulate them should be trivial.''
-
-\section*{Availability and Requirements} %{{{1
+  \subsection*{Yet another results sub-heading}
+    Text for this sub-section.  More results \ldots
+% }}}
+% {{{ Conclusions
+%%%%%%%%%%%%%%%%%%%%%%
+\section*{Conclusions}
+  Text for this section \ldots
+% }}}
+% {{{ Availability and Requirements
+\section*{Availability and Requirements}
 

### 2011-02-08 text/src/article.tex
Fix references for background

@@
 \emph{in silico} methods are preferable as they less costly in laboratory time
-and materials \cite{nagaryan2010}. \pb
+and materials \cite{nagarajan2010}. \pb
 
@@
 of the gaps between contigs may be reassembled using algorithms specifically
-for repeat resolution \cite{mulkov2002,koren2010}. Uniquely matching reads may
-be used to try and bridge gaps between scaffolds \cite{tsai2010}. Whole genome
-may use heuristic graph algorithms to build a path between contigs using
+for repeat resolution \cite{mulyukov2002,koren2010}. Uniquely matching reads
+may be used to try and bridge gaps between scaffolds \cite{tsai2010}. Whole
+genome may use heuristic graph algorithms to build a path between contigs using
 multiple input data \cite{pop2004,dayarian2010}. Finally when the scaffold
@@
 suggest the likely primers necessary for determining the sequence in gap
-regions \cite{gordon2001,nagaryan2010}. Taken together these tools provide
-a wide variety of possible options for scaffolding a genome should an
-assembler be unable to produce a continuous sequence from the sequencing
-reads. \pb
+regions \cite{gordon2001,nagarajan2010}. Taken together these tools provide
+a wide variety of possible options for scaffolding a genome should an assembler
+be unable to produce a continuous sequence from the sequencing reads. \pb
 

### 2011-02-08 text/src/article.tex
Two revisions of first background section

@@
 
-High throughput sequencing can be reads short regions of a genome sequence to
-produce hundreds of thousands to millions of lines of nucleotide sequence. The
-current status of parallel nucleotide sequencing is limited to producing reads
-less than $>$1000 nucleotides long. Therefore to obtain a contiguous complete
-genome sequence these sequence reads must therefore be joined together. \pb
+High throughput sequencing can produce hundreds of thousands to millions of
+sequence reads from a genome. Each read represents the machine-interpreted
+nucleotide sequence from a short region in the genome. At the time of writing
+high throughput sequencing is limited to producing reads less than $>$1000
+nucleotides. Therefore to produce to produce a sequence longer than this
+length, e.g. a complete genome, these numerous individual reads must be pieced
+together. The process of joining reads into longer sequences is the `assembly'
+stage in a genome project \cite{assembly-reviews}. \pb
 
-The process of joining sequencing reads together is described as the
-`assembly' stage in a genome project. Assembly software joins these numerous
-sequence reads together into contiguous regions (`contigs') composed of many
-overlapping individual reads. The goal of the assembly stage is produce
-a continuous genome sequence describing the complete genome from start to
-finish. \pb
+Assembly software takes the nucleotide reads produced by the sequencing
+hardware and, in the ideal case, outputs a complete genome sequence compose of
+these individual reads. An analogy to describe this is a jigsaw puzzle. Each
+nucleotide read represents a single piece and the final genome sequence is the
+completed puzzle. Repetitive nucleotide `repeat' regions within a genome or
+biased and incomplete sequencing data may however prevent the genome being
+assembled into a continuous sequence. This is analogous to missing jigsaw
+pieces or pieces that fit to multiple other pieces. \pb
 
-Often repetitive nucleotide `repeat' regions in the genome or biased and
-incomplete sequencing data may prevent the genome being assembled into
-a continuous sequence. This sequence data can however describe the majority of
-the genetic information in the genome and in the case of microbes and moving
-forward with the current sequence data at this stage may be the best
-investment of time and money \cite{branscomb2002}. \pb
+The recent advent of these current high throughput sequencing methods has lead
+to a renewed interest into algorithms for solving the problem of genome
+assembly \cite{assembly-reviews}. Genome assembly software is often unable to
+completely these sequence reads together due to problems outlined above.
+Instead assembly algorithms will generate large contiguous regions of sequence
+(`contigs') composed of many individual reads. These represent optimum assembly
+of reads into the longest sequences for the assembly algorithm. \pb
 
-A continuous high-quality `finished' genome sequence provides a much greater
-depth of information as compared to a draft sequence
-\cite{parkhill2002,fraser2002}. Examples of depth of information are complete
-resolution of repeat regions and the exact distances between genomic elements.
-The process of moving from a draft sequence to a finished continuous sequence
-involves the process of `scaffolding' the contigs produced by the assembly
-software together. \pb
+The process of ameliorating a draft sequence to a finished continuous sequence
+can be expensive in time and laboratory effort. Therefore in many cases a set
+of contigs may often describe enough of the genomic content for many research
+questions \cite{branscomb2002}. A continuous high-quality `finished' genome
+sequence does however provide further depth of information
+\cite{parkhill2002,fraser2002}. The benefits of a complete genome sequence
+include complete resolution of repeat regions and the exact distances between
+genomic elements. This process of joining contigs from assembly together to form
+a continuous sequence may be called `scaffolding' or `finishing'. Here on we
+use the term scaffolding. \pb
 

### 2011-02-08 text/src/article.tex
Fix quotation marks in abstract

--- 2011-02-08 text/src/article.tex
@@
 
-  \paragraph*{Methods:} The software outlined "Scaffolder" is implemented in
+  \paragraph*{Methods:} The software outlined ``Scaffolder'' is implemented in
   the Ruby programming language and can be installed via the RubyGems software
@@
 % }}}
-\section*{Background} %{{{1
-%{{{2
-
-High throughput sequencing can be reads short regions of a genome sequence to
-produce hundreds of thousands to millions of lines of nucleotide sequence. The
-current status of parallel nucleotide sequencing is limited to producing reads
-less than $>$1000 nucleotides long. Therefore to obtain a contiguous complete
-genome sequence these sequence reads must therefore be joined together. \pb
-
-The process of joining sequencing reads together is described as the
-`assembly' stage in a genome project. Assembly software joins these numerous
-sequence reads together into contiguous regions (`contigs') composed of many
-overlapping individual reads. The goal of the assembly stage is produce
-a continuous genome sequence describing the complete genome from start to
-finish. \pb
-
-Often repetitive nucleotide `repeat' regions in the genome or biased and
-incomplete sequencing data may prevent the genome being assembled into
-a continuous sequence. This sequence data can however describe the majority of
-the genetic information in the genome and in the case of microbes and moving
-forward with the current sequence data at this stage may be the best
-investment of time and money \cite{branscomb2002}. \pb
-
-A continuous high-quality `finished' genome sequence provides a much greater
-depth of information as compared to a draft sequence
-\cite{parkhill2002,fraser2002}. Examples of depth of information are complete
-resolution of repeat regions and the exact distances between genomic elements.
-The process of moving from a draft sequence to a finished continuous sequence
-involves the process of `scaffolding' the contigs produced by the assembly
-software together. \pb
-
-\subsection*{Scaffolding} %{{{2
-
-Scaffolding can be described as the process of moving from a series of
-disconnected contigs to complete continuous sequence of all the contigs joined
-together. The scaffolding process may not be completed but may still resolve
-and join some contigs together or add additional regions of sequence to fill
-gaps. In the case of a microbial genome, the process of scaffolding may
-include the following steps:
-
-\subsubsection*{Contig Orientation} %{{{3
-
-Sequence reads may be produced from either DNA strand during the sequencing
-process. Therefore the resulting contigs produced from these sequences may are
-likely to be representative from either of the genome strands. Contig
-orientation involves correcting the sequence data so that all contigs
-represent sequence data from the same DNA strand. In the case of microbes this
-orientation is often the 5`-3` strand from the origin of DNA replication.
-
-\subsubsection*{Contig Ordering} %{{{3
-
-The source location of each contig from the genome is unknown. Contig ordering
-represents determining the order of contigs which most accurately represent
-the actual genome sequence. This process can also determine if any contigs
-represent extra-genomic DNA such as plasmids and therefore should be separated
-from the genomic contigs. Contigs are usually ordered in reference to the
-origin of replication where the contig enclosing this region is the first
-contigs and subsequent contigs are labelled in the direction of DNA
-replication.
-
-\subsubsection*{Contig Distancing} %{{{3
-
-Once the order of contigs in the genome sequence is known it is useful to
-determine their approximate distance from each other. This allows the size of
-the gaps between contigs to be determined. Filling in these regions in with
-unknown nucleotide characters allows a draft continuous representation of the
-genome to determined.
-
-\subsubsection*{Gap Closing and Filling} %{{{3
-
-The final stage of producing a genome sequence is to close and fill gaps in
-the sequence. This may require returning to the laboratory to perform
-additional sequencing to close gaps or \emph{in silico} methods may be able to
-close gaps using existing data. Once all gaps in a scaffold have been closed
-the genome may be considered finished. Though additional sequencing may be
-performed to `proof' any low coverage regions for erroneous nucleotide
-insertions, deletions or substitutions.
-
-\subsection*{Computational Methods for Scaffolding} %{{{2
-
-Scaffolding may be considered a combination of \emph{in vitro} and \emph{in
-silico} methods. Several steps may be performed computationally by using
-available data to generate a scaffold. Examples may be the use of pair reads
-sequencing, where reads a known distance apart, or from comparing contigs to
-a reference genome. In the case where \emph{in silico} methods are unable to
-complete the scaffold in may be necessary to use PCR to amplify then use
-traditional Sanger sequencing to find the sequence and close gaps. Obviously
-\emph{in silico} methods are preferable as they less costly in laboratory time
-and materials \cite{nagaryan2010}. \pb
-
-For these reason may examples exist for software to generate genome scaffolds
-from assembly data. These software use a variety of approaches. Synteny
-approaches use a reference genome to order and orient contigs
-\cite{richter2007,zhao2008}. Repeat regions which may be responsible for many
-of the gaps between contigs may be reassembled using algorithms specifically
-for repeat resolution \cite{mulkov2002,koren2010}. Uniquely matching reads may
-be used to try and bridge gaps between scaffolds \cite{tsai2010}. Whole genome
-may use heuristic graph algorithms to build a path between contigs using
-multiple input data \cite{pop2004,dayarian2010}. Finally when the scaffold
-cannot be completely resolved \emph{in silico} software packages exist to
-suggest the likely primers necessary for determining the sequence in gap
-regions \cite{gordon2001,nagaryan2010}. Taken together these tools provide
-a wide variety of possible options for scaffolding a genome should an
-assembler be unable to produce a continuous sequence from the sequencing
-reads. \pb
-
-Scaffolding however will inevitably require manual editing if computational
-methods do not completely resolve the final genome sequence. This involves
-moving an joining larges blocks of nucleotide text together in what the genome
-finisher perceives to be the correct scaffold. This process may be expected to
-be error-prone. Further manual editing cannot be reliably reproduced which
-precludes reproducibility in a genome scaffolding process. \pb
-
-\subsection*{The Scaffolder Software} %{{{2
-
-The software outlined here ``Scaffolder'' is aimed at joining contigs together
-to build a genome scaffold. Scaffolder uses a specific data syntax to define
-how contigs are joined together. Furthermore scaffolder allows inserts into
-contigs and unknown regions in a scaffold to be defined. This syntax allows
-a scaffold to be edited in much easier way than editing large blocks of
-nucleotides. Further more this makes genome scaffolds reproducible.
-
-\section*{Implementation} %{{{1
-
+% {{{ Background
+\section*{Background}
+ Text for this section.\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
+% }}}
+% {{{ Implementation
+%%%%%%%%%%%%%%%%%%%%%%
+\section*{Implementation}
   Text for this section \ldots

### 2011-02-08 text/src/article.tex
Two revisions to second section of background

--- 2011-02-08 text/src/article.tex
@@
 
-Scaffolding can be described as the process of moving from a series of
-disconnected contigs to complete continuous sequence of all the contigs joined
-together. The scaffolding process may not be completed but may still resolve
-and join some contigs together or add additional regions of sequence to fill
-gaps. In the case of a microbial genome, the process of scaffolding may
-include the following steps:
+Scaffolding is the process of joining a series disconnected contigs into
+complete continuous sequence. Due to genomic complexity, missing data and
+diminishing returns scaffolding may not ultimately produce a final completed
+sequence. The process of scaffolding may still however joining a subset of
+contigs or additional regions of sequence to fill gaps.
 
@@
 
-Sequence reads may be produced from either DNA strand during the sequencing
-process. Therefore the resulting contigs produced from these sequences may are
-likely to be representative from either of the genome strands. Contig
-orientation involves correcting the sequence data so that all contigs
-represent sequence data from the same DNA strand. In the case of microbes this
-orientation is often the 5`-3` strand from the origin of DNA replication.
+The sequencing process produces reads from either strand of the DNA helix. The
+resulting contigs constructed from these reads therefore point in either of two
+DNA strand directions. Contig orientation involves reverse complementing
+sequences so that all contigs point in the same direction and represent the
+same DNA strand. In the case of microbes this orientation is often the 5`-3`
+strand leading away from the origin of replication.
 
@@
 
-The source location of each contig from the genome is unknown. Contig ordering
-represents determining the order of contigs which most accurately represent
-the actual genome sequence. This process can also determine if any contigs
-represent extra-genomic DNA such as plasmids and therefore should be separated
-from the genomic contigs. Contigs are usually ordered in reference to the
-origin of replication where the contig enclosing this region is the first
-contigs and subsequent contigs are labelled in the direction of DNA
-replication.
+Contig ordering determines the placement of contigs to best represent the
+actual nucleotides as appearing in the genome sequence. This may also determine
+if any contigs represent extra-genomic DNA, such as plasmids, and therefore
+should be treated separately from genomic contigs. The contig order is commonly
+referenced to the origin of replication where the contig containing the
+replication start point is first. All subsequent contigs then follow in the
+direction of DNA replication.
 
@@
 
-Once the order of contigs in the genome sequence is known it is useful to
-determine their approximate distance from each other. This allows the size of
-the gaps between contigs to be determined. Filling in these regions in with
-unknown nucleotide characters allows a draft continuous representation of the
-genome to determined.
+Given the order and orientation of contigs determining the distances between
+contigs may be useful for estimating genome size. The size of each gaps between
+contigs represents the length of an unknown region in the genome. Filling in
+these regions in with unknown nucleotide characters (`N') allows a draft
+continuous representation of the genome to produced. This draft scaffold
+sequence represents both the contigs and unknown regions of sequence between
+them.
 
@@
 
-The final stage of producing a genome sequence is to close and fill gaps in
-the sequence. This may require returning to the laboratory to perform
-additional sequencing to close gaps or \emph{in silico} methods may be able to
-close gaps using existing data. Once all gaps in a scaffold have been closed
-the genome may be considered finished. Though additional sequencing may be
-performed to `proof' any low coverage regions for erroneous nucleotide
-insertions, deletions or substitutions.
+During the scaffolding process closing and filling sequence gaps improves the
+accuracy of the genome scaffold. This may require returning to the laboratory
+to perform additional sequencing or using computational methods. Once all
+contigs have been joined and gaps in a scaffold closed the genome may be
+considered as finished. In the case of organisms or particular interest
+additional sequencing may be performed to further `proof' any low coverage
+regions or detect erroneous nucleotides resulting from insertions, deletions or
+substitutions.
 

### 2011-02-08 text/src/article.tex
Revise last section of background

@@
 
-Scaffolding may be considered a combination of \emph{in vitro} and \emph{in
-silico} methods. Several steps may be performed computationally by using
-available data to generate a scaffold. Examples may be the use of pair reads
-sequencing, where reads a known distance apart, or from comparing contigs to
-a reference genome. In the case where \emph{in silico} methods are unable to
-complete the scaffold in may be necessary to use PCR to amplify then use
-traditional Sanger sequencing to find the sequence and close gaps. Obviously
-\emph{in silico} methods are preferable as they less costly in laboratory time
-and materials \cite{nagarajan2010}. \pb
+The process of resolving a genome sequence and building a scaffold uses
+\emph{in vitro} methods, \emph{in silico} methods, or both in conjunction.
+Examples of computation methods is the use of pair reads data, where sequencing
+reads are a known distance apart in the genome, or from comparing contigs to
+a reference genome. Laboratory methods may use PCR to amplify the DNA in gap
+regions, then use traditional Sanger sequencing to determine this sequence.
+Computation methods are more preferable as they less costly in laboratory time
+and materials compared to manual gap resolution \cite{nagarajan2010}. \pb
 
-For these reason may examples exist for software to generate genome scaffolds
-from assembly data. These software use a variety of approaches. Synteny
-approaches use a reference genome to order and orient contigs
-\cite{richter2007,zhao2008}. Repeat regions which may be responsible for many
-of the gaps between contigs may be reassembled using algorithms specifically
-for repeat resolution \cite{mulyukov2002,koren2010}. Uniquely matching reads
-may be used to try and bridge gaps between scaffolds \cite{tsai2010}. Whole
-genome may use heuristic graph algorithms to build a path between contigs using
-multiple input data \cite{pop2004,dayarian2010}. Finally when the scaffold
-cannot be completely resolved \emph{in silico} software packages exist to
-suggest the likely primers necessary for determining the sequence in gap
-regions \cite{gordon2001,nagarajan2010}. Taken together these tools provide
-a wide variety of possible options for scaffolding a genome should an assembler
-be unable to produce a continuous sequence from the sequencing reads. \pb
+The need for computational scaffolding methods has resulted in a variety of
+algorithmic approaches. Synteny methods compare the assembled contigs to a more
+complete reference genome sequence. Areas of corresponding sequence between the
+the two sets are indicative of the possible contig organisation
+\cite{richter2007,zhao2008}. \pb
 
-Scaffolding however will inevitably require manual editing if computational
-methods do not completely resolve the final genome sequence. This involves
-moving an joining larges blocks of nucleotide text together in what the genome
-finisher perceives to be the correct scaffold. This process may be expected to
-be error-prone. Further manual editing cannot be reliably reproduced which
-precludes reproducibility in a genome scaffolding process. \pb
+Repeat regions may be responsible for many of the gaps in building a genome
+sequence. The existence of tandemly repeated nucleotide regions produces many
+reads of similar sequence. The assembly of these regions may result in the
+repeats being collapsed into a single instance or ignored by conservative
+assembly algorithms. These regions may be reassembled using algorithms
+specifically for repeat resolution \cite{mulyukov2002,koren2010}. A related
+approach uses uniquely matching reads try find overlaps and bridge across the
+gaps between contigs \cite{tsai2010}. \pb
 
-\subsection*{The Scaffolder Software} %{{{2
+As mentioned above, the genome may be prepared prior to sequencing so that
+reads are produced in pairs at a known distance apart. This paired-read data
+provides an extra level of information on how contigs may be scaffolded
+together. Heuristic graph algorithms may be used to search for optimal
+configuration of contigs in the scaffold to take into account this information
+\cite{dayarian2010}. Additional data using a reference genome may also be
+combined when estimating the best contig configuration \cite{pop2004}. \pb
 
-The software outlined here ``Scaffolder'' is aimed at joining contigs together
-to build a genome scaffold. Scaffolder uses a specific data syntax to define
-how contigs are joined together. Furthermore scaffolder allows inserts into
-contigs and unknown regions in a scaffold to be defined. This syntax allows
-a scaffold to be edited in much easier way than editing large blocks of
-nucleotides. Further more this makes genome scaffolds reproducible.
+Finally when the scaffold cannot be completely resolved \emph{in silico}
+software packages exist to suggest the likely primers necessary for determining
+the sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
+
+\subsection*{Summary} %{{{2
+
+Taken together these methods provide a wide variety of possible approaches for
+scaffolding a genome. Scaffolding however may still however require manual
+editing of the scaffold if computational methods do not resolve the final
+sequence.  This process involves moving and joining larges blocks of nucleotide
+text together. This however may be error-prone and manual editing also
+precludes reproducibility of the genome scaffold. \pb
+
+The software outlined here ``Scaffolder'' is aimed at creating a syntax and
+software framework for editing a genome scaffold. Scaffolder uses a specific
+data syntax to define how contigs are joined, additional sequences inserted,
+and specify unknown regions. This syntax allows a scaffold to be edited in much
+easier way and makes genome scaffolds reproducible. \pb
 

### 2011-02-09 text/src/article.tex

@@
 
-  Text for this section \ldots
-% }}}
-% {{{ Results and Discussion
+\subsection*{Code and Dependencies} %{{{2
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Results and Discussion %%
-%%
-\section*{Results and Discussion}
-  \subsection*{Results sub-heading}
-    \subsubsection*{This is a sub-sub-heading}
-      Sub-sub-sub-headings are made with the \textsl{\\subsubsection} command. \pb
-      pb at end of lines ensures correct paragraph spacing.\pb
-	  Text for this sub-sub-section \ldots
-    \subsubsection*{Another sub-sub-sub-heading}
-      Text for this sub-sub-section \ldots
+Scaffolder is written in the Ruby programming language, version 1.8.7
+\cite{ruby-lang}. Unit tests were implemented during development to maintain
+functionality of individual elements of source code. Integration tests were
+also written to test the scaffold software interface as a whole. These unit and
+integration tests were written using the RSpec and Cucumber libraries
+respectively\cite{rspec}.\pb
 
-  \subsection*{Another results sub-heading}
-    Text for this sub-section \ldots
+The Scaffolder source code is documented using the Yard library\cite{yard}.
+Unix manual pages for the command line interface were generated using the Ronn
+library\cite{ronn}. The manipulation of biological sequences in Scaffolder is
+performed using the BioRuby library\cite{bioruby}. A full list of the software
+dependencies in scaffolder can be found in the Gemfile in the root of source
+code directory.\pb
 
-  \subsection*{Yet another results sub-heading}
-    Text for this sub-section.  More results \ldots
-% }}}
-% {{{ Conclusions
-%%%%%%%%%%%%%%%%%%%%%%
-\section*{Conclusions}
-  Text for this section \ldots
-% }}}
-% {{{ Availability and Requirements
-\section*{Availability and Requirements}
+\subsection*{Scaffold file syntax} %{{{2
+
+The choice of nucleotide sequences comprising the genome scaffold is specified
+in YAML \cite{yaml}. YAML is a language format for data using white-space and
+indentation to produce a machine readable structure. As a standardised data
+format third-party developers have the option to generate a genome scaffold
+using any programming language for which a YAML parser exists. Furthermore YAML
+formatted scaffold files can be validated using third party tools such as
+Kwalify\cite{kwalify}.\pb
+
+The initial assembly of sequencing reads may result in an incomplete genome
+build. Comparison of contigs to reference genomes and the generation of
+additional sequence from PCR translates to genome finishing as an ongoing
+process. Therefore the scaffold file should be simple to edit manually in
+addition to being computationally tractable. This requirement for the scaffold
+files to be modifiable in a simple simple text editor is the second reason YAML
+was selected as the format for scaffold files.\pb
+
+The form of the scaffold file is a list of entries. Each entry corresponds to
+a length of sequence in the the draft genome sequence. When generating the
+genome sequence from the scaffold file each entry is processed in order. Each
+of the nucleotide sequence specified by the scaffold file entries are
+concatenated together to produce the draft genome sequence. Each entry in the
+scaffold file may have attributes that define whether a sub-sequence or the
+reverse complement of the sequence should be used. The types of attributes are
+outlined in the Results and Discussion section below.\pb
+
+\section*{Results and Discussion} %{{{1
+
+\subsection*{Scaffolder Simplifies Genome Finishing} %{{{2
+
+The Scaffolder software allows manual or computational joining of multiple
+discontiguous nucleotide sequences together into a contiguous super sequence.
+Simple plain-text configuration files are used to select the nucleotide
+sequences in the scaffold and how they should be joined together. This is
+called the ``scaffold file''. \pb
+
+The scaffold file is read by the scaffolder software to determine how the
+scaffold super sequence is generated. In addition to specifying which
+nucleotides should be used to generate the draft super sequence, the scaffold
+file allows the sequences to trimmed down to smaller sub-sequences and to
+reverse complement the sequence if necessary. \pb
+
+There may be unknown regions in the sequences used in the scaffold. Region of
+nucleotide sequence the assembler was unable to determine the nucleotide
+sequence. The process of genome finishing may involve determining the correct
+nucleotide sequence for these regions then filling the gaps with the additional
+correct sequence. The scaffold file specifically allows these inserts to be
+added to scaffold. These inserts can be treated the same way as a larger
+contig: trimmed and/or reversed to match the required gap region. \pb
+
+When scaffolding a genome there may be regions of sequence approximate size
+region may be estimated from mate-pair insert size or from mapping the contigs
+to reference genome. These unknown regions are useful to join non-contiguous
+sequences together by the estimated size of the region whilst still
+highlighting areas in the genome to be resolved. These unknown regions can be
+specified in the scaffold file for the above reasons. \pb
+
+The original raw nucleotide sequence for each assembled contig or insert
+sequence is maintained in a fasta file. The fasta file of contigs is a common
+output for assemblers and as such should be readily available after assembly
+without additional processing. The choice of the nucleotide sequence to use in
+the scaffold is specified using the first space-delimited word from the fasta
+header of the sequence. By maintaining the nucleotide sequences in a separate
+file this preserves the original raw sequence in the fasta file. Furthermore
+this separates the actual nucleotide sequence from the determination of how it
+is selected and edited in the final genome sequence. \pb
+
+\subsection*{Defining a scaffold in a text file} %{{{2
+
+The scaffold file is written using `YAML', a data format designed to be both
+machine readable and manually editable. An example of a scaffold file in YAML
+format is shown in Figure 1. This example scaffold file illustrates the
+available features in scaffolder. The basic layout of the scaffold file is as
+a list - a series of entries where each entry defines a region of sequence to
+be generated for the final super sequence. In the final super sequence each of
+the entries are joined together to generate a single continuous sequences. \pb
+
+\subsubsection*{Simple sequence region} %{{{3
+
+The first line of the scaffold file begins with three dashes. This is
+a requirement to indicate the start of a YAML formatted document. Line 2 begins
+with a dash character `-' denoting the first entry in the sequence scaffold.
+Line 3 is indented by two spaces. Whitespace is used to delimite attributes in
+the file. Lines with an incremented level of white space are considered linked
+to the first preceeding non-indented line. \pb
+
+The ``sequence'' tag used on line 3 indicates that this region is in the
+scaffold will contain nucleotide sequence from the fasta file. The following
+line indented by two spaces and indicating an attribute of the sequence entry
+defined on line 3. In this case the ``source'' keyword identifies the name of
+the sequence to be inserted into the scaffold. As described above the first
+space-delimited word of the fasta header should be used to identify the
+sequence. Taken together lines 2-4 describe the first entry in the scaffold is
+a simple sequence which the identifier `sequence'. \pb
+
+\subsubsection*{Unresolved sequence region} %{{{3
+
+Line 5 has no whitespace indentation and begins with a dash and therefore
+describes the next entry in the scaffold. This entry is identified by the
+``unresolved'' tag, and specifies a region of unknown nucleotides but known
+length. The following line specifies the size of this unknown region. Lines 5-6
+therefore insert a region of 20 `N' characters in the super sequence directly
+following the region defined by the first entry on lines 2-5. \pb
+
+\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3
+
+The last entry adds the sequence 'sequence2' in the scaffold file. This entry
+demonstrates three further attributes for a sequence entry.The `start' and
+`stop' tags are used to trim the sequence used in the scaffold to these
+coordinates inclusively. The ``reverse'' tag is used to reverse complement this
+region in the final super sequence. \pb
+
+This region further includes the ``inserts'' tag to update the inserted
+sequence with additional regions of sequence. These inserts are also added as
+a YAML list with each insert listed with a dash on one line followed by the
+remainder of the insert errata. \pb
+
+Lines 16-19 illustrate an insert using similar attributes to that of a sequence
+entry. The reverse, start and stop tags trim the inserted sequence and the
+`source' tag identifies the corresponding fasta sequence. The ``open'' and
+``close'' tags on lines 20-21 determine where the insert is added to the
+enclosing sequence.  These coordinates inclusively replace the region within
+these coordinates with the specified fasta sequence. The open and close
+positions need not be the same size as the inserted sequence. \pb
+
+The next insert on lines 23-24 specifies only the open position. This
+illustrate that only the one of the `open' or `close' tags is required.  When
+only open or close tag is used then the opposing position is calculated from
+the length on the insert fasta sequence. This allows inserts to partially fill
+gap regions in sequences. \pb
+
+\subsubsection*{Scaffolder file attribute processing} %{{{3
+
+The sequence entry in the scaffold file allows multiple sequence-editting
+attributes which when applied in different combinations will result in
+different output sequence. An example is reverse complementing the sequence
+then follwing adding sequence inserts. Performing these actions in the reverse
+order will produce a difference output sequence. This section therefore
+outlines the scaffolder protocol for processing sequence entries. \pb
+
+\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
+inserts these are reverse sorted by close position. The insert with the
+largest close position is first in the list and inserts with the smallest
+close position is last. This preserves the original indented coordinates for
+insert during the process of adding inserts to the sequence. \pb
+
+\paragraph{Update sequence with insert:} Each insert sequence is processed
+according to any specified attributes. The insert is added to the enclosing
+sequence in the region specified by the open and close coordinates. \pb
+
+If the insert added to the sequence is a different length than the region
+replaced the sequence start and stop co-ordinates are updated. For example if
+a 7bp insert replaced a 5bp region the stop coordinate of the enclosing
+sequence would be incremented by the 2bp difference. \pb
+
+\paragraph{Triming sequence:} If the sequence is has start or stop sequences
+specified these are used to trip the sequence to these co-ordinates
+inclusively. \pb
+
+\paragraph{Reverse complement sequence:} If the ``reverse'' keyword is
+specified the sequence is reverse complemented. This is final step in
+processing the sequence with results in the sequence being returned. \pb
+
+\subsection*{Scaffolder usage} %{{{2
+
+Scaffolder provides a standardised set of Ruby classes and methods (the
+application programming interface or API) allowing the scaffold to be
+manipulated programmatically. This allows scaffolder to be integrated into
+existing genomics workflows or used with Ruby build tools such as using Rake.
+\pb
+
+Scaffolder also provides a command line interface (CLI) through a scaffolder
+binary. This the scaffold sequence to be produced and validated without
+requiring any Ruby knowledge. The scaffolder CLI behaves as a standard Unix
+tool - returning appropriate exit statuses and providing appropriate manual
+pages. The use of both these scaffolder interfaces is outlined in the following
+sections. \pb
+
+\subsubsection*{Installation} %{{{3
+
+Scaffolder relies on both the ruby programming language and the RubyGems
+package management system to work. These are either already present easily
+installed on recent distributions of Mac OSX or Linux. The presences of these
+tools can be tested by typing the following on the command line. \pb
+
+\begin{verbatim}
+  ruby -v
+  gem -v
+\end{verbatim}
+
+If both these commands return without error this indicates the necessary
+system requirements to install and run scaffolder. The scaffolder API and CLI
+can then be installed through RubyGems with the following command.
+
+\begin{verbatim}
+  gem install scaffolder scaffolder-tools
+\end{verbatim}
+
+To ensure that the scaffolder CLI is accessible the RubyGems `bin' directory
+should be in the command line path. Assuming that RubyGems are installed in
+~/.gem this can be done by adding the following to the ~/.bashrc file. This
+should be adjusted for other shells such as zsh etc..
+
+\begin{verbatim}
+  export PATH=~/.gem/bin:$PATH
+\end{verbatim}
+
+\subsubsection*{Command Line Interface} %{{{3
+
+The Scaffolder CLI provides functions to validate the scaffold file and build the draft scaffold sequence. Assuming scaffolder has been installed and the
+binary is available in the command line path the following command should run without error and list the available functions. \pb
+
+\begin{verbatim}
+  scaffolder
+\end{verbatim}
+
+Help on a specific command can be found by typing `help' followed by the name
+of the command. This will return the corresponding Unix manual page. The
+following example will return the manual page for the sequence command. \pb
+
+\begin{verbatim}
+  scaffolder help sequence
+\end{verbatim}
+
+The ``sequence'' command builds the fasta sequence from the scaffold and fasta
+file. Running this command requires the locations of both the scaffold file and
+the fasta file of sequences. This will then return the complete scaffold
+sequence. \pb
+
+\begin{verbatim}
+  scaffolder sequence scaffold.yml sequences.fna
+\end{verbatim}
+
+Scaffolder provides a function to test the scaffold for overlapping insert
+coordinates. Overlapping inserts will result in unexpected behaviour as the
+coordinates for the first insert may change the insert location for subsequent
+inserts. The ``validate'' command will check the scaffold for any such inserts
+and return them as a YAML list. \pb
+
+\begin{verbatim}
+  scaffolder validate scaffold.yml sequences.fna
+\end{verbatim}
+
+\subsubsection*{Application programming interface} %{{{3
+
+Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
+This API is detailed in full on the scaffolder website. The following code
+snippet outlines how the API may be used to create a script to replicate the
+function of the `scaffolder sequence' command described above. \pb
+
+\begin{verbatim}
+  #!/usr/bin/env ruby
+
+  require 'rubygems'
+  require 'scaffolder'
+
+  scaffold_file = ARGV.shift
+  sequence_file = ARGV.shift
+
+  scaffold = Scaffolder.new(
+    YAML.load(File.read(scaffold_file)),
+    sequence_file)
+
+  sequence = scaffold.map{|entry| entry.sequence }.join
+
+  puts sequence
+\end{verbatim}
+
+This code snippet illustrates that the scaffold object is initialised with two
+arguments: the YAML loaded scaffold and the location of the fasta sequence
+file. The scaffold is an enumerable object as the next line illustrates the
+trivial code required to generate the scaffold sequence. Further details of
+the classes and method related to each scaffold entry can be found in the
+documentation for the Scaffolder::Region class. \pb
+
+\subsection*{Discussion of Scaffolder Approach} %{{{2
+
+The main aims of scaffolder are two foldfold. The first is produce easy to
+install and simplifies the task of genome finishing. The second aim is to
+allow to create a tool to facilitate reproducibility in the task of joining
+large nucleotide sequences together. \pb
+
+\subsubsection{Simple to use}
+
+Scaffolder is designed to be simple to use as possible. Assuming the existence
+of Ruby and RubyGems, scaffolder can be installed in a single command line
+step. Scaffolder uses a minimal and compact syntax to described the genome
+scaffold. This syntax is simple to construct and edit whilst being succinct
+and readable. This aims to allow creating a genome scaffold to relatively
+simple to do in a common text editor. Furthermore scaffolder provides
+extensive documentation both for the command line tools and for the Scaffolder
+Ruby API. Therefore should any part of the scaffolder tool be unclear the
+documentation should make an solution easy to find. \pb
+
+\subsubsection{Reproducible genome construction}
+
+Constructing a genome sequence using a plain text scaffold file means the same
+sequence is always output for the same file. Joining large nucleotide
+sequences manually cannot however be reliably reproduced. The layout of the
+scaffold file furthermore provides a human readable record of how the sequence
+was constructed. \pb
+
+Configuring the final sequence in the scaffold file means the build is easier
+to edit once already constructed compared with working on the sequence alone.
+The YAML text format is also much easier to compare differences using standard
+text comparison tools such as diff. This makes scaffold files easier to store
+in version control systems. This affords genome scaffolding the same benefits
+enjoyed by software developers. \pb
+
+\section*{Conclusions} %{{{1
+
+Scaffolder follows the Unix philosophy ``Do one thing and do it well''.
+Scaffolder focuses on creating a simple to use and install tool for generating
+a genome scaffold. The YAML data format for creating a scaffold file is
+standardised and easily manipulated programmatically. This thereby follows
+a second Unix tenet: ``If your data structures are good enough, the algorithm
+to manipulate them should be trivial.''
+
+\section*{Availability and Requirements} %{{{1
 
@@
 % }}}
+% {{{ Availability and Requirements
+\section*{List of Abbreviations Used}
+
+  \begin{description}
+    \item[YAML:] YAML Ain't Markup Language\cite{yaml}
+  \end{description}
+
+% }}}
 % {{{ Author Contributions
@@
   Text for this section \ldots
-%}}}
-% {{{ Bibliography
+%{{{1 Bibliography
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}

### 2011-02-09 text/src/article.tex

@@
-% {{{ Preamble
-
-%% BioMed_Central_Tex_Template_v1.05
-%%                                      %
-%  bmc_article.tex            ver: 1.05 %
-%                                       %
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%         <27 January 2006>           %%
-%%                                     %%
-%%                                     %%
-%% Uses:                               %%
-%% cite.sty, url.sty, bmc_article.cls  %%
-%% ifthen.sty. multicol.sty		       %%
-%%									   %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%	
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.pdf and the instructions for    %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% BioMed_Central_Tex_Template_v1.05 %{{{1
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%   
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %% 
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %% 
-%%  submitted article.                         %% 
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \def\includegraphic{}
@@
 \newboolean{publ}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                              %%
-%% You may change the following style settings  %%
-%% Should you wish to format your article       %%
-%% in a publication style for printing out and  %%
-%% sharing with colleagues, but ensure that     %%
-%% before submitting to BMC that the style is   %%
-%% returned to the Review style setting.        %%
-%%                                              %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{bmcformat}
-% }}}
-% {{{ Title
 
-\title{Scaffolder - Software for Microbial Genome Scaffolding.}
+\title{Scaffolder - Software for Microbial Genome Scaffolding.} %{{{1
 
@@
 \maketitle
-%}}}
-% {{{ Abstract
 
-\begin{abstract}
+\begin{abstract} %{{{1
 
@@
 
-% }}}
 \section*{Background} %{{{1
@@
 % }}}
-% {{{ Competing Interests
-\section*{Authors contributions}
+\section*{Authors contributions} %{{{1
   The authors declare no competing interests.
-% }}}
-% {{{ Availability and Requirements
-\section*{List of Abbreviations Used}
+
+\section*{List of Abbreviations Used} %{{{1
 
@@
 
-% }}}
-% {{{ Author Contributions
-\section*{Authors contributions}
+\section*{Authors contributions} %{{{1
 
@@
 
-% }}}
-% {{{ Author's Information
-\section*{Authors contributions}
-    Text for this section \ldots
-% }}}
-% {{{ Acknowledgements
-\section*{Acknowledgements}
+\section*{Acknowledgements} %{{{1
   \ifthenelse{\boolean{publ}}{\small}{}
   Text for this section \ldots
-%{{{1 Bibliography
+
+% Bibliography {{{1
 {\ifthenelse{\boolean{publ}}{\footnotesize}{\small}

### 2011-02-09 text/src/article.tex
Spell check

@@
 with a dash character `-' denoting the first entry in the sequence scaffold.
-Line 3 is indented by two spaces. Whitespace is used to delimite attributes in
+Line 3 is indented by two spaces. Whitespace is used to delimit attributes in
 the file. Lines with an incremented level of white space are considered linked
-to the first preceeding non-indented line. \pb
+to the first preceding non-indented line. \pb
 
@@
 
-The sequence entry in the scaffold file allows multiple sequence-editting
+The sequence entry in the scaffold file allows multiple sequence-editing
 attributes which when applied in different combinations will result in
 different output sequence. An example is reverse complementing the sequence
-then follwing adding sequence inserts. Performing these actions in the reverse
+then following adding sequence inserts. Performing these actions in the reverse
 order will produce a difference output sequence. This section therefore
@@
 
-\paragraph{Triming sequence:} If the sequence is has start or stop sequences
+\paragraph{Trimming sequence:} If the sequence is has start or stop sequences
 specified these are used to trip the sequence to these co-ordinates
@@
 ~/.gem this can be done by adding the following to the ~/.bashrc file. This
-should be adjusted for other shells such as zsh etc..
+should be adjusted for other shells such as Zsh etc..
 
@@
 
-The main aims of scaffolder are two foldfold. The first is produce easy to
+The main aims of scaffolder are two fold. The first is produce easy to
 install and simplifies the task of genome finishing. The second aim is to

### 2011-02-09 text/src/article.tex
Delete tables section
Add text for figure 1

@@
 
-  \subsection*{Figure 1 - Sample figure title}
+\subsection*{Figure 1 - Example of scaffolder file}
 
-  A short description of the figure content should go here.
-
-\section*{Tables} %{{{1
-
-  \subsection*{Table 1 - Sample table title}
-    Here is an example of a \emph{small} table in \LaTeX\ using  
-    \verb|\tabular{...}|. This is where the description of the table 
-    should go. \par \mbox{}
-    \par
-    \mbox{
-      \begin{tabular}{|c|c|c|}
-        \hline \multicolumn{3}{|c|}{My Table}\\ \hline
-        A1 & B2  & C3 \\ \hline
-        A2 & ... & .. \\ \hline
-        A3 & ..  & .  \\ \hline
-      \end{tabular}
-      }
-  \subsection*{Table 2 - Sample table title}
-    Large tables are attached as separate files but should
-    still be described here.
+An example scaffold file written using the YAML syntax \cite{yaml}. The
+scaffold file contains three entries. Each entry is delimited by a `-' on a new
+line. The first entry is an example adding a simple sequence. The second entry
+inserts an unresolved gap region into the scaffold. The final entry illustrates
+a more complex examplea where the sequece is trimmed, reversed and has
+additional sequence inserted \pb
 

### 2011-02-09 text/src/article.tex
Delete additional files section

@@
 a more complex examplea where the sequece is trimmed, reversed and has
-additional sequence inserted \pb
-
-\section*{Additional Files} %{{{1
-
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
-
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
+additional sequence inserted. \pb
 

### 2011-02-09 text/src/article.tex
Fix spacing in title address
Fix incorrect quotes

@@
 
-\address{\iid(1) Department of Biological Sciences, Northern Kentucky%
+\address{\iid(1) Department of Biological Sciences, Northern Kentucky %
 University, Nunn Drive, Highland Heights, KY 41076 }%
@@
 \maketitle
+
+\clearpage
 
@@
 \ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}
+
+\clearpage
 
@@
 
+\clearpage
+
 \section*{Implementation} %{{{1
@@
 outlined in the Results and Discussion section below.\pb
+
+\clearpage
 
@@
 
-The last entry adds the sequence 'sequence2' in the scaffold file. This entry
+The last entry adds the sequence `sequence2' in the scaffold file. This entry
 demonstrates three further attributes for a sequence entry.The `start' and
@@
 
+\clearpage
+
 \section*{Conclusions} %{{{1
@@
 to manipulate them should be trivial.''
+
+\clearpage
 
@@
 
-% }}}
+\clearpage
+
 \section*{Authors contributions} %{{{1
@@
 
+\clearpage
+
 % Bibliography {{{1
@@
 \ifthenelse{\boolean{publ}}{\end{multicols}}{}
+
+\clearpage
 

### 2011-02-09 text/src/article.tex
Remove numbering from sections
Revise implementation

@@
 Scaffolder is written in the Ruby programming language, version 1.8.7
-\cite{ruby-lang}. Unit tests were implemented during development to maintain
-functionality of individual elements of source code. Integration tests were
-also written to test the scaffold software interface as a whole. These unit and
-integration tests were written using the RSpec and Cucumber libraries
-respectively\cite{rspec}.\pb
+\cite{ruby-lang}. The scaffolder package is split into two libraries. The first
+called ``scaffolder'' contains the core scaffolder application programming
+interface (API). The second library ``scaffolder-tools'' contains the
+scaffolder command line interface (CLI). \pb
 
-The Scaffolder source code is documented using the Yard library\cite{yard}.
+Unit tests were implemented to develop and maintain individual elements of the
+source code.  Integration tests were written to test the scaffolder software
+interface as a whole. Unit tests were written using the Shoulda and RSpec
+\cite{rspec} libraries.  Integration tests were written using the Cucumber
+library \cite{rspec}. \pb
+
+The Scaffolder source code is documented using the Yard library \cite{yard}.
 Unix manual pages for the command line interface were generated using the Ronn
-library\cite{ronn}. The manipulation of biological sequences in Scaffolder is
-performed using the BioRuby library\cite{bioruby}. A full list of the software
-dependencies in scaffolder can be found in the Gemfile in the root of source
-code directory.\pb
+library \cite{ronn}. The manipulation of biological sequences in Scaffolder is
+performed using the BioRuby library \cite{goto2010}. A full list of the
+software dependencies in scaffolder can be found in the Gemfile in the root of
+each source code directory.\pb
 
@@
 
-The choice of nucleotide sequences comprising the genome scaffold is specified
-in YAML \cite{yaml}. YAML is a language format for data using white-space and
-indentation to produce a machine readable structure. As a standardised data
-format third-party developers have the option to generate a genome scaffold
-using any programming language for which a YAML parser exists. Furthermore YAML
-formatted scaffold files can be validated using third party tools such as
-Kwalify\cite{kwalify}.\pb
+The choice of nucleotide sequences making up the scaffold is specified using
+the YAML syntax \cite{yaml}. YAML is a language format using whitespace and
+indentation to produce a machine readable structure. As YAML is a standardised
+data format, third-party developers have the option to generate a genome
+scaffold using any programming language for which a YAML parser and generator
+exists. The YAML website lists parsers for languages including C/C++, Ruby,
+Python, Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
+supported YAML formatted scaffold files can be validated for correct syntax
+using third party tools such as Kwalify \cite{kwalify}. \pb
 
-The initial assembly of sequencing reads may result in an incomplete genome
-build. Comparison of contigs to reference genomes and the generation of
-additional sequence from PCR translates to genome finishing as an ongoing
-process. Therefore the scaffold file should be simple to edit manually in
-addition to being computationally tractable. This requirement for the scaffold
-files to be modifiable in a simple simple text editor is the second reason YAML
-was selected as the format for scaffold files.\pb
+As described in the Background section, the initial sequencing data assembly
+may result in an incomplete genome build. The use of computational software and
+generation of additional sequence using PCR translates into genome scaffolding
+as an on-going process. Therefore the scaffold file should be simple to edit
+update with new sequence manually in addition to being computationally
+tractable. This requirement was therefore second reason YAML was selected,
+where YAML is human-readable and should be relatively easy to update in a
+standard text editor. \pb
 
-The form of the scaffold file is a list of entries. Each entry corresponds to
-a length of sequence in the the draft genome sequence. When generating the
-genome sequence from the scaffold file each entry is processed in order. Each
-of the nucleotide sequence specified by the scaffold file entries are
+The scaffold file takes the form of a list of entries. Each entry corresponds
+to a region of sequence, most likely a contig, in the genome scaffold. The
+scaffolder software processes this list of entries in order, converting each to
+the corresponding nucleotide sequence. These nucleotide sequences are then
 concatenated together to produce the draft genome sequence. Each entry in the
 scaffold file may have attributes that define whether a sub-sequence or the
-reverse complement of the sequence should be used. The types of attributes are
-outlined in the Results and Discussion section below.\pb
+reverse complement of the sequence should be used. The types of attributes
+available and an example scaffold file are outlined in the Results and
+Discussion section below.  \pb
 
@@
 
-\subsubsection{Simple to use}
+\subsubsection*{Simple to use}
 
@@
 
-\subsubsection{Reproducible genome construction}
+\subsubsection*{Reproducible genome construction}
 
@@
   \begin{description}
+    \item[API:] Application programming interface
+    \item[CLI:] Command line interface
     \item[YAML:] YAML Ain't Markup Language\cite{yaml}

### 2011-02-09 src/article.tex
Move text directory structure up one directory

--- 2011-02-09 text/src/article.tex
@@
 
-\section*{Authors contributions} %{{{1
+\section*{Competing Interests} %{{{1
+
 The authors declare no competing interests.
@@
   \ifthenelse{\boolean{publ}}{\small}{}
-  Text for this section \ldots
+
+This work was supported by the National Institute for Health: IDeA Network of
+Biomedical Research Excellence (KY-INBRE) grant: NIH 2P20 RR016481-09 and the
+NIH R15 AREA Program grant: NIH R15GM079775.
 

### 2011-02-09 src/article.tex
Revise results section

@@
 
-The Scaffolder software allows manual or computational joining of multiple
-discontiguous nucleotide sequences together into a contiguous super sequence.
-Simple plain-text configuration files are used to select the nucleotide
-sequences in the scaffold and how they should be joined together. This is
-called the ``scaffold file''. \pb
+The Scaffolder software outlined here is designed to allows manual or
+computational joining of nucleotide sequences together into a single contiguous
+super sequence. Plain-text files written using the YAML syntax files are used
+to specify the nucleotide sequences in the scaffold and how they should be
+joined together. This file is described as the ``scaffold file''. \pb
 
 The scaffold file is read by the scaffolder software to determine how the
-scaffold super sequence is generated. In addition to specifying which
-nucleotides should be used to generate the draft super sequence, the scaffold
-file allows the sequences to trimmed down to smaller sub-sequences and to
-reverse complement the sequence if necessary. \pb
+scaffold sequence should be generated. In addition to specifying which
+nucleotides should be used to generate this draft sequence, the scaffold file
+allows the sequences to trimmed down to smaller sub-sequences and reverse
+complemented if necessary. \pb
 
-There may be unknown regions in the sequences used in the scaffold. Region of
-nucleotide sequence the assembler was unable to determine the nucleotide
-sequence. The process of genome finishing may involve determining the correct
-nucleotide sequence for these regions then filling the gaps with the additional
-correct sequence. The scaffold file specifically allows these inserts to be
-added to scaffold. These inserts can be treated the same way as a larger
-contig: trimmed and/or reversed to match the required gap region. \pb
+The process of genome finishing may involve producing additional nucleotide
+sequence to fill gaps in the scaffold. The scaffold file format allows these
+additional insert sequences to be added to scaffold to bridge the gaps inside
+contigs. These inserts can also be treated in a similar way to larger contig
+sequences: trimmed and/or reverse complemented to match the corresponding gap
+region. \pb
 
-When scaffolding a genome there may be regions of sequence approximate size
-region may be estimated from mate-pair insert size or from mapping the contigs
-to reference genome. These unknown regions are useful to join non-contiguous
-sequences together by the estimated size of the region whilst still
-highlighting areas in the genome to be resolved. These unknown regions can be
-specified in the scaffold file for the above reasons. \pb
+In the process of build a genome scaffold distances between contigs may be
+estimated from paired read insert size or from mapping the contigs to reference
+genome. These unknown gap regions are useful to join non-contiguous sequences
+together by the estimated size of the region whilst still highlighting areas in
+the scaffold to be resolved. The scaffold file also allows for the
+specification of these unknown regions. \pb
 
-The original raw nucleotide sequence for each assembled contig or insert
-sequence is maintained in a fasta file. The fasta file of contigs is a common
-output for assemblers and as such should be readily available after assembly
-without additional processing. The choice of the nucleotide sequence to use in
-the scaffold is specified using the first space-delimited word from the fasta
-header of the sequence. By maintaining the nucleotide sequences in a separate
-file this preserves the original raw sequence in the fasta file. Furthermore
-this separates the actual nucleotide sequence from the determination of how it
-is selected and edited in the final genome sequence. \pb
+The scaffold file specifies how the contigs nucleotide sequences are joined
+together, however the sequences are not present in the scaffold file. These
+nucleotide sequences are instead maintained in a fasta file. A fasta file of
+contigs is a common output for assemblers and as such should be readily
+available without further processing. Nucleotide sequence are specified in the
+scaffold using the first space-delimited word from the fasta header.
+Maintaining the nucleotide sequences in a separate file preserves the original
+unedited sequence and furthermore separates raw sequence from the markup
+describing how it should be used in the scaffold. \pb
 
@@
 
-The scaffold file is written using `YAML', a data format designed to be both
-machine readable and manually editable. An example of a scaffold file in YAML
-format is shown in Figure 1. This example scaffold file illustrates the
-available features in scaffolder. The basic layout of the scaffold file is as
-a list - a series of entries where each entry defines a region of sequence to
-be generated for the final super sequence. In the final super sequence each of
-the entries are joined together to generate a single continuous sequences. \pb
+The scaffold file is written using the YAML syntax. YAML is a data format
+designed to be both machine readable and easy to edit manually. An example
+scaffold file is shown in Figure 1. This file illustrates the text attributes
+used to describe a genome scaffold. The basic layout of the scaffold file is as
+a list - a series of entries where each entry insert to a region of sequence in
+the scaffold super sequence. \pb
 
@@
 a requirement to indicate the start of a YAML formatted document. Line 2 begins
-with a dash character `-' denoting the first entry in the sequence scaffold.
-Line 3 is indented by two spaces. Whitespace is used to delimit attributes in
-the file. Lines with an incremented level of white space are considered linked
-to the first preceding non-indented line. \pb
+with a dash character `-', dashes are used to denote entries in a YAML list.
+This first line therefore indicates indicate the first entry in the scaffold.
+Line 3 is indented by two spaces where whitespace is used to group similar
+attributes together. Lines indented by whitespace are linked to the first
+non-indented preceding line. \pb
 
-The ``sequence'' tag used on line 3 indicates that this region is in the
-scaffold will contain nucleotide sequence from the fasta file. The following
-line indented by two spaces and indicating an attribute of the sequence entry
-defined on line 3. In this case the ``source'' keyword identifies the name of
-the sequence to be inserted into the scaffold. As described above the first
-space-delimited word of the fasta header should be used to identify the
-sequence. Taken together lines 2-4 describe the first entry in the scaffold is
-a simple sequence which the identifier `sequence'. \pb
+The ``sequence'' tag on line 3 indicates this scaffold entry corresponds to an
+entry in the fasta file. The following line, indented by two spaces, indicates
+the name of the fasta sequence using the ``source'' keyword. As described above
+the first space-delimited word of the fasta header should be used to identify
+which sequence is used. Therefore lines 2-4 describe the first entry in the
+scaffold as a simple sequence using the fasta sequence named `sequence'. \pb
 
@@
 
-Line 5 has no whitespace indentation and begins with a dash and therefore
-describes the next entry in the scaffold. This entry is identified by the
+Line 5 has no whitespace indentation and begins with a dash therefore
+describing the next entry in the scaffold. This entry is identified by the
 ``unresolved'' tag, and specifies a region of unknown nucleotides but known
 length. The following line specifies the size of this unknown region. Lines 5-6
-therefore insert a region of 20 `N' characters in the super sequence directly
-following the region defined by the first entry on lines 2-5. \pb
+therefore insert a region of 20 `N' characters in the scaffold. \pb
 
@@
 
-The last entry adds the sequence `sequence2' in the scaffold file. This entry
-demonstrates three further attributes for a sequence entry.The `start' and
-`stop' tags are used to trim the sequence used in the scaffold to these
-coordinates inclusively. The ``reverse'' tag is used to reverse complement this
-region in the final super sequence. \pb
+The last entry adds the fasta entry `sequence2' to the scaffold. This entry
+demonstrates how sequence may be edited when added to the scaffold. The `start'
+and `stop' tags trim the sequence to these coordinates inclusively. The
+``reverse'' tag reverse complements this region. \pb
 
@@
 ``close'' tags on lines 20-21 determine where the insert is added to the
-enclosing sequence.  These coordinates inclusively replace the region within
-these coordinates with the specified fasta sequence. The open and close
-positions need not be the same size as the inserted sequence. \pb
+enclosing sequence.  These coordinates inclusively replace the sequence region
+within these coordinates with the specified insert fasta sequence. The open and
+close positions need not be the same size as the replaced region. \pb
 
-The next insert on lines 23-24 specifies only the open position. This
-illustrate that only the one of the `open' or `close' tags is required.  When
-only open or close tag is used then the opposing position is calculated from
-the length on the insert fasta sequence. This allows inserts to partially fill
-gap regions in sequences. \pb
+The next insert on lines 23-24 is specified by only the open position tag. This
+illustrates that only the one of the `open' or `close' tags is required.  When
+only one of these tag is used the corresponding opposing coordinate position is
+calculated from the length in the insert fasta sequence. This allows inserts to
+bridge into and partially fill gap regions. \pb
 
@@
 The sequence entry in the scaffold file allows multiple sequence-editing
-attributes which when applied in different combinations will result in
-different output sequence. An example is reverse complementing the sequence
-then following adding sequence inserts. Performing these actions in the reverse
-order will produce a difference output sequence. This section therefore
-outlines the scaffolder protocol for processing sequence entries. \pb
+operations to be applied. If applied in different combinations, these
+operations could result in different output sequence. An example would be
+reverse complementing the sequence then trimming the start of the sequence by
+five nucleotides. Performing these actions in the reverse order would produce
+a difference output sequence. This section therefore outlines the scaffolder
+protocol for processing sequence entries. \pb
 
 \paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts these are reverse sorted by close position. The insert with the
-largest close position is first in the list and inserts with the smallest
-close position is last. This preserves the original indented coordinates for
-insert during the process of adding inserts to the sequence. \pb
+inserts these are reverse sorted by the insert close coordinate. The insert
+with the largest close position is first in the list and the insert with the
+smallest close position is last. This preserves the original indented
+coordinates for each insert during the process of adding inserts to the
+sequence. \pb
 
 \paragraph{Update sequence with insert:} Each insert sequence is processed
-according to any specified attributes. The insert is added to the enclosing
-sequence in the region specified by the open and close coordinates. \pb
+according to any specified attributes. The insert is then added to the
+enclosing sequence in the region specified by the open and close coordinates.
+\pb
 
-If the insert added to the sequence is a different length than the region
-replaced the sequence start and stop co-ordinates are updated. For example if
-a 7bp insert replaced a 5bp region the stop coordinate of the enclosing
-sequence would be incremented by the 2bp difference. \pb
+If the insert added to the sequence is a different size compared to the region
+it replaces then the sequence start and stop co-ordinates are updated.  For
+example if a 7bp insert replaced a 5bp region in the sequence the stop
+coordinate of the enclosing sequence would be incremented by the 2bp
+difference. As inserts are added by reverse close position order the remaining
+inserts are unaffected by this change in the sequence size. \pb
 
-\paragraph{Trimming sequence:} If the sequence is has start or stop sequences
-specified these are used to trip the sequence to these co-ordinates
+\paragraph{Trimming sequence:} If the sequence has start or stop sequences
+specified these are used to trim the sequence to the specified length
 inclusively. \pb
 
-\paragraph{Reverse complement sequence:} If the ``reverse'' keyword is
-specified the sequence is reverse complemented. This is final step in
-processing the sequence with results in the sequence being returned. \pb
+\paragraph{Reverse complement sequence:} If the reverse tag is specified the
+sequence is reverse complemented. This is final step in processing the sequence
+and results in the sequence being returned for inclusion in the scaffold. \pb
 
@@
 
-Scaffolder provides a standardised set of Ruby classes and methods (the
-application programming interface or API) allowing the scaffold to be
-manipulated programmatically. This allows scaffolder to be integrated into
-existing genomics workflows or used with Ruby build tools such as using Rake.
-\pb
+Scaffolder provides a standardised set of Ruby classes and methods (API)
+allowing the scaffold to be manipulated programmatically. This allows
+scaffolder to be integrated into existing genomics workflows or used with Ruby
+build tools such as Rake. \pb
 
-Scaffolder also provides a command line interface (CLI) through a scaffolder
-binary. This the scaffold sequence to be produced and validated without
-requiring any Ruby knowledge. The scaffolder CLI behaves as a standard Unix
-tool - returning appropriate exit statuses and providing appropriate manual
-pages. The use of both these scaffolder interfaces is outlined in the following
-sections. \pb
+In addition Scaffolder provides a command line interface (CLI) through the
+scaffolder binary. This allows the scaffold sequence to be produced and
+validated without requiring any Ruby experience. The scaffolder CLI behaves as
+a standard Unix tool - returning appropriate exit statuses and providing
+appropriate manual pages. The use of both these scaffolder interfaces is
+outlined in the following sections. \pb
 
@@
 
-Scaffolder relies on both the ruby programming language and the RubyGems
-package management system to work. These are either already present easily
-installed on recent distributions of Mac OSX or Linux. The presences of these
+Scaffolder relies on both the Ruby programming language and the RubyGems
+package management system. These are either already present or easily
+installed on recent distributions of Mac OSX or Linux. The presence of these
 tools can be tested by typing the following on the command line. \pb
@@
 system requirements to install and run scaffolder. The scaffolder API and CLI
-can then be installed through RubyGems with the following command.
+can then be installed through RubyGems with the following command. \pb
 
@@
 
-To ensure that the scaffolder CLI is accessible the RubyGems `bin' directory
+To ensure that the scaffolder CLI is accessible, the RubyGems `bin' directory
 should be in the command line path. Assuming that RubyGems are installed in
-~/.gem this can be done by adding the following to the ~/.bashrc file. This
-should be adjusted for other shells such as Zsh etc..
+~/.gem this can be done by adding the following to the ~/.bashrc file. The
+corresponding login shell should be updated for other shells such as Zsh, etc..
+\pb
 
@@
 
-The Scaffolder CLI provides functions to validate the scaffold file and build the draft scaffold sequence. Assuming scaffolder has been installed and the
-binary is available in the command line path the following command should run without error and list the available functions. \pb
+The Scaffolder CLI provides functions to validate the scaffold file and build
+the draft scaffold sequence. Assuming scaffolder has been installed and the
+binary is available in the command line path the following command should run
+without error and list the available scaffolder functions. \pb
 
@@
 Help on a specific command can be found by typing `help' followed by the name
-of the command. This will return the corresponding Unix manual page. The
-following example will return the manual page for the sequence command. \pb
+of the command. This returns the corresponding Unix manual page. The following
+example will return the manual page for the sequence command. \pb
 
@@
 file. Running this command requires the locations of both the scaffold file and
-the fasta file of sequences. This will then return the complete scaffold
-sequence. \pb
+the fasta file to be passed as argument. This will then return the complete
+scaffold sequence in fasta format. \pb
 
@@
 snippet outlines how the API may be used to create a script to replicate the
-function of the `scaffolder sequence' command described above. \pb
+function of the `scaffolder sequence' command described above in the CLI
+section. \pb
 
@@
 arguments: the YAML loaded scaffold and the location of the fasta sequence
-file. The scaffold is an enumerable object as the next line illustrates the
-trivial code required to generate the scaffold sequence. Further details of
-the classes and method related to each scaffold entry can be found in the
+file. The scaffold is an enumerable object and the next line illustrates the
+trivial code required to generate the scaffold sequence. Further details of the
+classes and method related to each scaffold entry can be found in the
 documentation for the Scaffolder::Region class. \pb
@@
 
-The main aims of scaffolder are two fold. The first is produce easy to
-install and simplifies the task of genome finishing. The second aim is to
-allow to create a tool to facilitate reproducibility in the task of joining
-large nucleotide sequences together. \pb
+The aim of scaffolder are twofold. The first aim is to provide scaffolding
+software which is easy to install and simplifies the task of genome finishing.
+The second aim is to facilitate reproducibility when joining large nucleotide
+sequences together to create a genome scaffold. \pb
 
-\subsubsection*{Simple to use}
+\subsubsection*{Simple to use} %{{{3
 
@@
 of Ruby and RubyGems, scaffolder can be installed in a single command line
-step. Scaffolder uses a minimal and compact syntax to described the genome
-scaffold. This syntax is simple to construct and edit whilst being succinct
-and readable. This aims to allow creating a genome scaffold to relatively
-simple to do in a common text editor. Furthermore scaffolder provides
-extensive documentation both for the command line tools and for the Scaffolder
-Ruby API. Therefore should any part of the scaffolder tool be unclear the
-documentation should make an solution easy to find. \pb
+step. This should negate the possible barriers to use in requiring manual
+compilation of the source code. \pb
 
-\subsubsection*{Reproducible genome construction}
+Scaffolder uses a minimal and compact syntax to described the genome scaffold.
+This syntax is simple to construct and edit whilst being succinct and readable.
+The aim of this is genome scaffold creation to relatively simple in a common
+text editor. Furthermore scaffolder provides extensive documentation both for
+the command line tools and for the Scaffolder Ruby API. Therefore should any
+part of the scaffolding process be unclear the documentation should make an
+solution easy to find. \pb
 
-Constructing a genome sequence using a plain text scaffold file means the same
-sequence is always output for the same file. Joining large nucleotide
+\subsubsection*{Reproducible genome construction} %{{{3
+
+Constructing a genome sequence using a text file to specify the scaffold means
+the same sequence is always output for the same file. Joining large nucleotide
 sequences manually cannot however be reliably reproduced. The layout of the
 scaffold file furthermore provides a human readable record of how the sequence
-was constructed. \pb
+was constructed. The same cannot be said for large block of nucleotide
+sequences comprised of many source contigs. \pb
 
 Configuring the final sequence in the scaffold file means the build is easier
-to edit once already constructed compared with working on the sequence alone.
-The YAML text format is also much easier to compare differences using standard
-text comparison tools such as diff. This makes scaffold files easier to store
-in version control systems. This affords genome scaffolding the same benefits
-enjoyed by software developers. \pb
+to edit once already constructed compared with working with an already drafted
+sequence. The YAML text format is also facilitates easier comparison of
+differences between scaffold build using standard text comparison tools such as
+diff. This makes scaffold files easier to store in version control systems and
+affords genome finishers similar benefits enjoyed by software developers. \pb
 

### 2011-02-15 src/article.tex
Typo in figure 1 legend
Revise manuscript

@@
 \author{
-  Michael D Barton$^{1}$%
-  \email{Michael D Barton - mail@michaelbarton.me.uk}%
+  Michael D Barton$^{1}$\correspondingauthor%
+  \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
 \and
-  Hazel A Barton\correspondingauthor$^1$%
-  \email{Hazel A Barton\correspondingauthor - bartonh@nku.edu}%
+  Hazel A Barton$^1$%
+  \email{Hazel A Barton - bartonh@nku.edu}%
       }
@@
 
-  \paragraph*{Background:} Assembly of short read sequencing data can result in
-  a fragmented series of contigs. Therefore a common step in a genome project
-  is to join neighbouring sequence regions together and fill gaps using insert
-  data generated through PCR. This scaffolding step, however, is non-trivial
+  \paragraph*{Background:} Assembly of short read sequencing data can result
+  in a fragmented series of contigs. Therefore a common step in a genome
+  project is to join neighbouring sequence regions together and fill gaps
+  using additional sequence. This scaffolding step, however, is non-trivial
   and requires manually editing large blocks of nucleotide sequence. Joining
   sequence contigs together also hides the source of each region in the final
-  genome sequence. Taken together, these considerations may make reproducing or
-  editing an existing genome build difficult.
+  genome sequence. Taken together, these considerations may make reproducing
+  or editing an existing genome build difficult.
 
-  \paragraph*{Methods:} The software outlined ``Scaffolder'' is implemented in
-  the Ruby programming language and can be installed via the RubyGems software
-  management system. Genome scaffolds are defined using YAML - a markup
-  language, which is both human and machine-readable. Command line binaries and
-  extensive documentation are provided.
+  \paragraph*{Methods:} The software outlined here ``Scaffolder'' is
+  implemented in the Ruby programming language and can be installed via the
+  RubyGems software management system. Genome scaffolds are defined using
+  YAML - a markup language, which is both human and machine-readable. Command
+  line binaries and extensive documentation are available.
 
@@
   terms of the constituent contigs using a relatively simple to write and edit
-  scaffold file syntax. The Scaffolder syntax further allows inserts to be
-  added to the scaffold for the purpose of filling unknown sequence regions.
+  scaffold file syntax. This syntax further allows  unknown regions to be
+  defined, and inserts to be added to the scaffold to fill contig gaps.
   Defining the genome construction in a file makes the scaffolding process
-  reproducible and comparatively easier to edit than raw fasta.
+  reproducible and comparatively easier to edit and compare than raw fasta.
 
@@
 sequence reads from a genome. Each read represents the machine-interpreted
-nucleotide sequence from a short region in the genome. At the time of writing
+nucleotide sequence from a short region in the genome. At the time of writing,
 high throughput sequencing is limited to producing reads less than $>$1000
-nucleotides. Therefore to produce to produce a sequence longer than this
-length, e.g. a complete genome, these numerous individual reads must be pieced
-together. The process of joining reads into longer sequences is the `assembly'
-stage in a genome project \cite{assembly-reviews}. \pb
+nucleotides long. Therefore a sequence longer than this length, e.g.
+a complete genome, these numerous individual reads must be pieced together.
+The process of joining reads into longer sequences is the `assembly' stage in
+a genome project \cite{miller2010}. \pb
 
@@
 hardware and, in the ideal case, outputs a complete genome sequence compose of
-these individual reads. An analogy to describe this is a jigsaw puzzle. Each
+these individual reads. An analogy for this process is a jigsaw puzzle. Each
 nucleotide read represents a single piece and the final genome sequence is the
@@
 to a renewed interest into algorithms for solving the problem of genome
-assembly \cite{assembly-reviews}. Genome assembly software is often unable to
-completely these sequence reads together due to problems outlined above.
-Instead assembly algorithms will generate large contiguous regions of sequence
-(`contigs') composed of many individual reads. These represent optimum assembly
-of reads into the longest sequences for the assembly algorithm. \pb
+assembly \cite{pop2008,pop2009}. Current genome assembly software is however
+often unable to produce a complete sequence from these reads due to these
+problems. In many cases assembly algorithms will generate several large
+contiguous regions of sequence (`contigs') composed of many individual reads.
+These contigs represent the assembly of reads into the longest possible
+sequences. These contigs represent a fragmented picture of the genome and
+require additional work to piece together. These contigs represent
+a fragmented picture of the genome and require additional work to piece
+together. \pb
 
@@
 questions \cite{branscomb2002}. A continuous high-quality `finished' genome
-sequence does however provide further depth of information
-\cite{parkhill2002,fraser2002}. The benefits of a complete genome sequence
-include complete resolution of repeat regions and the exact distances between
-genomic elements. This process of joining contigs from assembly together to form
-a continuous sequence may be called `scaffolding' or `finishing'. Here on we
-use the term scaffolding. \pb
+sequence does however provide further depth of information including complete
+resolution of repeat regions and the exact distances between genomic elements
+\cite{parkhill2002,fraser2002}. This process of joining contigs together to
+form a continuous sequence may be called the `scaffolding' or `finishing'
+stage. \pb
 
@@
 Scaffolding is the process of joining a series disconnected contigs into
-complete continuous sequence. Due to genomic complexity, missing data and
-diminishing returns scaffolding may not ultimately produce a final completed
-sequence. The process of scaffolding may still however joining a subset of
-contigs or additional regions of sequence to fill gaps.
+complete continuous sequence. Due to genomic complexity and missing data,
+scaffolding may not ultimately produce a final completed sequence. The process
+of scaffolding may still however succeed in joining a subset of contigs
+together or using additional regions of sequence to fill gaps. An overview of
+the scaffolding process is outlined in the following section.
 
@@
 
-The sequencing process produces reads from either strand of the DNA helix. The
-resulting contigs constructed from these reads therefore point in either of two
-DNA strand directions. Contig orientation involves reverse complementing
-sequences so that all contigs point in the same direction and represent the
-same DNA strand. In the case of microbes this orientation is often the 5`-3`
-strand leading away from the origin of replication.
+The sequencing process produces reads from either strand of the DNA helix.
+Therefore the resulting contigs constructed from these reads may point in
+either direction of the DNA strand. Contig orientation reverse complements
+sequences where necessary so that all contigs point in the same direction and
+represent the same DNA strand. In the case of microbes this orientation may be
+the 5'-3' strand leading away from the origin of replication.
 
@@
 
-Contig ordering determines the placement of contigs to best represent the
-actual nucleotides as appearing in the genome sequence. This may also determine
-if any contigs represent extra-genomic DNA, such as plasmids, and therefore
-should be treated separately from genomic contigs. The contig order is commonly
-referenced to the origin of replication where the contig containing the
-replication start point is first. All subsequent contigs then follow in the
-direction of DNA replication.
+Contig ordering determines the placement of contigs to best represent their
+order in the true genome sequence. Correctly placing each contig may also
+highlight any extra-genomic DNA, such as plasmids, and which should be treated
+separately from genomic sequence. The contig order is commonly referenced
+against the origin of replication where the contig containing the replication
+start point is first. All subsequent contigs are then ordered in the direction
+of DNA replication.
 
@@
 
-Given the order and orientation of contigs determining the distances between
-contigs may be useful for estimating genome size. The size of each gaps between
-contigs represents the length of an unknown region in the genome. Filling in
-these regions in with unknown nucleotide characters (`N') allows a draft
-continuous representation of the genome to produced. This draft scaffold
-sequence represents both the contigs and unknown regions of sequence between
-them.
+Given the order and orientation of contigs, determining the correct distances
+between contigs leads to an estimate of the genome size. The size of inter
+contig gaps represents the length of an unknown region in the genome. Filling
+in these regions in with unknown nucleotide characters (`N') allows a draft
+continuous representation of the genome. This draft scaffold sequence is
+useful for representing both the sequenced and unknown regions in the genome.
 
-\subsubsection*{Gap Closing and Filling} %{{{3
+\subsubsection*{Gap Closing} %{{{3
 
@@
 accuracy of the genome scaffold. This may require returning to the laboratory
-to perform additional sequencing or using computational methods. Once all
-contigs have been joined and gaps in a scaffold closed the genome may be
-considered as finished. In the case of organisms or particular interest
-additional sequencing may be performed to further `proof' any low coverage
-regions or detect erroneous nucleotides resulting from insertions, deletions or
-substitutions.
+to perform additional sequencing or using computational methods of estimating
+the gap sequence. These additional sequences can be used to bridge and close
+the gaps between contigs. Once all contigs have been joined and gaps in
+a scaffold closed the genome may be considered finished.
 
@@
 The process of resolving a genome sequence and building a scaffold uses
-\emph{in vitro} methods, \emph{in silico} methods, or both in conjunction.
-Examples of computation methods is the use of pair reads data, where sequencing
-reads are a known distance apart in the genome, or from comparing contigs to
-a reference genome. Laboratory methods may use PCR to amplify the DNA in gap
-regions, then use traditional Sanger sequencing to determine this sequence.
-Computation methods are more preferable as they less costly in laboratory time
-and materials compared to manual gap resolution \cite{nagarajan2010}. \pb
+\emph{in vitro} methods, \emph{in silico} methods, or a combination of both.
+An example of a computational method uses pair read data, where the sequencing
+reads are a known distance apart in the genome, when paired reads are in
+separate contigs this can to estimate the order and distance between them.
+Laboratory methods may use PCR to amplify the unknown DNA in gap regions,
+followed by traditional Sanger sequencing. Computation methods are more
+preferable as they less costly in laboratory time and materials compared to
+manual gap resolution \cite{nagarajan2010}. The need for computational
+scaffolding methods has resulted in a variety of algorithmic approaches. \pb
 
-The need for computational scaffolding methods has resulted in a variety of
-algorithmic approaches. Synteny methods compare the assembled contigs to a more
-complete reference genome sequence. Areas of corresponding sequence between the
-the two sets are indicative of the possible contig organisation
-\cite{richter2007,zhao2008}. \pb
+Synteny methods use sequence similarity to compare the assembled contigs to
+a more complete reference genome sequence. Areas of corresponding sequence
+between the the two sets are indicative of the possible contig organisation in
+the true genome \cite{richter2007,zhao2008}. \pb
 
@@
 specifically for repeat resolution \cite{mulyukov2002,koren2010}. A related
-approach uses uniquely matching reads try find overlaps and bridge across the
-gaps between contigs \cite{tsai2010}. \pb
+approach uses uniquely matching reads to try and bridge across contig gaps
+between contigs \cite{tsai2010}. \pb
 
 As mentioned above, the genome may be prepared prior to sequencing so that
-reads are produced in pairs at a known distance apart. This paired-read data
+reads are produced in pairs at a known distance. This paired-read data
 provides an extra level of information on how contigs may be scaffolded
-together. Heuristic graph algorithms may be used to search for optimal
-configuration of contigs in the scaffold to take into account this information
+together. Heuristic graph algorithms may take advantage of this data to search
+for optimal configuration of contigs in the scaffold to take this into account
 \cite{dayarian2010}. Additional data using a reference genome may also be
@@
 Taken together these methods provide a wide variety of possible approaches for
-scaffolding a genome. Scaffolding however may still however require manual
-editing of the scaffold if computational methods do not resolve the final
-sequence.  This process involves moving and joining larges blocks of nucleotide
-text together. This however may be error-prone and manual editing also
-precludes reproducibility of the genome scaffold. \pb
+scaffolding a genome. The scaffold may however still require manual editing
+and addition of sequence if computational methods are not completely
+successful. This process involves moving and joining larges blocks of
+nucleotide text together. This process may be expected to be error-prone and
+precludes any reproducibility of the scaffold. \pb
 
 The software outlined here ``Scaffolder'' is aimed at creating a syntax and
-software framework for editing a genome scaffold. Scaffolder uses a specific
-data syntax to define how contigs are joined, additional sequences inserted,
-and specify unknown regions. This syntax allows a scaffold to be edited in much
-easier way and makes genome scaffolds reproducible. \pb
+software framework for producing and editing a genome scaffold. Scaffolder
+uses a specific data syntax to define how contigs are joined, additional
+sequences inserted, and for the specification of unknown regions. This syntax
+allows a scaffold to be edited in a simpler way and allows the scaffolding
+process to be reproducible and deterministic. \pb
 
@@
 \cite{ruby-lang}. The scaffolder package is split into two libraries. The first
-called ``scaffolder'' contains the core scaffolder application programming
+called ``scaffolder'' contains the core Scaffolder application programming
 interface (API). The second library ``scaffolder-tools'' contains the
-scaffolder command line interface (CLI). \pb
+Scaffolder command line interface (CLI). \pb
 
-Unit tests were implemented to develop and maintain individual elements of the
-source code.  Integration tests were written to test the scaffolder software
-interface as a whole. Unit tests were written using the Shoulda and RSpec
-\cite{rspec} libraries.  Integration tests were written using the Cucumber
-library \cite{rspec}. \pb
+Unit tests were implemented to develop and maintain the individual elements of
+the source code.  Integration tests were written to test the Scaffolder
+software interface as a whole. Unit tests were written using the Shoulda and
+RSpec \cite{rspec} libraries.  Integration tests were written using the
+Cucumber library \cite{rspec}. \pb
 
@@
 performed using the BioRuby library \cite{goto2010}. A full list of the
-software dependencies in scaffolder can be found in the Gemfile in the root of
-each source code directory.\pb
+software dependencies in Scaffolder can be found in the Gemfile in the root of
+each source code directory. \pb
 
@@
 
-The choice of nucleotide sequences making up the scaffold is specified using
+The choice of nucleotide sequences comprising the scaffold is specified using
 the YAML syntax \cite{yaml}. YAML is a language format using whitespace and
@@
 scaffold using any programming language for which a YAML parser and generator
-exists. The YAML website lists parsers for languages including C/C++, Ruby,
-Python, Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
-supported YAML formatted scaffold files can be validated for correct syntax
-using third party tools such as Kwalify \cite{kwalify}. \pb
+exists. The YAML website lists current parsers for languages including C/C++,
+Ruby, Python, Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being
+widely supported YAML formatted scaffold files can be validated for correct
+syntax using third party tools such as Kwalify \cite{kwalify}. \pb
 
@@
 as an on-going process. Therefore the scaffold file should be simple to edit
-update with new sequence manually in addition to being computationally
-tractable. This requirement was therefore second reason YAML was selected,
-where YAML is human-readable and should be relatively easy to update in a
-standard text editor. \pb
+update manually in addition to being computationally tractable. This
+requirement was therefore the second reason YAML was selected, where YAML is
+human-readable and should be relatively easy to manipulate in a standard text
+editor. \pb
 
@@
 to a region of sequence, most likely a contig, in the genome scaffold. The
-scaffolder software processes this list of entries in order, converting each to
+Scaffolder software processes this list of entries in order, converting each to
 the corresponding nucleotide sequence. These nucleotide sequences are then
-concatenated together to produce the draft genome sequence. Each entry in the
+concatenated together to produce the draft super sequence. Each entry in the
 scaffold file may have attributes that define whether a sub-sequence or the
@@
 
-The Scaffolder software outlined here is designed to allows manual or
-computational joining of nucleotide sequences together into a single contiguous
-super sequence. Plain-text files written using the YAML syntax files are used
-to specify the nucleotide sequences in the scaffold and how they should be
-joined together. This file is described as the ``scaffold file''. \pb
+The Scaffolder software outlined here facilitates or computational joining of
+nucleotide sequences together into a single contiguous super sequence.
+Plain-text scaffold files written in the YAML syntax are used to specify how
+nucleotide sequences should be joined together. This scaffold file is read by
+the software to determine how the draft scaffold sequence should be generated.
+In addition to specifying which contigs should be used, the specification of
+the scaffold file allows the contigs to trimmed down to smaller sub-sequences
+and reverse complemented if necessary. \pb
 
-The scaffold file is read by the scaffolder software to determine how the
-scaffold sequence should be generated. In addition to specifying which
-nucleotides should be used to generate this draft sequence, the scaffold file
-allows the sequences to trimmed down to smaller sub-sequences and reverse
-complemented if necessary. \pb
+The process of genome finishing may involve producing additional
+oligonucleotide sequence to fill areas of unknown sequence in the scaffold.
+The scaffold file format allows these additional insert sequences to be added
+and bridge gaps. These inserts can also be treated in a similar way to larger
+contig sequences: trimmed and/or reverse complemented to match the
+corresponding gap region. \pb
 
-The process of genome finishing may involve producing additional nucleotide
-sequence to fill gaps in the scaffold. The scaffold file format allows these
-additional insert sequences to be added to scaffold to bridge the gaps inside
-contigs. These inserts can also be treated in a similar way to larger contig
-sequences: trimmed and/or reverse complemented to match the corresponding gap
-region. \pb
+In the process of building a genome scaffold distances between contigs may be
+estimated from paired read data or from mapping the contigs to a reference
+genome. These inter contig gap regions are useful to join regions of sequence
+together by an estimated size. The scaffold file also allows for the
+specification of these unknown regions to highlight areas in the genome
+sequence still to be resolved. \pb
 
-In the process of build a genome scaffold distances between contigs may be
-estimated from paired read insert size or from mapping the contigs to reference
-genome. These unknown gap regions are useful to join non-contiguous sequences
-together by the estimated size of the region whilst still highlighting areas in
-the scaffold to be resolved. The scaffold file also allows for the
-specification of these unknown regions. \pb
+The scaffold file is used to specify how sequence regions are joined together,
+however the sequences themselves are not stored in the scaffold file. The
+sequences are instead maintained in a fasta file. A fasta file of contigs and
+scaffolded contigs is a common output for assemblers and as such should be
+readily available without further processing required. Nucleotide sequence are
+reverenced in the scaffold using the first space-delimited word from each fasta
+header. Maintaining the nucleotide sequences in a separate file preserves the
+original sequence and furthermore decouples the raw sequence from the markup
+describing how it should be used to generate the genome sequence. \pb
 
-The scaffold file specifies how the contigs nucleotide sequences are joined
-together, however the sequences are not present in the scaffold file. These
-nucleotide sequences are instead maintained in a fasta file. A fasta file of
-contigs is a common output for assemblers and as such should be readily
-available without further processing. Nucleotide sequence are specified in the
-scaffold using the first space-delimited word from the fasta header.
-Maintaining the nucleotide sequences in a separate file preserves the original
-unedited sequence and furthermore separates raw sequence from the markup
-describing how it should be used in the scaffold. \pb
-
-\subsection*{Defining a scaffold in a text file} %{{{2
+\subsection*{Defining a Scaffold in a Text File} %{{{2
 
@@
 scaffold file is shown in Figure 1. This file illustrates the text attributes
-used to describe a genome scaffold. The basic layout of the scaffold file is as
-a list - a series of entries where each entry insert to a region of sequence in
-the scaffold super sequence. \pb
+used to describe a scaffold and the corresponding genome build. The basic
+layout of the scaffold file is a list - a series of entries where each entry
+corresponds to a region of sequence in the scaffold super sequence. \pb
 
@@
 The first line of the scaffold file begins with three dashes. This is
-a requirement to indicate the start of a YAML formatted document. Line 2 begins
-with a dash character `-', dashes are used to denote entries in a YAML list.
-This first line therefore indicates indicate the first entry in the scaffold.
-Line 3 is indented by two spaces where whitespace is used to group similar
-attributes together. Lines indented by whitespace are linked to the first
-non-indented preceding line. \pb
+a requirement to indicate the start of a YAML formatted document. The first
+entry in the scaffold begins on Line 2. The first line of this entry,
+highlighted in green, begins with a dash character `-', dashes are used to
+denote entries in a YAML list. This first line therefore indicates indicate
+the first entry in the scaffold. The first line of each entry in the scaffold
+begins with a dash line. The second line of this entry indented by two spaces
+where whitespace is used to group similar attributes together. Lines indented
+by whitespace are linked to the first non-indented preceding line in the file.
+\pb
 
-The ``sequence'' tag on line 3 indicates this scaffold entry corresponds to an
-entry in the fasta file. The following line, indented by two spaces, indicates
-the name of the fasta sequence using the ``source'' keyword. As described above
-the first space-delimited word of the fasta header should be used to identify
-which sequence is used. Therefore lines 2-4 describe the first entry in the
-scaffold as a simple sequence using the fasta sequence named `sequence'. \pb
+The ``sequence'' tag indicates this entry corresponds to a sequence in the
+fasta file. The following line indicates the name of the fasta sequence using
+the ``source'' tag. As described above the first space-delimited word of the
+fasta header is used to identify which sequence. Therefore these three lines
+describe the first entry in the scaffold as a simple sequence using the fasta
+sequence named `sequence1'. \pb
 
@@
 
-Line 5 has no whitespace indentation and begins with a dash therefore
-describing the next entry in the scaffold. This entry is identified by the
+The second entry in the scaffold, highlighted in orange, is identified by the
 ``unresolved'' tag, and specifies a region of unknown nucleotides but known
-length. The following line specifies the size of this unknown region. Lines 5-6
-therefore insert a region of 20 `N' characters in the scaffold. \pb
+length. The following line in the entry specifies the size of this unknown
+region. This entry therefore inserts a region of 20 `N' characters in the
+scaffold. \pb
 
@@
 
-The last entry adds the fasta entry `sequence2' to the scaffold. This entry
-demonstrates how sequence may be edited when added to the scaffold. The `start'
-and `stop' tags trim the sequence to these coordinates inclusively. The
-``reverse'' tag reverse complements this region. \pb
+The last entry in the scaffold, highlighted in blue, adds the fasta entry
+`sequence2' to the scaffold. This entry demonstrates how the nucleotide
+sequence may be manipulated prior to addition to the scaffold. The `start' and
+`stop' tags trim the sequence to these coordinates inclusively. The
+``reverse'' tag reverse complements the sequence. \pb
 
-This region further includes the ``inserts'' tag to update the inserted
-sequence with additional regions of sequence. These inserts are also added as
-a YAML list with each insert listed with a dash on one line followed by the
-remainder of the insert errata. \pb
+This region further includes the ``inserts'' tag to update this sequence with
+regions of sequence. These inserts are also added as a YAML list with each
+insert beginning with a dash line. The first insert, shown in purple,
+illustrates using similar attributes to that of a sequence entry. The reverse,
+start and stop tags, are used to trim and reverse complement the insert.
+Similarly the `source' tag identifies the corresponding fasta sequence. The
+``open'' and ``close'' tags determine where the insert is added in the
+enclosing sequence. These coordinates inclusively replace the sequence region
+within these coordinates with the specified insert fasta sequence. \pb
 
-Lines 16-19 illustrate an insert using similar attributes to that of a sequence
-entry. The reverse, start and stop tags trim the inserted sequence and the
-`source' tag identifies the corresponding fasta sequence. The ``open'' and
-``close'' tags on lines 20-21 determine where the insert is added to the
-enclosing sequence.  These coordinates inclusively replace the sequence region
-within these coordinates with the specified insert fasta sequence. The open and
-close positions need not be the same size as the replaced region. \pb
-
-The next insert on lines 23-24 is specified by only the open position tag. This
-illustrates that only the one of the `open' or `close' tags is required.  When
-only one of these tag is used the corresponding opposing coordinate position is
-calculated from the length in the insert fasta sequence. This allows inserts to
-bridge into and partially fill gap regions. \pb
+The next insert, shown in brown, specifies only the open position tag. This
+illustrates that only one of either `open' or `close' tags is required. When
+only one of these tag is used the corresponding opposing coordinate position
+is calculated from the length in the insert fasta sequence. This allows
+inserts to bridge into, and partially fill gap regions. \pb
 
@@
 five nucleotides. Performing these actions in the reverse order would produce
-a difference output sequence. This section therefore outlines the scaffolder
+a difference output sequence. This section therefore outlines the Scaffolder
 protocol for processing sequence entries. \pb
@@
 \paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts these are reverse sorted by the insert close coordinate. The insert
-with the largest close position is first in the list and the insert with the
-smallest close position is last. This preserves the original indented
-coordinates for each insert during the process of adding inserts to the
+inserts, these inserts are reverse sorted by the insert close coordinate. The
+insert with the largest close position is first in the list and the insert
+with the smallest close position is last. This preserves the original intended
+coordinates for each insert during the process of adding each insert to the
 sequence. \pb
 
-\paragraph{Update sequence with insert:} Each insert sequence is processed
+\paragraph{Update sequence with inserts:} Each insert sequence is processed
 according to any specified attributes. The insert is then added to the
@@
 
-If the insert added to the sequence is a different size compared to the region
-it replaces then the sequence start and stop co-ordinates are updated.  For
+If the insert added to the sequence is a different size to that of the
+replaced region then the sequence start and stop co-ordinates are updated. For
 example if a 7bp insert replaced a 5bp region in the sequence the stop
@@
 difference. As inserts are added by reverse close position order the remaining
-inserts are unaffected by this change in the sequence size. \pb
+inserts are unaffected by changes in the encoding sequence size. \pb
 
-\paragraph{Trimming sequence:} If the sequence has start or stop sequences
-specified these are used to trim the sequence to the specified length
-inclusively. \pb
+\paragraph{Trimming sequence:} If the sequence has start or stop tags
+specified these are used as coordinates to trim the sequence to the specified
+length inclusively. \pb
 
 \paragraph{Reverse complement sequence:} If the reverse tag is specified the
-sequence is reverse complemented. This is final step in processing the sequence
-and results in the sequence being returned for inclusion in the scaffold. \pb
+sequence is reverse complemented. This is final step in processing the
+sequence and the sequence is then returned for inclusion in the scaffold. \pb
 
-\subsection*{Scaffolder usage} %{{{2
+\subsection*{Scaffolder Usage} %{{{2
 
-Scaffolder provides a standardised set of Ruby classes and methods (API)
-allowing the scaffold to be manipulated programmatically. This allows
-scaffolder to be integrated into existing genomics workflows or used with Ruby
-build tools such as Rake. \pb
+Scaffolder provides a standardised set of Ruby classes and methods (API) for
+manipulation of the scaffold. This allows Scaffolder to be integrated into
+existing genomics workflows or used with Ruby build tools such as Rake. \pb
 
-In addition Scaffolder provides a command line interface (CLI) through the
-scaffolder binary. This allows the scaffold sequence to be produced and
-validated without requiring any Ruby experience. The scaffolder CLI behaves as
-a standard Unix tool - returning appropriate exit statuses and providing
-appropriate manual pages. The use of both these scaffolder interfaces is
-outlined in the following sections. \pb
+In addition Scaffolder provides a command line interface (CLI). This allows
+the scaffold sequence to be generated and validated without any Ruby
+experience. The Scaffolder CLI behaves as a standard Unix tool - returning
+appropriate exit codes and providing manual pages. The use of both these
+Scaffolder interfaces is outlined in the following sections. \pb
 
@@
 If both these commands return without error this indicates the necessary
-system requirements to install and run scaffolder. The scaffolder API and CLI
-can then be installed through RubyGems with the following command. \pb
+components are available to install and run Scaffolder. The Scaffolder API and
+CLI can then be installed using the following command. \pb
 
@@
 
-To ensure that the scaffolder CLI is accessible, the RubyGems `bin' directory
+To ensure that the Scaffolder CLI is accessible, the RubyGems `bin' directory
 should be in the command line path. Assuming that RubyGems are installed in
-~/.gem this can be done by adding the following to the ~/.bashrc file. The
-corresponding login shell should be updated for other shells such as Zsh, etc..
-\pb
+\~/.gem this can be done by adding the following to the \~/.bashrc file. The
+corresponding login script should be updated for other shells such as Zsh,
+etc.. \pb
 
@@
 The Scaffolder CLI provides functions to validate the scaffold file and build
-the draft scaffold sequence. Assuming scaffolder has been installed and the
+the draft scaffold sequence. Assuming Scaffolder has been installed and the
 binary is available in the command line path the following command should run
-without error and list the available scaffolder functions. \pb
+without error. This will list the available Scaffolder functions. \pb
 
@@
 
-Help on a specific command can be found by typing `help' followed by the name
-of the command. This returns the corresponding Unix manual page. The following
-example will return the manual page for the sequence command. \pb
+Help on a specific command can be found by typing ``help'' followed by the
+name of the command. This prints the corresponding Unix manual page to the
+terminal. The following example will return the manual page for the sequence
+command. \pb
 
@@
 The ``sequence'' command builds the fasta sequence from the scaffold and fasta
-file. Running this command requires the locations of both the scaffold file and
-the fasta file to be passed as argument. This will then return the complete
-scaffold sequence in fasta format. \pb
+file. Running this command requires the locations of both the scaffold file
+and the fasta file to be passed as arguments. This will then return the
+complete scaffold sequence in fasta format. \pb
 
@@
 Scaffolder provides a function to test the scaffold for overlapping insert
-coordinates. Overlapping inserts will result in unexpected behaviour as the
-coordinates for the first insert may change the insert location for subsequent
-inserts. The ``validate'' command will check the scaffold for any such inserts
-and return them as a YAML list. \pb
+coordinates. Overlapping inserts will result in unexpected behaviour when
+generating the scaffold sequence as the coordinates for the first insert may
+change the insert location for subsequent inserts. The ``validate'' command
+will check the scaffold for any such inserts and return them as a YAML list.
+\pb
 
@@
 
-\subsubsection*{Application programming interface} %{{{3
+\subsubsection*{Application Programming Interface} %{{{3
 
 Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
-This API is detailed in full on the scaffolder website. The following code
+This API is detailed in full on the Scaffolder website. The following code
 snippet outlines how the API may be used to create a script to replicate the
-function of the `scaffolder sequence' command described above in the CLI
-section. \pb
+function of the `scaffolder sequence' command described in the CLI section
+above. \pb
 
@@
 
-    require 'rubygems'
-    require 'scaffolder'
+    require `rubygems'
+    require `scaffolder'
 
@@
 arguments: the YAML loaded scaffold and the location of the fasta sequence
-file. The scaffold is an enumerable object and the next line illustrates the
-trivial code required to generate the scaffold sequence. Further details of the
-classes and method related to each scaffold entry can be found in the
-documentation for the Scaffolder::Region class. \pb
+file. The scaffold is an enumerable array and the next line illustrates the
+trivial code required to generate the scaffold sequence. Further details of
+the classes and method related to each scaffold entry can be found in the
+documentation for the Scaffolder::Region class on the Scaffolder website. \pb
 
@@
 
-The aim of scaffolder are twofold. The first aim is to provide scaffolding
-software which is easy to install and simplifies the task of genome finishing.
-The second aim is to facilitate reproducibility when joining large nucleotide
-sequences together to create a genome scaffold. \pb
+The aims of Scaffolder are twofold. The first is providing software which is
+easy to install and simplifies the task of genome finishing. The second aim is
+to facilitate reproducibility when joining large nucleotide sequences together
+to create a genome scaffold. \pb
 
@@
 
-Scaffolder is designed to be simple to use as possible. Assuming the existence
-of Ruby and RubyGems, scaffolder can be installed in a single command line
-step. This should negate the possible barriers to use in requiring manual
-compilation of the source code. \pb
+Scaffolder is designed to be as simple to use as possible. Assuming the
+existence of Ruby and RubyGems, Scaffolder can be installed in a single
+command line step. This should negate any possible barriers from requiring
+manual compilation of the source code. \pb
 
 Scaffolder uses a minimal and compact syntax to described the genome scaffold.
-This syntax is simple to construct and edit whilst being succinct and readable.
-The aim of this is genome scaffold creation to relatively simple in a common
-text editor. Furthermore scaffolder provides extensive documentation both for
-the command line tools and for the Scaffolder Ruby API. Therefore should any
-part of the scaffolding process be unclear the documentation should make an
+This syntax is simple to construct and edit whilst being succinct and
+readable. The goal of this syntax is to make genome scaffolds easy to
+construct a common text editor. Furthermore Scaffolder provides extensive
+documentation both for the CLI and for the Ruby API. Therefore should any part
+of the scaffolding process be unclear the documentation should make an
 solution easy to find. \pb
@@
 
-Constructing a genome sequence using a text file to specify the scaffold means
-the same sequence is always output for the same file. Joining large nucleotide
-sequences manually cannot however be reliably reproduced. The layout of the
-scaffold file furthermore provides a human readable record of how the sequence
-was constructed. The same cannot be said for large block of nucleotide
-sequences comprised of many source contigs. \pb
+Constructing a genome sequence using a text file means the generation the
+scaffold super sequence both reproducible and deterministic. In comparison
+joining large nucleotide sequences by hand cannot be reliably reproduced. The
+layout of the scaffold file furthermore provides a human readable record of
+how the sequence was constructed and decouples the layout of the contigs in
+the scaffold from their source sequence. The same cannot be said for large
+block of nucleotide sequences comprised of many source contigs. \pb
 
 Configuring the final sequence in the scaffold file means the build is easier
-to edit once already constructed compared with working with an already drafted
-sequence. The YAML text format is also facilitates easier comparison of
-differences between scaffold build using standard text comparison tools such as
-diff. This makes scaffold files easier to store in version control systems and
-affords genome finishers similar benefits enjoyed by software developers. \pb
+to edit once already constructed. The YAML text format also allows comparison
+of differences between scaffold builds using standard text comparison tools
+such as diff. This makes scaffold files easier to store in version control
+systems and affords genome finishers similar tools used in software
+development. \pb
 
@@
 
-Scaffolder follows the Unix philosophy ``Do one thing and do it well''.
-Scaffolder focuses on creating a simple to use and install tool for generating
-a genome scaffold. The YAML data format for creating a scaffold file is
+Scaffolder follows the Unix philosophy of ``Do one thing and do it well''.
+Scaffolder focuses on creating a simple to use and install tool for creating
+a genome scaffold. The YAML data format for creating the scaffold file is
 standardised and easily manipulated programmatically. This thereby follows
@@
 
-MDB developed and maintains the scaffolder tool. MDB and HAB wrote the
+MDB developed and maintains the Scaffolder tool. MDB and HAB wrote the
 manuscript.
@@
 
-\subsection*{Figure 1 - Example of scaffolder file}
+\subsection*{Figure 1 - Example of Scaffolder file and the resulting build}
 
@@
 inserts an unresolved gap region into the scaffold. The final entry illustrates
-a more complex examplea where the sequece is trimmed, reversed and has
+a more complex example where the sequence is trimmed, reversed and has
 additional sequence inserted. \pb

### 2011-02-16 src/article.tex
Revise manuscript for readability again

@@
 % Scaffolder url
-\urldef{\scaffolder}\url{http://www.michaelbarton.me.uk/projects/scaffolder/}
+\urldef{\scaffolder}\url{http://next.gs}
 
@@
   Michael D Barton$^{1}$\correspondingauthor%
-  \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
+  \email{Michael D Barton\correspondingauthor  - mail@nucleotid.es}%
 \and
@@
 
-  \paragraph*{Background:} Assembly of short read sequencing data can result
-  in a fragmented series of contigs. Therefore a common step in a genome
-  project is to join neighbouring sequence regions together and fill gaps
-  using additional sequence. This scaffolding step, however, is non-trivial
-  and requires manually editing large blocks of nucleotide sequence. Joining
-  sequence contigs together also hides the source of each region in the final
-  genome sequence. Taken together, these considerations may make reproducing
-  or editing an existing genome build difficult.
+  \paragraph*{Background:} Assembly of short read sequencing data can result in
+  a fragmented series of contigs. Therefore a common step in a genome project
+  is to join neighbouring sequence regions together and fill gaps in the
+  assembly using additional sequence. This scaffolding step, however, is
+  non-trivial and requires manually editing large blocks of nucleotide
+  sequence. Joining sequence contigs together also hides the source of each
+  region in the final genome sequence. Taken together, these considerations may
+  make reproducing or editing an existing genome build difficult.
 
@@
   scaffold file syntax. This syntax further allows  unknown regions to be
-  defined, and inserts to be added to the scaffold to fill contig gaps.
-  Defining the genome construction in a file makes the scaffolding process
-  reproducible and comparatively easier to edit and compare than raw fasta.
+  defined, and inserts to be added to the scaffold to fill gaps. Defining the
+  genome construction in a file makes the scaffolding process reproducible and
+  comparatively easier to edit and compare than fasta sequence.
 
   \paragraph*{Conclusions:} Scaffolder is easy to use genome scaffolding
-  software. This tool promotes reproducibility and maintenance in building
-  a genome. Scaffolder can be found at \scaffolder.
+  software. This tool promotes reproducibility and continuous development in a
+  genome project. Scaffolder can be found at \scaffolder.
 
@@
 High throughput sequencing can produce hundreds of thousands to millions of
-sequence reads from a genome. Each read represents the machine-interpreted
-nucleotide sequence from a short region in the genome. At the time of writing,
-high throughput sequencing is limited to producing reads less than $>$1000
-nucleotides long. Therefore a sequence longer than this length, e.g.
-a complete genome, these numerous individual reads must be pieced together.
+sequence reads from a genome. Each sequence represents the machine-read
+nucleotides from a short region in the genome. At the time of writing, high
+throughput sequencing is limited to producing reads less than $>$1000
+nucleotides in length. Therefore to determine a sequence longer than this, such
+as a complete genome, these numerous individual reads must be pieced together.
 The process of joining reads into longer sequences is the `assembly' stage in
@@
 nucleotide read represents a single piece and the final genome sequence is the
-completed puzzle. Repetitive nucleotide `repeat' regions within a genome or
-biased and incomplete sequencing data may however prevent the genome being
-assembled into a continuous sequence. This is analogous to missing jigsaw
-pieces or pieces that fit to multiple other pieces. \pb
+completed puzzle. Sequences of repetitive nucleotide `repeat' regions within
+the genome, or biased and incomplete sequencing data may however prevent the
+genome being assembled into a continuous sequence. This is analogous to missing
+jigsaw pieces or pieces that fit to multiple other pieces. \pb
 
@@
 to a renewed interest into algorithms for solving the problem of genome
-assembly \cite{pop2008,pop2009}. Current genome assembly software is however
-often unable to produce a complete sequence from these reads due to these
-problems. In many cases assembly algorithms will generate several large
-contiguous regions of sequence (`contigs') composed of many individual reads.
-These contigs represent the assembly of reads into the longest possible
-sequences. These contigs represent a fragmented picture of the genome and
-require additional work to piece together. These contigs represent
-a fragmented picture of the genome and require additional work to piece
-together. \pb
+assembly \cite{pop2008,pop2009}. Genome assembly software may however be unable
+to produce a complete sequence from these reads. In many cases assembly
+algorithms will generate several large contiguous regions of sequence
+(`contigs') composed of many individual reads given the available data and
+assembly algorithm. These contigs represent the assembly of reads into the
+longest possible sequences. These contigs represent a fragmented picture of the
+genome and require additional work to join together into larger sequence
+lengths. \pb
 
-The process of ameliorating a draft sequence to a finished continuous sequence
-can be expensive in time and laboratory effort. Therefore in many cases a set
-of contigs may often describe enough of the genomic content for many research
+The process of ameliorating a draft sequence into a finished continuous
+sequence can be expensive in both time and laboratory effort. Therefore the
+genomic data present in a set of contigs may often be enough for many research
 questions \cite{branscomb2002}. A continuous high-quality `finished' genome
-sequence does however provide further depth of information including complete
-resolution of repeat regions and the exact distances between genomic elements
-\cite{parkhill2002,fraser2002}. This process of joining contigs together to
-form a continuous sequence may be called the `scaffolding' or `finishing'
-stage. \pb
+sequence does however provide a greater depth of information, such as complete
+resolution of repeat regions and more precise estimates of distances between
+genomic elements \cite{parkhill2002,fraser2002}. This process of joining
+contigs together to form a continuous sequence may be called the `scaffolding'
+or `finishing' stage. \pb
 
@@
 complete continuous sequence. Due to genomic complexity and missing data,
-scaffolding may not ultimately produce a final completed sequence. The process
-of scaffolding may still however succeed in joining a subset of contigs
-together or using additional regions of sequence to fill gaps. An overview of
-the scaffolding process is outlined in the following section.
+scaffolding may not ultimately produce a final completed sequence, but may
+still succeed in joining a subset of contigs together or filling gaps. An
+overview of the scaffolding process is outlined in the following section.
 
@@
 
-The sequencing process produces reads from either strand of the DNA helix.
+The sequencing process generates reads from either strand of the DNA helix.
 Therefore the resulting contigs constructed from these reads may point in
-either direction of the DNA strand. Contig orientation reverse complements
-sequences where necessary so that all contigs point in the same direction and
-represent the same DNA strand. In the case of microbes this orientation may be
-the 5'-3' strand leading away from the origin of replication.
+either direction of the DNA strand .The orientation of contigs to point in the
+same direction and represent the same DNA strand requires reverse complementing
+sequences where necessary. In the case of microbes this orientation may be the
+5'-3' strand leading away from the origin of replication.
 
@@
 
-Contig ordering determines the placement of contigs to best represent their
-order in the true genome sequence. Correctly placing each contig may also
-highlight any extra-genomic DNA, such as plasmids, and which should be treated
+Contig ordering determines the placement of observed contigs to best represent
+their order in the true genome sequence. Correctly placing each contig also
+highlights any extra-genomic DNA, such as plasmids, and should be scaffolded
 separately from genomic sequence. The contig order is commonly referenced
-against the origin of replication where the contig containing the replication
-start point is first. All subsequent contigs are then ordered in the direction
-of DNA replication.
+against the origin of replication where the containing contig is split at the
+replication start point. All subsequent contigs are then ordered in the
+direction of DNA replication.
 
@@
 
-Given the order and orientation of contigs, determining the correct distances
-between contigs leads to an estimate of the genome size. The size of inter
+Given the correct order and orientation, determining the distances between
+contigs results in an estimate of the complete genome size. The size of inter
 contig gaps represents the length of an unknown region in the genome. Filling
-in these regions in with unknown nucleotide characters (`N') allows a draft
-continuous representation of the genome. This draft scaffold sequence is
-useful for representing both the sequenced and unknown regions in the genome.
+these regions with unknown nucleotide characters (`N') allows a draft
+continuous representation of the genome. This draft scaffold sequence is useful
+for representing both the sequenced and unknown regions in the genome.
 
@@
 
-During the scaffolding process closing and filling sequence gaps improves the
-accuracy of the genome scaffold. This may require returning to the laboratory
-to perform additional sequencing or using computational methods of estimating
-the gap sequence. These additional sequences can be used to bridge and close
-the gaps between contigs. Once all contigs have been joined and gaps in
+During the scaffolding process closing and filling gaps between contigs
+improves the accuracy of the genome scaffold. This may require returning to the
+laboratory to perform additional sequencing or using computational methods of
+estimating the gap sequence. This additional sequence is used to replace the
+gap between two contigs joining them into a single sequence if they are not
+already scaffolded together. Once all contigs have been joined and gaps in
 a scaffold closed the genome may be considered finished.
@@
 The process of resolving a genome sequence and building a scaffold uses
-\emph{in vitro} methods, \emph{in silico} methods, or a combination of both.
-An example of a computational method uses pair read data, where the sequencing
-reads are a known distance apart in the genome, when paired reads are in
-separate contigs this can to estimate the order and distance between them.
-Laboratory methods may use PCR to amplify the unknown DNA in gap regions,
-followed by traditional Sanger sequencing. Computation methods are more
-preferable as they less costly in laboratory time and materials compared to
-manual gap resolution \cite{nagarajan2010}. The need for computational
-scaffolding methods has resulted in a variety of algorithmic approaches. \pb
+\emph{in vitro} methods, \emph{in silico} methods, or a combination of both. An
+example of a computational method might use paired-read data, where the
+sequencing reads were generated at a known distance apart in the genome. The
+occurrence of a pair of reds in separate contigs this can be used to estimate
+the order and distance. Laboratory methods may use PCR to amplify the unknown
+DNA in a gap region, then use traditional Sanger sequencing to determine the
+sequence. Computation methods are more preferable as they are less costly in
+laboratory time and materials compared to manual gap resolution
+\cite{nagarajan2010}. This need for computational scaffolding methods has
+resulted in a variety of algorithmic approaches. \pb
 
-Synteny methods use sequence similarity to compare the assembled contigs to
-a more complete reference genome sequence. Areas of corresponding sequence
-between the the two sets are indicative of the possible contig organisation in
-the true genome \cite{richter2007,zhao2008}. \pb
+Synteny methods compare the assembled contigs to a more complete reference
+genome sequence searching for ares of sequence similarity. Areas of
+corresponding sequence between the the two sets are can be used to infer contig
+organisation \cite{richter2007,zhao2008}. Recombination between the two genomes
+can however reduce the effectiveness of this analysis. \pb
 
@@
 sequence. The existence of tandemly repeated nucleotide regions produces many
-reads of similar sequence. The assembly of these regions may result in the
-repeats being collapsed into a single instance or ignored by conservative
-assembly algorithms. These regions may be reassembled using algorithms
-specifically for repeat resolution \cite{mulyukov2002,koren2010}. A related
-approach uses uniquely matching reads to try and bridge across contig gaps
-between contigs \cite{tsai2010}. \pb
+reads of similar sequence. The assembly of reads generated from repeats can
+result in the assemble collapsing into a single repeat instance or ignored by
+conservative assembly algorithms. These regions may be resolved by using
+reassembly algorithms specifically for repeat resolution
+\cite{mulyukov2002,koren2010}. Similarly a related approach may use uniquely
+matching reads to try and bridge across gaps between contigs \cite{tsai2010}.
+\pb
 
@@
 together. Heuristic graph algorithms may take advantage of this data to search
-for optimal configuration of contigs in the scaffold to take this into account
+for optimal configuration of contigs in the scaffold using paired-read distances
 \cite{dayarian2010}. Additional data using a reference genome may also be
@@
 Taken together these methods provide a wide variety of possible approaches for
-scaffolding a genome. The scaffold may however still require manual editing
-and addition of sequence if computational methods are not completely
-successful. This process involves moving and joining larges blocks of
-nucleotide text together. This process may be expected to be error-prone and
-precludes any reproducibility of the scaffold. \pb
+scaffolding a genome. The scaffold may still however require manual editing and
+addition of sequence if computational methods are not completely successful.
+This involves moving and joining larges blocks of nucleotide text together by
+hand which may be expected to be error-prone and precludes any reproducibility.
+\pb
 
 The software outlined here ``Scaffolder'' is aimed at creating a syntax and
-software framework for producing and editing a genome scaffold. Scaffolder
-uses a specific data syntax to define how contigs are joined, additional
-sequences inserted, and for the specification of unknown regions. This syntax
-allows a scaffold to be edited in a simpler way and allows the scaffolding
-process to be reproducible and deterministic. \pb
+software framework for producing and editing a genome scaffold. Scaffolder uses
+a specific file format to define how contigs are joined, additional sequences
+inserted, and for the specification of unknown regions. This syntax allows
+a scaffold to be edited in a simpler way and allows the scaffolding process to
+be both reproducible and deterministic. \pb
 
@@
 
-Unit tests were implemented to develop and maintain the individual elements of
-the source code.  Integration tests were written to test the Scaffolder
+Unit tests were implemented to maintain individual elements of the source code
+during development. Integration tests were written to test the Scaffolder
 software interface as a whole. Unit tests were written using the Shoulda and
@@
 The Scaffolder source code is documented using the Yard library \cite{yard}.
-Unix manual pages for the command line interface were generated using the Ronn
-library \cite{ronn}. The manipulation of biological sequences in Scaffolder is
+Unix manual pages for the command line were generated using the Ronn library
+\cite{ronn}. The manipulation of biological sequences in Scaffolder is
 performed using the BioRuby library \cite{goto2010}. A full list of the
@@
 
-\subsection*{Scaffold file syntax} %{{{2
+\subsection*{Scaffold File Syntax} %{{{2
 
@@
 data format, third-party developers have the option to generate a genome
-scaffold using any programming language for which a YAML parser and generator
-exists. The YAML website lists current parsers for languages including C/C++,
-Ruby, Python, Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being
-widely supported YAML formatted scaffold files can be validated for correct
-syntax using third party tools such as Kwalify \cite{kwalify}. \pb
+scaffold using any programming language for which a YAML library exists. The
+YAML website lists current parsers for languages including C/C++, Ruby, Python,
+Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
+supported YAML formatted scaffold files can be validated for correct syntax
+using third party tools such as Kwalify \cite{kwalify}. \pb
 
@@
 generation of additional sequence using PCR translates into genome scaffolding
-as an on-going process. Therefore the scaffold file should be simple to edit
-update manually in addition to being computationally tractable. This
-requirement was therefore the second reason YAML was selected, where YAML is
-human-readable and should be relatively easy to manipulate in a standard text
-editor. \pb
+as an on-going process. Therefore the scaffold file should be simple to update
+manually in addition to being computationally tractable. This requirement was
+therefore the second reason YAML was selected, where YAML is human-readable and
+easy to edit in a standard text editor. \pb
 
 The scaffold file takes the form of a list of entries. Each entry corresponds
-to a region of sequence, most likely a contig, in the genome scaffold. The
-Scaffolder software processes this list of entries in order, converting each to
-the corresponding nucleotide sequence. These nucleotide sequences are then
-concatenated together to produce the draft super sequence. Each entry in the
-scaffold file may have attributes that define whether a sub-sequence or the
-reverse complement of the sequence should be used. The types of attributes
-available and an example scaffold file are outlined in the Results and
-Discussion section below.  \pb
+to a region of sequence, most likely a contig or previously scaffolded set of
+contigs. The Scaffolder software processes this list in order, converting each
+to the corresponding nucleotide sequence. These nucleotide sequences are then
+concatenated to produce a draft super sequence. Each entry in the scaffold file
+may have attributes that define whether a sub-sequence or the reverse
+complement of the sequence should be used. The types of attributes available
+and an example scaffold file are outlined in the Results and Discussion section
+below.  \pb
 
@@
 
-The Scaffolder software outlined here facilitates or computational joining of
+The Scaffolder software outlined here allows computational joining of
 nucleotide sequences together into a single contiguous super sequence.
-Plain-text scaffold files written in the YAML syntax are used to specify how
-nucleotide sequences should be joined together. This scaffold file is read by
-the software to determine how the draft scaffold sequence should be generated.
-In addition to specifying which contigs should be used, the specification of
-the scaffold file allows the contigs to trimmed down to smaller sub-sequences
-and reverse complemented if necessary. \pb
+Plain-text scaffold files written in the YAML syntax specify how these
+sequences should be joined. This scaffold file is read by the scaffolder
+software and used to generate the scaffold sequence. In addition to specifying
+which contigs, the scaffold file allows the contigs to be edited to produce
+smaller sub-sequences or reverse complemented if necessary. \pb
 
 The process of genome finishing may involve producing additional
-oligonucleotide sequence to fill areas of unknown sequence in the scaffold.
-The scaffold file format allows these additional insert sequences to be added
-and bridge gaps. These inserts can also be treated in a similar way to larger
+oligonucleotide sequence to fill areas of unknown regions in the scaffold. The
+scaffold file format allows these additional insert sequences to be added to
+bridge these gaps. These inserts can also be treated in a similar way to larger
 contig sequences: trimmed and/or reverse complemented to match the
@@
 
-In the process of building a genome scaffold distances between contigs may be
-estimated from paired read data or from mapping the contigs to a reference
-genome. These inter contig gap regions are useful to join regions of sequence
-together by an estimated size. The scaffold file also allows for the
-specification of these unknown regions to highlight areas in the genome
-sequence still to be resolved. \pb
+Distances between contigs may be estimated from paired read data or from
+mapping the contigs to a reference genome. These inter-contig gap regions are
+useful to join regions of sequence together by an estimated size. The scaffold
+file allows for the specification of these unknown regions to join contigs
+together. The use of these regions in the scaffold highlights areas still to be
+resolved. \pb
 
-The scaffold file is used to specify how sequence regions are joined together,
-however the sequences themselves are not stored in the scaffold file. The
-sequences are instead maintained in a fasta file. A fasta file of contigs and
+The scaffold file is used to specify how sequence regions are joined together
+but the sequences themselves are not stored in the scaffold file. The sequences
+are instead maintained as a separate fasta file. A fasta file of contigs and
 scaffolded contigs is a common output for assemblers and as such should be
-readily available without further processing required. Nucleotide sequence are
-reverenced in the scaffold using the first space-delimited word from each fasta
-header. Maintaining the nucleotide sequences in a separate file preserves the
-original sequence and furthermore decouples the raw sequence from the markup
-describing how it should be used to generate the genome sequence. \pb
+readily available. The nucleotide sequences in the fasta file are referenced in
+the scaffold using the first space-delimited word from each fasta header.
+Maintaining the nucleotide sequences in a separate file preserves the original
+sequence and furthermore decouples the sequence from the markup describing how
+it should be used to generate the genome sequence. \pb
 
-\subsection*{Defining a Scaffold in a Text File} %{{{2
+\subsection*{Defining a Scaffold as a Text File} %{{{2
 
@@
 scaffold file is shown in Figure 1. This file illustrates the text attributes
-used to describe a scaffold and the corresponding genome build. The basic
-layout of the scaffold file is a list - a series of entries where each entry
-corresponds to a region of sequence in the scaffold super sequence. \pb
+used to describe a scaffold and the joining of sequences in the genome build.
+The basic layout of the scaffold file is a list - a series of entries where
+each entry corresponds to a region of sequence in the generated scaffold super
+sequence. The right hand side indicates hows the entries are joined together.
+This figure is for illustrative purposes and not to scale. \pb
 
@@
 a requirement to indicate the start of a YAML formatted document. The first
-entry in the scaffold begins on Line 2. The first line of this entry,
+entry in the scaffold begins on the second line. The first line of this entry,
 highlighted in green, begins with a dash character `-', dashes are used to
-denote entries in a YAML list. This first line therefore indicates indicate
-the first entry in the scaffold. The first line of each entry in the scaffold
-begins with a dash line. The second line of this entry indented by two spaces
-where whitespace is used to group similar attributes together. Lines indented
-by whitespace are linked to the first non-indented preceding line in the file.
-\pb
+denote entries in a YAML list. The first line of each entry in the scaffold
+therefore begins with a dash line. The next line is indented by two spaces
+- whitespace is required to group similar attributes together. Lines indented
+by whitespace are part to the first non-indented preceding line. \pb
 
@@
 the ``source'' tag. As described above the first space-delimited word of the
-fasta header is used to identify which sequence. Therefore these three lines
-describe the first entry in the scaffold as a simple sequence using the fasta
-sequence named `sequence1'. \pb
+fasta header is used to identify which sequence is selected from the fasta
+file. Together these three lines describe the first entry in the scaffold as
+a simple sequence using the fasta sequence named `sequence1'. On the right hand
+side of Figure 1 this produces the first region in the scaffold, also shown in
+green. \pb
 
@@
 The second entry in the scaffold, highlighted in orange, is identified by the
-``unresolved'' tag, and specifies a region of unknown nucleotides but known
-length. The following line in the entry specifies the size of this unknown
-region. This entry therefore inserts a region of 20 `N' characters in the
-scaffold. \pb
+``unresolved'' tag, indicating a region of unknown sequence but known length.
+The line below specifies the size of this unknown region. This entry therefore
+inserts a region of 20 `N' characters in the scaffold. \pb
 
@@
 sequence may be manipulated prior to addition to the scaffold. The `start' and
-`stop' tags trim the sequence to these coordinates inclusively. The
-``reverse'' tag reverse complements the sequence. \pb
+`stop' tags trim the sequence to these coordinates inclusively. The ``reverse''
+tag instructs scaffolder to also reverse complement the sequence. On the right
+hand side of the sequence of Figure 1 this completes the scaffold sequence. \pb
 
-This region further includes the ``inserts'' tag to update this sequence with
-regions of sequence. These inserts are also added as a YAML list with each
-insert beginning with a dash line. The first insert, shown in purple,
-illustrates using similar attributes to that of a sequence entry. The reverse,
-start and stop tags, are used to trim and reverse complement the insert.
-Similarly the `source' tag identifies the corresponding fasta sequence. The
-``open'' and ``close'' tags determine where the insert is added in the
-enclosing sequence. These coordinates inclusively replace the sequence region
-within these coordinates with the specified insert fasta sequence. \pb
+This entry includes the ``inserts'' tag to update this sequence with regions of
+sequence. These inserts are also added as a YAML list with each insert
+beginning with a dash line. The first insert, shown in purple, illustrates
+using similar attributes to that of a sequence entry. The reverse, start and
+stop tags, are used to trim and reverse complement the insert. Similarly the
+`source' tag identifies the corresponding fasta sequence. The ``open'' and
+``close'' tags are specific to inserts and determine where the insert is added
+in the enclosing sequence. The region of the sequence inside these coordinates
+inclusively is replaced with the specified insert fasta sequence. This is
+highlighted in the putative scaffold in Figure 1 by the black lines bisecting
+the blue sequence. \pb
 
-The next insert, shown in brown, specifies only the open position tag. This
-illustrates that only one of either `open' or `close' tags is required. When
-only one of these tag is used the corresponding opposing coordinate position
-is calculated from the length in the insert fasta sequence. This allows
-inserts to bridge into, and partially fill gap regions. \pb
+The next insert, shown in brown, is specified using only the open position tag.
+This illustrates that only one of either `open' or `close' tags is required.
+When only one of these tag is used the corresponding opposing coordinate
+position is calculated from the length in the insert fasta sequence. This
+allows inserts to bridge into, and partially fill gap regions with requiring an
+end co-ordinate position. \pb
 
@@
 
-The sequence entry in the scaffold file allows multiple sequence-editing
-operations to be applied. If applied in different combinations, these
-operations could result in different output sequence. An example would be
-reverse complementing the sequence then trimming the start of the sequence by
-five nucleotides. Performing these actions in the reverse order would produce
-a difference output sequence. This section therefore outlines the Scaffolder
-protocol for processing sequence entries. \pb
+The the scaffold file format allows multiple sequence-editing operations to be
+applied to a sequence or insert. If applied in different combinations, these
+operations will produce different output sequences. An example would be reverse
+complementing the sequence then trimming the start of the sequence by five
+nucleotides. Performing these actions in the reverse order would result in five
+nucleotides being removed from the opposite end. This section therefore
+outlines the Scaffolder protocol for processing sequence entries. \pb
 
@@
 with the smallest close position is last. This preserves the original intended
-coordinates for each insert during the process of adding each insert to the
+coordinates for each insert during the process of adding inserts to the
 sequence. \pb
@@
 coordinate of the enclosing sequence would be incremented by the 2bp
-difference. As inserts are added by reverse close position order the remaining
+difference. As inserts are added by reverse close position the remaining
 inserts are unaffected by changes in the encoding sequence size. \pb
@@
 Scaffolder provides a standardised set of Ruby classes and methods (API) for
-manipulation of the scaffold. This allows Scaffolder to be integrated into
+interacting with the scaffold. This allows Scaffolder to be integrated into
 existing genomics workflows or used with Ruby build tools such as Rake. \pb
@@
 Scaffolder relies on both the Ruby programming language and the RubyGems
-package management system. These are either already present or easily
-installed on recent distributions of Mac OSX or Linux. The presence of these
-tools can be tested by typing the following on the command line. \pb
+package management system. These are either already present or easily installed
+on recent distributions of Mac OS X or Linux. The presence of both these tools
+can be tested by typing the following on the command line. \pb
 
@@
 
-If both these commands return without error this indicates the necessary
-components are available to install and run Scaffolder. The Scaffolder API and
-CLI can then be installed using the following command. \pb
+If these commands return without error this indicates the necessary components
+are available to install and run Scaffolder. The Scaffolder API and CLI can
+then be installed using the following RubyGems command. \pb
 
@@
 To ensure that the Scaffolder CLI is accessible, the RubyGems `bin' directory
-should be in the command line path. Assuming that RubyGems are installed in
-\~/.gem this can be done by adding the following to the \~/.bashrc file. The
-corresponding login script should be updated for other shells such as Zsh,
-etc.. \pb
+should be in the command line path. Assuming that RubyGems are installed to
+$\sim$/.gem this can be done by adding the following to the $\sim$/.bashrc
+file. The corresponding login script should be updated for other shells such as
+Zsh, etc.. \pb
 
@@
 The Scaffolder CLI provides functions to validate the scaffold file and build
-the draft scaffold sequence. Assuming Scaffolder has been installed and the
-binary is available in the command line path the following command should run
-without error. This will list the available Scaffolder functions. \pb
+a draft super sequence. Assuming Scaffolder has been installed and the binary
+is available in the command line path the following should run without error.
+This command will list the available Scaffolder functions. \pb
 
@@
 generating the scaffold sequence as the coordinates for the first insert may
-change the insert location for subsequent inserts. The ``validate'' command
-will check the scaffold for any such inserts and return them as a YAML list.
-\pb
+change the intended coordinates for subsequent inserts. The ``validate''
+command will check the scaffold for any such inserts and return them as a YAML
+list. \pb
 
@@
 Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
-This API is detailed in full on the Scaffolder website. The following code
-snippet outlines how the API may be used to create a script to replicate the
-function of the `scaffolder sequence' command described in the CLI section
-above. \pb
+This API is detailed in full on the Scaffolder website (\scaffolder). The
+following code snippet outlines how the API may be used to create a script to
+replicate the function of the `scaffolder sequence' command described in the
+CLI section above. \pb
 
@@
 arguments: the YAML loaded scaffold and the location of the fasta sequence
-file. The scaffold is an enumerable array and the next line illustrates the
-trivial code required to generate the scaffold sequence. Further details of
-the classes and method related to each scaffold entry can be found in the
+file. The scaffold is an enumerable array and the penultimate line illustrates
+the trivial code required to generate the scaffold sequence. Further details of
+the classes and methods related to each scaffold entry can be found in the
 documentation for the Scaffolder::Region class on the Scaffolder website. \pb
@@
 
-Scaffolder is designed to be as simple to use as possible. Assuming the
-existence of Ruby and RubyGems, Scaffolder can be installed in a single
+Scaffolder is designed to be as simple to use as possible. Assuming the Ruby
+and RubyGems software are present, Scaffolder can be installed in a single
 command line step. This should negate any possible barriers from requiring
@@
 Scaffolder uses a minimal and compact syntax to described the genome scaffold.
-This syntax is simple to construct and edit whilst being succinct and
-readable. The goal of this syntax is to make genome scaffolds easy to
-construct a common text editor. Furthermore Scaffolder provides extensive
-documentation both for the CLI and for the Ruby API. Therefore should any part
-of the scaffolding process be unclear the documentation should make an
-solution easy to find. \pb
+This syntax is simple to construct and edit whilst being succinct and readable.
+The goal of this syntax is to make genome scaffolds easy to construct in
+a common text editor. Furthermore Scaffolder provides extensive documentation
+both for the CLI and for the Ruby API. Therefore should any part of the
+scaffolding process be unclear the documentation should make an solution easy
+to find. \pb
 
@@
 
-Constructing a genome sequence using a text file means the generation the
-scaffold super sequence both reproducible and deterministic. In comparison
-joining large nucleotide sequences by hand cannot be reliably reproduced. The
-layout of the scaffold file furthermore provides a human readable record of
-how the sequence was constructed and decouples the layout of the contigs in
-the scaffold from their source sequence. The same cannot be said for large
-block of nucleotide sequences comprised of many source contigs. \pb
+Constructing a genome sequence using a text file makes generating a scaffold
+super sequence both reproducible and deterministic. In comparison joining large
+nucleotide sequences by hand cannot be reliably reproduced. The layout of the
+scaffold file furthermore provides a human readable record of how the scaffold
+was constructed and decouples the layout of contigs from their source sequence.
+\pb
 
@@
 to edit once already constructed. The YAML text format also allows comparison
-of differences between scaffold builds using standard text comparison tools
-such as diff. This makes scaffold files easier to store in version control
-systems and affords genome finishers similar tools used in software
-development. \pb
+of differences between scaffold builds using standard Unix tools such as diff.
+This makes scaffold files easier to store in version control systems and allows
+genome finishers methods similar to that of software development. \pb
 
@@
 a genome scaffold. The YAML data format for creating the scaffold file is
-standardised and easily manipulated programmatically. This thereby follows
-a second Unix tenet: ``If your data structures are good enough, the algorithm
-to manipulate them should be trivial.''
+standardised and easily manipulated programmatically. This thereby means the
+scaffolding process follows a second Unix tenet: ``If your data structures are
+good enough, the algorithm to manipulate them should be trivial.''
 
@@
 
-\subsection*{Figure 1 - Example of Scaffolder file and the resulting build}
+\subsection*{Figure 1 - Example of Scaffolder File and the Resulting Build}
 
-An example scaffold file written using the YAML syntax \cite{yaml}. The
-scaffold file contains three entries. Each entry is delimited by a `-' on a new
-line. The first entry is an example adding a simple sequence. The second entry
-inserts an unresolved gap region into the scaffold. The final entry illustrates
-a more complex example where the sequence is trimmed, reversed and has
-additional sequence inserted. \pb
+An example scaffold file written using the YAML syntax \cite{yaml} (left) and
+the resulting genome build (right). The scaffold contains three entries and two
+inserts. Each entry in the scaffold file text is delimited by a `-' on a new
+line and highlighted using separate colours. The scaffold diagram on the right
+illustrates how the scaffold sequences are joined and is not to scale. \pb
 

### 2011-02-17 src/article.tex
Revision of manuscript for readability

@@
   genome construction in a file makes the scaffolding process reproducible and
-  comparatively easier to edit and compare than fasta sequence.
+  easier to edit and compare than fasta sequence.
 
@@
 nucleotides from a short region in the genome. At the time of writing, high
-throughput sequencing is limited to producing reads less than $>$1000
-nucleotides in length. Therefore to determine a sequence longer than this, such
-as a complete genome, these numerous individual reads must be pieced together.
-The process of joining reads into longer sequences is the `assembly' stage in
+throughput sequencing is limited to producing reads less than 1000 nucleotides
+in length. Therefore to determine a sequence longer than this, such as
+a complete genome, these numerous individual reads must be pieced together. The
+process of joining reads into longer sequences is the `assembly' stage in
 a genome project \cite{miller2010}. \pb
@@
 Assembly software takes the nucleotide reads produced by the sequencing
-hardware and, in the ideal case, outputs a complete genome sequence compose of
+hardware and, in the ideal case, outputs a complete genome sequence composed of
 these individual reads. An analogy for this process is a jigsaw puzzle. Each
@@
 algorithms will generate several large contiguous regions of sequence
-(`contigs') composed of many individual reads given the available data and
-assembly algorithm. These contigs represent the assembly of reads into the
-longest possible sequences. These contigs represent a fragmented picture of the
+(`contigs') composed of many individual reads. These contigs represent the
+assembly of reads into the longest possible sequences given the available data
+and assembly algorithm. These contigs represent a fragmented picture of the
 genome and require additional work to join together into larger sequence
@@
 sequence does however provide a greater depth of information, such as complete
-resolution of repeat regions and more precise estimates of distances between
-genomic elements \cite{parkhill2002,fraser2002}. This process of joining
-contigs together to form a continuous sequence may be called the `scaffolding'
-or `finishing' stage. \pb
+resolution of repeat regions and precise estimates of distances between genomic
+elements \cite{parkhill2002,fraser2002}. This process of joining contigs
+together to form a continuous sequence may be called the `scaffolding' or
+`finishing' stage. \pb
 
@@
 still succeed in joining a subset of contigs together or filling gaps. An
-overview of the scaffolding process is outlined in the following section.
+overview of the scaffolding process is outlined below.
 
@@
 Therefore the resulting contigs constructed from these reads may point in
-either direction of the DNA strand .The orientation of contigs to point in the
-same direction and represent the same DNA strand requires reverse complementing
-sequences where necessary. In the case of microbes this orientation may be the
-5'-3' strand leading away from the origin of replication.
+either direction of the DNA. The orientation of contigs to point in the same
+direction and represent the same strand requires reverse complementing
+sequences where necessary. In the case of microbes this orientation may be to
+the 5'-3' direction leading away from the origin of replication.
 
@@
 their order in the true genome sequence. Correctly placing each contig also
-highlights any extra-genomic DNA, such as plasmids, and should be scaffolded
-separately from genomic sequence. The contig order is commonly referenced
-against the origin of replication where the containing contig is split at the
-replication start point. All subsequent contigs are then ordered in the
-direction of DNA replication.
+highlights any extra-genomic DNA, such as plasmids which should be scaffolded
+separately from the genomic sequence. Contig order is commonly started at the
+origin of replication where the first contig is split at the replication start
+point. All subsequent contigs are then ordered in the direction of DNA
+replication.
 
@@
 these regions with unknown nucleotide characters (`N') allows a draft
-continuous representation of the genome. This draft scaffold sequence is useful
-for representing both the sequenced and unknown regions in the genome.
+continuous sequence. This draft scaffold sequence is useful for representing
+both the sequenced and unknown regions in the genome.
 
@@
 During the scaffolding process closing and filling gaps between contigs
-improves the accuracy of the genome scaffold. This may require returning to the
-laboratory to perform additional sequencing or using computational methods of
-estimating the gap sequence. This additional sequence is used to replace the
-gap between two contigs joining them into a single sequence if they are not
-already scaffolded together. Once all contigs have been joined and gaps in
-a scaffold closed the genome may be considered finished.
+improves the completeness of the genome scaffold. Closing gaps may require
+returning to the laboratory to perform additional sequencing or using
+computational methods of estimating the unknown sequence. This additional
+sequence is used to replace the gap between two contigs joining them into
+a single sequence if they are not already scaffolded together. Once all contigs
+have been joined and gaps in a scaffold closed the genome may be considered
+finished.
 
@@
 example of a computational method might use paired-read data, where the
-sequencing reads were generated at a known distance apart in the genome. The
-occurrence of a pair of reds in separate contigs this can be used to estimate
-the order and distance. Laboratory methods may use PCR to amplify the unknown
-DNA in a gap region, then use traditional Sanger sequencing to determine the
-sequence. Computation methods are more preferable as they are less costly in
-laboratory time and materials compared to manual gap resolution
-\cite{nagarajan2010}. This need for computational scaffolding methods has
-resulted in a variety of algorithmic approaches. \pb
+sequencing reads are produced in pairs at a known distance apart in the genome.
+The occurrence of a pair of reads in separate contigs can be used to estimate
+the order and distance between the contigs. Alternatively laboratory methods
+may use PCR to amplify the unknown DNA in a gap region, then use traditional
+Sanger sequencing to determine the sequence. Computational methods are more
+preferable as they are less costly in laboratory time and materials compared to
+manual gap resolution \cite{nagarajan2010}. This need for computational
+scaffolding methods has resulted in a variety of algorithmic approaches. \pb
 
 Synteny methods compare the assembled contigs to a more complete reference
-genome sequence searching for ares of sequence similarity. Areas of
-corresponding sequence between the the two sets are can be used to infer contig
-organisation \cite{richter2007,zhao2008}. Recombination between the two genomes
-can however reduce the effectiveness of this analysis. \pb
+genome sequence searching for areas of sequence similarity between the two.
+Areas of corresponding sequence between the the two sets are can be used to
+infer contig placement and build the scaffold \cite{richter2007,zhao2008}.
+Recombination between the two genomes can however reduce the effectiveness of
+this analysis. \pb
 
@@
 sequence. The existence of tandemly repeated nucleotide regions produces many
-reads of similar sequence. The assembly of reads generated from repeats can
-result in the assemble collapsing into a single repeat instance or ignored by
-conservative assembly algorithms. These regions may be resolved by using
-reassembly algorithms specifically for repeat resolution
+similar reads in the sequence process. The assembly of reads generated from
+repeat regions can result in the assemble collapsing into a single repeat
+instance or ignored by conservative assembly algorithms. These regions may be
+resolved by using reassembly algorithms specifically for repeat resolution
 \cite{mulyukov2002,koren2010}. Similarly a related approach may use uniquely
@@
 \cite{dayarian2010}. Additional data using a reference genome may also be
-combined when estimating the best contig configuration \cite{pop2004}. \pb
-
-Finally when the scaffold cannot be completely resolved \emph{in silico}
-software packages exist to suggest the likely primers necessary for determining
-the sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
+combined when estimating the best contig configuration \cite{pop2004}. Finally
+when the scaffold cannot be completely resolved \emph{in silico} software
+packages exist to suggest the likely primers necessary for PCR amplifying the
+sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
 
@@
 This involves moving and joining larges blocks of nucleotide text together by
-hand which may be expected to be error-prone and precludes any reproducibility.
+hand which may be error-prone and precludes any reproducibility.
 \pb
@@
 inserted, and for the specification of unknown regions. This syntax allows
-a scaffold to be edited in a simpler way and allows the scaffolding process to
-be both reproducible and deterministic. \pb
+a scaffold to be updated by simply editing the scaffold file and allows the
+scaffolding process to be both reproducible and deterministic. \pb
 
@@
 Unix manual pages for the command line were generated using the Ronn library
-\cite{ronn}. The manipulation of biological sequences in Scaffolder is
-performed using the BioRuby library \cite{goto2010}. A full list of the
-software dependencies in Scaffolder can be found in the Gemfile in the root of
-each source code directory. \pb
+\cite{ronn}. The manipulation of biological sequences in Scaffolder uses the
+BioRuby library \cite{goto2010}. A full list of the software dependencies in
+Scaffolder can be found in the Gemfile in the root of each source code
+directory. \pb
 
@@
 may result in an incomplete genome build. The use of computational software and
-generation of additional sequence using PCR translates into genome scaffolding
-as an on-going process. Therefore the scaffold file should be simple to update
+generation of additional sequence using PCR means genome scaffolding may be an
+on-going process. Therefore the scaffold file should be simple to update
 manually in addition to being computationally tractable. This requirement was
 therefore the second reason YAML was selected, where YAML is human-readable and
-easy to edit in a standard text editor. \pb
+simple to edit in a standard text editor. \pb
 
@@
 may have attributes that define whether a sub-sequence or the reverse
-complement of the sequence should be used. The types of attributes available
+complement of the sequence should be used. The types of attributes available,
 and an example scaffold file are outlined in the Results and Discussion section
@@
 software and used to generate the scaffold sequence. In addition to specifying
-which contigs, the scaffold file allows the contigs to be edited to produce
-smaller sub-sequences or reverse complemented if necessary. \pb
+which contigs are required, the scaffold file allows the contigs to be edited
+to produce smaller sub-sequences or reverse complemented if necessary. \pb
 
 The process of genome finishing may involve producing additional
-oligonucleotide sequence to fill areas of unknown regions in the scaffold. The
-scaffold file format allows these additional insert sequences to be added to
-bridge these gaps. These inserts can also be treated in a similar way to larger
-contig sequences: trimmed and/or reverse complemented to match the
-corresponding gap region. \pb
+oligonucleotide sequence to fill unknown regions in the scaffold. The scaffold
+file format allows these additional insert sequences to be added to bridge
+these gaps. These inserts can also be treated in a similar way to larger contig
+sequences: trimmed and/or reverse complemented to match the corresponding gap
+region. \pb
 
@@
 The basic layout of the scaffold file is a list - a series of entries where
-each entry corresponds to a region of sequence in the generated scaffold super
-sequence. The right hand side indicates hows the entries are joined together.
-This figure is for illustrative purposes and not to scale. \pb
+each entry corresponds to a region of sequence in the generated scaffold
+super-sequence. The right hand side indicates how the entries are joined
+together. This figure is illustrates the putative scaffold and is not to scale.
+\pb
 
@@
 a requirement to indicate the start of a YAML formatted document. The first
-entry in the scaffold begins on the second line. The first line of this entry,
-highlighted in green, begins with a dash character `-', dashes are used to
+entry, highlighted in green, in the scaffold begins on the second line. The
+first line of this entry begins with a dash character `-', dashes are used to
 denote entries in a YAML list. The first line of each entry in the scaffold
 therefore begins with a dash line. The next line is indented by two spaces
-- whitespace is required to group similar attributes together. Lines indented
-by whitespace are part to the first non-indented preceding line. \pb
+- whitespace is required to group similar attributes together. \pb
 
@@
 ``unresolved'' tag, indicating a region of unknown sequence but known length.
-The line below specifies the size of this unknown region. This entry therefore
+The second line specifies the size of this unknown region. This entry therefore
 inserts a region of 20 `N' characters in the scaffold. \pb
@@
 
-This entry includes the ``inserts'' tag to update this sequence with regions of
+This final entry also includes the ``inserts'' tag to add additional regions of
 sequence. These inserts are also added as a YAML list with each insert
@@
 in the enclosing sequence. The region of the sequence inside these coordinates
-inclusively is replaced with the specified insert fasta sequence. This is
+is inclusively replaced by the specified insert fasta sequence. This is
 highlighted in the putative scaffold in Figure 1 by the black lines bisecting
@@
 When only one of these tag is used the corresponding opposing coordinate
-position is calculated from the length in the insert fasta sequence. This
-allows inserts to bridge into, and partially fill gap regions with requiring an
-end co-ordinate position. \pb
+position is calculated from the length of the insert fasta sequence. This
+allows inserts to bridge into, and partially fill gap regions without requiring
+an end coordinate position. \pb
 
@@
 \paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts, these inserts are reverse sorted by the insert close coordinate. The
-insert with the largest close position is first in the list and the insert
-with the smallest close position is last. This preserves the original intended
-coordinates for each insert during the process of adding inserts to the
-sequence. \pb
+inserts, these inserts are reverse sorted by the insert close coordinate
+position. The insert with the largest close position is first in the list and
+the insert with the smallest close position is last. This preserves the
+original intended coordinates as each insert is added to the sequence. \pb
 
@@
 on recent distributions of Mac OS X or Linux. The presence of both these tools
-can be tested by typing the following on the command line. \pb
+can be tested by typing the following commands on the terminal. \pb
 
@@
 file. The scaffold is an enumerable array and the penultimate line illustrates
-the trivial code required to generate the scaffold sequence. Further details of
+the trivial code required to generate the super-sequence. Further details of
 the classes and methods related to each scaffold entry can be found in the
-documentation for the Scaffolder::Region class on the Scaffolder website. \pb
+documentation for the Scaffolder::Region class in the documentation on the
+Scaffolder website. \pb
 
@@
 to facilitate reproducibility when joining large nucleotide sequences together
-to create a genome scaffold. \pb
+to create a genome sequence. \pb
 
@@
 and RubyGems software are present, Scaffolder can be installed in a single
-command line step. This should negate any possible barriers from requiring
-manual compilation of the source code. \pb
+command line step. This should negate any possible barriers to entry which
+require manual compilation of source code. \pb
 
-Scaffolder uses a minimal and compact syntax to described the genome scaffold.
+Scaffolder uses a minimal and compact syntax to describe the genome scaffold.
 This syntax is simple to construct and edit whilst being succinct and readable.
@@
 This makes scaffold files easier to store in version control systems and allows
-genome finishers methods similar to that of software development. \pb
+genome finishers methods similar to those in software development. \pb
 
@@
 
-Scaffolder follows the Unix philosophy of ``Do one thing and do it well''.
-Scaffolder focuses on creating a simple to use and install tool for creating
-a genome scaffold. The YAML data format for creating the scaffold file is
-standardised and easily manipulated programmatically. This thereby means the
-scaffolding process follows a second Unix tenet: ``If your data structures are
-good enough, the algorithm to manipulate them should be trivial.''
+Scaffolder follows the Unix philosophy of ``Do one thing and do it well''. The
+focus of Scaffolder is a simple to use tool for creating a genome scaffold. The
+YAML data format for creating the scaffold file is standardised and easily
+manipulated programmatically. This thereby means the scaffolding process
+follows a second Unix tenet: ``If your data structures are good enough, the
+algorithm to manipulate them should be trivial.''
 
@@
     \item[CLI:] Command line interface
-    \item[YAML:] YAML Ain't Markup Language\cite{yaml}
+    \item[PCR:] Polymerase chain reaction
+    \item[YAML:] YAML ain't markup language\cite{yaml}
   \end{description}
@@
 An example scaffold file written using the YAML syntax \cite{yaml} (left) and
-the resulting genome build (right). The scaffold contains three entries and two
-inserts. Each entry in the scaffold file text is delimited by a `-' on a new
-line and highlighted using separate colours. The scaffold diagram on the right
-illustrates how the scaffold sequences are joined and is not to scale. \pb
+the resulting putative scaffold (right). The scaffold contains three entries
+and two inserts. Each entry in the scaffold file text is delimited by a `-' on
+a new line and highlighted using separate colours. The scaffold diagram on the
+right is not to scale but illustrates how the scaffold sequences are joined.
+\pb
 

### 2011-02-22 src/article.tex
Two additonal scaffolding references

@@
 Areas of corresponding sequence between the the two sets are can be used to
-infer contig placement and build the scaffold \cite{richter2007,zhao2008}.
-Recombination between the two genomes can however reduce the effectiveness of
-this analysis. \pb
+infer contig placement and build the scaffold
+\cite{richter2007,zhao2008,assefa2009}. Recombination between the two genomes
+can however reduce the effectiveness of this analysis. \pb
 
@@
 provides an extra level of information on how contigs may be scaffolded
-together. Heuristic graph algorithms may take advantage of this data to search
-for optimal configuration of contigs in the scaffold using paired-read distances
-\cite{dayarian2010}. Additional data using a reference genome may also be
-combined when estimating the best contig configuration \cite{pop2004}. Finally
-when the scaffold cannot be completely resolved \emph{in silico} software
-packages exist to suggest the likely primers necessary for PCR amplifying the
-sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
+together. Heuristic algorithms may take advantage of this data to search for
+optimal configuration of contigs in the scaffold using paired-read distances
+\cite{dayarian2010,boetzer2011}. Additional data using a reference genome may
+also be combined when estimating the best contig configuration \cite{pop2004}.
+Finally when the scaffold cannot be completely resolved \emph{in silico}
+software packages exist to suggest the likely primers necessary for PCR
+amplifying the sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
 

### 2011-02-25 src/article.tex
Add Hazel corrections to introduction

--- 2011-02-17 src/article.tex
@@
 throughput sequencing is limited to producing reads less than 1000 nucleotides
-in length. Therefore to determine a sequence longer than this, such as
-a complete genome, these numerous individual reads must be pieced together. The
-process of joining reads into longer sequences is the `assembly' stage in
-a genome project \cite{miller2010}. \pb
+in length. Therefore to resolve a sequence longer than this, such as a complete
+genome, these numerous smaller fragments must be pieced together. The process
+of joining reads into longer sequences is the `assembly' stage of a genome
+project \cite{miller2010}. \pb
 
@@
 hardware and, in the ideal case, outputs a complete genome sequence composed of
-these individual reads. An analogy for this process is a jigsaw puzzle. Each
-nucleotide read represents a single piece and the final genome sequence is the
-completed puzzle. Sequences of repetitive nucleotide `repeat' regions within
-the genome, or biased and incomplete sequencing data may however prevent the
-genome being assembled into a continuous sequence. This is analogous to missing
-jigsaw pieces or pieces that fit to multiple other pieces. \pb
+these individual fragments. An analogy for this process is a jigsaw puzzle.
+Each nucleotide read represents a single piece and the final genome sequence is
+the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
+within the genome, or biased and incomplete sequencing data, may prevent the
+genome being assembled into a continuous sequence. This may be due to not
+enough or multiple overlaps between sequence reads and is analogous to missing
+pieces in the jigsaw or pieces that fit to multiple other pieces. \pb
 
@@
 to a renewed interest into algorithms for solving the problem of genome
-assembly \cite{pop2008,pop2009}. Genome assembly software may however be unable
+assembly \cite{pop2008,pop2009}; genome assembly software may however be unable
 to produce a complete sequence from these reads. In many cases assembly
 algorithms will generate several large contiguous regions of sequence
-(`contigs') composed of many individual reads. These contigs represent the
-assembly of reads into the longest possible sequences given the available data
-and assembly algorithm. These contigs represent a fragmented picture of the
-genome and require additional work to join together into larger sequence
-lengths. \pb
+(`contigs') composed of the many available individual reads, which represent the
+assembly of reads into the longest possible sequences given the data.
+Nonetheless, these contigs represent a fragmented picture of the genome and
+require additional work to join together into larger sequence lengths. \pb
 
 The process of ameliorating a draft sequence into a finished continuous
-sequence can be expensive in both time and laboratory effort. Therefore the
-genomic data present in a set of contigs may often be enough for many research
-questions \cite{branscomb2002}. A continuous high-quality `finished' genome
-sequence does however provide a greater depth of information, such as complete
-resolution of repeat regions and precise estimates of distances between genomic
-elements \cite{parkhill2002,fraser2002}. This process of joining contigs
-together to form a continuous sequence may be called the `scaffolding' or
-`finishing' stage. \pb
+sequence can be expensive in both time and laboratory effort, although the
+genomic data present in a set of contigs may often be sufficient for many
+research questions \cite{branscomb2002}. Nevertheless, a continuous
+high-quality `finished' genome sequence does provide a greater depth of
+information, such as complete resolution of repeat regions and precise
+estimates of distances between genomic elements \cite{parkhill2002,fraser2002}.
+This process of joining contigs together to form a continuous sequence is
+called the `scaffolding' or `finishing' stage. \pb
 
@@
 Scaffolding is the process of joining a series disconnected contigs into
-complete continuous sequence. Due to genomic complexity and missing data,
-scaffolding may not ultimately produce a final completed sequence, but may
-still succeed in joining a subset of contigs together or filling gaps. An
-overview of the scaffolding process is outlined below.
+complete continuous genome sequence. Due to genomic complexity and missing
+data, scaffolding may not ultimately produce a final completed sequence, but
+may still succeed in joining a greater subset of contigs together or resolving
+gaps. An overview of required steps in the scaffolding process is outlined
+below.
 
@@
 
-The sequencing process generates reads from either strand of the DNA helix.
-Therefore the resulting contigs constructed from these reads may point in
-either direction of the DNA. The orientation of contigs to point in the same
-direction and represent the same strand requires reverse complementing
-sequences where necessary. In the case of microbes this orientation may be to
-the 5'-3' direction leading away from the origin of replication.
+The sequencing process generates reads from either strand of the DNA helix, the
+resulting contigs constructed from these reads may point in either direction of
+the DNA. Orientating contigs to point in the same direction and represent the
+same strand requires producing the reverse complement of sequences where
+necessary. In the case of microbial genomes this orientation may be to the
+5'-3' direction leading away from the origin of replication.
 
@@
 Contig ordering determines the placement of observed contigs to best represent
-their order in the true genome sequence. Correctly placing each contig also
-highlights any extra-genomic DNA, such as plasmids which should be scaffolded
-separately from the genomic sequence. Contig order is commonly started at the
-origin of replication where the first contig is split at the replication start
-point. All subsequent contigs are then ordered in the direction of DNA
-replication.
+their order in the true genome sequence. The Correct placement of each contig
+also highlights any extra-genomic DNA, such as plasmids which are scaffolded
+separately from the genomic sequence. The contig order is commonly started at
+the contig containing the origin of replication. All subsequent contigs are
+then ordered in the 5' - 3' direction of DNA replication.
 
@@
 
-Given the correct order and orientation, determining the distances between
-contigs results in an estimate of the complete genome size. The size of inter
-contig gaps represents the length of an unknown region in the genome. Filling
-these regions with unknown nucleotide characters (`N') allows a draft
-continuous sequence. This draft scaffold sequence is useful for representing
-both the sequenced and unknown regions in the genome.
+Given the correct order and orientation, determining the distance between
+contigs results in an estimate of the complete genome size. The size of
+inter-contig gaps represents the length of an unknown region in the genome.
+Filling these regions with unknown nucleotide characters (`N') allows a draft
+continuous sequence, which is useful for representing both the sequenced and
+still to be resolved area in the genome sequence.
 
@@
 a single sequence if they are not already scaffolded together. Once all contigs
-have been joined and gaps in a scaffold closed the genome may be considered
+have been joined and gaps in a scaffold closed, the genome may be considered
 finished.
@@
 The occurrence of a pair of reads in separate contigs can be used to estimate
-the order and distance between the contigs. Alternatively laboratory methods
+the order and distance between the contigs. Alternatively, laboratory methods
 may use PCR to amplify the unknown DNA in a gap region, then use traditional
-Sanger sequencing to determine the sequence. Computational methods are more
+Sanger sequencing to determine the gap sequence. Computational methods are more
 preferable as they are less costly in laboratory time and materials compared to
-manual gap resolution \cite{nagarajan2010}. This need for computational
-scaffolding methods has resulted in a variety of algorithmic approaches. \pb
+manual gap resolution and uses the available data that exists following
+sequencing \cite{nagarajan2010}. \pb
 
@@
 genome sequence searching for areas of sequence similarity between the two.
-Areas of corresponding sequence between the the two sets are can be used to
-infer contig placement and build the scaffold \cite{richter2007,zhao2008}.
+Areas of corresponding sequence between the two sets are can be used to infer
+contig placement and build the scaffold \cite{richter2007,zhao2008,assefa2009}.
 Recombination between the two genomes can however reduce the effectiveness of
-this analysis. \pb
+this analysis. Repeat regions may be responsible for many of the gaps in
+building a genome sequence. The existence of tandemly repeated nucleotide
+regions produces many similar reads in the sequence process. The assembly of
+reads generated from repeat regions can result in the assemble collapsing into
+a single repeat instance or ignored by conservative assembly algorithms. These
+regions may be resolved by using reassembly algorithms specifically for repeat
+resolution \cite{mulyukov2002,koren2010}. Similarly a related approach may use
+unassembled sequence reads matching the regions around a scaffold gap to
+construct a uniquely overlapping set of reads across the region
+\cite{tsai2010}. \pb
 
-Repeat regions may be responsible for many of the gaps in building a genome
-sequence. The existence of tandemly repeated nucleotide regions produces many
-similar reads in the sequence process. The assembly of reads generated from
-repeat regions can result in the assemble collapsing into a single repeat
-instance or ignored by conservative assembly algorithms. These regions may be
-resolved by using reassembly algorithms specifically for repeat resolution
-\cite{mulyukov2002,koren2010}. Similarly a related approach may use uniquely
-matching reads to try and bridge across gaps between contigs \cite{tsai2010}.
-\pb
+Paired-read data provides an extra level of information on how contigs may be
+scaffolded together. Heuristic algorithms may take advantage of this data to
+search for optimal configuration of contigs in the scaffold using paired-read
+distances \cite{dayarian2010,boetzer2011}. Additional data using a reference
+genome may also be combined when estimating the best contig configuration
+\cite{pop2004}. Finally when the scaffold cannot be completely resolved
+\emph{in silico} software packages exist to suggest the likely primers
+necessary for PCR amplifying the sequence in gap regions
+\cite{gordon2001,nagarajan2010}. \pb
 
-As mentioned above, the genome may be prepared prior to sequencing so that
-reads are produced in pairs at a known distance. This paired-read data
-provides an extra level of information on how contigs may be scaffolded
-together. Heuristic graph algorithms may take advantage of this data to search
-for optimal configuration of contigs in the scaffold using paired-read distances
-\cite{dayarian2010}. Additional data using a reference genome may also be
-combined when estimating the best contig configuration \cite{pop2004}. Finally
-when the scaffold cannot be completely resolved \emph{in silico} software
-packages exist to suggest the likely primers necessary for PCR amplifying the
-sequence in gap regions \cite{gordon2001,nagarajan2010}. \pb
+These methods provide a wide variety of possible approaches for scaffolding
+a genome. The scaffold may still however require manual editing and addition of
+sequence if computational methods are not completely successful. This involves
+moving and joining larges blocks of nucleotide text together by hand which may
+be error-prone and precludes any reproducibility. \pb
 
-\subsection*{Summary} %{{{2
-
-Taken together these methods provide a wide variety of possible approaches for
-scaffolding a genome. The scaffold may still however require manual editing and
-addition of sequence if computational methods are not completely successful.
-This involves moving and joining larges blocks of nucleotide text together by
-hand which may be error-prone and precludes any reproducibility.
-\pb
+\subsection*{Overview} %{{{2
 

### 2011-02-28 src/article.tex
Christina revisions to abstract

@@
   \paragraph*{Background:} Assembly of short read sequencing data can result in
-  a fragmented series of contigs. Therefore a common step in a genome project
+  a fragmented series of sequences. Therefore a common step in a genome project
   is to join neighbouring sequence regions together and fill gaps in the
@@
   \paragraph*{Results:} This software allows a genome build to be defined in
-  terms of the constituent contigs using a relatively simple to write and edit
-  scaffold file syntax. This syntax further allows  unknown regions to be
-  defined, and inserts to be added to the scaffold to fill gaps. Defining the
-  genome construction in a file makes the scaffolding process reproducible and
-  easier to edit and compare than fasta sequence.
+  terms of the constituent contigs using a relatively simple format to write
+  and edit scaffold file syntax. This syntax further allows  unknown regions to
+  be defined, and inserts to be added to the scaffold to fill gaps. Defining
+  the genome construction in a file makes the scaffolding process reproducible
+  and easier to edit and compare than fasta sequence.
 
-  \paragraph*{Conclusions:} Scaffolder is easy to use genome scaffolding
+  \paragraph*{Conclusions:} Scaffolder is easy-to-use genome scaffolding
   software. This tool promotes reproducibility and continuous development in a

### 2011-02-28 src/article.tex
Christina changes to introduction

@@
 nucleotides from a short region in the genome. At the time of writing, high
-throughput sequencing is limited to producing reads less than 1000 nucleotides
+throughput sequencing is limited to producing reads less than 1,000 nucleotides
 in length. Therefore to resolve a sequence longer than this, such as a complete
@@
 these individual fragments. An analogy for this process is a jigsaw puzzle.
-Each nucleotide read represents a single piece and the final genome sequence is
-the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
+Each nucleotide read represents a single piece, and the final genome sequence
+is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
 within the genome, or biased and incomplete sequencing data, may prevent the
@@
 
-The recent advent of these current high throughput sequencing methods has lead
+The recent advent of these current high throughput sequencing methods has led
 to a renewed interest into algorithms for solving the problem of genome
@@
 algorithms will generate several large contiguous regions of sequence
-(`contigs') composed of the many available individual reads, which represent the
-assembly of reads into the longest possible sequences given the data.
+(`contigs') composed of the many available individual reads, which represent
+the assembly of reads into the longest possible sequences given the data.
 Nonetheless, these contigs represent a fragmented picture of the genome and
@@
 
-Scaffolding is the process of joining a series disconnected contigs into
-complete continuous genome sequence. Due to genomic complexity and missing
+Scaffolding is the process of joining a series of disconnected contigs into
+a complete continuous genome sequence. Due to genomic complexity and missing
 data, scaffolding may not ultimately produce a final completed sequence, but
@@
 
-During the scaffolding process closing and filling gaps between contigs
+During the scaffolding process, closing and filling gaps between contigs
 improves the completeness of the genome scaffold. Closing gaps may require
@@
 Sanger sequencing to determine the gap sequence. Computational methods are more
-preferable as they are less costly in laboratory time and materials compared to
-manual gap resolution and uses the available data that exists following
+preferable, as they are less costly in laboratory time and materials compared
+to manual gap resolution and uses the available data that exists following
 sequencing \cite{nagarajan2010}. \pb
@@
 genome sequence searching for areas of sequence similarity between the two.
-Areas of corresponding sequence between the two sets are can be used to infer
-contig placement and build the scaffold \cite{richter2007,zhao2008,assefa2009}.
-Recombination between the two genomes can however reduce the effectiveness of
-this analysis. Repeat regions may be responsible for many of the gaps in
-building a genome sequence. The existence of tandemly repeated nucleotide
-regions produces many similar reads in the sequence process. The assembly of
-reads generated from repeat regions can result in the assemble collapsing into
-a single repeat instance or ignored by conservative assembly algorithms. These
-regions may be resolved by using reassembly algorithms specifically for repeat
-resolution \cite{mulyukov2002,koren2010}. Similarly a related approach may use
-unassembled sequence reads matching the regions around a scaffold gap to
-construct a uniquely overlapping set of reads across the region
-\cite{tsai2010}. \pb
+Areas of corresponding sequence between the two sets can be used to infer
+contig placement and to build the scaffold
+\cite{richter2007,zhao2008,assefa2009}. Recombination between the two genomes
+can, however, reduce the effectiveness of this analysis. Repeat regions may be
+responsible for many of the gaps in building a genome sequence. The existence
+of tandemly repeated nucleotide regions produces many similar reads in the
+sequence process. The assembly of reads generated from repeat regions can
+result in the assemble collapsing into a single repeat instance or ignored by
+conservative assembly algorithms. These regions may be resolved by using
+reassembly algorithms specifically for repeat resolution
+\cite{mulyukov2002,koren2010}. Similarly a related approach may use unassembled
+sequence reads matching the regions around a scaffold gap to construct
+a uniquely overlapping set of reads across the region \cite{tsai2010}. \pb
 
-Paired-read data provides an extra level of information on how contigs may be
-scaffolded together. Heuristic algorithms may take advantage of this data to
+Paired-read data provides an extra level of information about how contigs may
+be scaffolded together. Heuristic algorithms may take advantage of this data to
 search for optimal configuration of contigs in the scaffold using paired-read
@@
 genome may also be combined when estimating the best contig configuration
-\cite{pop2004}. Finally when the scaffold cannot be completely resolved
+\cite{pop2004}. Finally when the scaffold cannot be completely resolved,
 \emph{in silico} software packages exist to suggest the likely primers
@@
 
-The software outlined here ``Scaffolder'' is aimed at creating a syntax and
+The software outlined here, ``Scaffolder'', is aimed at creating a syntax and
 software framework for producing and editing a genome scaffold. Scaffolder uses

### 2011-02-28 src/article.tex
Revise acknowledgements

@@
 This work was supported by the National Institute for Health: IDeA Network of
-Biomedical Research Excellence (KY-INBRE) grant: NIH 2P20 RR016481-09 and the
-NIH R15 AREA Program grant: NIH R15GM079775.
+Biomedical Research Excellence (KY-INBRE) grant (NIH 2P20 RR016481-09) and the
+NIH R15 AREA Program grant (R15GM079775).
 

### 2011-02-28 src/article.tex
Add HB + CW edits to methods section

@@
 Unit tests were implemented to maintain individual elements of the source code
-during development. Integration tests were written to test the Scaffolder
-software interface as a whole. Unit tests were written using the Shoulda and
-RSpec \cite{rspec} libraries.  Integration tests were written using the
-Cucumber library \cite{rspec}. \pb
+during development and were written using the Shoulda and RSpec \cite{rspec}
+libraries. Integration tests were written to test the Scaffolder software
+interface as a whole and were written using the Cucumber library \cite{rspec}.
+\pb
 
@@
 Java, Perl, C\#/.NET, PHP, and JavaScript. In addition to being widely
-supported YAML formatted scaffold files can be validated for correct syntax
-using third party tools such as Kwalify \cite{kwalify}. \pb
+supported, YAML-formatted scaffold files can be validated for correct syntax
+using third-party tools such as Kwalify \cite{kwalify}. \pb
 
-As described in the Background section, the initial sequencing data assembly
-may result in an incomplete genome build. The use of computational software and
-generation of additional sequence using PCR means genome scaffolding may be an
-on-going process. Therefore the scaffold file should be simple to update
-manually in addition to being computationally tractable. This requirement was
-therefore the second reason YAML was selected, where YAML is human-readable and
-simple to edit in a standard text editor. \pb
+Initial sequencing data assembly may result in an incomplete genome build.
+Generation of additional sequences either through PCR or computational methods
+means that genome scaffolding may be an on-going process. The scaffold file
+should therefore be simple to update manually in addition to being
+computationally tractable. This requirement was also best suited to YAML which
+is human-readable and simple to edit in a standard text editor. \pb
 
 The scaffold file takes the form of a list of entries. Each entry corresponds
-to a region of sequence, most likely a contig or previously scaffolded set of
-contigs. The Scaffolder software processes this list in order, converting each
-to the corresponding nucleotide sequence. These nucleotide sequences are then
-concatenated to produce a draft super sequence. Each entry in the scaffold file
-may have attributes that define whether a sub-sequence or the reverse
-complement of the sequence should be used. The types of attributes available,
-and an example scaffold file are outlined in the Results and Discussion section
-below.  \pb
+to a region of sequence inserted from the corresponding fasta file. The
+Scaffolder software processes this list in order, converting each to the
+corresponding nucleotide sequences, which are then concatenated to produce
+a draft super sequence. Each entry in the scaffold file may have attributes
+that define whether a sub-sequence or the reverse complement of the sequence
+should be used. The types of attributes available, and an example scaffold file
+are outlined below. \pb
 

### 2011-02-28 src/article.tex
Correct punctutation in conclusions

@@
 
-Scaffolder follows the Unix philosophy of ``Do one thing and do it well''. The
+Scaffolder follows the Unix philosophy of ``Do one thing and do it well.''. The
 focus of Scaffolder is a simple to use tool for creating a genome scaffold. The
@@
 follows a second Unix tenet: ``If your data structures are good enough, the
-algorithm to manipulate them should be trivial.''
+algorithm to manipulate them should be trivial.''.
 

### 2011-02-28 src/article.tex
Revision to results and discussion section

@@
   the genome construction in a file makes the scaffolding process reproducible
-  and easier to edit and compare than fasta sequence.
+  and easier to edit and compare than FASTA sequence.
 
@@
 The scaffold file takes the form of a list of entries. Each entry corresponds
-to a region of sequence inserted from the corresponding fasta file. The
+to a region of sequence inserted from the corresponding FASTA file. The
 Scaffolder software processes this list in order, converting each to the
@@
 
-The Scaffolder software outlined here allows computational joining of
-nucleotide sequences together into a single contiguous super sequence.
-Plain-text scaffold files written in the YAML syntax specify how these
-sequences should be joined. This scaffold file is read by the scaffolder
-software and used to generate the scaffold sequence. In addition to specifying
-which contigs are required, the scaffold file allows the contigs to be edited
-to produce smaller sub-sequences or reverse complemented if necessary. \pb
+The Scaffolder software allows computational joining of nucleotide sequences
+together into a single contiguous scaffolded super-sequence. Plain-text
+scaffold files written in the YAML syntax specify how these sequences should be
+joined and the scaffold file as read by the scaffolder software is used to
+generate the scaffold sequence. In addition to specifying which contigs are
+required, the scaffold file allows the contigs to be edited to produce smaller
+sub-sequences or reverse complemented if necessary. The process of genome
+finishing may involve producing additional oligonucleotide sequences to fill
+unknown regions in the scaffold. The scaffold file format allows these
+additional insert sequences to bridge these gaps. These inserts can also be
+treated in a similar way to larger contig sequences: trimmed and/or reverse
+complemented to match the corresponding gap region. \pb
 
-The process of genome finishing may involve producing additional
-oligonucleotide sequence to fill unknown regions in the scaffold. The scaffold
-file format allows these additional insert sequences to be added to bridge
-these gaps. These inserts can also be treated in a similar way to larger contig
-sequences: trimmed and/or reverse complemented to match the corresponding gap
-region. \pb
+The distances between contigs may be estimated from paired read data or from
+mapping the contigs to a reference genome. These inter-contig gap regions are
+useful to join separate sequences together by an estimated size and the
+scaffold file allows for the specification of these unknown regions. The use of
+these regions in the scaffold which areas in the scaffold and thier approximate
+size are still to be resolved. \pb
 
-Distances between contigs may be estimated from paired read data or from
-mapping the contigs to a reference genome. These inter-contig gap regions are
-useful to join regions of sequence together by an estimated size. The scaffold
-file allows for the specification of these unknown regions to join contigs
-together. The use of these regions in the scaffold highlights areas still to be
-resolved. \pb
-
-The scaffold file is used to specify how sequence regions are joined together
-but the sequences themselves are not stored in the scaffold file. The sequences
-are instead maintained as a separate fasta file. A fasta file of contigs and
-scaffolded contigs is a common output for assemblers and as such should be
-readily available. The nucleotide sequences in the fasta file are referenced in
-the scaffold using the first space-delimited word from each fasta header.
-Maintaining the nucleotide sequences in a separate file preserves the original
-sequence and furthermore decouples the sequence from the markup describing how
-it should be used to generate the genome sequence. \pb
+The scaffold file specifies how sequence regions are joined together; however
+the sequences themselves are not stored in the scaffold file and are instead
+maintained as a separate FASTA file. A FASTA file of contigs and scaffolded
+contigs is a common output for assemblers. The nucleotide sequences stored in
+the corresponding FASTA file are referenced in the scaffold using the first
+word in the FASTA header. Maintaining the nucleotide sequences in a separate
+file preserves the original sequence and decouples the original sequence data
+from the scaffold file describing how it should be used to in the genome
+sequence. \pb
 
@@
 
-The scaffold file is written using the YAML syntax. YAML is a data format
-designed to be both machine readable and easy to edit manually. An example
-scaffold file is shown in Figure 1. This file illustrates the text attributes
-used to describe a scaffold and the joining of sequences in the genome build.
-The basic layout of the scaffold file is a list - a series of entries where
-each entry corresponds to a region of sequence in the generated scaffold
-super-sequence. The right hand side indicates how the entries are joined
-together. This figure is illustrates the putative scaffold and is not to scale.
-\pb
+The scaffold file is written using the YAML syntax. An example scaffold file is
+shown in Figure 1. This file illustrates the text attributes used to describe
+a scaffold and the joining of sequences in the genome build. The basic layout
+of the scaffold file is a list of entries, where each entry corresponds to
+a region of sequence in the generated scaffold super-sequence. \pb
 
@@
 
-The first line of the scaffold file begins with three dashes. This is
-a requirement to indicate the start of a YAML formatted document. The first
-entry, highlighted in green, in the scaffold begins on the second line. The
-first line of this entry begins with a dash character `-', dashes are used to
-denote entries in a YAML list. The first line of each entry in the scaffold
-therefore begins with a dash line. The next line is indented by two spaces
-- whitespace is required to group similar attributes together. \pb
-
-The ``sequence'' tag indicates this entry corresponds to a sequence in the
-fasta file. The following line indicates the name of the fasta sequence using
-the ``source'' tag. As described above the first space-delimited word of the
-fasta header is used to identify which sequence is selected from the fasta
-file. Together these three lines describe the first entry in the scaffold as
-a simple sequence using the fasta sequence named `sequence1'. On the right hand
-side of Figure 1 this produces the first region in the scaffold, also shown in
-green. \pb
+The first line of the scaffold file begins with three dashes to indicate the
+start of a YAML-formatted document. The first entry (highlighted in green)
+begins with a dash character `-' to denote an entry in the YAML list. The first
+line of each entry in the scaffold therefore begins with a dash line. The next
+line is indented by two spaces as whitespace is required to group similar
+attributes together. The ``sequence'' tag indicates that this entry corresponds
+to a sequence in the FASTA file and the following line indicates the name of
+this FASTA sequence using the ``source'' tag. As described the first
+space-delimited word of the FASTA header is used to identify which sequence is
+selected from the FASTA file. Together these three lines describe the first
+entry in the scaffold as a simple sequence using the FASTA sequence named
+`sequence1'. On the right hand side of Figure 1 this produces the first region
+in the scaffold, also shown in green. \pb
 
@@
 
-The last entry in the scaffold, highlighted in blue, adds the fasta entry
-`sequence2' to the scaffold. This entry demonstrates how the nucleotide
-sequence may be manipulated prior to addition to the scaffold. The `start' and
-`stop' tags trim the sequence to these coordinates inclusively. The ``reverse''
-tag instructs scaffolder to also reverse complement the sequence. On the right
-hand side of the sequence of Figure 1 this completes the scaffold sequence. \pb
+The last entry in the scaffold, highlighted in blue, adds the FASTA entry
+`sequence2' to the scaffold. This entry demonstrates how this sequence may be
+manipulated prior to addition to the scaffold. The `start' and `stop' tags trim
+the sequence to these coordinates inclusively, while the ``reverse'' tag
+instructs scaffolder to also reverse complement the sequence. In the scaffolded
+genome shown in Figure 1 this completes the scaffold sequence. \pb
 
-This final entry also includes the ``inserts'' tag to add additional regions of
-sequence. These inserts are also added as a YAML list with each insert
-beginning with a dash line. The first insert, shown in purple, illustrates
-using similar attributes to that of a sequence entry. The reverse, start and
-stop tags, are used to trim and reverse complement the insert. Similarly the
-`source' tag identifies the corresponding fasta sequence. The ``open'' and
+The final entry in the scaffold also includes the ``inserts'' tag to add
+additional regions of sequence. These inserts are also added as a YAML list,
+with each insert beginning with a dash line. The first insert, shown in purple,
+similar attributes to that of a sequence entry; the reverse, start and stop
+tags are used to trim and reverse complement the insert. Similarly the `source'
+tag identifies the corresponding FASTA sequence as `insert'. The ``open'' and
 ``close'' tags are specific to inserts and determine where the insert is added
 in the enclosing sequence. The region of the sequence inside these coordinates
-is inclusively replaced by the specified insert fasta sequence. This is
-highlighted in the putative scaffold in Figure 1 by the black lines bisecting
-the blue sequence. \pb
+is inclusively replaced by the specified insert FASTA sequence. This is
+visualised in the putative gene Figure 1 by the black lines bisecting the blue
+sequence. \pb
 
-The next insert, shown in brown, is specified using only the open position tag.
-This illustrates that only one of either `open' or `close' tags is required.
-When only one of these tag is used the corresponding opposing coordinate
-position is calculated from the length of the insert fasta sequence. This
-allows inserts to bridge into, and partially fill gap regions without requiring
-an end coordinate position. \pb
+The next insert, shown in brown, is specified using only the `open' tag. This
+illustrates that only one of either `open' or `close' tags is required when
+adding an insert sequence. If only only one of the `open' or `close' tags is
+used the corresponding opposing position distance is calculated from the length
+of the insert FASTA sequence. This allows inserts to bridge into, and partially
+fill gap regions without requiring an end coordinate position. \pb
 
@@
 The the scaffold file format allows multiple sequence-editing operations to be
-applied to a sequence or insert. If applied in different combinations, these
-operations will produce different output sequences. An example would be reverse
-complementing the sequence then trimming the start of the sequence by five
-nucleotides. Performing these actions in the reverse order would result in five
-nucleotides being removed from the opposite end. This section therefore
-outlines the Scaffolder protocol for processing sequence entries. \pb
+applied to a sequence or insert. If applied in different combinations, then
+these operations will produce different output sequences. An example would be
+reverse complementing the sequence then trimming the start of the sequence by
+five low base-call quality nucleotides. Performing these actions in the reverse
+order would result in five nucleotides being removed from the opposite end.
+This section therefore outlines the Scaffolder protocol for processing sequence
+entries. \pb
 
 \paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts, these inserts are reverse sorted by the insert close coordinate
-position. The insert with the largest close position is first in the list and
+inserts, these inserts are sorted by decesding value of the insert `close' tag
+value. The insert with the furthest `close' coordinate is first in the list and
 the insert with the smallest close position is last. This preserves the
-original intended coordinates as each insert is added to the sequence. \pb
+intended location for each insert if the size of the encoding sequence is
+changed by the addition of inserts. \pb
 
 \paragraph{Update sequence with inserts:} Each insert sequence is processed
-according to any specified attributes. The insert is then added to the
-enclosing sequence in the region specified by the open and close coordinates.
-\pb
-
-If the insert added to the sequence is a different size to that of the
-replaced region then the sequence start and stop co-ordinates are updated. For
-example if a 7bp insert replaced a 5bp region in the sequence the stop
-coordinate of the enclosing sequence would be incremented by the 2bp
-difference. As inserts are added by reverse close position the remaining
-inserts are unaffected by changes in the encoding sequence size. \pb
+according to any specified attributes, before being added to the enclosing
+sequence in the region specified by the open and close coordinates. If the
+insert added to the sequence is a different size to that of the replaced region
+then the sequence start and stop coordinates are updated. For example if a 7bp
+insert replaces a 5bp region in the encoding sequence, then the stop coordinate
+of the enclosing sequence would be incremented by the 2bp difference. As
+inserts are added by reverse close position the remaining inserts are
+unaffected by changes in the encoding sequence size. \pb
 
 \paragraph{Trimming sequence:} If the sequence has start or stop tags
-specified these are used as coordinates to trim the sequence to the specified
-length inclusively. \pb
+specified, then these are used as coordinates to trim the sequence to the
+specified length inclusively of the `start' and `stop' coordinate tags. \pb
 
-\paragraph{Reverse complement sequence:} If the reverse tag is specified the
-sequence is reverse complemented. This is final step in processing the
-sequence and the sequence is then returned for inclusion in the scaffold. \pb
+\paragraph{Reverse complement sequence:} If the reverse tag is specified, then
+the sequence is reverse complemented, which is the final is the final step in
+processing the sequence. \pb
 
@@
 
-The ``sequence'' command builds the fasta sequence from the scaffold and fasta
-file. Running this command requires the locations of both the scaffold file
-and the fasta file to be passed as arguments. This will then return the
-complete scaffold sequence in fasta format. \pb
+The ``sequence'' command builds the FASTA sequence from the scaffold and FASTA
+file. Running this command requires the locations of both the scaffold file and
+the FASTA file to be passed as arguments. This will then return the complete
+scaffold sequence in FASTA format. \pb
 
@@
 This code snippet illustrates that the scaffold object is initialised with two
-arguments: the YAML loaded scaffold and the location of the fasta sequence
+arguments: the YAML loaded scaffold and the location of the FASTA sequence
 file. The scaffold is an enumerable array and the penultimate line illustrates
@@
 
-The aims of Scaffolder are twofold. The first is providing software which is
+The aims of Scaffolder are twofold. The first is providing software that is
 easy to install and simplifies the task of genome finishing. The second aim is
@@
 
-\subsubsection*{Simple to use} %{{{3
+Scaffolder was designed to be as simple to use as possible and assuming the
+Ruby and RubyGems software are present, Scaffolder can be installed in a single
+command line step. This should negate any possible barriers to entry which
+require manual compilation of source code. Scaffolder also uses a minimal and
+compact syntax to describe the genome scaffold. This syntax is simple to
+construct and edit whilst being succinct and readable. The goal of this syntax
+is to make genome scaffolds easy to construct in a common text editor. \pb
 
-Scaffolder is designed to be as simple to use as possible. Assuming the Ruby
-and RubyGems software are present, Scaffolder can be installed in a single
-command line step. This should negate any possible barriers to entry which
-require manual compilation of source code. \pb
-
-Scaffolder uses a minimal and compact syntax to describe the genome scaffold.
-This syntax is simple to construct and edit whilst being succinct and readable.
-The goal of this syntax is to make genome scaffolds easy to construct in
-a common text editor. Furthermore Scaffolder provides extensive documentation
-both for the CLI and for the Ruby API. Therefore should any part of the
-scaffolding process be unclear the documentation should make an solution easy
-to find. \pb
-
-\subsubsection*{Reproducible genome construction} %{{{3
-
-Constructing a genome sequence using a text file makes generating a scaffold
-super sequence both reproducible and deterministic. In comparison joining large
-nucleotide sequences by hand cannot be reliably reproduced. The layout of the
-scaffold file furthermore provides a human readable record of how the scaffold
-was constructed and decouples the layout of contigs from their source sequence.
-\pb
+Constructing a genome using a text file makes generating a scaffold super
+sequence both reproducible and deterministic for the same scaffold and FASTA
+sequence file. In comparison, joining large nucleotide sequences by hand cannot
+be reliably reproduced. The scaffold file furthermore provides a human readable
+description of how the scaffold is constructed. \pb
 

### 2011-02-28 src/article.tex
Remove scaffolder usage subsection

@@
 
-\subsection*{Scaffolder Usage} %{{{2
-
-Scaffolder provides a standardised set of Ruby classes and methods (API) for
-interacting with the scaffold. This allows Scaffolder to be integrated into
-existing genomics workflows or used with Ruby build tools such as Rake. \pb
-
-In addition Scaffolder provides a command line interface (CLI). This allows
-the scaffold sequence to be generated and validated without any Ruby
-experience. The Scaffolder CLI behaves as a standard Unix tool - returning
-appropriate exit codes and providing manual pages. The use of both these
-Scaffolder interfaces is outlined in the following sections. \pb
-
-\subsubsection*{Installation} %{{{3
-
-Scaffolder relies on both the Ruby programming language and the RubyGems
-package management system. These are either already present or easily installed
-on recent distributions of Mac OS X or Linux. The presence of both these tools
-can be tested by typing the following commands on the terminal. \pb
-
-  \begin{verbatim}
-    ruby -v
-    gem -v
-  \end{verbatim}
-
-If these commands return without error this indicates the necessary components
-are available to install and run Scaffolder. The Scaffolder API and CLI can
-then be installed using the following RubyGems command. \pb
-
-  \begin{verbatim}
-    gem install scaffolder scaffolder-tools
-  \end{verbatim}
-
-To ensure that the Scaffolder CLI is accessible, the RubyGems `bin' directory
-should be in the command line path. Assuming that RubyGems are installed to
-$\sim$/.gem this can be done by adding the following to the $\sim$/.bashrc
-file. The corresponding login script should be updated for other shells such as
-Zsh, etc.. \pb
-
-  \begin{verbatim}
-    export PATH=~/.gem/bin:$PATH
-  \end{verbatim}
-
-\subsubsection*{Command Line Interface} %{{{3
-
-The Scaffolder CLI provides functions to validate the scaffold file and build
-a draft super sequence. Assuming Scaffolder has been installed and the binary
-is available in the command line path the following should run without error.
-This command will list the available Scaffolder functions. \pb
-
-  \begin{verbatim}
-    scaffolder
-  \end{verbatim}
-
-Help on a specific command can be found by typing ``help'' followed by the
-name of the command. This prints the corresponding Unix manual page to the
-terminal. The following example will return the manual page for the sequence
-command. \pb
-
-  \begin{verbatim}
-    scaffolder help sequence
-  \end{verbatim}
-
-The ``sequence'' command builds the FASTA sequence from the scaffold and FASTA
-file. Running this command requires the locations of both the scaffold file and
-the FASTA file to be passed as arguments. This will then return the complete
-scaffold sequence in FASTA format. \pb
-
-  \begin{verbatim}
-    scaffolder sequence scaffold.yml sequences.fna
-  \end{verbatim}
-
-Scaffolder provides a function to test the scaffold for overlapping insert
-coordinates. Overlapping inserts will result in unexpected behaviour when
-generating the scaffold sequence as the coordinates for the first insert may
-change the intended coordinates for subsequent inserts. The ``validate''
-command will check the scaffold for any such inserts and return them as a YAML
-list. \pb
-
-  \begin{verbatim}
-    scaffolder validate scaffold.yml sequences.fna
-  \end{verbatim}
-
-\subsubsection*{Application Programming Interface} %{{{3
-
-Scaffolder provides an API allowing Ruby programs to manipulate the scaffold.
-This API is detailed in full on the Scaffolder website (\scaffolder). The
-following code snippet outlines how the API may be used to create a script to
-replicate the function of the `scaffolder sequence' command described in the
-CLI section above. \pb
-
-  \begin{verbatim}
-    #!/usr/bin/env ruby
-
-    require `rubygems'
-    require `scaffolder'
-
-    scaffold_file = ARGV.shift
-    sequence_file = ARGV.shift
-
-    scaffold = Scaffolder.new(
-      YAML.load(File.read(scaffold_file)),
-      sequence_file)
-
-    sequence = scaffold.map{|entry| entry.sequence }.join
-
-    puts sequence
-  \end{verbatim}
-
-This code snippet illustrates that the scaffold object is initialised with two
-arguments: the YAML loaded scaffold and the location of the FASTA sequence
-file. The scaffold is an enumerable array and the penultimate line illustrates
-the trivial code required to generate the super-sequence. Further details of
-the classes and methods related to each scaffold entry can be found in the
-documentation for the Scaffolder::Region class in the documentation on the
-Scaffolder website. \pb
-
 \subsection*{Discussion of Scaffolder Approach} %{{{2

### 2011-02-28 src/article.tex
Remove section on scaffolder attribute processing

@@
 
-\subsubsection*{Scaffolder file attribute processing} %{{{3
-
-The the scaffold file format allows multiple sequence-editing operations to be
-applied to a sequence or insert. If applied in different combinations, then
-these operations will produce different output sequences. An example would be
-reverse complementing the sequence then trimming the start of the sequence by
-five low base-call quality nucleotides. Performing these actions in the reverse
-order would result in five nucleotides being removed from the opposite end.
-This section therefore outlines the Scaffolder protocol for processing sequence
-entries. \pb
-
-\paragraph{Reverse sort inserts:} If the sequence entry contains a list of
-inserts, these inserts are sorted by decesding value of the insert `close' tag
-value. The insert with the furthest `close' coordinate is first in the list and
-the insert with the smallest close position is last. This preserves the
-intended location for each insert if the size of the encoding sequence is
-changed by the addition of inserts. \pb
-
-\paragraph{Update sequence with inserts:} Each insert sequence is processed
-according to any specified attributes, before being added to the enclosing
-sequence in the region specified by the open and close coordinates. If the
-insert added to the sequence is a different size to that of the replaced region
-then the sequence start and stop coordinates are updated. For example if a 7bp
-insert replaces a 5bp region in the encoding sequence, then the stop coordinate
-of the enclosing sequence would be incremented by the 2bp difference. As
-inserts are added by reverse close position the remaining inserts are
-unaffected by changes in the encoding sequence size. \pb
-
-\paragraph{Trimming sequence:} If the sequence has start or stop tags
-specified, then these are used as coordinates to trim the sequence to the
-specified length inclusively of the `start' and `stop' coordinate tags. \pb
-
-\paragraph{Reverse complement sequence:} If the reverse tag is specified, then
-the sequence is reverse complemented, which is the final is the final step in
-processing the sequence. \pb
-
 \subsection*{Discussion of Scaffolder Approach} %{{{2

### 2011-03-01 src/article.tex
Spell check
Split results and discussion sections
Add further description of the type of input data

@@
 
-\section*{Results and Discussion} %{{{1
+\section*{Results} %{{{1
 
@@
 scaffold file allows for the specification of these unknown regions. The use of
-these regions in the scaffold which areas in the scaffold and thier approximate
+these regions in the scaffold which areas in the scaffold and their approximate
 size are still to be resolved. \pb
@@
 the sequences themselves are not stored in the scaffold file and are instead
-maintained as a separate FASTA file. A FASTA file of contigs and scaffolded
-contigs is a common output for assemblers. The nucleotide sequences stored in
-the corresponding FASTA file are referenced in the scaffold using the first
-word in the FASTA header. Maintaining the nucleotide sequences in a separate
-file preserves the original sequence and decouples the original sequence data
-from the scaffold file describing how it should be used to in the genome
-sequence. \pb
+maintained as a separate FASTA file. Series of nucleotide sequences in
+a FASTA-formatted file is standard output for a genome assembler. These may be
+contigs, sets of contigs scaffolded into larger sequences, or both. These
+nucleotide sequences stored in the corresponding FASTA file are referenced in
+the scaffold using the first word form the FASTA header of the corresponding
+nucleotide sequence. Maintaining the nucleotide sequences in a separate file
+preserves the original sequence and decouples the original sequence data from
+the scaffold file describing how it should be used to in the genome sequence.
+\pb
 
@@
 to a sequence in the FASTA file and the following line indicates the name of
-this FASTA sequence using the ``source'' tag. As described the first
-space-delimited word of the FASTA header is used to identify which sequence is
-selected from the FASTA file. Together these three lines describe the first
-entry in the scaffold as a simple sequence using the FASTA sequence named
-`sequence1'. On the right hand side of Figure 1 this produces the first region
-in the scaffold, also shown in green. \pb
+this FASTA sequence using the ``source'' tag. As described the first word of
+the FASTA header is used to identify which sequence is selected from the file.
+Together these three lines describe the first entry in the scaffold as a simple
+sequence using the FASTA sequence named `sequence1'. On the right hand side of
+Figure 1 this produces the first region in the scaffold, also shown in green.
+\pb
 
@@
 
-\subsection*{Discussion of Scaffolder Approach} %{{{2
+\section*{Discussion} %{{{1
 

### 2011-03-01 src/article.tex
Add reason for scaffolder at top of discussion

@@
 
-The aims of Scaffolder are twofold. The first is providing software that is
-easy to install and simplifies the task of genome finishing. The second aim is
-to facilitate reproducibility when joining large nucleotide sequences together
-to create a genome sequence. \pb
+Scaffolding is a intermediate stage in a genome project. Additional sequences
+are produced using a combination of computational and laboratory methods. These
+new sequences are then added to existing sequencing data to produce
+a continuous sequence of the target organism genome. The scaffolding process
+however is a laborious manual process involving joining and adding large
+regions of nucleotide sequence together. Any manual process is inherently hard
+to reproduce and introduces the possibility of human error. The aims of
+Scaffolder are therefore twofold. The first is providing software that is easy
+to install and simplifies the task of genome finishing. The second aim is to
+facilitate reproducibility when joining large nucleotide sequences together to
+create a genome sequence. \pb
 

### 2011-03-01 src/article.tex
Who is scaffolder for in conclusions section

@@
 
-Scaffolder follows the Unix philosophy of ``Do one thing and do it well.''. The
-focus of Scaffolder is a simple to use tool for creating a genome scaffold. The
-YAML data format for creating the scaffold file is standardised and easily
-manipulated programmatically. This thereby means the scaffolding process
-follows a second Unix tenet: ``If your data structures are good enough, the
-algorithm to manipulate them should be trivial.''.
+Scaffolder is software aimed at bioinformaticians and biologists familar with
+the command line who wish to build scaffold a genome sequence in a concise
+readable way. The Scaffolder file format allows other biologists to see how the
+genome sequence was assembled and allows a broad overview of how the sequence
+was assembled not possible not possible from a megabase-sized string of
+nucleotide characters. Scaffolder also furthers the cause for reporducibility
+in genome sequence by allowing the scaffold super-sequence to be reliably
+reproduced from the scaffold file. \pb
+
+The design of the Scaffolder software follows the Unix philosophy of ``Do one
+thing and do it well.''. The focus of Scaffolder is a simple to use tool for
+creating a genome scaffold. The YAML data format for creating the scaffold file
+is standardised and easily manipulated programmatically. This thereby means the
+scaffolding process follows a second Unix tenet: ``If your data structures are
+good enough, the algorithm to manipulate them should be trivial.''.
 

### 2011-03-01 src/article.tex
Provide short description of CLI and API in results

@@
 
+\subsection{Scaffolder Sofware Interface} %{{{2
+
+Scaffolder provides a standardised set of Ruby classes and methods (API) for
+interacting with the scaffold. This allows Scaffolder to be integrated into
+existing genomics workflows or used with Ruby build tools such as Rake. In
+addition Scaffolder provides a command line interface (CLI) providing functions
+to validate the scaffold file and build a draft super sequence. This allows the
+scaffold sequence to be generated and validated without any Ruby experience.
+The Scaffolder CLI behaves as a standard Unix tool and returns appropriate exit
+codes and manual pages. The use of both these Scaffolder interfaces is outlined
+in detail on the scaffolder website (\scaffolder). \pb
+
 \section*{Discussion} %{{{1

### 2011-03-01 src/article.tex
Non-restrictive clause in introduction

@@
 
-  \paragraph*{Methods:} The software outlined here ``Scaffolder'' is
+  \paragraph*{Methods:} The software outlined here, ``Scaffolder'', is
   implemented in the Ruby programming language and can be installed via the
-  RubyGems software management system. Genome scaffolds are defined using
-  YAML - a markup language, which is both human and machine-readable. Command
-  line binaries and extensive documentation are available.
+  RubyGems software management system. Genome scaffolds are defined using YAML
+  - a markup language, which is both human and machine-readable. Command line
+  binaries and extensive documentation are available.
 

### 2011-03-01 src/article.tex
Description of scaffolder problem in background

@@
 
-These methods provide a wide variety of possible approaches for scaffolding
-a genome. The scaffold may still however require manual editing and addition of
-sequence if computational methods are not completely successful. This involves
-moving and joining larges blocks of nucleotide text together by hand which may
-be error-prone and precludes any reproducibility. \pb
+These methods provide a wide array of approaches for scaffolding contigs into
+a continious super-sequence. The scaffolding process may still however require
+joining and addition of sequence by hand in the case manually derived
+scaffolding data such as through PCR. This intermediate manual stage in the
+scaffolding process involves moving and joining larges blocks of nucleotide
+text together by hand which may be error-prone and precludes any
+reproducibility. \pb
 
-\subsection*{Overview} %{{{2
-
-The software outlined here, ``Scaffolder'', is aimed at creating a syntax and
-software framework for producing and editing a genome scaffold. Scaffolder uses
-a specific file format to define how contigs are joined, additional sequences
-inserted, and for the specification of unknown regions. This syntax allows
-a scaffold to be updated by simply editing the scaffold file and allows the
-scaffolding process to be both reproducible and deterministic. \pb
+The software outlined here, ``Scaffolder'', aims to solve this problem by
+creating at creating a syntax and software framework for producing and editing
+a genome scaffold. Scaffolder uses a specific file format to define how contigs
+are joined, additional sequences inserted, and for the specification of unknown
+regions. This syntax allows a scaffold to be updated by simply editing the
+scaffold file wich allows the scaffolding process to be both reproducible and
+provides a concise overview of how the scaffold was constructed. \pb
 

### 2011-03-02 src/article.tex
Proof read new sections

--- 2011-03-01 src/article.tex
@@
 
-These methods provide a wide variety of possible approaches for scaffolding
-a genome. The scaffold may still however require manual editing and addition of
-sequence if computational methods are not completely successful. This involves
-moving and joining larges blocks of nucleotide text together by hand which may
-be error-prone and precludes any reproducibility. \pb
+These methods provide a wide array of approaches for merging contigs into
+larger a continuous scaffolds. The scaffolding process may still however
+require inserting additional sequence or joining contigs using PCR-derived
+information from the laboratory. This manual stage in the finishing process
+will moving and editing large blocks of nucleotide text by hand which is subject
+to human error-prone and precludes any reproducibility for the steps taken. \pb
 
-\subsection*{Overview} %{{{2
-
-The software outlined here, ``Scaffolder'', is aimed at creating a syntax and
-software framework for producing and editing a genome scaffold. Scaffolder uses
-a specific file format to define how contigs are joined, additional sequences
-inserted, and for the specification of unknown regions. This syntax allows
-a scaffold to be updated by simply editing the scaffold file and allows the
-scaffolding process to be both reproducible and deterministic. \pb
+The software outlined here, ``Scaffolder'', aims to solve address these
+problems by creating a file syntax and software framework for editing a genome
+scaffold. Scaffolder uses a specific file format to define how contigs are
+joined, additional sequences inserted, and for the specification of unknown
+regions. This syntax allows a scaffold to be updated by simply editing the
+scaffold file facilitating a reproducible scaffolding process and providing
+a concise overview of how the scaffold was constructed. \pb
 
@@
 
-\subsection{Scaffolder Sofware Interface} %{{{2
+\subsection{Scaffolder Software Interface} %{{{2
 
@@
 addition Scaffolder provides a command line interface (CLI) providing functions
-to validate the scaffold file and build a draft super sequence. This allows the
-scaffold sequence to be generated and validated without any Ruby experience.
-The Scaffolder CLI behaves as a standard Unix tool and returns appropriate exit
-codes and manual pages. The use of both these Scaffolder interfaces is outlined
-in detail on the scaffolder website (\scaffolder). \pb
+to validate the scaffold file and build a draft super sequence. The Scaffolder
+CLI behaves as a standard Unix tool and returns appropriate exit codes and
+manual pages. The use of both these Scaffolder interfaces is outlined in detail
+on the scaffolder website (\scaffolder). \pb
 
@@
 
-Scaffolding is a intermediate stage in a genome project. Additional sequences
-are produced using a combination of computational and laboratory methods. These
-new sequences are then added to existing sequencing data to produce
-a continuous sequence of the target organism genome. The scaffolding process
-however is a laborious manual process involving joining and adding large
-regions of nucleotide sequence together. Any manual process is inherently hard
-to reproduce and introduces the possibility of human error. The aims of
-Scaffolder are therefore twofold. The first is providing software that is easy
-to install and simplifies the task of genome finishing. The second aim is to
-facilitate reproducibility when joining large nucleotide sequences together to
-create a genome sequence. \pb
+Scaffolding requires joining assembled contigs and insert additional sequence
+data using a combination of computational and laboratory methods. The process
+of manually editing a scaffold is inherently hard to reproduce and introduces
+the possibility of human error. The aims of this Scaffolder software are
+therefore twofold. The first is providing software that is easy to install and
+simplifies the task of genome finishing. The second aim is to facilitate
+reproducibility when joining large nucleotide sequences together to create
+a genome sequence. \pb
 
@@
 
-Scaffolder is software aimed at bioinformaticians and biologists familar with
-the command line who wish to build scaffold a genome sequence in a concise
-readable way. The Scaffolder file format allows other biologists to see how the
-genome sequence was assembled and allows a broad overview of how the sequence
-was assembled not possible not possible from a megabase-sized string of
-nucleotide characters. Scaffolder also furthers the cause for reporducibility
-in genome sequence by allowing the scaffold super-sequence to be reliably
-reproduced from the scaffold file. \pb
-
-The design of the Scaffolder software follows the Unix philosophy of ``Do one
-thing and do it well.''. The focus of Scaffolder is a simple to use tool for
-creating a genome scaffold. The YAML data format for creating the scaffold file
-is standardised and easily manipulated programmatically. This thereby means the
-scaffolding process follows a second Unix tenet: ``If your data structures are
-good enough, the algorithm to manipulate them should be trivial.''.
+Scaffolder is software aimed at bioinformaticians and biologists familiar with
+the command line building genome sequence scaffold. The file format maintains
+the scaffold in a concise and readable way allowing third parties to see how
+the genome sequence was assembled. This file format allows a broad overview of
+of which sequences were included and how they are ordered possible not possible
+from a megabase-sized strings of nucleotide characters. Scaffolder also
+furthers the cause for reproducibility in genome projects by allowing the
+scaffold super-sequence to be reliably reproduced from the scaffold file. The
+YAML syntax for writing the scaffold file is also standardised and to
+manipulate programmatically. This thereby means the scaffolding process follows
+the Unix tenet of: ``If your data structures are good enough, the algorithm to
+manipulate them should be trivial.''.
 

### 2011-03-02 src/article.tex
Add version number
Revise title and contact details
Proof read abstract

@@
 
-\title{Scaffolder - Software for Microbial Genome Scaffolding.} %{{{1
+\title{Scaffolder - Software for Reproducible Genome Scaffolding.} %{{{1
 
@@
   Michael D Barton$^{1}$\correspondingauthor%
-  \email{Michael D Barton\correspondingauthor  - mail@nucleotid.es}%
+  \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
 \and
@@
 
-  \paragraph*{Background:} Assembly of short read sequencing data can result in
-  a fragmented series of sequences. Therefore a common step in a genome project
-  is to join neighbouring sequence regions together and fill gaps in the
-  assembly using additional sequence. This scaffolding step, however, is
-  non-trivial and requires manually editing large blocks of nucleotide
-  sequence. Joining sequence contigs together also hides the source of each
-  region in the final genome sequence. Taken together, these considerations may
-  make reproducing or editing an existing genome build difficult.
+  \paragraph*{Background:} Assembly of short-read sequencing data can result in
+  a fragmented non-contiguous series of genomic sequences. Therefore a common
+  step in a genome project is to join neighbouring sequence regions together
+  and fill gaps in the assembly using additional sequence. This scaffolding
+  step, however, is non-trivial and requires manually editing large blocks of
+  nucleotide sequence. Joining sequence contigs together also hides the source
+  of each region in the final genome sequence. Taken together, these
+  considerations may make reproducing or editing an existing genome build
+  difficult.
 
@@
   RubyGems software management system. Genome scaffolds are defined using YAML
-  - a markup language, which is both human and machine-readable. Command line
+  - a data format, which is both human and machine-readable. Command line
   binaries and extensive documentation are available.
@@
   \paragraph*{Results:} This software allows a genome build to be defined in
-  terms of the constituent contigs using a relatively simple format to write
+  terms of the constituent sequences using a relatively simple format to write
   and edit scaffold file syntax. This syntax further allows  unknown regions to
@@
   the genome construction in a file makes the scaffolding process reproducible
-  and easier to edit and compare than FASTA sequence.
+  and easier to edit and compare than FASTA nucleotide sequence.
 
@@
   \begin{description}
-    \item[Project name:] Scaffolder
+    \item[Project name:] Scaffolder, version 0.4.1
     \item[Project home page:] \scaffolder

### 2011-03-02 src/article.tex
Proof read introduction

@@
 
-High throughput sequencing can produce hundreds of thousands to millions of
+High-throughput sequencing can produce hundreds of thousands to millions of
 sequence reads from a genome. Each sequence represents the machine-read
-nucleotides from a short region in the genome. At the time of writing, high
-throughput sequencing is limited to producing reads less than 1,000 nucleotides
-in length. Therefore to resolve a sequence longer than this, such as a complete
-genome, these numerous smaller fragments must be pieced together. The process
-of joining reads into longer sequences is the `assembly' stage of a genome
-project \cite{miller2010}. \pb
+nucleotides from a short region in the genome. At the time of writing,
+high-throughput sequencing is limited to producing reads less than 1,000
+nucleotides in length. Therefore to resolve a sequence longer than this, such
+as a complete genome, these numerous smaller fragments must be pieced together.
+The process of joining reads into longer sequences is the `assembly' stage of
+a genome project \cite{miller2010}. \pb
 
@@
 hardware and, in the ideal case, outputs a complete genome sequence composed of
-these individual fragments. An analogy for this process is a jigsaw puzzle.
-Each nucleotide read represents a single piece, and the final genome sequence
+these individual fragments. An analogy for this process is a jigsaw puzzle:
+each nucleotide read represents a single piece, and the final genome sequence
 is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
@@
 genome being assembled into a continuous sequence. This may be due to not
-enough or multiple overlaps between sequence reads and is analogous to missing
-pieces in the jigsaw or pieces that fit to multiple other pieces. \pb
+enough, or multiple different overlaps between reads and is analogous to
+missing pieces in the jigsaw or pieces that fit to multiple other pieces. \pb
 
@@
 assembly \cite{pop2008,pop2009}; genome assembly software may however be unable
-to produce a complete sequence from these reads. In many cases assembly
-algorithms will generate several large contiguous regions of sequence
-(`contigs') composed of the many available individual reads, which represent
-the assembly of reads into the longest possible sequences given the data.
-Nonetheless, these contigs represent a fragmented picture of the genome and
-require additional work to join together into larger sequence lengths. \pb
+to produce a complete sequence from these reads. In many cases the algorithm
+will generate several large assembled regions of sequence (`contigs') composed
+of the many available individual reads, which represent the assembly of reads
+into the longest possible sequences given the data. Nonetheless, these contigs
+represent a fragmented picture of the genome and require additional work to
+join together into larger lengths. \pb
 
@@
 may still succeed in joining a greater subset of contigs together or resolving
-gaps. An overview of required steps in the scaffolding process is outlined
+gaps. An overview of the required steps in the scaffolding process is outlined
 below.
@@
 
-The sequencing process generates reads from either strand of the DNA helix, the
+The sequencing process generates reads from either strand of the DNA helix the
 resulting contigs constructed from these reads may point in either direction of
 the DNA. Orientating contigs to point in the same direction and represent the
-same strand requires producing the reverse complement of sequences where
-necessary. In the case of microbial genomes this orientation may be to the
-5'-3' direction leading away from the origin of replication.
+same strand requires reverse complementing sequences where necessary. In the
+case of microbial genomes this orientation may be to the 5' $\rightarrow$ 3'
+direction leading away from the origin of replication.
 
@@
 Contig ordering determines the placement of observed contigs to best represent
-their order in the true genome sequence. The Correct placement of each contig
+their order in the true genome sequence. The correct placement of each contig
 also highlights any extra-genomic DNA, such as plasmids which are scaffolded
-separately from the genomic sequence. The contig order is commonly started at
-the contig containing the origin of replication. All subsequent contigs are
-then ordered in the 5' - 3' direction of DNA replication.
+separately from the genomic sequence. The order is commonly started at the
+contig containing the origin of replication. All subsequent contigs are then
+ordered in the 5' $\rightarrow$ 3' direction of DNA replication.
 
@@
 Given the correct order and orientation, determining the distance between
-contigs results in an estimate of the complete genome size. The size of
+contigs results in an estimate of the complete genome size. The size of any
 inter-contig gaps represents the length of an unknown region in the genome.
@@
 continuous sequence, which is useful for representing both the sequenced and
-still to be resolved area in the genome sequence.
+still to be resolved areas in the genome sequence.
 
@@
 computational methods of estimating the unknown sequence. This additional
-sequence is used to replace the gap between two contigs joining them into
+sequence is then used to replace the gap between two contigs joining them into
 a single sequence if they are not already scaffolded together. Once all contigs
@@
 Areas of corresponding sequence between the two sets can be used to infer
-contig placement and to build the scaffold
+contig placement and build the contigs into a scaffold
 \cite{richter2007,zhao2008,assefa2009}. Recombination between the two genomes
-can, however, reduce the effectiveness of this analysis. Repeat regions may be
-responsible for many of the gaps in building a genome sequence. The existence
-of tandemly repeated nucleotide regions produces many similar reads in the
-sequence process. The assembly of reads generated from repeat regions can
-result in the assemble collapsing into a single repeat instance or ignored by
-conservative assembly algorithms. These regions may be resolved by using
-reassembly algorithms specifically for repeat resolution
+can, however, reduce the effectiveness of this analysis. Repeat regions may
+also be responsible for many of the gaps in building a genome sequence. The
+existence of tandemly repeated nucleotide regions produces many similar reads
+fromgenome sequencing. The assembly of reads generated from repeat regions can
+result in the assembly collapsing into a single repeat instance or ignored by
+more conservative assembly algorithms. These regions may however be resolved by
+using reassembly algorithms specifically for repeat resolution
 \cite{mulyukov2002,koren2010}. Similarly a related approach may use unassembled
 sequence reads matching the regions around a scaffold gap to construct
-a uniquely overlapping set of reads across the region \cite{tsai2010}. \pb
+a uniquely overlapping set of reads across the gap \cite{tsai2010}. \pb
 
@@
 These methods provide a wide array of approaches for merging contigs into
-larger a continuous scaffolds. The scaffolding process may still however
+larger continuous scaffold sequences. The scaffolding process may still however
 require inserting additional sequence or joining contigs using PCR-derived
 information from the laboratory. This manual stage in the finishing process
-will moving and editing large blocks of nucleotide text by hand which is subject
-to human error-prone and precludes any reproducibility for the steps taken. \pb
+comprises moving and editing large blocks of nucleotide text by hand which is
+subject to human error and precludes any reproducibility of the steps taken.
+\pb
 
-The software outlined here, ``Scaffolder'', aims to solve address these
-problems by creating a file syntax and software framework for editing a genome
-scaffold. Scaffolder uses a specific file format to define how contigs are
-joined, additional sequences inserted, and for the specification of unknown
-regions. This syntax allows a scaffold to be updated by simply editing the
-scaffold file facilitating a reproducible scaffolding process and providing
+The software outlined here, ``Scaffolder'', aims to address these problems by
+creating a file syntax and software framework for editing a genome scaffold.
+Scaffolder uses a specific file format to define how contigs are joined,
+additional sequences inserted, and for the specification of unknown regions.
+This syntax allows a scaffold to be updated by simply editing the scaffold
+file. Scaffolder facilites a reproducible scaffolding process and provides
 a concise overview of how the scaffold was constructed. \pb

### 2011-03-02 src/article.tex
Proof read results section

@@
 The scaffold file takes the form of a list of entries. Each entry corresponds
-to a region of sequence inserted from the corresponding FASTA file. The
-Scaffolder software processes this list in order, converting each to the
-corresponding nucleotide sequences, which are then concatenated to produce
+to a region of sequence used in the final scaffold sequence. The Scaffolder
+software processes this list in order, converting each entry to the
+corresponding nucleotide sequence, all entries are then concatenated to produce
 a draft super sequence. Each entry in the scaffold file may have attributes

### 2011-03-02 src/article.tex
Proof read results

@@
 scaffold files written in the YAML syntax specify how these sequences should be
-joined and the scaffold file as read by the scaffolder software is used to
-generate the scaffold sequence. In addition to specifying which contigs are
-required, the scaffold file allows the contigs to be edited to produce smaller
-sub-sequences or reverse complemented if necessary. The process of genome
-finishing may involve producing additional oligonucleotide sequences to fill
-unknown regions in the scaffold. The scaffold file format allows these
-additional insert sequences to bridge these gaps. These inserts can also be
-treated in a similar way to larger contig sequences: trimmed and/or reverse
-complemented to match the corresponding gap region. \pb
+joined and the scaffolder software is used to generate the scaffold sequence.
+In addition to specifying which contigs are required, the scaffold file allows
+the contigs to be edited into smaller sub-sequences or reverse complemented if
+necessary. The process of genome finishing may involve producing additional
+oligonucleotide sequences to fill unknown regions in the scaffold. The scaffold
+file format allows these additional insert sequences to bridge gaps. These
+inserts can also be treated in a similar way to larger contig sequences:
+trimmed and/or reverse complemented to match the corresponding gap region. \pb
 
-The distances between contigs may be estimated from paired read data or from
+The distances between contigs may be estimated from paired-read data or from
 mapping the contigs to a reference genome. These inter-contig gap regions are
 useful to join separate sequences together by an estimated size and the
-scaffold file allows for the specification of these unknown regions. The use of
-these regions in the scaffold which areas in the scaffold and their approximate
-size are still to be resolved. \pb
+scaffold file allows for the specification of such unresolved regions. The use
+of these regions in the scaffold indicates the unresolved regions in the build
+and their approximate size. \pb
 
@@
 the sequences themselves are not stored in the scaffold file and are instead
-maintained as a separate FASTA file. Series of nucleotide sequences in
-a FASTA-formatted file is standard output for a genome assembler. These may be
-contigs, sets of contigs scaffolded into larger sequences, or both. These
-nucleotide sequences stored in the corresponding FASTA file are referenced in
-the scaffold using the first word form the FASTA header of the corresponding
-nucleotide sequence. Maintaining the nucleotide sequences in a separate file
-preserves the original sequence and decouples the original sequence data from
-the scaffold file describing how it should be used to in the genome sequence.
-\pb
+maintained as a separate FASTA file. Nucleotide sequences in a FASTA-formatted
+file are the standard output for a genome assembler, and these may be contigs,
+sets of contigs scaffolded into larger sequences, or both. These nucleotide
+sequences stored in the FASTA file are referenced in the scaffold using the
+first word from in the FASTA header of the corresponding sequence. Maintaining
+the nucleotide sequences in a separate file preserves the unedited sequence and
+decouples the data from the scaffold file describing how it should be used to
+in the genome sequence. \pb
 
@@
 
-The scaffold file is written using the YAML syntax. An example scaffold file is
-shown in Figure 1. This file illustrates the text attributes used to describe
-a scaffold and the joining of sequences in the genome build. The basic layout
+The scaffold file is written using the YAML syntax and an example is shown in
+Figure 1. This file illustrates the text attributes used to describe a scaffold
+and how the sequences are joined together in the genome build. The basic layout
 of the scaffold file is a list of entries, where each entry corresponds to
@@
 start of a YAML-formatted document. The first entry (highlighted in green)
-begins with a dash character `-' to denote an entry in the YAML list. The first
-line of each entry in the scaffold therefore begins with a dash line. The next
-line is indented by two spaces as whitespace is required to group similar
+begins with a dash character `-' to denote an entry in the YAML list. This is
+a requirement of the YAML format: that each entry begins with a dash line. The
+next line is indented by two spaces as whitespace is required to group similar
 attributes together. The ``sequence'' tag indicates that this entry corresponds
 to a sequence in the FASTA file and the following line indicates the name of
-this FASTA sequence using the ``source'' tag. As described the first word of
-the FASTA header is used to identify which sequence is selected from the file.
+this sequence using the ``source'' tag. As described the first word of the
+FASTA header is used to identify which sequence is selected from the file.
 Together these three lines describe the first entry in the scaffold as a simple
-sequence using the FASTA sequence named `sequence1'. On the right hand side of
-Figure 1 this produces the first region in the scaffold, also shown in green.
-\pb
+sequence using the sequence named `sequence1'. On the right hand side of Figure
+1 this produces the first region in the scaffold, also shown in green. \pb
 
@@
 ``unresolved'' tag, indicating a region of unknown sequence but known length.
-The second line specifies the size of this unknown region. This entry therefore
-inserts a region of 20 `N' characters in the scaffold. \pb
+The second line specifies the size of this unknown region. This entry produces
+a region of 20 `N' characters in the scaffold. \pb
 
@@
 
-The last entry in the scaffold, highlighted in blue, adds the FASTA entry
+The last entry in the scaffold, highlighted in blue, adds the sequence named
 `sequence2' to the scaffold. This entry demonstrates how this sequence may be
@@
 additional regions of sequence. These inserts are also added as a YAML list,
-with each insert beginning with a dash line. The first insert, shown in purple,
+with each insert beginning with a dash. The first insert, shown in purple,
 similar attributes to that of a sequence entry; the reverse, start and stop
 tags are used to trim and reverse complement the insert. Similarly the `source'
-tag identifies the corresponding FASTA sequence as `insert'. The ``open'' and
+tag identifies the corresponding FASTA sequence as `insert1'. The ``open'' and
 ``close'' tags are specific to inserts and determine where the insert is added
 in the enclosing sequence. The region of the sequence inside these coordinates
-is inclusively replaced by the specified insert FASTA sequence. This is
-visualised in the putative gene Figure 1 by the black lines bisecting the blue
-sequence. \pb
+is inclusively replaced by the specified insert sequence. This is visualised in
+the putative gene Figure 1 by the black lines bisecting the blue sequence. \pb
 
@@
 illustrates that only one of either `open' or `close' tags is required when
-adding an insert sequence. If only only one of the `open' or `close' tags is
-used the corresponding opposing position distance is calculated from the length
-of the insert FASTA sequence. This allows inserts to bridge into, and partially
+adding an insert sequence. If only one of the `open' or `close' tags is used
+the corresponding opposing position distance is calculated from the length of
+the insert FASTA sequence. This allows inserts to bridge into, and partially
 fill gap regions without requiring an end coordinate position. \pb
 
-\subsection{Scaffolder Software Interface} %{{{2
+\subsection*{Scaffolder Software Interface} %{{{2
 
@@
 existing genomics workflows or used with Ruby build tools such as Rake. In
-addition Scaffolder provides a command line interface (CLI) providing functions
-to validate the scaffold file and build a draft super sequence. The Scaffolder
-CLI behaves as a standard Unix tool and returns appropriate exit codes and
-manual pages. The use of both these Scaffolder interfaces is outlined in detail
-on the scaffolder website (\scaffolder). \pb
+addition Scaffolder provides a command line interface (CLI) to validate the
+scaffold file and build a draft super sequence. The Scaffolder CLI behaves as
+a standard Unix tool and returns appropriate exit codes and manual pages. The
+use of both these Scaffolder interfaces is outlined in detail on the scaffolder
+website (\scaffolder). \pb
 
@@
 
-Scaffolder was designed to be as simple to use as possible and assuming the
-Ruby and RubyGems software are present, Scaffolder can be installed in a single
-command line step. This should negate any possible barriers to entry which
-require manual compilation of source code. Scaffolder also uses a minimal and
-compact syntax to describe the genome scaffold. This syntax is simple to
-construct and edit whilst being succinct and readable. The goal of this syntax
-is to make genome scaffolds easy to construct in a common text editor. \pb
+Scaffolder was designed to be as simple to use and, assuming the Ruby and
+RubyGems software are present, Scaffolder can be installed in a single command
+line step. This should negate any possible barriers to entry which require
+manual compilation of source code. Scaffolder also uses a minimal and compact
+syntax to describe the genome scaffold. This syntax is simple to construct and
+edit whilst being succinct and readable. The goal of this syntax is to make
+genome scaffolds easy to make in a common text editor. The use of word
+processors for this task is however discouraged. \pb
 
@@
 be reliably reproduced. The scaffold file furthermore provides a human readable
-description of how the scaffold is constructed. \pb
-
-Configuring the final sequence in the scaffold file means the build is easier
-to edit once already constructed. The YAML text format also allows comparison
-of differences between scaffold builds using standard Unix tools such as diff.
-This makes scaffold files easier to store in version control systems and allows
-genome finishers methods similar to those in software development. \pb
+description of how the scaffold is constructed. Configuring the final sequence
+in the scaffold file means the build is easier to edit once already
+constructed. The YAML text format allows comparison of differences between
+scaffold builds using standard Unix tools such as diff. This makes scaffold
+files easier to store in version control systems and allows genome finishers to
+use methods similar to those in software development. \pb
 

### 2011-03-02 src/article.tex
Revise conclusions

@@
 Scaffolder is software aimed at bioinformaticians and biologists familiar with
-the command line building genome sequence scaffold. The file format maintains
-the scaffold in a concise and readable way allowing third parties to see how
-the genome sequence was assembled. This file format allows a broad overview of
-of which sequences were included and how they are ordered possible not possible
-from a megabase-sized strings of nucleotide characters. Scaffolder also
-furthers the cause for reproducibility in genome projects by allowing the
-scaffold super-sequence to be reliably reproduced from the scaffold file. The
-YAML syntax for writing the scaffold file is also standardised and to
-manipulate programmatically. This thereby means the scaffolding process follows
-the Unix tenet of: ``If your data structures are good enough, the algorithm to
-manipulate them should be trivial.''.
+the command line who wish to build a genome sequence scaffold. The file format
+maintains the scaffold in a concise and readable way allowing third parties to
+see how the genome sequence was assembled. This file format allows a broad
+overview of of which sequences were included and how they are ordered,
+something not possible from a megabase-length strings of nucleotide characters.
+Scaffolder furthers the case for reproducibility in genome projects by allowing
+the scaffold super-sequence to be reliably reproduced from the same scaffold
+file. The YAML syntax for writing the scaffold file is also standardised and
+simple to manipulate programmatically. This thereby means the scaffolding
+process follows the Unix tenet of: ``If your data structures are good enough,
+the algorithm to manipulate them should be trivial.''.
 

### 2011-03-02 src/article.tex
Add version to requirements
Revise conclusions

@@
   \begin{description}
-    \item[Project name:] Scaffolder, version 0.4.1
+    \item[Project name:] Scaffolder v0.4.1, Scaffolder Tools v0.1.2
     \item[Project home page:] \scaffolder
@@
 An example scaffold file written using the YAML syntax \cite{yaml} (left) and
-the resulting putative scaffold (right). The scaffold contains three entries
-and two inserts. Each entry in the scaffold file text is delimited by a `-' on
-a new line and highlighted using separate colours. The scaffold diagram on the
-right is not to scale but illustrates how the scaffold sequences are joined.
-\pb
+the resulting putative scaffold sequence (right). The scaffold contains three
+entries and two inserts. Each entry in the scaffold file text is delimited by
+a `-' on a new line and highlighted using separate colours. The scaffold
+diagram on the right is not to scale and instead illustrates how the scaffold
+sequences are joined. \pb
 

### 2011-03-02 src/article.tex
Spell check

@@
 existence of tandemly repeated nucleotide regions produces many similar reads
-fromgenome sequencing. The assembly of reads generated from repeat regions can
+from genome sequencing. The assembly of reads generated from repeat regions can
 result in the assembly collapsing into a single repeat instance or ignored by
@@
 This syntax allows a scaffold to be updated by simply editing the scaffold
-file. Scaffolder facilites a reproducible scaffolding process and provides
+file. Scaffolder facilitates a reproducible scaffolding process and provides
 a concise overview of how the scaffold was constructed. \pb

### 2011-03-07 src/article.tex
Revise background

--- 2011-03-02 src/article.tex
@@
 High-throughput sequencing can produce hundreds of thousands to millions of
-sequence reads from a genome. Each sequence represents the machine-read
-nucleotides from a short region in the genome. At the time of writing,
-high-throughput sequencing is limited to producing reads less than 1,000
-nucleotides in length. Therefore to resolve a sequence longer than this, such
-as a complete genome, these numerous smaller fragments must be pieced together.
-The process of joining reads into longer sequences is the `assembly' stage of
-a genome project \cite{miller2010}. \pb
+sequence reads from a genome. At the time of writing, high-throughput
+sequencing is limited to producing reads less than 1,000 nucleotides in length.
+Therefore to resolve a sequence longer than this, such as a complete genome,
+these numerous smaller fragments must be pieced together. The process of
+joining reads into longer sequences is the `assembly' stage of a genome project
+\cite{miller2010}. \pb
 
-Assembly software takes the nucleotide reads produced by the sequencing
-hardware and, in the ideal case, outputs a complete genome sequence composed of
+Assembly software takes the nucleotide reads produced by sequencing hardware
+and, in the ideal case, outputs a single complete genome sequence composed of
 these individual fragments. An analogy for this process is a jigsaw puzzle:
@@
 
-The recent advent of these current high throughput sequencing methods has led
-to a renewed interest into algorithms for solving the problem of genome
-assembly \cite{pop2008,pop2009}; genome assembly software may however be unable
-to produce a complete sequence from these reads. In many cases the algorithm
-will generate several large assembled regions of sequence (`contigs') composed
-of the many available individual reads, which represent the assembly of reads
-into the longest possible sequences given the data. Nonetheless, these contigs
-represent a fragmented picture of the genome and require additional work to
-join together into larger lengths. \pb
+The advent of high throughput sequencing methods has led to a renewed interest
+in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
+Nonetheless, due to the complexity of merging large numbers of reads genome
+assembly software may be unable to produce a complete sequence from these
+reads. Instead the algorithm may generate several large assembled regions of
+sequence (`contigs') composed of the many available individual reads that
+represent the assembly of reads into the longest possible sequences, given the
+data. Nonetheless, these contigs represent a fragmented picture of the genome
+and require additional work to join together into a complete sequence. \pb
 
 The process of ameliorating a draft sequence into a finished continuous
-sequence can be expensive in both time and laboratory effort, although the
-genomic data present in a set of contigs may often be sufficient for many
+sequence can be expensive in terms of time and laboratory effort, although the
+genomic data present in a set of generated contigs may be sufficient for many
 research questions \cite{branscomb2002}. Nevertheless, a continuous
@@
 estimates of distances between genomic elements \cite{parkhill2002,fraser2002}.
-This process of joining contigs together to form a continuous sequence is
-called the `scaffolding' or `finishing' stage. \pb
+The process of joining these contigs together to form a continuous genome
+sequence is called the `scaffolding' or `finishing' stage. \pb
 
@@
 data, scaffolding may not ultimately produce a final completed sequence, but
-may still succeed in joining a greater subset of contigs together or resolving
-gaps. An overview of the required steps in the scaffolding process is outlined
-below.
+may still succeed in a greater subset of contigs together or resolving gaps. An
+overview of the required steps in the scaffolding process is outlined below:
 
@@
 
-The sequencing process generates reads from either strand of the DNA helix the
-resulting contigs constructed from these reads may point in either direction of
-the DNA. Orientating contigs to point in the same direction and represent the
-same strand requires reverse complementing sequences where necessary. In the
-case of microbial genomes this orientation may be to the 5' $\rightarrow$ 3'
-direction leading away from the origin of replication.
+The sequencing process generates reads from either strand of the DNA helix and
+the resulting contigs constructed from these reads may represent either the
+5'~$\rightarrow$~3' or 3'~$\rightarrow$~5' orientation with respect to the
+origin of replication. Orientating all contigs to point in the same direction
+requires reverse complementing sequences where necessary. In the case of
+microbial genomes this orientation will be to the 5'~$\rightarrow$~3' direction
+leading away from the origin of replication.
 
@@
 continuous sequence, which is useful for representing both the sequenced and
-still to be resolved areas in the genome sequence.
+to-be-resolved areas in the genome sequence.
 
@@
 During the scaffolding process, closing and filling gaps between contigs
-improves the completeness of the genome scaffold. Closing gaps may require
-returning to the laboratory to perform additional sequencing or using
-computational methods of estimating the unknown sequence. This additional
-sequence is then used to replace the gap between two contigs joining them into
-a single sequence if they are not already scaffolded together. Once all contigs
-have been joined and gaps in a scaffold closed, the genome may be considered
-finished.
+completes and improves the genome scaffold. Closing gaps may require returning
+to the laboratory to perform additional sequencing or using computational
+methods of estimating the unknown sequence. This additional sequence is then
+used to replace the gap between two contigs, joining them into a single
+sequence. Once all contigs have been joined and gaps in a scaffold closed, the
+genome may be considered finished.
 
@@
 
-The process of resolving a genome sequence and building a scaffold uses
-\emph{in vitro} methods, \emph{in silico} methods, or a combination of both. An
-example of a computational method might use paired-read data, where the
-sequencing reads are produced in pairs at a known distance apart in the genome.
-The occurrence of a pair of reads in separate contigs can be used to estimate
-the order and distance between the contigs. Alternatively, laboratory methods
-may use PCR to amplify the unknown DNA in a gap region, then use traditional
-Sanger sequencing to determine the gap sequence. Computational methods are more
-preferable, as they are less costly in laboratory time and materials compared
-to manual gap resolution and uses the available data that exists following
-sequencing \cite{nagarajan2010}. \pb
+The process of resolving a genome sequence and building a scaffold uses wet
+laboratory methods, \emph{in silico} methods, or a combination of both. An
+example of a computational method might use the existing paired-read data from
+the sequencing stage where reads were produced in pairs at a known distance
+apart in the genome. The occurrence of paired reads in separate contigs can be
+used to probabalistically estimate the order and distance between these
+contigs. Alternatively, laboratory methods may use PCR to amplify the unknown
+DNA in a gap region, then use traditional Sanger sequencing to determine the
+sequence of this gap. Computational methods are more preferable as they are
+less costly in laboratory time and materials compared to manual gap resolution,
+and use the available data that exists following sequencing
+\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
+\emph{in silico} software packages exist to suggest the likely primers
+necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb
 
-Synteny methods compare the assembled contigs to a more complete reference
-genome sequence searching for areas of sequence similarity between the two.
-Areas of corresponding sequence between the two sets can be used to infer
-contig placement and build the contigs into a scaffold
-\cite{richter2007,zhao2008,assefa2009}. Recombination between the two genomes
-can, however, reduce the effectiveness of this analysis. Repeat regions may
-also be responsible for many of the gaps in building a genome sequence. The
-existence of tandemly repeated nucleotide regions produces many similar reads
-fromgenome sequencing. The assembly of reads generated from repeat regions can
-result in the assembly collapsing into a single repeat instance or ignored by
-more conservative assembly algorithms. These regions may however be resolved by
-using reassembly algorithms specifically for repeat resolution
-\cite{mulyukov2002,koren2010}. Similarly a related approach may use unassembled
-sequence reads matching the regions around a scaffold gap to construct
-a uniquely overlapping set of reads across the gap \cite{tsai2010}. \pb
+Synteny methods for scaffolding compare the assembled contigs to a more
+complete reference genome sequence by searching for areas of sequence
+similarity between the two. Any areas of corresponding sequence with the
+reference genome can be used to infer contig placement and build the contigs
+into a scaffold \cite{richter2007,zhao2008,assefa2009}. Recombination within
+each of the two compared genomes can, however, reduce the effectiveness of this
+analysis. \pb
 
-Paired-read data provides an extra level of information about how contigs may
-be scaffolded together. Heuristic algorithms may take advantage of this data to
-search for optimal configuration of contigs in the scaffold using paired-read
-distances \cite{dayarian2010,boetzer2011}. Additional data using a reference
-genome may also be combined when estimating the best contig configuration
-\cite{pop2004}. Finally when the scaffold cannot be completely resolved,
-\emph{in silico} software packages exist to suggest the likely primers
-necessary for PCR amplifying the sequence in gap regions
-\cite{gordon2001,nagarajan2010}. \pb
+Repeat regions may also be responsible for multiple gaps when building a genome
+sequence; tandemly repeated nucleotide regions in the genome produces multiple
+reads with similar sequence. As many assembly algorithms rely on sequence
+overlaps between reads to build a contig the similarity between repeat-region
+reads can result in the assembly collapsing into an artifically short sequence
+at this region or being ignored by more conservative assembly algorithms. Such
+regions can be resolved by using algorithms to specifically reassemble the
+collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
+a related approach uses unassembled sequence reads matching the regions around
+a scaffold gap to construct a uniquely overlapping set of reads across it
+\cite{tsai2010}. \pb
 
-These methods provide a wide array of approaches for merging contigs into
-larger continuous scaffold sequences. The scaffolding process may still however
-require inserting additional sequence or joining contigs using PCR-derived
-information from the laboratory. This manual stage in the finishing process
-comprises moving and editing large blocks of nucleotide text by hand which is
-subject to human error and precludes any reproducibility of the steps taken.
-\pb
+Paired-read data can provide an extra level of information about how contigs
+may be scaffolded together. Heuristic scaffolding algorithms take advantage of
+this data to search for optimal configuration of contigs in the scaffold using
+paired-read distances \cite{dayarian2010,boetzer2011}. Synteny data from
+a reference genome can also be combined with this paired-read data to estimate
+the best contig configuration \cite{pop2004}. \pb
 
-The software outlined here, ``Scaffolder'', aims to address these problems by
-creating a file syntax and software framework for editing a genome scaffold.
-Scaffolder uses a specific file format to define how contigs are joined,
-additional sequences inserted, and for the specification of unknown regions.
-This syntax allows a scaffold to be updated by simply editing the scaffold
-file. Scaffolder facilites a reproducible scaffolding process and provides
-a concise overview of how the scaffold was constructed. \pb
+These described methods provide a wide array of approaches for merging contigs
+into a larger, continuous scaffold sequence. Nevertheless the scaffolding
+process may still require inserting additional sequences or further joining
+contigs using PCR-derived sequence. This manual stage in the finishing
+processmay require moving and editing large blocks of nucleotide text by hand
+which is subject to human error and precludes any reproducibility of the steps
+in this process. \pb
+
+The software outlined here, ``Scaffolder'', aims to address these problems of
+reproducibility by creating a file syntax and software framework for editing
+a genome scaffold. Scaffolder uses a specific file format to define how contigs
+are joined, additional sequences are inserted, and for the specification of
+unknown regions. This syntax allows a scaffold to be updated by simply editing
+the scaffold file. As such scaffolder facilitates a reproducible scaffolding
+process and provides a concise overview of how the final genomic scaffold was
+constructed. \pb
 

### 2011-03-07 src/article.tex
Revise results section

@@
 necessary. The process of genome finishing may involve producing additional
-oligonucleotide sequences to fill unknown regions in the scaffold. The scaffold
-file format allows these additional insert sequences to bridge gaps. These
-inserts can also be treated in a similar way to larger contig sequences:
-trimmed and/or reverse complemented to match the corresponding gap region. \pb
+oligonucleotide sequences to fill unknown regions in the scaffold and The
+scaffold file format allows these additional insert sequences to bridge gaps.
+These inserts can also be treated in the same manner as larger contig
+sequences: trimmed and/or reverse complemented to match the corresponding gap
+region. \pb
 
@@
 useful to join separate sequences together by an estimated size and the
-scaffold file allows for the specification of such unresolved regions. The use
-of these regions in the scaffold indicates the unresolved regions in the build
-and their approximate size. \pb
+scaffold file allows for the specification of such unresolved regions by
+inserting regions of `N' nucleotides into the scaffold. The use of these
+regions in the scaffold indicates the unresolved regions in the build and their
+approximate size. \pb
 
@@
 the sequences themselves are not stored in the scaffold file and are instead
-maintained as a separate FASTA file. Nucleotide sequences in a FASTA-formatted
+maintained as a separate FASTA file. Nucleotide sequences in a FASTA-format
 file are the standard output for a genome assembler, and these may be contigs,
@@
 is inclusively replaced by the specified insert sequence. This is visualised in
-the putative gene Figure 1 by the black lines bisecting the blue sequence. \pb
+the putative genome in Figure 1 by the black lines bisecting the blue sequence.
+\pb
 
@@
 adding an insert sequence. If only one of the `open' or `close' tags is used
-the corresponding opposing position distance is calculated from the length of
-the insert FASTA sequence. This allows inserts to bridge into, and partially
-fill gap regions without requiring an end coordinate position. \pb
+the corresponding opposing `open'/`close' coordinate is calculated from the
+length of the insert FASTA sequence. This allows inserts to bridge into, and
+partially fill gap regions, without requiring an end coordinate position. \pb
 

### 2011-03-07 src/article.tex
Revise discussion section

@@
 of manually editing a scaffold is inherently hard to reproduce and introduces
-the possibility of human error. The aims of this Scaffolder software are
-therefore twofold. The first is providing software that is easy to install and
-simplifies the task of genome finishing. The second aim is to facilitate
+the possibility of irreporducible techniques and/or human error. The aims of
+the Scaffolder software are twofold: 1) to provide software that is easy to
+install and simplifies the task of genome finishing; and 2) to facilitate
 reproducibility when joining large nucleotide sequences together to create
@@
 edit whilst being succinct and readable. The goal of this syntax is to make
-genome scaffolds easy to make in a common text editor. The use of word
-processors for this task is however discouraged. \pb
+genome scaffolds easy to make in a common text editor; however the use of word
+processors for this task is discouraged. \pb
 

### 2011-03-07 src/article.tex
Revise discussion

@@
 Scaffolder is software aimed at bioinformaticians and biologists familiar with
-the command line who wish to build a genome sequence scaffold. The file format
-maintains the scaffold in a concise and readable way allowing third parties to
-see how the genome sequence was assembled. This file format allows a broad
-overview of of which sequences were included and how they are ordered,
-something not possible from a megabase-length strings of nucleotide characters.
-Scaffolder furthers the case for reproducibility in genome projects by allowing
-the scaffold super-sequence to be reliably reproduced from the same scaffold
-file. The YAML syntax for writing the scaffold file is also standardised and
-simple to manipulate programmatically. This thereby means the scaffolding
-process follows the Unix tenet of: ``If your data structures are good enough,
-the algorithm to manipulate them should be trivial.''.
+the command line who wish to build a genome sequence scaffold from a set of
+assembled contigs. The Scaffolder file format maintains the genome scaffold as
+a concise and readable text representation that allows third parties to see how
+the genome sequence was scaffolded. This file format also allows a broad
+overview of which sequences were included and how they are ordered in the
+genome scaffold, something not possible to deduce from a megabase-length
+strings of nucleotide characters. Scaffolder furthers increases the ease of
+reproducibility in genome projects by allowing the scaffold super-sequence to
+be reliably reproduced from the same scaffold file. The YAML syntax for writing
+the scaffold file is also standardised and simple to manipulate
+programmatically. This thereby means the scaffolding process follows the Unix
+tenet of: ``If your data structures are good enough, the algorithm to
+manipulate them should be trivial.''.
 

### 2011-03-07 src/article.tex
Spell check

@@
 apart in the genome. The occurrence of paired reads in separate contigs can be
-used to probabalistically estimate the order and distance between these
+used to probabilistically estimate the order and distance between these
 contigs. Alternatively, laboratory methods may use PCR to amplify the unknown
@@
 overlaps between reads to build a contig the similarity between repeat-region
-reads can result in the assembly collapsing into an artifically short sequence
+reads can result in the assembly collapsing into an artificially short sequence
 at this region or being ignored by more conservative assembly algorithms. Such
@@
 contigs using PCR-derived sequence. This manual stage in the finishing
-processmay require moving and editing large blocks of nucleotide text by hand
+process may require moving and editing large blocks of nucleotide text by hand
 which is subject to human error and precludes any reproducibility of the steps
@@
 of manually editing a scaffold is inherently hard to reproduce and introduces
-the possibility of irreporducible techniques and/or human error. The aims of
+the possibility of irreducible techniques and/or human error. The aims of
 the Scaffolder software are twofold: 1) to provide software that is easy to

### 2011-03-07 src/article.tex
Add section on scaffolder input data and algorithm

@@
 
+\subsection*{Input sequence data and scaffolding process} %{{{2
+
+The input data for Scaffolder is nucleotide sequence in FASTA format file.
+These nucleotide sequences can be of any length and for example may be
+individual sequences, assembled contigs or contigs which have been joined to
+into larger scaffolds. The case in which Scaffolder may be most useful is using
+the contigs and scaffolded contigs produced from assembling a set of sequencing
+reads. \pb
+
+The scaffolder algorithm reads through the scaffolder file parsing each entry
+in the file. Each entry is converted to corresponding sequence by fetching from
+the FASTA file and then performing any designated sequence processing
+operations. Each of the sequence entries are then joined as a s single super
+sequence and returned to the console output. \pb
+
 \clearpage

### 2011-03-08 src/article.tex
Revise abstract

@@
   step in a genome project is to join neighbouring sequence regions together
-  and fill gaps in the assembly using additional sequence. This scaffolding
+  and fill gaps in the assembly using additional sequences. This scaffolding
   step, however, is non-trivial and requires manually editing large blocks of
-  nucleotide sequence. Joining sequence contigs together also hides the source
+  nucleotide sequence. Joining these sequences together also hides the source
   of each region in the final genome sequence. Taken together, these
@@
   \paragraph*{Results:} This software allows a genome build to be defined in
-  terms of the constituent sequences using a relatively simple format to write
-  and edit scaffold file syntax. This syntax further allows  unknown regions to
-  be defined, and inserts to be added to the scaffold to fill gaps. Defining
-  the genome construction in a file makes the scaffolding process reproducible
-  and easier to edit and compare than FASTA nucleotide sequence.
+  terms of the constituent sequences using a relatively simple syntax to
+  define the scaffold. This syntax further allows  unknown regions to be
+  defined, and add inserts to fill gaps in the scaffold. Defining the genome
+  construction in a file makes the scaffolding process reproducible and easier
+  to edit compared with FASTA nucleotide sequence.
 

### 2011-03-08 src/article.tex
Revise background section

@@
 in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
-Nonetheless, due to the complexity of merging large numbers of reads genome
-assembly software may be unable to produce a complete sequence from these
-reads. Instead the algorithm may generate several large assembled regions of
-sequence (`contigs') composed of the many available individual reads that
-represent the assembly of reads into the longest possible sequences, given the
-data. Nonetheless, these contigs represent a fragmented picture of the genome
-and require additional work to join together into a complete sequence. \pb
+Nonetheless, due to the complexity of merging large numbers of overlapping
+reads genome assembly software may be unable to produce a complete sequence.
+Instead the algorithm may generate several large assembled regions of sequence
+(`contigs') composed of the many individual reads that represent assembly into
+the longest possible sequences, given the data. Nonetheless, these contigs
+represent a fragmented picture of the genome and require additional work to
+join together into a complete sequence. \pb
 
@@
 The process of joining these contigs together to form a continuous genome
-sequence is called the `scaffolding' or `finishing' stage. \pb
+sequence is called the `scaffolding' or `finishing' stage and is the focus of
+the software described in this article. \pb
 
@@
 data, scaffolding may not ultimately produce a final completed sequence, but
-may still succeed in a greater subset of contigs together or resolving gaps. An
-overview of the required steps in the scaffolding process is outlined below:
+may still succeed in joining a subset of contigs together or resolving gaps
+between contigs. An overview of the required steps in the scaffolding process
+is outlined below:
 
@@
 the resulting contigs constructed from these reads may represent either the
-5'~$\rightarrow$~3' or 3'~$\rightarrow$~5' orientation with respect to the
+5'~$\rightarrow$~3' or 3'~$\rightarrow$~5' DNA strands with respect to the
 origin of replication. Orientating all contigs to point in the same direction
 requires reverse complementing sequences where necessary. In the case of
-microbial genomes this orientation will be to the 5'~$\rightarrow$~3' direction
-leading away from the origin of replication.
+microbial genomes this orientation will be to the 5'~$\rightarrow$~3'
+direction following the direction of genome replication.
 
@@
 Filling these regions with unknown nucleotide characters (`N') allows a draft
-continuous sequence, which is useful for representing both the sequenced and
+continuous sequence, which is useful for representing both the known and
 to-be-resolved areas in the genome sequence.
@@
 
-Synteny methods for scaffolding compare the assembled contigs to a more
-complete reference genome sequence by searching for areas of sequence
-similarity between the two. Any areas of corresponding sequence with the
-reference genome can be used to infer contig placement and build the contigs
-into a scaffold \cite{richter2007,zhao2008,assefa2009}. Recombination within
-each of the two compared genomes can, however, reduce the effectiveness of this
-analysis. \pb
+Examples of \emph{in silico} methods include using synteny to compare the
+assembled contigs to a complete reference genome sequence by searching for
+areas of sequence similarity between the two. Any areas of corresponding
+sequence with the reference genome can be used to infer contig placement and
+build the contigs into a scaffold \cite{richter2007,zhao2008,assefa2009}.
+Recombination within each of the two compared genomes can, however, reduce the
+effectiveness of this analysis. \pb
 
-Repeat regions may also be responsible for multiple gaps when building a genome
-sequence; tandemly repeated nucleotide regions in the genome produces multiple
-reads with similar sequence. As many assembly algorithms rely on sequence
-overlaps between reads to build a contig the similarity between repeat-region
-reads can result in the assembly collapsing into an artificially short sequence
-at this region or being ignored by more conservative assembly algorithms. Such
-regions can be resolved by using algorithms to specifically reassemble the
+Repeat regions may also be responsible for multiple gaps when building
+a genome sequence; tandemly repeated nucleotide regions in the genome produce
+multiple reads with similar sequence. As many assembly algorithms rely on
+sequence overlaps between reads to build a contig, the similarity between
+repeat-region reads can result in the assembly collapsing into an artificially
+short sequence or being ignored by more conservative assembly algorithms. Such
+regions can be resolved by using algorithms specifically reassemble the
 collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
@@
 may be scaffolded together. Heuristic scaffolding algorithms take advantage of
-this data to search for optimal configuration of contigs in the scaffold using
-paired-read distances \cite{dayarian2010,boetzer2011}. Synteny data from
-a reference genome can also be combined with this paired-read data to estimate
-the best contig configuration \cite{pop2004}. \pb
+this data to search for the optimal configuration of contigs in the scaffold
+that matches these paired-read distances \cite{dayarian2010,boetzer2011}.
+Synteny data from a reference genome can also be combined with this
+paired-read data to estimate the best contig configuration \cite{pop2004}. \pb
 
-These described methods provide a wide array of approaches for merging contigs
-into a larger, continuous scaffold sequence. Nevertheless the scaffolding
-process may still require inserting additional sequences or further joining
-contigs using PCR-derived sequence. This manual stage in the finishing
-process may require moving and editing large blocks of nucleotide text by hand
-which is subject to human error and precludes any reproducibility of the steps
-in this process. \pb
+These described \emph{in silico} methods provide a wide array of approaches
+for merging contigs into a larger, continuous scaffold sequence. Nevertheless
+the scaffolding process may still require manually inserting additional
+sequences or further joining contigs using PCR-derived sequence. Moving and
+editing large blocks of nucleotide text by hand however possibly introduces
+human error and precludes any reproducibility of the steps in this process. \pb
 

### 2011-03-08 src/article.tex
Revise implementation section

@@
 
-Scaffolder is written in the Ruby programming language, version 1.8.7
-\cite{ruby-lang}. The scaffolder package is split into two libraries. The first
-called ``scaffolder'' contains the core Scaffolder application programming
-interface (API). The second library ``scaffolder-tools'' contains the
-Scaffolder command line interface (CLI). \pb
+Scaffolder is written in the Ruby programming language using version 1.8.7
+\cite{ruby-lang}. The scaffolder package is split into two libraries. The
+first called ``scaffolder'' provides the core Scaffolder application
+programming interface (API). The second library ``scaffolder-tools'' provides
+the Scaffolder command line interface (CLI). \pb
 
@@
 should therefore be simple to update manually in addition to being
-computationally tractable. This requirement was also best suited to YAML which
-is human-readable and simple to edit in a standard text editor. \pb
+computationally tractable. This requirement was also best suited to YAML
+syntax which is human-readable and simple to edit in a standard text editor.
+\pb
 
 The scaffold file takes the form of a list of entries. Each entry corresponds
-to a region of sequence used in the final scaffold sequence. The Scaffolder
-software processes this list in order, converting each entry to the
-corresponding nucleotide sequence, all entries are then concatenated to produce
-a draft super sequence. Each entry in the scaffold file may have attributes
-that define whether a sub-sequence or the reverse complement of the sequence
-should be used. The types of attributes available, and an example scaffold file
-are outlined below. \pb
+to a region of sequence used in the final scaffold sequence. Each entry in the
+scaffold file may have attributes that define whether a sub-sequence or the
+reverse complement of the sequence should be used. The types of attributes
+available, and an example scaffold file are outlined in the Results section.
+\pb
 
@@
 individual sequences, assembled contigs or contigs which have been joined to
-into larger scaffolds. The case in which Scaffolder may be most useful is using
-the contigs and scaffolded contigs produced from assembling a set of sequencing
-reads. \pb
+into larger scaffolds. The case in which Scaffolder may be most useful is
+using the contigs and scaffolded contigs produced from the previously
+assembled set of sequencing reads, combined with additional gap filling
+sequences produced by PCR or \emph{in silico} methods as outlined in the
+Background. \pb
 
-The scaffolder algorithm reads through the scaffolder file parsing each entry
-in the file. Each entry is converted to corresponding sequence by fetching from
-the FASTA file and then performing any designated sequence processing
-operations. Each of the sequence entries are then joined as a s single super
-sequence and returned to the console output. \pb
+The scaffolder algorithm reads through the scaffolder file parsing. Each entry
+is then converted to corresponding nucleotide sequence by fetching from the
+FASTA file and then performing any designated sequence trimming or reverse
+complenting. Each of the sequence entries are then joined into a continuous
+single super-sequence and returned to the console output. \pb
 

### 2011-03-08 src/article.tex
Revise results and discussion

@@
 
-The Scaffolder software allows computational joining of nucleotide sequences
+The Scaffolder software facilitates reproducibly joining nucleotide sequences
 together into a single contiguous scaffolded super-sequence. Plain-text
-scaffold files written in the YAML syntax specify how these sequences should be
-joined and the scaffolder software is used to generate the scaffold sequence.
-In addition to specifying which contigs are required, the scaffold file allows
-the contigs to be edited into smaller sub-sequences or reverse complemented if
-necessary. The process of genome finishing may involve producing additional
-oligonucleotide sequences to fill unknown regions in the scaffold and The
-scaffold file format allows these additional insert sequences to bridge gaps.
-These inserts can also be treated in the same manner as larger contig
-sequences: trimmed and/or reverse complemented to match the corresponding gap
-region. \pb
+scaffold files written in the YAML syntax specify how these sequences should
+be joined and the scaffolder software is used to generate the scaffold
+sequence from these instructions. In addition to specifying which contigs are
+required, the scaffold file allows the contigs to be edited into smaller
+sub-sequences or reverse complemented if necessary. The process of genome
+finishing may involve producing additional oligonucleotide sequences to fill
+unknown regions in the scaffold and the Scaffolder allows these additional
+insert sequences to be used to fill these gaps. These inserts can also be
+treated in the same manner as larger contig sequences: trimmed and/or reverse
+complemented to match the corresponding gap region size and orientation. \pb
 
@@
 first word from in the FASTA header of the corresponding sequence. Maintaining
-the nucleotide sequences in a separate file preserves the unedited sequence and
-decouples the data from the scaffold file describing how it should be used to
-in the genome sequence. \pb
+the nucleotide sequences in a separate file preserves the unedited sequence
+and decouples the data from the specification of how it should be used to
+produce the genome sequence. \pb
 
@@
 The scaffold file is written using the YAML syntax and an example is shown in
-Figure 1. This file illustrates the text attributes used to describe a scaffold
-and how the sequences are joined together in the genome build. The basic layout
-of the scaffold file is a list of entries, where each entry corresponds to
-a region of sequence in the generated scaffold super-sequence. \pb
+Figure 1. This file illustrates the text attributes used to describe
+a scaffold and how the sequences are correspondingly joined together in the
+genome build. The basic layout of the scaffold file is a list of entries,
+where each entry corresponds to a region of sequence in the generated scaffold
+super-sequence. \pb
 
@@
 
-The first line of the scaffold file begins with three dashes to indicate the
-start of a YAML-formatted document. The first entry (highlighted in green)
-begins with a dash character `-' to denote an entry in the YAML list. This is
-a requirement of the YAML format: that each entry begins with a dash line. The
-next line is indented by two spaces as whitespace is required to group similar
-attributes together. The ``sequence'' tag indicates that this entry corresponds
-to a sequence in the FASTA file and the following line indicates the name of
-this sequence using the ``source'' tag. As described the first word of the
-FASTA header is used to identify which sequence is selected from the file.
-Together these three lines describe the first entry in the scaffold as a simple
-sequence using the sequence named `sequence1'. On the right hand side of Figure
+The first line of the scaffold file in Figure 1 begins with three dashes to
+indicate the start of a YAML-formatted document. The first entry (highlighted
+in green) begins with a dash character `-' to denote an entry in the YAML
+list. This is a requirement of the YAML format: that each entry begins with
+a dash line. The next line is indented by two spaces where whitespace is
+required to group similar attributes together. The ``sequence'' tag indicates
+that this entry corresponds to a sequence in the FASTA file and the following
+line indicates the name of this sequence using the ``source'' tag. The first
+word of the FASTA header is used to identify which sequence is selected from
+the file. Together these three lines describe the first entry in the scaffold
+as a simple sequence using the sequence named `sequence1'. On the right hand
+side of Figure
 1 this produces the first region in the scaffold, also shown in green. \pb
@@
 the sequence to these coordinates inclusively, while the ``reverse'' tag
-instructs scaffolder to also reverse complement the sequence. In the scaffolded
-genome shown in Figure 1 this completes the scaffold sequence. \pb
+instructs scaffolder to also reverse complement the sequence. In the putative
+scaffold shown in Figure 1 this completes the scaffold sequence. \pb
 
@@
 with each insert beginning with a dash. The first insert, shown in purple,
-similar attributes to that of a sequence entry; the reverse, start and stop
-tags are used to trim and reverse complement the insert. Similarly the `source'
-tag identifies the corresponding FASTA sequence as `insert1'. The ``open'' and
-``close'' tags are specific to inserts and determine where the insert is added
-in the enclosing sequence. The region of the sequence inside these coordinates
-is inclusively replaced by the specified insert sequence. This is visualised in
-the putative genome in Figure 1 by the black lines bisecting the blue sequence.
-\pb
+uses similar attributes to that of a sequence entry; the reverse, start and
+stop tags are used to trim and reverse complement the insert. Similarly the
+`source' tag identifies the corresponding FASTA sequence as `insert1'. The
+``open'' and ``close'' tags are specific to inserts and determine where the
+insert is added in the enclosing sequence. The region of the sequence inside
+these coordinates is inclusively replaced by the specified insert sequence.
+This is visualised in the putative scaffold in Figure 1 by the black lines
+bisecting the blue sequence. \pb
 
@@
 addition Scaffolder provides a command line interface (CLI) to validate the
-scaffold file and build a draft super sequence. The Scaffolder CLI behaves as
-a standard Unix tool and returns appropriate exit codes and manual pages. The
-use of both these Scaffolder interfaces is outlined in detail on the scaffolder
-website (\scaffolder). \pb
+scaffold file and build the draft super sequence. The Scaffolder CLI behaves
+as a standard Unix tool and returns appropriate exit codes and manual pages.
+The use of both these Scaffolder interfaces is outlined in detail on the
+scaffolder website (\scaffolder). \pb
+
+\clearpage
 
@@
 
-Scaffolding requires joining assembled contigs and insert additional sequence
-data using a combination of computational and laboratory methods. The process
-of manually editing a scaffold is inherently hard to reproduce and introduces
-the possibility of irreducible techniques and/or human error. The aims of
-the Scaffolder software are twofold: 1) to provide software that is easy to
-install and simplifies the task of genome finishing; and 2) to facilitate
-reproducibility when joining large nucleotide sequences together to create
-a genome sequence. \pb
+Scaffolding an incomplete genome assembly requires joining contigs and
+additional gap-filling sequence using a combination of computational and
+laboratory methods. The process of manually editing a scaffold is inherently
+hard to reproduce and introduces the possibility of irreducible techniques
+and/or human error. In respect to these drawbacks the aims of the Scaffolder
+software are twofold: 1) to provide software that is easy to install and
+simplifies the task of genome finishing; and 2) to facilitate reproducibility
+in the scaffolding and finishing stage of a genome project. \pb
 
 Scaffolder was designed to be as simple to use and, assuming the Ruby and
-RubyGems software are present, Scaffolder can be installed in a single command
-line step. This should negate any possible barriers to entry which require
-manual compilation of source code. Scaffolder also uses a minimal and compact
-syntax to describe the genome scaffold. This syntax is simple to construct and
-edit whilst being succinct and readable. The goal of this syntax is to make
-genome scaffolds easy to make in a common text editor; however the use of word
-processors for this task is discouraged. \pb
+RubyGems software are present, can be installed in a single command line step.
+This should negate any possible barriers to entry which require manual
+compilation of source code. Scaffolder also uses a minimal and compact syntax
+to describe how the genome scaffold sequence should be generated. This syntax
+is simple to construct and edit whilst being succinct and readable. The goal
+of this syntax is to make genome scaffolds easy to write in a common text
+editor. \pb
 
-Constructing a genome using a text file makes generating a scaffold super
-sequence both reproducible and deterministic for the same scaffold and FASTA
-sequence file. In comparison, joining large nucleotide sequences by hand cannot
-be reliably reproduced. The scaffold file furthermore provides a human readable
-description of how the scaffold is constructed. Configuring the final sequence
-in the scaffold file means the build is easier to edit once already
-constructed. The YAML text format allows comparison of differences between
-scaffold builds using standard Unix tools such as diff. This makes scaffold
-files easier to store in version control systems and allows genome finishers to
-use methods similar to those in software development. \pb
+Constructing a genome using by specifying the scaffold organisation in text
+file makes generating a scaffold super sequence both reproducible and
+deterministic for the same file and set of FASTA sequences. In comparison,
+joining large nucleotide sequences by hand cannot be reliably reproduced. The
+scaffold file furthermore provides a human readable description of how the
+scaffold is constructed. Configuring the final sequence in the scaffold file
+means the build is easier to edit once already constructed. The YAML text
+format allows comparison of differences between scaffold builds using standard
+Unix tools such as diff. This makes scaffold files easier to store in version
+control systems and allows genome finishers to use methods similar to those in
+software development. \pb
 

### 2011-03-08 src/article.tex
Revise conclusions

@@
 
-Scaffolder is software aimed at bioinformaticians and biologists familiar with
-the command line who wish to build a genome sequence scaffold from a set of
-assembled contigs. The Scaffolder file format maintains the genome scaffold as
-a concise and readable text representation that allows third parties to see how
-the genome sequence was scaffolded. This file format also allows a broad
-overview of which sequences were included and how they are ordered in the
-genome scaffold, something not possible to deduce from a megabase-length
-strings of nucleotide characters. Scaffolder furthers increases the ease of
+Scaffolder is software aimed at both bioinformaticians and biologists familiar
+with the command line who wish to build a genome scaffold from a set of
+contigs. The Scaffolder file format maintains the genome scaffold as a concise
+and readable text representation that allows third parties to see how the
+genome sequence was scaffolded. This file format also allows a broad overview
+of which sequences were included and how they are ordered in the genome
+scaffold, something not possible to deduce from a megabase-length string of
+nucleotide characters. Scaffolder furthers increases the ease of
 reproducibility in genome projects by allowing the scaffold super-sequence to
-be reliably reproduced from the same scaffold file. The YAML syntax for writing
-the scaffold file is also standardised and simple to manipulate
+be reliably reproduced from the same scaffold file. The YAML syntax for
+writing the scaffold file is also standardised and simple to manipulate
 programmatically. This thereby means the scaffolding process follows the Unix

### 2011-03-08 src/article.tex
Spell check

@@
 FASTA file and then performing any designated sequence trimming or reverse
-complenting. Each of the sequence entries are then joined into a continuous
+complementing. Each of the sequence entries are then joined into a continuous
 single super-sequence and returned to the console output. \pb

### 2011-03-09 src/article.tex
Minor corrections

@@
 
-  \paragraph*{Methods:} The software outlined here, ``Scaffolder'', is
+  \paragraph*{Methods:} The software outlined here, ``Scaffolder,'' is
   implemented in the Ruby programming language and can be installed via the
-  RubyGems software management system. Genome scaffolds are defined using YAML
-  - a data format, which is both human and machine-readable. Command line
+  RubyGems software management system. Genome scaffolds are defined using
+  YAML - a data format, which is both human and machine-readable. Command line
   binaries and extensive documentation are available.
@@
   define the scaffold. This syntax further allows  unknown regions to be
-  defined, and add inserts to fill gaps in the scaffold. Defining the genome
-  construction in a file makes the scaffolding process reproducible and easier
-  to edit compared with FASTA nucleotide sequence.
+  defined, and add additional sequences to fill gaps in the scaffold. Defining
+  the genome construction in a file makes the scaffolding process reproducible
+  and easier to edit compared with FASTA nucleotide sequence.
 
@@
 reads genome assembly software may be unable to produce a complete sequence.
-Instead the algorithm may generate several large assembled regions of sequence
-(`contigs') composed of the many individual reads that represent assembly into
-the longest possible sequences, given the data. Nonetheless, these contigs
-represent a fragmented picture of the genome and require additional work to
-join together into a complete sequence. \pb
+Instead, the algorithm may generate several large assembled regions of
+sequence (`contigs') composed of the many individual reads that represent
+assembly into the longest possible sequences, given the data. Nonetheless,
+these contigs represent a fragmented picture of the genome and require
+additional work to join together into a complete sequence. \pb
 
@@
 
-The software outlined here, ``Scaffolder'', aims to address these problems of
+The software outlined here, ``Scaffolder,'' aims to address these problems of
 reproducibility by creating a file syntax and software framework for editing
-a genome scaffold. Scaffolder uses a specific file format to define how contigs
-are joined, additional sequences are inserted, and for the specification of
-unknown regions. This syntax allows a scaffold to be updated by simply editing
-the scaffold file. As such scaffolder facilitates a reproducible scaffolding
-process and provides a concise overview of how the final genomic scaffold was
-constructed. \pb
+a genome scaffold. Scaffolder uses a specific file format to define how
+contigs are joined, additional sequences are inserted, and for the
+specification of unknown regions. This syntax allows a scaffold to be updated
+by simply editing the scaffold file. As such, scaffolder facilitates
+a reproducible scaffolding process and provides a concise overview of how the
+final genomic scaffold was constructed. \pb
 
@@
 
-The scaffold file specifies how sequence regions are joined together; however
+The scaffold file specifies how sequence regions are joined together; however,
 the sequences themselves are not stored in the scaffold file and are instead
@@
 programmatically. This thereby means the scaffolding process follows the Unix
-tenet of: ``If your data structures are good enough, the algorithm to
-manipulate them should be trivial.''.
+tenet of ``If your data structures are good enough, the algorithm to
+manipulate them should be trivial.''
 

### 2011-08-10 src/article.tex
Tested on Ubuntu
Add track changes LaTeX package
List software dependencies in description section

@@
 \usepackage{multicol}   %Columns
+\usepackage{trackchanges}
 \usepackage[utf8]{inputenc} %unicode support
@@
     \item[Operating system:] Platform Independent. Tested on Mac OS X and
-    Debian.
+    Ubuntu.
     \item[Programming language:] Ruby
-    \item[Other requirements:] RubyGems
+    \item[Other requirements:] \change{RubyGems}{RubyGems package management
+    software and the following libraries: BioRuby and confligliere. A full
+    list of development dependencies can be found in the Gemfile in the base
+    directory of each project.}
     \item[License:] MIT

### 2011-08-13 src/article.tex
Change manuacript title
Text fixes

@@
 
-\title{Scaffolder - Software for Reproducible Genome Scaffolding.} %{{{1
+\title{Scaffolder - Software for \change{Reproducible}{Manual} Genome
+Scaffolding.}
+%{{{1
 
@@
 within the genome, or biased and incomplete sequencing data, may prevent the
-genome being assembled into a continuous sequence. This may be due to not
-enough, or multiple different overlaps between reads and is analogous to
-missing pieces in the jigsaw or pieces that fit to multiple other pieces. \pb
+genome being assembled into a continuous sequence. This may be due to
+\change{not enough}{insufficient}, or multiple different overlaps between reads
+and is analogous to missing pieces in the jigsaw or pieces that fit to multiple
+other pieces. \pb
 
-The advent of high throughput sequencing methods has led to a renewed interest
-in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
-Nonetheless, due to the complexity of merging large numbers of overlapping
-reads genome assembly software may be unable to produce a complete sequence.
-Instead, the algorithm may generate several large assembled regions of
-sequence (`contigs') composed of the many individual reads that represent
-assembly into the longest possible sequences, given the data. Nonetheless,
-these contigs represent a fragmented picture of the genome and require
-additional work to join together into a complete sequence. \pb
+The advent of high\add{-}throughput sequencing methods has led to a renewed
+interest in algorithms to solve the problem of genome assembly
+\cite{pop2008,pop2009}. Nonetheless, due to the complexity of merging large
+numbers of overlapping reads genome assembly software may be unable to produce
+a complete sequence. Instead, the algorithm may generate several large
+assembled regions of sequence (`contigs') composed of the many individual reads
+that represent assembly into the longest possible sequences, given the data.
+Nonetheless, these contigs represent a fragmented picture of the genome and
+require additional work to join together into a complete sequence. \pb
 
-The process of ameliorating a draft sequence into a finished continuous
-sequence can be expensive in terms of time and laboratory effort, although the
-genomic data present in a set of generated contigs may be sufficient for many
-research questions \cite{branscomb2002}. Nevertheless, a continuous
-high-quality `finished' genome sequence does provide a greater depth of
-information, such as complete resolution of repeat regions and precise
-estimates of distances between genomic elements \cite{parkhill2002,fraser2002}.
-The process of joining these contigs together to form a continuous genome
-sequence is called the `scaffolding' or `finishing' stage and is the focus of
-the software described in this article. \pb
+The process of \change{ameliorating a draft sequence into a finished continuous
+sequence}{finishing a genome sequence} can be expensive in terms of time and
+laboratory effort, although the genomic data present in a set of generated
+contigs may be sufficient for many research questions \cite{branscomb2002}.
+Nevertheless, a continuous high-quality `finished' genome sequence does provide
+a greater depth of information, such as complete resolution of repeat regions
+and precise estimates of distances between genomic elements
+\cite{parkhill2002,fraser2002}. The process of joining these contigs together
+to form a continuous genome sequence is called the `scaffolding' or `finishing'
+stage and is the focus of the software described in this article. \pb
 
@@
 requires reverse complementing sequences where necessary. In the case of
-microbial genomes this orientation will be to the 5'~$\rightarrow$~3'
-direction following the direction of genome replication.
+\change{microbial}{archaeal and bacterial} genomes this orientation will be to
+the 5'~$\rightarrow$~3' direction following the direction of genome
+replication.
 
@@
 apart in the genome. The occurrence of paired reads in separate contigs can be
-used to probabilistically estimate the order and distance between these
-contigs. Alternatively, laboratory methods may use PCR to amplify the unknown
-DNA in a gap region, then use traditional Sanger sequencing to determine the
-sequence of this gap. Computational methods are more preferable as they are
-less costly in laboratory time and materials compared to manual gap resolution,
-and use the available data that exists following sequencing
-\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
-\emph{in silico} software packages exist to suggest the likely primers
-necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb
+used to \change{probabilistically estimate}{estimate probabilistically} the
+order and distance between these contigs. Alternatively, laboratory methods may
+use PCR to amplify the unknown DNA in a gap region, then use traditional Sanger
+sequencing to determine the sequence of this gap. Computational methods are
+\remove{more} preferable as they are less costly in laboratory time and
+materials compared to manual gap resolution, and use the available data that
+exists following sequencing \cite{nagarajan2010}. Finally when the scaffold
+cannot be completely resolved, \emph{in silico} software packages exist to
+suggest the likely primers necessary for PCR amplifying the sequence in gap
+regions \cite{gordon2001}. \pb
 
@@
 short sequence or being ignored by more conservative assembly algorithms. Such
-regions can be resolved by using algorithms specifically reassemble the
-collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
+regions can be resolved by using algorithms \add{which} specifically reassemble
+the collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
 a related approach uses unassembled sequence reads matching the regions around

### 2011-08-13 src/article.tex
Update version dependency requirements
Add link to scaffolder getting started

@@
 The use of both these Scaffolder interfaces is outlined in detail on the
-scaffolder website (\scaffolder). \pb
+scaffolder website (\scaffolder). \add{This website also contains a ``getting
+started'' guide which introduces using Scaffolder.} \pb
 
@@
   \begin{description}
-    \item[Project name:] Scaffolder v0.4.1, Scaffolder Tools v0.1.2
+    \item[Project name:] Scaffolder v0.4.\change{1}{4}, Scaffolder Tools
+    v0.1.\change{2}{3}
     \item[Project home page:] \scaffolder
@@
     Ubuntu.
-    \item[Programming language:] Ruby
+    \item[Programming language:] Ruby \add{1.8.7 or 1.9.2}
     \item[Other requirements:] \change{RubyGems}{RubyGems package management
-    software and the following libraries: BioRuby and confligliere. A full
-    list of development dependencies can be found in the Gemfile in the base
-    directory of each project.}
+    software and the following libraries: BioRuby \add{1.4.x}, confligliere
+    \add{0.1.x}\add{, ronn 0.7.x}. A full list of development dependencies
+    can be found in the Gemfile in the base directory of each project.}
     \item[License:] MIT

### 2011-08-14 src/article.tex
Fix broken track change in requirements section

@@
     \item[Other requirements:] \change{RubyGems}{RubyGems package management
-    software and the following libraries: BioRuby \add{1.4.x}, confligliere
-    \add{0.1.x}\add{, ronn 0.7.x}. A full list of development dependencies
-    can be found in the Gemfile in the base directory of each project.}
-    \item[License:] MIT
-    \item[Any restrictions to use by non-academics:] None
+    software and the following libraries: BioRuby 1.4.x, confligliere
+    0.1.x, ronn 0.7.x. A full list of development dependencies can be found in
+    the Gemfile in the base directory of each project.}
+    \item[License:] MIT \item[Any restrictions to use by non-academics:] None
     \end{description}

### 2011-08-14 src/article.tex
Revise discussion section

@@
 Scaffolding an incomplete genome assembly requires joining contigs and
-additional gap-filling sequence using a combination of computational and
+additional gap-filling sequence\add{s} using a combination of computational and
 laboratory methods. The process of manually editing a scaffold is inherently
-hard to reproduce and introduces the possibility of irreducible techniques
-and/or human error. In respect to these drawbacks the aims of the Scaffolder
-software are twofold: 1) to provide software that is easy to install and
-simplifies the task of genome finishing; and 2) to facilitate reproducibility
-in the scaffolding and finishing stage of a genome project. \pb
+hard to reproduce and introduces \remove{the possibility of }irreducible
+\change{techniques}{edits} and/or human error. In respect to \change{these
+drawbacks}{this} the aims of the Scaffolder software are twofold: 1) to provide
+software that is easy to install and simplifies the task of genome finishing;
+and 2) to facilitate reproducibility in the scaffolding and finishing stage of
+a genome project. \pb
 
-Scaffolder was designed to be as simple to use and, assuming the Ruby and
-RubyGems software are present, can be installed in a single command line step.
-This should negate any possible barriers to entry which require manual
-compilation of source code. Scaffolder also uses a minimal and compact syntax
-to describe how the genome scaffold sequence should be generated. This syntax
-is simple to construct and edit whilst being succinct and readable. The goal
-of this syntax is to make genome scaffolds easy to write in a common text
-editor. \pb
+\remove{Scaffolder was designed to be as simple to use and, assuming the Ruby
+and RubyGems software are present, can be installed in a single command line
+step. This should negate any possible barriers to entry which require manual
+compilation of source code.} Scaffolder \remove{also} uses a minimal and
+compact syntax to describe how \change{the}{a} genome scaffold sequence should
+be generated. This syntax is simple to \change{construct}{write} and edit
+whilst being succinct and readable. \change{The goal of this syntax is to make
+genome scaffolds easy to write in a common text editor.}{The AGP format is
+a similar format for describing scaffolds. This format can be used to define
+contig order and N-filled gap regions in a scaffold. The advantages of the AGP
+format is that each contig entry is defined on a single line which allows
+searching the scaffold using UNIX command line tools. The scaffolder format in
+comparison in contrast is written in the standardised YAML format and therefore
+accessable to the many languages which provide parsers to this format. The
+scaffolder format is provided with a tool explicitly to produce the fasta
+sequence of scaffold specified by in the file.} \pb
 
@@
 scaffold is constructed. Configuring the final sequence in the scaffold file
-means the build is easier to edit once already constructed. The YAML text
-format allows comparison of differences between scaffold builds using standard
-Unix tools such as diff. This makes scaffold files easier to store in version
-control systems and allows genome finishers to use methods similar to those in
-software development. \pb
+means the build is easier to edit once already constructed. \pb
+
+\subsection*{Example Use Case}
+
+\add{The example use case for scaffolder is a combination of compuation and
+manual editing of a genome scaffold. As discussed many programming languages
+provide libraries for reading and writing YAML. We have used scaffolder in our
+genome projects create a scaffold file from parsing the output of \emph{in
+silico} scaffolding tools into the scaffold file format. This scaffold was then
+manually updated in a text editor with the results of using PCR to identify
+gaps sequences. We see this as the advantage of describing a scaffold in both
+a human and machine readable format.} \pb
+
+The YAML text format \add{also} allows comparison of differences between
+scaffold builds using standard Unix tools such as diff. This \add{therefore}
+makes scaffold files \add{ameniable} to storage in version control systems and
+allows genome finishers to use methods similar to those in software
+development. \pb
 

### 2011-08-14 src/article.tex
Results: One scaffold per scaffold file

@@
 where each entry corresponds to a region of sequence in the generated scaffold
-super-sequence. \pb
+super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
+sequence and separate scaffolds are defined in separate files.} \pb
 
@@
 comparison in contrast is written in the standardised YAML format and therefore
-accessable to the many languages which provide parsers to this format. The
+accessible to the many languages which provide parsers to this format. The
 scaffolder format is provided with a tool explicitly to produce the fasta
@@
 
-\add{The example use case for scaffolder is a combination of compuation and
+\add{The example use case for scaffolder is a combination of computation and
 manual editing of a genome scaffold. As discussed many programming languages
@@
 scaffold builds using standard Unix tools such as diff. This \add{therefore}
-makes scaffold files \add{ameniable} to storage in version control systems and
+makes scaffold files \add{amenable} to storage in version control systems and
 allows genome finishers to use methods similar to those in software

### 2011-08-14 src/article.tex
fix
Proof read abstract

@@
   step in a genome project is to join neighbouring sequence regions together
-  and fill gaps in the assembly using additional sequences. This scaffolding
-  step, however, is non-trivial and requires manually editing large blocks of
-  nucleotide sequence. Joining these sequences together also hides the source
-  of each region in the final genome sequence. Taken together, these
+  and fill gaps \remove{in the assembly using additional sequences}. This
+  scaffolding step, however, is non-trivial and requires manually editing large
+  blocks of nucleotide sequence. Joining these sequences together also hides
+  the source of each region in the final genome sequence. Taken together, these
   considerations may make reproducing or editing an existing genome build
@@
   implemented in the Ruby programming language and can be installed via the
-  RubyGems software management system. Genome scaffolds are defined using
-  YAML - a data format, which is both human and machine-readable. Command line
-  binaries and extensive documentation are available.
+  RubyGems software management system. Genome scaffolds are defined using YAML
+  - a data format\remove{,} which is both human and machine-readable. Command
+  line binaries and extensive documentation are available.
 
-  \paragraph*{Results:} This software allows a genome build to be defined in
-  terms of the constituent sequences using a relatively simple syntax to
-  define the scaffold. This syntax further allows  unknown regions to be
-  defined, and add additional sequences to fill gaps in the scaffold. Defining
-  the genome construction in a file makes the scaffolding process reproducible
-  and easier to edit compared with FASTA nucleotide sequence.
+  \paragraph*{Results:} This software allows a genome \change{scaffold}{build}
+  to be defined in terms of the constituent sequences using a relatively simple
+  syntax\remove{to define the scaffold}. This syntax further allows unknown
+  regions to be \change{defined}{specified} and add additional sequences to
+  fill gaps in the scaffold. Defining the genome construction in a file makes
+  the scaffolding process reproducible and easier to edit compared with
+  \add{large} FASTA nucleotide sequences.
 

### 2011-08-14 src/article.tex
Revise background section

@@
 is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
-within the genome, or biased and incomplete sequencing data, may prevent the
-genome being assembled into a continuous sequence. This may be due to
-\change{not enough}{insufficient}, or multiple different overlaps between reads
-and is analogous to missing pieces in the jigsaw or pieces that fit to multiple
-other pieces. \pb
+\remove{within the genome, }or biased and incomplete sequencing data, may
+prevent the genome being assembled into a continuous sequence. This may be due
+to \change{not enough}{insufficient}, or multiple different overlaps between
+reads and is analogous to missing pieces in the jigsaw or pieces that fit to
+multiple other pieces. \pb
 
@@
 interest in algorithms to solve the problem of genome assembly
-\cite{pop2008,pop2009}. Nonetheless, due to the complexity of merging large
-numbers of overlapping reads genome assembly software may be unable to produce
-a complete sequence. Instead, the algorithm may generate several large
-assembled regions of sequence (`contigs') composed of the many individual reads
-that represent assembly into the longest possible sequences, given the data.
-Nonetheless, these contigs represent a fragmented picture of the genome and
-require additional work to join together into a complete sequence. \pb
+\cite{pop2008,pop2009}. \remove{Nonetheless, due to }The complexity of merging
+large numbers of overlapping reads \add{can lead to} genome assembly software
+may be unable to produce a complete sequence. Instead, the algorithm may
+generate several large assembled regions of sequence (`contigs') composed of
+the many individual reads\remove{ that represent assembly into the longest
+possible sequences, given the data. Nonetheless, }These contigs represent
+a fragmented picture of the genome and \add{therefore} require additional work
+to join together into a complete sequence. \pb
 
@@
 sequence}{finishing a genome sequence} can be expensive in terms of time and
-laboratory effort, although the genomic data present in a set of generated
-contigs may be sufficient for many research questions \cite{branscomb2002}.
-Nevertheless, a continuous high-quality `finished' genome sequence does provide
-a greater depth of information, such as complete resolution of repeat regions
-and precise estimates of distances between genomic elements
-\cite{parkhill2002,fraser2002}. The process of joining these contigs together
-to form a continuous genome sequence is called the `scaffolding' or `finishing'
-stage and is the focus of the software described in this article. \pb
+laboratory effort. \change{although}{In some cases} the genomic data present in
+a set of generated contigs may be sufficient for many research questions
+\cite{branscomb2002}. Nevertheless, a continuous high-quality `finished' genome
+sequence does provide a greater depth of information, such as complete
+resolution of repeat regions and precise estimates of distances between genomic
+elements \cite{parkhill2002,fraser2002}. The process of joining these contigs
+together to form a continuous genome sequence is called the `scaffolding' or
+`finishing' stage and is the focus of the software described in this article.
+\pb
 
@@
 Filling these regions with unknown nucleotide characters (`N') allows a draft
-continuous sequence, which is useful for representing both the known and
-to-be-resolved areas in the genome sequence.
+continuous sequence. \change{which}{This sequence} is useful for representing
+both the known and to-be-resolved areas in the genome sequence.
 
@@
 to the laboratory to perform additional sequencing or using computational
-methods of estimating the unknown sequence. This additional sequence is then
-used to replace the gap between two contigs, joining them into a single
-sequence. Once all contigs have been joined and gaps in a scaffold closed, the
-genome may be considered finished.
+methods of estimating the unknown sequence. This additional sequence is
+\remove{then }used to replace the gap between two contigs, joining them into
+a single sequence. Once all contigs have been joined and gaps in a scaffold
+closed, the genome may be considered finished.
 
@@
 
-The process of resolving a genome sequence and building a scaffold uses wet
-laboratory methods, \emph{in silico} methods, or a combination of both. An
-example of a computational method might use the existing paired-read data from
-the sequencing stage where reads were produced in pairs at a known distance
-apart in the genome. The occurrence of paired reads in separate contigs can be
-used to \change{probabilistically estimate}{estimate probabilistically} the
-order and distance between these contigs. Alternatively, laboratory methods may
-use PCR to amplify the unknown DNA in a gap region, then use traditional Sanger
-sequencing to determine the sequence of this gap. Computational methods are
-\remove{more} preferable as they are less costly in laboratory time and
+The process of \change{resolving a genome sequence and building a}{finishing
+a genome} scaffold \change{may} use wet laboratory methods, \emph{in silico}
+methods, or a combination of both. An example of a computational method might
+use the \remove{existing} paired-read data from the sequencing stage.
+\remove{where reads were produced in pairs at a known distance apart in the
+genome.} The occurrence of paired reads in separate contigs can be used to
+\change{probabilistically estimate}{estimate probabilistically} the order and
+distance between these contigs. Alternatively, laboratory methods may use PCR
+to amplify the unknown DNA in a gap region, then use \remove{traditional}
+Sanger sequencing to determine the sequence of this gap. Computational methods
+are \remove{more} preferable as they are less costly in laboratory time and
 materials compared to manual gap resolution, and use the available data that
@@
 
-Examples of \emph{in silico} methods include using synteny to compare the
-assembled contigs to a complete reference genome sequence by searching for
-areas of sequence similarity between the two. Any areas of corresponding
-sequence with the reference genome can be used to infer contig placement and
-build the contigs into a scaffold \cite{richter2007,zhao2008,assefa2009}.
-Recombination within each of the two compared genomes can, however, reduce the
-effectiveness of this analysis. \pb
+Examples of \emph{in silico} methods include \remove{using synteny to}
+comparing the assembled contigs to a complete reference genome sequence
+\change{by}{to} search\remove{ing} for areas of sequence similarity between the
+two. Any areas of corresponding sequence with the reference genome can be used
+to infer contig placement and build the contigs into a scaffold
+\cite{richter2007,zhao2008,assefa2009}. Recombination within each of the two
+compared genomes can, however, reduce the efficacy of this. \pb
 
@@
 regions can be resolved by using algorithms \add{which} specifically reassemble
-the collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
-a related approach uses unassembled sequence reads matching the regions around
-a scaffold gap to construct a uniquely overlapping set of reads across it
-\cite{tsai2010}. \pb
+the collapsed repeat region correctly \cite{mulyukov2002,koren2010}.
+\change{while a}{A} related approach uses unassembled sequence reads matching
+the regions around a scaffold gap to construct a uniquely overlapping set of
+reads across \change{it}{the gap}. \cite{tsai2010}. \pb
 
@@
 
-These described \emph{in silico} methods provide a wide array of approaches
-for merging contigs into a larger, continuous scaffold sequence. Nevertheless
-the scaffolding process may still require manually inserting additional
-sequences or further joining contigs using PCR-derived sequence. Moving and
-editing large blocks of nucleotide text by hand however possibly introduces
-human error and precludes any reproducibility of the steps in this process. \pb
+These described \emph{in silico} methods provide a wide array of approaches for
+merging contigs into a larger, continuous scaffold sequence.
+\remove{Nevertheless} The scaffolding process may still require manually
+inserting additional sequences or further joining contigs using PCR-derived
+sequence. Moving and editing large blocks of nucleotide text by hand however
+\remove{possibly} introduces human error and precludes any reproducibility of
+the steps in this process. \pb
 
@@
 by simply editing the scaffold file. As such, scaffolder facilitates
-a reproducible scaffolding process and provides a concise overview of how the
-final genomic scaffold was constructed. \pb
+a reproducible \change{scaffolding}{finishing} process and provides a concise
+overview of how the final genomic scaffold was constructed. \pb
 

### 2011-08-14 src/article.tex
Revise implementation section

@@
 
-Scaffolder is written in the Ruby programming language using version 1.8.7
-\cite{ruby-lang}. The scaffolder package is split into two libraries. The
-first called ``scaffolder'' provides the core Scaffolder application
-programming interface (API). The second library ``scaffolder-tools'' provides
-the Scaffolder command line interface (CLI). \pb
+Scaffolder is written in the Ruby programming language \change{using version
+1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
+scaffolder package is split into two libraries. The first called ``scaffolder''
+which provides the core Scaffolder application programming interface (API). The
+second library ``scaffolder-tools'' provides the Scaffolder command line
+interface (CLI). \pb
 
@@
 
-The scaffolder algorithm reads through the scaffolder file parsing. Each entry
-is then converted to corresponding nucleotide sequence by fetching from the
-FASTA file and then performing any designated sequence trimming or reverse
-complementing. Each of the sequence entries are then joined into a continuous
-single super-sequence and returned to the console output. \pb
+\remove{The scaffolder algorithm reads through the scaffolder file parsing.
+Each entry is then converted to corresponding nucleotide sequence by fetching
+from the FASTA file and then performing any designated sequence trimming or
+reverse complementing. Each of the sequence entries are then joined into
+a continuous single super-sequence and returned to the console output.}
 

### 2011-08-15 src/article.tex
Revise discussion

@@
 
-  \paragraph*{Background:} Assembly of short-read sequencing data can result in
-  a fragmented non-contiguous series of genomic sequences. Therefore a common
-  step in a genome project is to join neighbouring sequence regions together
-  and fill gaps \remove{in the assembly using additional sequences}. This
-  scaffolding step, however, is non-trivial and requires manually editing large
-  blocks of nucleotide sequence. Joining these sequences together also hides
-  the source of each region in the final genome sequence. Taken together, these
-  considerations may make reproducing or editing an existing genome build
+  \paragraph*{Background:} \add{The} assembly of \add{next-generation}
+  short-read sequencing data can result in a fragmented non-contiguous series
+  of genomic sequences. Therefore a common step in a genome project is to join
+  neighbouring sequence regions together and fill gaps \remove{in the assembly
+  using additional sequences}. This scaffolding step\remove{, however,} is
+  non-trivial and requires manually editing large blocks of nucleotide
+  sequence. Joining these sequences together also hides the source of each
+  region in the final genome sequence. Taken together, these considerations
+  may make reproducing or editing an existing genome \change{build}{scaffold}
   difficult.
@@
   regions to be \change{defined}{specified} and add additional sequences to
-  fill gaps in the scaffold. Defining the genome construction in a file makes
-  the scaffolding process reproducible and easier to edit compared with
-  \add{large} FASTA nucleotide sequences.
+  fill \add{known} gaps in the scaffold. Defining the genome construction in
+  a file makes the scaffolding process reproducible and easier to edit
+  compared with \add{large} FASTA nucleotide sequences.
 
   \paragraph*{Conclusions:} Scaffolder is easy-to-use genome scaffolding
-  software. This tool promotes reproducibility and continuous development in a
-  genome project. Scaffolder can be found at \scaffolder.
+  software\change{. This tool}{ which} promotes reproducibility and continuous
+  development in a genome project. Scaffolder can be found at \scaffolder.
 

### 2011-08-15 src/article.tex
Revise results section

@@
 together into a single contiguous scaffolded super-sequence. Plain-text
-scaffold files written in the YAML syntax specify how these sequences should
-be joined and the scaffolder software is used to generate the scaffold
-sequence from these instructions. In addition to specifying which contigs are
-required, the scaffold file allows the contigs to be edited into smaller
-sub-sequences or reverse complemented if necessary. The process of genome
-finishing may involve producing additional oligonucleotide sequences to fill
-unknown regions in the scaffold and the Scaffolder allows these additional
-insert sequences to be used to fill these gaps. These inserts can also be
-treated in the same manner as larger contig sequences: trimmed and/or reverse
-complemented to match the corresponding gap region size and orientation. \pb
+scaffold files written in \remove{the} YAML \remove{syntax} specify how these
+sequences should be joined\change{. and the}{The} scaffolder software is used
+to generate the scaffold sequence from these instructions. In addition to
+specifying which contigs are required, the scaffold file allows the contigs to
+be edited into smaller sub-sequences or reverse complemented if necessary.\pb
+
+The process of genome finishing may involve producing additional
+oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
+\remove{and} The Scaffolder \change{allows}{format provides for using} these
+additional insert sequences to \remove{be used} to fill \remove{these} gaps.
+These inserts can also be treated in the same manner as larger contig
+sequences: trimmed and/or reverse complemented to match the corresponding gap
+region size and orientation. \pb
 
@@
 mapping the contigs to a reference genome. These inter-contig gap regions are
-useful to join separate sequences together by an estimated size and the
-scaffold file allows for the specification of such unresolved regions by
-inserting regions of `N' nucleotides into the scaffold. The use of these
-regions in the scaffold indicates the unresolved regions in the build and their
-approximate size. \pb
+useful to join separate sequences together by \change{an}{the} estimated
+\change{size and }{distance.} The scaffold file allows for the specification
+of such unresolved regions by inserting regions of `N' nucleotides into the
+scaffold. The use of these regions in the scaffold indicates the unresolved
+regions in the build and their approximate size. \pb
 
-The scaffold file specifies how sequence regions are joined together; however,
-the sequences themselves are not stored in the scaffold file and are instead
-maintained as a separate FASTA file. Nucleotide sequences in a FASTA-format
-file are the standard output for a genome assembler, and these may be contigs,
-sets of contigs scaffolded into larger sequences, or both. These nucleotide
-sequences stored in the FASTA file are referenced in the scaffold using the
+\change{The scaffold file specifies how sequence regions are joined together;
+however, the sequences themselves are no  stored in the scaffold file and are
+instead}{The nucleotide sequences used in the scaffold are} maintained as
+a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
+the standard output for a genome assembler, and these may be contigs, sets of
+contigs scaffolded into larger sequences, or both. }The nucleotide sequences
+\remove{stored in the FASTA file} are referenced in the scaffold using the
 first word from in the FASTA header of the corresponding sequence. Maintaining
@@
 super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
-sequence and separate scaffolds are defined in separate files.} \pb
+sequence and separate scaffolds should be defined in separate files.} \pb
 
@@
 as a simple sequence using the sequence named `sequence1'. On the right hand
-side of Figure
-1 this produces the first region in the scaffold, also shown in green. \pb
+side of Figure 1 this produces the first region in the scaffold, also shown in green. \pb
 
@@
 `sequence2' to the scaffold. This entry demonstrates how this sequence may be
-manipulated prior to addition to the scaffold. The `start' and `stop' tags trim
-the sequence to these coordinates inclusively, while the ``reverse'' tag
-instructs scaffolder to also reverse complement the sequence. In the putative
-scaffold shown in Figure 1 this completes the scaffold sequence. \pb
+manipulated prior to addition to the scaffold. The `start' and `stop' tags
+trim the sequence to these coordinates inclusively.\remove{while} The
+``reverse'' tag instructs scaffolder to also reverse complement the sequence.
+In the putative scaffold shown in Figure 1 this completes the \remove{scaffold
+} sequence. \pb
 
-The final entry in the scaffold also includes the ``inserts'' tag to add
-additional regions of sequence. These inserts are also added as a YAML list,
-with each insert beginning with a dash. The first insert, shown in purple,
-uses similar attributes to that of a sequence entry; the reverse, start and
-stop tags are used to trim and reverse complement the insert. Similarly the
-`source' tag identifies the corresponding FASTA sequence as `insert1'. The
-``open'' and ``close'' tags are specific to inserts and determine where the
-insert is added in the enclosing sequence. The region of the sequence inside
-these coordinates is inclusively replaced by the specified insert sequence.
-This is visualised in the putative scaffold in Figure 1 by the black lines
-bisecting the blue sequence. \pb
+The final entry in the scaffold \remove{also} includes the ``inserts'' tag to
+add additional regions of sequence. These inserts are also added as a YAML
+list, with each insert beginning with a dash. The first insert, shown in
+purple, uses similar attributes to that of a sequence entry; the reverse,
+start and stop tags are used to trim and reverse complement the insert.
+Similarly the `source' tag identifies the corresponding FASTA sequence as
+`insert1'. The ``open'' and ``close'' tags are specific to inserts and
+determine where the insert is added in the enclosing sequence. The region of
+the sequence inside these coordinates is inclusively replaced by the specified
+insert sequence. This is visualised in the putative scaffold in Figure 1 by
+the black lines bisecting the blue sequence. \pb
 
@@
 The use of both these Scaffolder interfaces is outlined in detail on the
-scaffolder website (\scaffolder). \add{This website also contains a ``getting
-started'' guide which introduces using Scaffolder.} \pb
+scaffolder website (\scaffolder). \add{This website provides a ``getting
+started'' guide as an introduction to using Scaffolder to build a genome
+scaffold.} \pb
 

### 2011-08-15 src/article.tex
Revise remainder of manuscript

@@
 
+\subsection*{Comparison with other scaffold formats} %{{{2
+
 \remove{Scaffolder was designed to be as simple to use and, assuming the Ruby
@@
 be generated. This syntax is simple to \change{construct}{write} and edit
-whilst being succinct and readable. \change{The goal of this syntax is to make
-genome scaffolds easy to write in a common text editor.}{The AGP format is
-a similar format for describing scaffolds. This format can be used to define
-contig order and N-filled gap regions in a scaffold. The advantages of the AGP
-format is that each contig entry is defined on a single line which allows
-searching the scaffold using UNIX command line tools. The scaffolder format in
-comparison in contrast is written in the standardised YAML format and therefore
-accessible to the many languages which provide parsers to this format. The
-scaffolder format is provided with a tool explicitly to produce the fasta
-sequence of scaffold specified by in the file.} \pb
+whilst being succinct and readable. \remove{The goal of this syntax is to make
+genome scaffolds easy to write in a common text editor.}
+
+\add{AGP is a similar format for describing scaffolds. This format can be used
+to define contig order and N-filled gap regions in a scaffold. The advantages
+of the AGP format is that each contig entry is defined on a single line which
+allows searching the scaffold using UNIX line-based tools. The scaffolder
+format in comparison in contrast is written in the standardised YAML format
+and therefore accessible to the many languages which provide parsers to this
+format. The scaffolder format is provided with a tool explicitly to produce
+the fasta sequence of scaffold specified by in the file.} \pb
 
@@
 
-\subsection*{Example Use Case}
+\subsection*{Example Use Case} %{{{2
 
 \add{The example use case for scaffolder is a combination of computation and
-manual editing of a genome scaffold. As discussed many programming languages
-provide libraries for reading and writing YAML. We have used scaffolder in our
-genome projects create a scaffold file from parsing the output of \emph{in
-silico} scaffolding tools into the scaffold file format. This scaffold was then
-manually updated in a text editor with the results of using PCR to identify
-gaps sequences. We see this as the advantage of describing a scaffold in both
-a human and machine readable format.} \pb
+manual editing of a genome scaffold. We have used scaffolder in our genome
+projects to create a scaffold file from parsing the output of \emph{in silico}
+scaffolding tools into the scaffold file format. This scaffold was then
+manually updated in a text editor to fill gap regions with laboratory-produced
+sequences.} \pb
 

### 2011-08-15 src/article.tex
Revise abstract

@@
   \paragraph*{Background:} \add{The} assembly of \add{next-generation}
-  short-read sequencing data can result in a fragmented non-contiguous series
-  of genomic sequences. Therefore a common step in a genome project is to join
-  neighbouring sequence regions together and fill gaps \remove{in the assembly
-  using additional sequences}. This scaffolding step\remove{, however,} is
-  non-trivial and requires manually editing large blocks of nucleotide
-  sequence. Joining these sequences together also hides the source of each
-  region in the final genome sequence. Taken together, these considerations
-  may make reproducing or editing an existing genome \change{build}{scaffold}
-  difficult.
+  short-read sequencing data can result in a fragmented non-contiguous
+  \change{series}{set} of genomic sequences. Therefore a common step in
+  a genome project is to join neighbouring sequence regions together and fill
+  gaps \remove{in the assembly using additional sequences}. This scaffolding
+  step\remove{, however,} is non-trivial and requires manually editing large
+  blocks of nucleotide sequence. Joining these sequences together also hides
+  the source of each region in the final genome sequence. Taken together,
+  these considerations may make reproducing or editing an existing genome
+  \change{build}{scaffold} difficult.
 
@@
   syntax\remove{to define the scaffold}. This syntax further allows unknown
-  regions to be \change{defined}{specified} and add additional sequences to
-  fill \add{known} gaps in the scaffold. Defining the genome construction in
-  a file makes the scaffolding process reproducible and easier to edit
-  compared with \add{large} FASTA nucleotide sequences.
+  regions to be \change{defined}{specified} and \remove{add} additional
+  sequence to \add{be used} fill \add{known} gaps in the scaffold. Defining
+  the genome construction in a file makes the scaffolding process reproducible
+  and easier to edit compared with \add{large} FASTA nucleotide sequences.
 

### 2011-08-15 src/article.tex
Revise Background

@@
 large numbers of overlapping reads \add{can lead to} genome assembly software
-may be unable to produce a complete sequence. Instead, the algorithm may
-generate several large assembled regions of sequence (`contigs') composed of
-the many individual reads\remove{ that represent assembly into the longest
-possible sequences, given the data. Nonetheless, }These contigs represent
-a fragmented picture of the genome and \add{therefore} require additional work
-to join together into a complete sequence. \pb
+\remove{may} being unable to produce a complete sequence. Instead, the
+algorithm may generate several large assembled regions of sequence (`contigs')
+composed of the many individual reads\remove{ that represent assembly into the
+longest possible sequences, given the data. Nonetheless, }These contigs
+represent a fragmented picture of the genome and \add{therefore} require
+additional work to join together into a complete sequence. \pb
 
@@
 The process of \change{resolving a genome sequence and building a}{finishing
-a genome} scaffold \change{may} use wet laboratory methods, \emph{in silico}
+a genome} scaffold \remove{may} uses wet laboratory methods, \emph{in silico}
 methods, or a combination of both. An example of a computational method might
@@
 distance between these contigs. Alternatively, laboratory methods may use PCR
-to amplify the unknown DNA in a gap region, then use \remove{traditional}
+to amplify the unknown DNA in a gap region then use \remove{traditional}
 Sanger sequencing to determine the sequence of this gap. Computational methods
@@
 to infer contig placement and build the contigs into a scaffold
-\cite{richter2007,zhao2008,assefa2009}. Recombination within each of the two
-compared genomes can, however, reduce the efficacy of this. \pb
+\cite{richter2007,zhao2008,assefa2009}. \add{Genomic} recombination
+\remove{within each of the two compared genomes} can however reduce the
+\change{efficiency}{efficacy} of this. \pb
 
@@
 short sequence or being ignored by more conservative assembly algorithms. Such
-regions can be resolved by using algorithms \add{which} specifically reassemble
-the collapsed repeat region correctly \cite{mulyukov2002,koren2010}.
+regions can be resolved by using algorithms \add{that} specifically reassemble
+the collapsed repeat region \remove{correctly} \cite{mulyukov2002,koren2010}.
 \change{while a}{A} related approach uses unassembled sequence reads matching

### 2011-08-15 src/article.tex
Revise implementation

@@
 
-\subsection*{Input sequence data and scaffolding process} %{{{2
-
 The input data for Scaffolder is nucleotide sequence in FASTA format file.
 These nucleotide sequences can be of any length and for example may be
-individual sequences, assembled contigs or contigs which have been joined to
-into larger scaffolds. The case in which Scaffolder may be most useful is
-using the contigs and scaffolded contigs produced from the previously
-assembled set of sequencing reads, combined with additional gap filling
-sequences produced by PCR or \emph{in silico} methods as outlined in the
-Background. \pb
+individual \change{sequences}{reads}, assembled contigs or contigs which have
+been joined to into larger scaffolds. The case in which Scaffolder may be most
+useful is using the contigs and scaffolded contigs produced from the
+previously assembled set of sequencing reads, combined with additional gap
+filling sequences produced by PCR or \emph{in silico} methods as outlined in
+the Background. \pb
 

### 2011-08-15 src/article.tex
Revise Results section

@@
 specifying which contigs are required, the scaffold file allows the contigs to
-be edited into smaller sub-sequences or reverse complemented if necessary.\pb
+be edited into smaller sub-sequences or reverse complemented if necessary.
+\add{Each scaffold file represents one scaffolded nucleotide sequence and
+separate scaffolds should be defined in separate files.} \pb
 
@@
 oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
-\remove{and} The Scaffolder \change{allows}{format provides for using} these
-additional insert sequences to \remove{be used} to fill \remove{these} gaps.
-These inserts can also be treated in the same manner as larger contig
+\remove{and} The Scaffolder \change{allows}{format provides functionality to
+use} these additional insert sequences \remove{be used} to fill \remove{these}
+gaps. These inserts can also be treated in the same manner as larger contig
 sequences: trimmed and/or reverse complemented to match the corresponding gap
@@
 \remove{stored in the FASTA file} are referenced in the scaffold using the
-first word from in the FASTA header of the corresponding sequence. Maintaining
-the nucleotide sequences in a separate file preserves the unedited sequence
-and decouples the data from the specification of how it should be used to
-produce the genome sequence. \pb
+first word from \remove{in} the FASTA header of the corresponding sequence.
+Maintaining the nucleotide sequences in a separate file preserves the unedited
+sequence and decouples the data from the specification of how it should be
+used to produce the genome sequence. \pb
 
@@
 where each entry corresponds to a region of sequence in the generated scaffold
-super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
-sequence and separate scaffolds should be defined in separate files.} \pb
+super-sequence. \pb
 
@@
 in green) begins with a dash character `-' to denote an entry in the YAML
-list. This is a requirement of the YAML format: that each entry begins with
-a dash line. The next line is indented by two spaces where whitespace is
-required to group similar attributes together. The ``sequence'' tag indicates
-that this entry corresponds to a sequence in the FASTA file and the following
-line indicates the name of this sequence using the ``source'' tag. The first
-word of the FASTA header is used to identify which sequence is selected from
-the file. Together these three lines describe the first entry in the scaffold
-as a simple sequence using the sequence named `sequence1'. On the right hand
-side of Figure 1 this produces the first region in the scaffold, also shown in green. \pb
+list. This is a requirement of the YAML format: \remove{that} each entry
+begins with a dash line. The next line is indented by two spaces where
+whitespace is \change{required}{used} to group similar attributes together.
+The ``sequence'' tag indicates that this entry corresponds to a sequence in
+the FASTA file and the following line indicates the name of this sequence
+using the ``source'' tag. The first word of the FASTA header is used to
+identify which sequence is selected from the file. Together these three lines
+describe the first entry in the scaffold as a simple sequence using the
+sequence named `sequence1'. On the right hand side of Figure 1 this produces
+the first region in the scaffold, also shown in green. \pb
 
@@
 trim the sequence to these coordinates inclusively.\remove{while} The
-``reverse'' tag instructs scaffolder to also reverse complement the sequence.
-In the putative scaffold shown in Figure 1 this completes the \remove{scaffold
-} sequence. \pb
+``reverse'' tag \add{also} instructs scaffolder to \remove{also} reverse
+complement the sequence. In the putative scaffold shown in Figure 1 this
+completes the \remove{scaffold } sequence. \pb
 
-The final entry in the scaffold \remove{also} includes the ``inserts'' tag to
-add additional regions of sequence. These inserts are also added as a YAML
-list, with each insert beginning with a dash. The first insert, shown in
-purple, uses similar attributes to that of a sequence entry; the reverse,
+The final entry in the scaffold \change{also includes}{uses} the ``inserts''
+tag to add additional regions of sequence. These inserts are also added as
+a YAML list, with each insert beginning with a dash. The first insert, shown
+in purple, uses similar attributes to that of a sequence entry; the reverse,
 start and stop tags are used to trim and reverse complement the insert.

### 2011-08-15 src/article.tex
Revise discussion

@@
 laboratory methods. The process of manually editing a scaffold is inherently
-hard to reproduce and introduces \remove{the possibility of }irreducible
-\change{techniques}{edits} and/or human error. In respect to \change{these
-drawbacks}{this} the aims of the Scaffolder software are twofold: 1) to provide
-software that is easy to install and simplifies the task of genome finishing;
-and 2) to facilitate reproducibility in the scaffolding and finishing stage of
-a genome project. \pb
-
-\subsection*{Comparison with other scaffold formats} %{{{2
+hard to reproduce and introduces \change{the possibility of irreducible of
+techniques}{irreproducible edits} and/or human error. In respect to
+\change{these drawbacks}{this} the aims of the Scaffolder software are
+twofold: 1) to provide software that is easy to install and simplifies the
+task of genome finishing; and 2) to facilitate reproducibility in the
+scaffolding and finishing stage of a genome project. \pb
 
@@
 \add{AGP is a similar format for describing scaffolds. This format can be used
-to define contig order and N-filled gap regions in a scaffold. The advantages
-of the AGP format is that each contig entry is defined on a single line which
-allows searching the scaffold using UNIX line-based tools. The scaffolder
-format in comparison in contrast is written in the standardised YAML format
-and therefore accessible to the many languages which provide parsers to this
-format. The scaffolder format is provided with a tool explicitly to produce
-the fasta sequence of scaffold specified by in the file.} \pb
+to describe contig order and N-filled gap regions in a scaffold. The
+advantages of the AGP format is that each contig entry is defined on a single
+line which allows searching the scaffold using UNIX line-based tools. The
+scaffolder format in comparison in contrast is written in the standardised
+YAML format and therefore accessible to the many languages which provide
+parsers to this format. The scaffolder format is provided with a tool
+explicitly to produce the fasta sequence of scaffold specified by in the file.
+The scaffold format further provides functions for trimming and replacing
+regions of sequence using inserts.} \pb
 
@@
 
-\subsection*{Example Use Case} %{{{2
-
-\add{The example use case for scaffolder is a combination of computation and
-manual editing of a genome scaffold. We have used scaffolder in our genome
-projects to create a scaffold file from parsing the output of \emph{in silico}
-scaffolding tools into the scaffold file format. This scaffold was then
-manually updated in a text editor to fill gap regions with laboratory-produced
-sequences.} \pb
+\add{An example use case for scaffolder is a combination of computational and
+manual editing of a genome scaffold. We have used scaffolder in our own genome
+projects to create an initial scaffold from compuationally parsing the output
+of \emph{in silico} scaffolding tools into YAML. This scaffold was then
+manually updated as the scaffold was finished with additional gap-filling
+sequences generated in the laboratory.} \pb
 

### 2011-08-15 src/article.tex

@@
 manual editing of a genome scaffold. We have used scaffolder in our own genome
-projects to create an initial scaffold from compuationally parsing the output
+projects to create an initial scaffold from computationally parsing the output
 of \emph{in silico} scaffolding tools into YAML. This scaffold was then

### 2011-08-15 src/article.tex
Capitalise Scaffolder consistently

--- 2011-08-15 src/article.tex
@@
 specification of unknown regions. This syntax allows a scaffold to be updated
-by simply editing the scaffold file. As such, scaffolder facilitates
+by simply editing the scaffold file. As such, Scaffolder facilitates
 a reproducible \change{scaffolding}{finishing} process and provides a concise
@@
 1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
-scaffolder package is split into two libraries. The first called ``scaffolder''
+Scaffolder package is split into two libraries. The first called ``scaffolder''
 which provides the core Scaffolder application programming interface (API). The
@@
 scaffold files written in \remove{the} YAML \remove{syntax} specify how these
-sequences should be joined\change{. and the}{The} scaffolder software is used
+sequences should be joined\change{. and the}{The} Scaffolder software is used
 to generate the scaffold sequence from these instructions. In addition to
@@
 trim the sequence to these coordinates inclusively.\remove{while} The
-``reverse'' tag \add{also} instructs scaffolder to \remove{also} reverse
+``reverse'' tag \add{also} instructs Scaffolder to \remove{also} reverse
 complement the sequence. In the putative scaffold shown in Figure 1 this
@@
 The use of both these Scaffolder interfaces is outlined in detail on the
-scaffolder website (\scaffolder). \add{This website provides a ``getting
+Scaffolder website (\scaffolder). \add{This website provides a ``getting
 started'' guide as an introduction to using Scaffolder to build a genome
@@
 line which allows searching the scaffold using UNIX line-based tools. The
-scaffolder format in comparison in contrast is written in the standardised
+Scaffolder format in comparison in contrast is written in the standardised
 YAML format and therefore accessible to the many languages which provide
-parsers to this format. The scaffolder format is provided with a tool
+parsers to this format. The Scaffolder format is provided with a tool
 explicitly to produce the fasta sequence of scaffold specified by in the file.
@@
 
-\add{An example use case for scaffolder is a combination of computational and
-manual editing of a genome scaffold. We have used scaffolder in our own genome
+\add{An example use case for Scaffolder is a combination of computational and
+manual editing of a genome scaffold. We have used Scaffolder in our own genome
 projects to create an initial scaffold from compuationally parsing the output

### 2011-08-15 src/article.tex
Add sentence to discussion

@@
 
-\add{An example use case for Scaffolder is a combination of computational and
-manual editing of a genome scaffold. We have used Scaffolder in our own genome
+\add{A use case for Scaffolder is a combination of computational and manual
+editing of a genome scaffold. We have used Scaffolder in our own genome
 projects to create an initial scaffold from compuationally parsing the output
@@
 manually updated as the scaffold was finished with additional gap-filling
-sequences generated in the laboratory.} \pb
+sequences generated in the laboratory. This is example of the scaffolder
+format being both compuationally tractable while being simple to edit
+manually.} \pb
 

### 2011-09-11 src/article.tex
Use all track changes
Add akron university affliation

@@
 \usepackage{multicol}   %Columns
-\usepackage{trackchanges}
+\usepackage[finalnew]{trackchanges}
 \usepackage[utf8]{inputenc} %unicode support
@@
 \author{
-  Michael D Barton$^{1}$\correspondingauthor%
+  Michael D Barton$^{1,2}$\correspondingauthor%
   \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
 \and
-  Hazel A Barton$^1$%
+  Hazel A Barton$^{1,2}$%
   \email{Hazel A Barton - bartonh@nku.edu}%
@@
 
-\address{\iid(1) Department of Biological Sciences, Northern Kentucky %
-University, Nunn Drive, Highland Heights, KY 41076 }%
+\address{
+  \iid(1) Biology Department, The University of Akron, Akron, %
+  OH 44325-3908, USA \\
+  \iid(2) Department of Biological Sciences, Northern Kentucky %
+  University, Nunn Drive, Highland Heights, KY 41076, USA }%
 

### 2011-09-12 src/article.tex
Make list of abbreviations into paragraph
Add approval statement
Merge in all track changes

@@
 
-\title{Scaffolder - Software for \change{Reproducible}{Manual} Genome
+\title{Scaffolder - Software for Manual Genome
 Scaffolding.}
@@
 \iid(2) Department of Biological Sciences, Northern Kentucky %
-  University, Nunn Drive, Highland Heights, KY 41076, USA }%
+University, Nunn Drive, Highland Heights, KY 41076, USA}
 
@@
 
-  \paragraph*{Background:} \add{The} assembly of \add{next-generation}
-  short-read sequencing data can result in a fragmented non-contiguous
-  \change{series}{set} of genomic sequences. Therefore a common step in
-  a genome project is to join neighbouring sequence regions together and fill
-  gaps \remove{in the assembly using additional sequences}. This scaffolding
-  step\remove{, however,} is non-trivial and requires manually editing large
-  blocks of nucleotide sequence. Joining these sequences together also hides
-  the source of each region in the final genome sequence. Taken together,
-  these considerations may make reproducing or editing an existing genome
-  \change{build}{scaffold} difficult.
+  \paragraph*{Background:} The assembly of next-generation short-read
+  sequencing data can result in a fragmented non-contiguous set of genomic
+  sequences. Therefore a common step in a genome project is to join
+  neighbouring sequence regions together and fill gaps . This scaffolding step
+  is non-trivial and requires manually editing large blocks of nucleotide
+  sequence. Joining these sequences together also hides the source of each
+  region in the final genome sequence. Taken together, these considerations
+  may make reproducing or editing an existing genome scaffold difficult.
 
@@
   implemented in the Ruby programming language and can be installed via the
-  RubyGems software management system. Genome scaffolds are defined using YAML
-  - a data format\remove{,} which is both human and machine-readable. Command
-  line binaries and extensive documentation are available.
+  RubyGems software management system. Genome scaffolds are defined using
+  YAML - a data format which is both human and machine-readable. Command line
+  binaries and extensive documentation are available.
 
-  \paragraph*{Results:} This software allows a genome \change{scaffold}{build}
-  to be defined in terms of the constituent sequences using a relatively simple
-  syntax\remove{to define the scaffold}. This syntax further allows unknown
-  regions to be \change{defined}{specified} and \remove{add} additional
-  sequence to \add{be used} fill \add{known} gaps in the scaffold. Defining
-  the genome construction in a file makes the scaffolding process reproducible
-  and easier to edit compared with \add{large} FASTA nucleotide sequences.
+  \paragraph*{Results:} This software allows a genome build to be defined in
+  terms of the constituent sequences using a relatively simple syntax. This
+  syntax further allows unknown regions to be specified and additional sequence
+  to be used fill known gaps in the scaffold. Defining the genome construction
+  in a file makes the scaffolding process reproducible and easier to edit
+  compared with large FASTA nucleotide sequences.
 
   \paragraph*{Conclusions:} Scaffolder is easy-to-use genome scaffolding
-  software\change{. This tool}{ which} promotes reproducibility and continuous
-  development in a genome project. Scaffolder can be found at \scaffolder.
+  software which promotes reproducibility and continuous development in a
+  genome project. Scaffolder can be found at \scaffolder.
 
@@
 sequence reads from a genome. At the time of writing, high-throughput
-sequencing is limited to producing reads less than 1,000 nucleotides in length.
-Therefore to resolve a sequence longer than this, such as a complete genome,
-these numerous smaller fragments must be pieced together. The process of
-joining reads into longer sequences is the `assembly' stage of a genome project
-\cite{miller2010}. \pb
+sequencing is limited to producing reads less than 1,000 nucleotides in
+length. Therefore to resolve a sequence longer than this, such as a complete
+genome, these numerous smaller fragments must be pieced together. The process
+of joining reads into longer sequences is the `assembly' stage of a genome
+project \cite{miller2010}. \pb
 
@@
 is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
-\remove{within the genome, }or biased and incomplete sequencing data, may
-prevent the genome being assembled into a continuous sequence. This may be due
-to \change{not enough}{insufficient}, or multiple different overlaps between
-reads and is analogous to missing pieces in the jigsaw or pieces that fit to
-multiple other pieces. \pb
+or biased and incomplete sequencing data, may prevent the genome being
+assembled into a continuous sequence. This may be due to insufficient, or
+multiple different overlaps between reads and is analogous to missing pieces
+in the jigsaw or pieces that fit to multiple other pieces. \pb
 
-The advent of high\add{-}throughput sequencing methods has led to a renewed
-interest in algorithms to solve the problem of genome assembly
-\cite{pop2008,pop2009}. \remove{Nonetheless, due to }The complexity of merging
-large numbers of overlapping reads \add{can lead to} genome assembly software
-\remove{may} being unable to produce a complete sequence. Instead, the
-algorithm may generate several large assembled regions of sequence (`contigs')
-composed of the many individual reads\remove{ that represent assembly into the
-longest possible sequences, given the data. Nonetheless, }These contigs
-represent a fragmented picture of the genome and \add{therefore} require
-additional work to join together into a complete sequence. \pb
+The advent of high-throughput sequencing methods has led to a renewed interest
+in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
+The complexity of merging large numbers of overlapping reads can lead to
+genome assembly software being unable to produce a complete sequence. Instead,
+the algorithm may generate several large assembled regions of sequence
+(`contigs') composed of the many individual readsThese contigs represent a
+fragmented picture of the genome and therefore require additional work to join
+together into a complete sequence. \pb
 
-The process of \change{ameliorating a draft sequence into a finished continuous
-sequence}{finishing a genome sequence} can be expensive in terms of time and
-laboratory effort. \change{although}{In some cases} the genomic data present in
-a set of generated contigs may be sufficient for many research questions
-\cite{branscomb2002}. Nevertheless, a continuous high-quality `finished' genome
-sequence does provide a greater depth of information, such as complete
-resolution of repeat regions and precise estimates of distances between genomic
-elements \cite{parkhill2002,fraser2002}. The process of joining these contigs
-together to form a continuous genome sequence is called the `scaffolding' or
-`finishing' stage and is the focus of the software described in this article.
-\pb
+The process of finishing a genome sequence can be expensive in terms of time
+and laboratory effort. In some cases the genomic data present in a set of
+generated contigs may be sufficient for many research questions
+\cite{branscomb2002}. Nevertheless, a continuous high-quality `finished'
+genome sequence does provide a greater depth of information, such as complete
+resolution of repeat regions and precise estimates of distances between
+genomic elements \cite{parkhill2002,fraser2002}. The process of joining these
+contigs together to form a continuous genome sequence is called the
+`scaffolding' or `finishing' stage and is the focus of the software described
+in this article. \pb
 
@@
 
-Scaffolding is the process of joining a series of disconnected contigs into
-a complete continuous genome sequence. Due to genomic complexity and missing
+Scaffolding is the process of joining a series of disconnected contigs into a
+complete continuous genome sequence. Due to genomic complexity and missing
 data, scaffolding may not ultimately produce a final completed sequence, but
@@
 requires reverse complementing sequences where necessary. In the case of
-\change{microbial}{archaeal and bacterial} genomes this orientation will be to
-the 5'~$\rightarrow$~3' direction following the direction of genome
-replication.
+archaeal and bacterial genomes this orientation will be to the
+5'~$\rightarrow$~3' direction following the direction of genome replication.
 
@@
 Filling these regions with unknown nucleotide characters (`N') allows a draft
-continuous sequence. \change{which}{This sequence} is useful for representing
-both the known and to-be-resolved areas in the genome sequence.
+continuous sequence. This sequence is useful for representing both the known
+and to-be-resolved areas in the genome sequence.
 
@@
 to the laboratory to perform additional sequencing or using computational
-methods of estimating the unknown sequence. This additional sequence is
-\remove{then }used to replace the gap between two contigs, joining them into
-a single sequence. Once all contigs have been joined and gaps in a scaffold
-closed, the genome may be considered finished.
+methods of estimating the unknown sequence. This additional sequence is used
+to replace the gap between two contigs, joining them into a single sequence.
+Once all contigs have been joined and gaps in a scaffold closed, the genome
+may be considered finished.
 
@@
 
-The process of \change{resolving a genome sequence and building a}{finishing
-a genome} scaffold \remove{may} uses wet laboratory methods, \emph{in silico}
-methods, or a combination of both. An example of a computational method might
-use the \remove{existing} paired-read data from the sequencing stage.
-\remove{where reads were produced in pairs at a known distance apart in the
-genome.} The occurrence of paired reads in separate contigs can be used to
-\change{probabilistically estimate}{estimate probabilistically} the order and
-distance between these contigs. Alternatively, laboratory methods may use PCR
-to amplify the unknown DNA in a gap region then use \remove{traditional}
-Sanger sequencing to determine the sequence of this gap. Computational methods
-are \remove{more} preferable as they are less costly in laboratory time and
-materials compared to manual gap resolution, and use the available data that
-exists following sequencing \cite{nagarajan2010}. Finally when the scaffold
-cannot be completely resolved, \emph{in silico} software packages exist to
-suggest the likely primers necessary for PCR amplifying the sequence in gap
-regions \cite{gordon2001}. \pb
+The process of finishing a genome scaffold uses wet laboratory methods,
+\emph{in silico} methods, or a combination of both. An example of a
+computational method might use the paired-read data from the sequencing stage.
+The occurrence of paired reads in separate contigs can be used to estimate
+probabilistically the order and distance between these contigs. Alternatively,
+laboratory methods may use PCR to amplify the unknown DNA in a gap region then
+use Sanger sequencing to determine the sequence of this gap. Computational
+methods are preferable as they are less costly in laboratory time and materials
+compared to manual gap resolution, and use the available data that exists
+following sequencing \cite{nagarajan2010}. Finally when the scaffold cannot be
+completely resolved, \emph{in silico} software packages exist to suggest the
+likely primers necessary for PCR amplifying the sequence in gap regions
+\cite{gordon2001}. \pb
 
-Examples of \emph{in silico} methods include \remove{using synteny to}
-comparing the assembled contigs to a complete reference genome sequence
-\change{by}{to} search\remove{ing} for areas of sequence similarity between the
-two. Any areas of corresponding sequence with the reference genome can be used
-to infer contig placement and build the contigs into a scaffold
-\cite{richter2007,zhao2008,assefa2009}. \add{Genomic} recombination
-\remove{within each of the two compared genomes} can however reduce the
-\change{efficiency}{efficacy} of this. \pb
+Examples of \emph{in silico} methods include comparing the assembled contigs
+to a complete reference genome sequence to search for areas of sequence
+similarity between the two. Any areas of corresponding sequence with the
+reference genome can be used to infer contig placement and build the contigs
+into a scaffold \cite{richter2007,zhao2008,assefa2009}. Genomic recombination
+can however reduce the efficacy of this. \pb
 
-Repeat regions may also be responsible for multiple gaps when building
-a genome sequence; tandemly repeated nucleotide regions in the genome produce
+Repeat regions may also be responsible for multiple gaps when building a
+genome sequence; tandemly repeated nucleotide regions in the genome produce
 multiple reads with similar sequence. As many assembly algorithms rely on
@@
 short sequence or being ignored by more conservative assembly algorithms. Such
-regions can be resolved by using algorithms \add{that} specifically reassemble
-the collapsed repeat region \remove{correctly} \cite{mulyukov2002,koren2010}.
-\change{while a}{A} related approach uses unassembled sequence reads matching
-the regions around a scaffold gap to construct a uniquely overlapping set of
-reads across \change{it}{the gap}. \cite{tsai2010}. \pb
+regions can be resolved by using algorithms that specifically reassemble the
+collapsed repeat region \cite{mulyukov2002,koren2010}. A related approach uses
+unassembled sequence reads matching the regions around a scaffold gap to
+construct a uniquely overlapping set of reads across the gap. \cite{tsai2010}.
+\pb
 
@@
 that matches these paired-read distances \cite{dayarian2010,boetzer2011}.
-Synteny data from a reference genome can also be combined with this
-paired-read data to estimate the best contig configuration \cite{pop2004}. \pb
+Synteny data from a reference genome can also be combined with this paired-read
+data to estimate the best contig configuration \cite{pop2004}. \pb
 
 These described \emph{in silico} methods provide a wide array of approaches for
-merging contigs into a larger, continuous scaffold sequence.
-\remove{Nevertheless} The scaffolding process may still require manually
-inserting additional sequences or further joining contigs using PCR-derived
-sequence. Moving and editing large blocks of nucleotide text by hand however
-\remove{possibly} introduces human error and precludes any reproducibility of
-the steps in this process. \pb
+merging contigs into a larger, continuous scaffold sequence. The scaffolding
+process may still require manually inserting additional sequences or further
+joining contigs using PCR-derived sequence. Moving and editing large blocks of
+nucleotide text by hand however introduces human error and precludes any
+reproducibility of the steps in this process. \pb
 
 The software outlined here, ``Scaffolder,'' aims to address these problems of
-reproducibility by creating a file syntax and software framework for editing
-a genome scaffold. Scaffolder uses a specific file format to define how
-contigs are joined, additional sequences are inserted, and for the
-specification of unknown regions. This syntax allows a scaffold to be updated
-by simply editing the scaffold file. As such, Scaffolder facilitates
-a reproducible \change{scaffolding}{finishing} process and provides a concise
-overview of how the final genomic scaffold was constructed. \pb
+reproducibility by creating a file syntax and software framework for editing a
+genome scaffold. Scaffolder uses a specific file format to define how contigs
+are joined, additional sequences are inserted, and for the specification of
+unknown regions. This syntax allows a scaffold to be updated by simply editing
+the scaffold file. As such, Scaffolder facilitates a reproducible finishing
+process and provides a concise overview of how the final genomic scaffold was
+constructed. \pb
 
@@
 
-Scaffolder is written in the Ruby programming language \change{using version
-1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
-Scaffolder package is split into two libraries. The first called ``scaffolder''
-which provides the core Scaffolder application programming interface (API). The
-second library ``scaffolder-tools'' provides the Scaffolder command line
-interface (CLI). \pb
+Scaffolder is written in the Ruby programming language and tested against
+versions 1.8.7 and 1.9.2 \cite{ruby-lang}. The Scaffolder package is split into
+two libraries. The first called ``scaffolder'' which provides the core
+Scaffolder application programming interface (API). The second library
+``scaffolder-tools'' provides the Scaffolder command line interface (CLI). \pb
 
@@
 should therefore be simple to update manually in addition to being
-computationally tractable. This requirement was also best suited to YAML
-syntax which is human-readable and simple to edit in a standard text editor.
-\pb
+computationally tractable. This requirement was also best suited to YAML syntax
+which is human-readable and simple to edit in a standard text editor. \pb
 
@@
 These nucleotide sequences can be of any length and for example may be
-individual \change{sequences}{reads}, assembled contigs or contigs which have
-been joined to into larger scaffolds. The case in which Scaffolder may be most
-useful is using the contigs and scaffolded contigs produced from the
-previously assembled set of sequencing reads, combined with additional gap
-filling sequences produced by PCR or \emph{in silico} methods as outlined in
-the Background. \pb
-
-\remove{The scaffolder algorithm reads through the scaffolder file parsing.
-Each entry is then converted to corresponding nucleotide sequence by fetching
-from the FASTA file and then performing any designated sequence trimming or
-reverse complementing. Each of the sequence entries are then joined into
-a continuous single super-sequence and returned to the console output.}
+individual reads, assembled contigs or contigs which have been joined to into
+larger scaffolds. The case in which Scaffolder may be most useful is using the
+contigs and scaffolded contigs produced from the previously assembled set of
+sequencing reads, combined with additional gap filling sequences produced by
+PCR or \emph{in silico} methods as outlined in the Background. \pb
 
@@
 together into a single contiguous scaffolded super-sequence. Plain-text
-scaffold files written in \remove{the} YAML \remove{syntax} specify how these
-sequences should be joined\change{. and the}{The} Scaffolder software is used
-to generate the scaffold sequence from these instructions. In addition to
-specifying which contigs are required, the scaffold file allows the contigs to
-be edited into smaller sub-sequences or reverse complemented if necessary.
-\add{Each scaffold file represents one scaffolded nucleotide sequence and
-separate scaffolds should be defined in separate files.} \pb
+scaffold files written in YAML specify how these sequences should be joinedThe
+Scaffolder software is used to generate the scaffold sequence from these
+instructions. In addition to specifying which contigs are required, the
+scaffold file allows the contigs to be edited into smaller sub-sequences or
+reverse complemented if necessary. Each scaffold file represents one scaffolded
+nucleotide sequence and separate scaffolds should be defined in separate files.
+\pb
 
 The process of genome finishing may involve producing additional
-oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
-\remove{and} The Scaffolder \change{allows}{format provides functionality to
-use} these additional insert sequences \remove{be used} to fill \remove{these}
+oligonucleotide sequences to fill unknown regions in a scaffold. The Scaffolder
+format provides functionality to use these additional insert sequences to fill
 gaps. These inserts can also be treated in the same manner as larger contig
@@
 mapping the contigs to a reference genome. These inter-contig gap regions are
-useful to join separate sequences together by \change{an}{the} estimated
-\change{size and }{distance.} The scaffold file allows for the specification
-of such unresolved regions by inserting regions of `N' nucleotides into the
-scaffold. The use of these regions in the scaffold indicates the unresolved
-regions in the build and their approximate size. \pb
+useful to join separate sequences together by the estimated distance. The
+scaffold file allows for the specification of such unresolved regions by
+inserting regions of `N' nucleotides into the scaffold. The use of these
+regions in the scaffold indicates the unresolved regions in the build and their
+approximate size. \pb
 
-\change{The scaffold file specifies how sequence regions are joined together;
-however, the sequences themselves are no  stored in the scaffold file and are
-instead}{The nucleotide sequences used in the scaffold are} maintained as
-a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
-the standard output for a genome assembler, and these may be contigs, sets of
-contigs scaffolded into larger sequences, or both. }The nucleotide sequences
-\remove{stored in the FASTA file} are referenced in the scaffold using the
-first word from \remove{in} the FASTA header of the corresponding sequence.
-Maintaining the nucleotide sequences in a separate file preserves the unedited
-sequence and decouples the data from the specification of how it should be
-used to produce the genome sequence. \pb
+The nucleotide sequences used in the scaffold are maintained as a separate
+FASTA file. The nucleotide sequences are referenced in the scaffold using the
+first word from the FASTA header of the corresponding sequence. Maintaining the
+nucleotide sequences in a separate file preserves the unedited sequence and
+decouples the data from the specification of how it should be used to produce
+the genome sequence. \pb
 
@@
 The scaffold file is written using the YAML syntax and an example is shown in
-Figure 1. This file illustrates the text attributes used to describe
-a scaffold and how the sequences are correspondingly joined together in the
-genome build. The basic layout of the scaffold file is a list of entries,
-where each entry corresponds to a region of sequence in the generated scaffold
-super-sequence. \pb
+Figure 1. This file illustrates the text attributes used to describe a scaffold
+and how the sequences are correspondingly joined together in the genome build.
+The basic layout of the scaffold file is a list of entries, where each entry
+corresponds to a region of sequence in the generated scaffold super-sequence.
+\pb
 
@@
 indicate the start of a YAML-formatted document. The first entry (highlighted
-in green) begins with a dash character `-' to denote an entry in the YAML
-list. This is a requirement of the YAML format: \remove{that} each entry
-begins with a dash line. The next line is indented by two spaces where
-whitespace is \change{required}{used} to group similar attributes together.
-The ``sequence'' tag indicates that this entry corresponds to a sequence in
-the FASTA file and the following line indicates the name of this sequence
-using the ``source'' tag. The first word of the FASTA header is used to
-identify which sequence is selected from the file. Together these three lines
-describe the first entry in the scaffold as a simple sequence using the
-sequence named `sequence1'. On the right hand side of Figure 1 this produces
-the first region in the scaffold, also shown in green. \pb
+in green) begins with a dash character `-' to denote an entry in the YAML list.
+This is a requirement of the YAML format: each entry begins with a dash line.
+The next line is indented by two spaces where whitespace is used to group
+similar attributes together. The ``sequence'' tag indicates that this entry
+corresponds to a sequence in the FASTA file and the following line indicates
+the name of this sequence using the ``source'' tag. The first word of the FASTA
+header is used to identify which sequence is selected from the file. Together
+these three lines describe the first entry in the scaffold as a simple sequence
+using the sequence named `sequence1'. On the right hand side of Figure 1 this
+produces the first region in the scaffold, also shown in green. \pb
 
@@
 `sequence2' to the scaffold. This entry demonstrates how this sequence may be
-manipulated prior to addition to the scaffold. The `start' and `stop' tags
-trim the sequence to these coordinates inclusively.\remove{while} The
-``reverse'' tag \add{also} instructs Scaffolder to \remove{also} reverse
-complement the sequence. In the putative scaffold shown in Figure 1 this
-completes the \remove{scaffold } sequence. \pb
+manipulated prior to addition to the scaffold. The `start' and `stop' tags trim
+the sequence to these coordinates inclusively. The ``reverse'' tag also
+instructs Scaffolder to reverse complement the sequence. In the putative
+scaffold shown in Figure 1 this completes the sequence. \pb
 
-The final entry in the scaffold \change{also includes}{uses} the ``inserts''
-tag to add additional regions of sequence. These inserts are also added as
-a YAML list, with each insert beginning with a dash. The first insert, shown
-in purple, uses similar attributes to that of a sequence entry; the reverse,
-start and stop tags are used to trim and reverse complement the insert.
-Similarly the `source' tag identifies the corresponding FASTA sequence as
-`insert1'. The ``open'' and ``close'' tags are specific to inserts and
-determine where the insert is added in the enclosing sequence. The region of
-the sequence inside these coordinates is inclusively replaced by the specified
-insert sequence. This is visualised in the putative scaffold in Figure 1 by
-the black lines bisecting the blue sequence. \pb
+The final entry in the scaffold uses the ``inserts'' tag to add additional
+regions of sequence. These inserts are also added as a YAML list, with each
+insert beginning with a dash. The first insert, shown in purple, uses similar
+attributes to that of a sequence entry; the reverse, start and stop tags are
+used to trim and reverse complement the insert. Similarly the `source' tag
+identifies the corresponding FASTA sequence as `insert1'. The ``open'' and
+``close'' tags are specific to inserts and determine where the insert is added
+in the enclosing sequence. The region of the sequence inside these coordinates
+is inclusively replaced by the specified insert sequence. This is visualised in
+the putative scaffold in Figure 1 by the black lines bisecting the blue
+sequence. \pb
 
@@
 addition Scaffolder provides a command line interface (CLI) to validate the
-scaffold file and build the draft super sequence. The Scaffolder CLI behaves
-as a standard Unix tool and returns appropriate exit codes and manual pages.
-The use of both these Scaffolder interfaces is outlined in detail on the
-Scaffolder website (\scaffolder). \add{This website provides a ``getting
-started'' guide as an introduction to using Scaffolder to build a genome
-scaffold.} \pb
+scaffold file and build the draft super sequence. The Scaffolder CLI behaves as
+a standard Unix tool and returns appropriate exit codes and manual pages. The
+use of both these Scaffolder interfaces is outlined in detail on the Scaffolder
+website (\scaffolder). This website provides a ``getting started'' guide as an
+introduction to using Scaffolder to build a genome scaffold. \pb
 
@@
 Scaffolding an incomplete genome assembly requires joining contigs and
-additional gap-filling sequence\add{s} using a combination of computational and
+additional gap-filling sequences using a combination of computational and
 laboratory methods. The process of manually editing a scaffold is inherently
-hard to reproduce and introduces \change{the possibility of irreducible of
-techniques}{irreproducible edits} and/or human error. In respect to
-\change{these drawbacks}{this} the aims of the Scaffolder software are
-twofold: 1) to provide software that is easy to install and simplifies the
-task of genome finishing; and 2) to facilitate reproducibility in the
-scaffolding and finishing stage of a genome project. \pb
+hard to reproduce and introduces irreproducible edits and/or human error. In
+respect to this the aims of the Scaffolder software are twofold: 1) to provide
+software that is easy to install and simplifies the task of genome finishing;
+and 2) to facilitate reproducibility in the scaffolding and finishing stage of
+a genome project. \pb
 
-\remove{Scaffolder was designed to be as simple to use and, assuming the Ruby
-and RubyGems software are present, can be installed in a single command line
-step. This should negate any possible barriers to entry which require manual
-compilation of source code.} Scaffolder \remove{also} uses a minimal and
-compact syntax to describe how \change{the}{a} genome scaffold sequence should
-be generated. This syntax is simple to \change{construct}{write} and edit
-whilst being succinct and readable. \remove{The goal of this syntax is to make
-genome scaffolds easy to write in a common text editor.}
+Scaffolder uses a minimal and compact syntax to describe how a genome scaffold
+sequence should be generated. This syntax is simple to write and edit whilst
+being succinct and readable.
 
-\add{AGP is a similar format for describing scaffolds. This format can be used
-to describe contig order and N-filled gap regions in a scaffold. The
-advantages of the AGP format is that each contig entry is defined on a single
-line which allows searching the scaffold using UNIX line-based tools. The
-Scaffolder format in comparison in contrast is written in the standardised
-YAML format and therefore accessible to the many languages which provide
-parsers to this format. The Scaffolder format is provided with a tool
-explicitly to produce the fasta sequence of scaffold specified by in the file.
-The scaffold format further provides functions for trimming and replacing
-regions of sequence using inserts.} \pb
+AGP is a similar format for describing scaffolds. This format can be used to
+describe contig order and N-filled gap regions in a scaffold. The advantages of
+the AGP format is that each contig entry is defined on a single line which
+allows searching the scaffold using UNIX line-based tools. The Scaffolder
+format in comparison in contrast is written in the standardised YAML format and
+therefore accessible to the many languages which provide parsers to this
+format. The Scaffolder format is provided with a tool explicitly to produce the
+fasta sequence of scaffold specified by in the file. The scaffold format
+further provides functions for trimming and replacing regions of sequence using
+inserts. \pb
 
@@
 
-\add{A use case for Scaffolder is a combination of computational and manual
-editing of a genome scaffold. We have used Scaffolder in our own genome
-projects to create an initial scaffold from compuationally parsing the output
-of \emph{in silico} scaffolding tools into YAML. This scaffold was then
-manually updated as the scaffold was finished with additional gap-filling
-sequences generated in the laboratory. This is example of the scaffolder
-format being both compuationally tractable while being simple to edit
-manually.} \pb
+A use case for Scaffolder is a combination of computational and manual editing
+of a genome scaffold. We have used Scaffolder in our own genome projects to
+create an initial scaffold from compuationally parsing the output of \emph{in
+silico} scaffolding tools into YAML. This scaffold was then manually updated as
+the scaffold was finished with additional gap-filling sequences generated in
+the laboratory. This is example of the scaffolder format being both
+compuationally tractable while being simple to edit manually. \pb
 
-The YAML text format \add{also} allows comparison of differences between
-scaffold builds using standard Unix tools such as diff. This \add{therefore}
-makes scaffold files \add{amenable} to storage in version control systems and
-allows genome finishers to use methods similar to those in software
-development. \pb
+The YAML text format also allows comparison of differences between scaffold
+builds using standard Unix tools such as diff. This therefore makes scaffold
+files amenable to storage in version control systems and allows genome
+finishers to use methods similar to those in software development. \pb
 
@@
 reproducibility in genome projects by allowing the scaffold super-sequence to
-be reliably reproduced from the same scaffold file. The YAML syntax for
-writing the scaffold file is also standardised and simple to manipulate
+be reliably reproduced from the same scaffold file. The YAML syntax for writing
+the scaffold file is also standardised and simple to manipulate
 programmatically. This thereby means the scaffolding process follows the Unix
-tenet of ``If your data structures are good enough, the algorithm to
-manipulate them should be trivial.''
+tenet of ``If your data structures are good enough, the algorithm to manipulate
+them should be trivial.''
 
@@
   \begin{description}
-    \item[Project name:] Scaffolder v0.4.\change{1}{4}, Scaffolder Tools
-    v0.1.\change{2}{3}
+    \item[Project name:] Scaffolder v0.4.4, Scaffolder Tools
+    v0.1.3
     \item[Project home page:] \scaffolder
@@
     Ubuntu.
-    \item[Programming language:] Ruby \add{1.8.7 or 1.9.2}
-    \item[Other requirements:] \change{RubyGems}{RubyGems package management
+    \item[Programming language:] Ruby 1.8.7 or 1.9.2
+    \item[Other requirements:] RubyGems package management
     software and the following libraries: BioRuby 1.4.x, confligliere
     0.1.x, ronn 0.7.x. A full list of development dependencies can be found in
-    the Gemfile in the base directory of each project.}
+    the Gemfile in the base directory of each project.
     \item[License:] MIT \item[Any restrictions to use by non-academics:] None
@@
 
-  \begin{description}
-    \item[API:] Application programming interface
-    \item[CLI:] Command line interface
-    \item[PCR:] Polymerase chain reaction
-    \item[YAML:] YAML ain't markup language\cite{yaml}
-  \end{description}
+API: Application programming interface, CLI: Command line interface, PCR:
+Polymerase chain reaction, YAML: YAML ain't markup language\cite{yaml}.
 
@@
 MDB developed and maintains the Scaffolder tool. MDB and HAB wrote the
-manuscript.
+manuscript. All authors have read and approved the manuscript.
 
@@
 the resulting putative scaffold sequence (right). The scaffold contains three
-entries and two inserts. Each entry in the scaffold file text is delimited by
-a `-' on a new line and highlighted using separate colours. The scaffold
-diagram on the right is not to scale and instead illustrates how the scaffold
-sequences are joined. \pb
+entries and two inserts. Each entry in the scaffold file text is delimited by a
+`-' on a new line and highlighted using separate colours. The scaffold diagram
+on the right is not to scale and instead illustrates how the scaffold sequences
+are joined. \pb
 

### 2011-09-12 src/article.tex
Proof read

@@
   sequence. Joining these sequences together also hides the source of each
-  region in the final genome sequence. Taken together, these considerations
+  region in the final genome sequence. Taken together these considerations
   may make reproducing or editing an existing genome scaffold difficult.
@@
   syntax further allows unknown regions to be specified and additional sequence
-  to be used fill known gaps in the scaffold. Defining the genome construction
+  to be used to fill known gaps in the scaffold. Defining the genome construction
   in a file makes the scaffolding process reproducible and easier to edit
@@
 the algorithm may generate several large assembled regions of sequence
-(`contigs') composed of the many individual readsThese contigs represent a
+(`contigs') composed of the many individual reads. These contigs represent a
 fragmented picture of the genome and therefore require additional work to join
@@
 use Sanger sequencing to determine the sequence of this gap. Computational
-methods are preferable as they are less costly in laboratory time and materials
-compared to manual gap resolution, and use the available data that exists
-following sequencing \cite{nagarajan2010}. Finally when the scaffold cannot be
-completely resolved, \emph{in silico} software packages exist to suggest the
-likely primers necessary for PCR amplifying the sequence in gap regions
-\cite{gordon2001}. \pb
+methods using the available sequencing data are preferable as they are less
+costly in laboratory time and materials compared to manual gap resolution
+\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
+\emph{in silico} software packages exist to suggest the likely primers
+necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb
 
@@
 to a complete reference genome sequence to search for areas of sequence
-similarity between the two. Any areas of corresponding sequence with the
+similarity between the two. Any areas of corresponding sequence in the
 reference genome can be used to infer contig placement and build the contigs
@@
 nucleotide text by hand however introduces human error and precludes any
-reproducibility of the steps in this process. \pb
+reproducibility. \pb
 
@@
 The choice of nucleotide sequences comprising the scaffold is specified using
-the YAML syntax \cite{yaml}. YAML is a language format using whitespace and
+the YAML syntax \cite{yaml}. YAML is a data format using whitespace and
 indentation to produce a machine readable structure. As YAML is a standardised
@@
 larger scaffolds. The case in which Scaffolder may be most useful is using the
-contigs and scaffolded contigs produced from the previously assembled set of
-sequencing reads, combined with additional gap filling sequences produced by
-PCR or \emph{in silico} methods as outlined in the Background. \pb
+contigs and scaffolded contigs, combined with additional gap filling sequences
+produced by PCR or \emph{in silico} methods as outlined in the Background. \pb
 
@@
 together into a single contiguous scaffolded super-sequence. Plain-text
-scaffold files written in YAML specify how these sequences should be joinedThe
-Scaffolder software is used to generate the scaffold sequence from these
+scaffold files written in YAML specify how these sequences should be joined.
+The Scaffolder software is used to generate the scaffold sequence from these
 instructions. In addition to specifying which contigs are required, the
@@
 sequence should be generated. This syntax is simple to write and edit whilst
-being succinct and readable.
+being succinct and readable. \pb
 
@@
 format. The Scaffolder format is provided with a tool explicitly to produce the
-fasta sequence of scaffold specified by in the file. The scaffold format
+FASTA sequence of scaffold specified by in the file. The Scaffolder format
 further provides functions for trimming and replacing regions of sequence using
@@
 contigs. The Scaffolder file format maintains the genome scaffold as a concise
-and readable text representation that allows third parties to see how the
+and readable text representation that allows third-parties to see how the
 genome sequence was scaffolded. This file format also allows a broad overview

### 2011-09-12 src/article.tex
Proof read again

@@
 The sequencing process generates reads from either strand of the DNA helix and
-the resulting contigs constructed from these reads may represent either the
-5'~$\rightarrow$~3' or 3'~$\rightarrow$~5' DNA strands with respect to the
-origin of replication. Orientating all contigs to point in the same direction
-requires reverse complementing sequences where necessary. In the case of
-archaeal and bacterial genomes this orientation will be to the
-5'~$\rightarrow$~3' direction following the direction of genome replication.
+the resulting contigs constructed from these reads may represent either DNA
+strand. Orientating all contigs to point in the same direction requires reverse
+complementing sequences where necessary. In the case of archaeal and bacterial
+genomes this orientation will be to the 5'~$\rightarrow$~3' direction following
+the direction of genome replication.
 
@@
 to the laboratory to perform additional sequencing or using computational
-methods of estimating the unknown sequence. This additional sequence is used
-to replace the gap between two contigs, joining them into a single sequence.
-Once all contigs have been joined and gaps in a scaffold closed, the genome
-may be considered finished.
+methods to estimate the unknown sequence. This additional sequence is used to
+replace the gap between two contigs, joining them into a single sequence. Once
+all contigs have been joined and gaps in a scaffold closed, the genome may be
+considered finished.
 
@@
 use Sanger sequencing to determine the sequence of this gap. Computational
-methods using the available sequencing data are preferable as they are less
+methods, using available sequencing data, are preferable as they are less
 costly in laboratory time and materials compared to manual gap resolution
@@
 Initial sequencing data assembly may result in an incomplete genome build.
-Generation of additional sequences either through PCR or computational methods
-means that genome scaffolding may be an on-going process. The scaffold file
-should therefore be simple to update manually in addition to being
-computationally tractable. This requirement was also best suited to YAML syntax
-which is human-readable and simple to edit in a standard text editor. \pb
+Adding further sequences from either PCR or computational methods also means
+that genome scaffolding may be an on-going process. The scaffold file should
+therefore be simple to update manually in addition to being computationally
+tractable. This requirement was also best suited to YAML syntax which is
+human-readable and simple to edit in a standard text editor. \pb
 
@@
 These nucleotide sequences can be of any length and for example may be
-individual reads, assembled contigs or contigs which have been joined to into
+individual reads, assembled contigs or contigs which have been joined into
 larger scaffolds. The case in which Scaffolder may be most useful is using the
@@
 these three lines describe the first entry in the scaffold as a simple sequence
-using the sequence named `sequence1'. On the right hand side of Figure 1 this
+using a FASTA entry named `sequence1'. On the right hand side of Figure 1 this
 produces the first region in the scaffold, also shown in green. \pb
@@
 
-The last entry in the scaffold, highlighted in blue, adds the sequence named
+The last entry in the scaffold, highlighted in blue, adds a FASTA entry named
 `sequence2' to the scaffold. This entry demonstrates how this sequence may be
@@
 
-The final entry in the scaffold uses the ``inserts'' tag to add additional
+This final entry in the scaffold uses the ``inserts'' tag to add additional
 regions of sequence. These inserts are also added as a YAML list, with each
@@
 The next insert, shown in brown, is specified using only the `open' tag. This
-illustrates that only one of either `open' or `close' tags is required when
+illustrates that only one of either the `open' or `close' tags is required when
 adding an insert sequence. If only one of the `open' or `close' tags is used
@@
 and 2) to facilitate reproducibility in the scaffolding and finishing stage of
-a genome project. \pb
-
-Scaffolder uses a minimal and compact syntax to describe how a genome scaffold
-sequence should be generated. This syntax is simple to write and edit whilst
-being succinct and readable. \pb
+a genome project. Scaffolder uses a minimal and compact syntax to describe how
+a genome scaffold sequence should be generated. This syntax is simple to write
+and edit whilst being succinct and readable. \pb
 
 AGP is a similar format for describing scaffolds. This format can be used to
-describe contig order and N-filled gap regions in a scaffold. The advantages of
+describe contig order and N-filled gap regions in a scaffold. The advantage of
 the AGP format is that each contig entry is defined on a single line which
 allows searching the scaffold using UNIX line-based tools. The Scaffolder
-format in comparison in contrast is written in the standardised YAML format and
-therefore accessible to the many languages which provide parsers to this
-format. The Scaffolder format is provided with a tool explicitly to produce the
-FASTA sequence of scaffold specified by in the file. The Scaffolder format
-further provides functions for trimming and replacing regions of sequence using
+format in contrast is written in the standardised YAML format and therefore
+accessible to the many languages which provide parsers to this format. The
+Scaffolder format is provided with a tool explicitly to produce the FASTA
+sequence of scaffold specified by the file. The Scaffolder format further
+provides functions for trimming and replacing regions of sequence using
 inserts. \pb
 
-Constructing a genome using by specifying the scaffold organisation in text
-file makes generating a scaffold super sequence both reproducible and
-deterministic for the same file and set of FASTA sequences. In comparison,
-joining large nucleotide sequences by hand cannot be reliably reproduced. The
-scaffold file furthermore provides a human readable description of how the
-scaffold is constructed. Configuring the final sequence in the scaffold file
-means the build is easier to edit once already constructed. \pb
+Constructing a genome by specifying the scaffold organisation in text file
+makes generating a scaffold super sequence both reproducible and deterministic
+for the same file and set of FASTA sequences. In comparison, joining large
+nucleotide sequences by hand cannot be reliably reproduced. The scaffold file
+furthermore provides a human readable description of how the scaffold is
+constructed. Configuring the final sequence in the scaffold file means the
+build is easier to edit once already constructed. \pb
 

### 2011-09-12 src/article.tex
Spell check

@@
 of a genome scaffold. We have used Scaffolder in our own genome projects to
-create an initial scaffold from compuationally parsing the output of \emph{in
+create an initial scaffold from computationally parsing the output of \emph{in
 silico} scaffolding tools into YAML. This scaffold was then manually updated as
@@
 the laboratory. This is example of the scaffolder format being both
-compuationally tractable while being simple to edit manually. \pb
+computationally tractable while being simple to edit manually. \pb
 

### 2011-09-13 src/article.tex
Update author details
Further proof read

@@
 \author{
-  Michael D Barton$^{1,2}$\correspondingauthor%
+  Michael D Barton$^{1}$\correspondingauthor%
   \email{Michael D Barton\correspondingauthor  - mail@michaelbarton.me.uk}%
 \and
-  Hazel A Barton$^{1,2}$%
-  \email{Hazel A Barton - bartonh@nku.edu}%
+  Hazel A Barton$^{1}$%
+  \email{Hazel A Barton - bartonh@uakron.edu}%
       }
@@
 \iid(1) Biology Department, The University of Akron, Akron, %
-OH 44325-3908, USA \\
-\iid(2) Department of Biological Sciences, Northern Kentucky %
-University, Nunn Drive, Highland Heights, KY 41076, USA}
+OH 44325-3908, USA}
 
@@
   sequences. Therefore a common step in a genome project is to join
-  neighbouring sequence regions together and fill gaps . This scaffolding step
+  neighbouring sequence regions together and fill gaps. This scaffolding step
   is non-trivial and requires manually editing large blocks of nucleotide
@@
 is the completed puzzle. Sequences of repetitive nucleotide `repeat' regions
-or biased and incomplete sequencing data, may prevent the genome being
-assembled into a continuous sequence. This may be due to insufficient, or
+or biased and incomplete sequencing data may prevent the genome being
+assembled into a continuous sequence. This may be due to insufficient or
 multiple different overlaps between reads and is analogous to missing pieces
@@
 inter-contig gaps represents the length of an unknown region in the genome.
-Filling these regions with unknown nucleotide characters (`N') allows a draft
+Filling these regions with unknown nucleotide characters `N' allows a draft
 continuous sequence. This sequence is useful for representing both the known
@@
 
-The input data for Scaffolder is nucleotide sequence in FASTA format file.
-These nucleotide sequences can be of any length and for example may be
-individual reads, assembled contigs or contigs which have been joined into
-larger scaffolds. The case in which Scaffolder may be most useful is using the
-contigs and scaffolded contigs, combined with additional gap filling sequences
-produced by PCR or \emph{in silico} methods as outlined in the Background. \pb
+The input data for Scaffolder are nucleotide sequences in FASTA format file.
+These nucleotide sequences can be of any length and may be individual reads,
+assembled contigs or contigs which have been joined into larger scaffolds. The
+case in which Scaffolder may be most useful is using the contigs and scaffolded
+contigs, combined with additional gap filling sequences produced by PCR or
+\emph{in silico} methods as outlined in the Background. \pb
 
@@
 reverse complemented if necessary. Each scaffold file represents one scaffolded
-nucleotide sequence and separate scaffolds should be defined in separate files.
-\pb
+nucleotide sequence and as such separate scaffolds should be defined in
+separate files. \pb
 
@@
 The nucleotide sequences used in the scaffold are maintained as a separate
-FASTA file. The nucleotide sequences are referenced in the scaffold using the
+FASTA file: the nucleotide sequences are referenced in the scaffold using the
 first word from the FASTA header of the corresponding sequence. Maintaining the
@@
 ``unresolved'' tag, indicating a region of unknown sequence but known length.
-The second line specifies the size of this unknown region. This entry produces
-a region of 20 `N' characters in the scaffold. \pb
+The second line specifies the size of this unknown region. In this example this
+entry produces a region of 20 `N' characters in the scaffold. \pb
 
@@
 length of the insert FASTA sequence. This allows inserts to bridge into, and
-partially fill gap regions, without requiring an end coordinate position. \pb
+partially fill, gap regions without requiring an end coordinate position. \pb
 
@@
 for the same file and set of FASTA sequences. In comparison, joining large
-nucleotide sequences by hand cannot be reliably reproduced. The scaffold file
-furthermore provides a human readable description of how the scaffold is
+nucleotide sequences by hand cannot be reliably reproduced, while the scaffold
+file also provides a human readable description of how the scaffold is
 constructed. Configuring the final sequence in the scaffold file means the
-build is easier to edit once already constructed. \pb
+build is easier to edit, once constructed. \pb
 
-A use case for Scaffolder is a combination of computational and manual editing
-of a genome scaffold. We have used Scaffolder in our own genome projects to
-create an initial scaffold from computationally parsing the output of \emph{in
-silico} scaffolding tools into YAML. This scaffold was then manually updated as
-the scaffold was finished with additional gap-filling sequences generated in
-the laboratory. This is example of the scaffolder format being both
-computationally tractable while being simple to edit manually. \pb
+An example use case for Scaffolder is a combination of computational and manual
+editing of a genome scaffold. We have used Scaffolder in our own genome
+projects to create an initial scaffold from computationally parsing the output
+of \emph{in silico} scaffolding tools into YAML. This scaffold was then
+manually updated as the scaffold was finished with additional gap-filling
+sequences generated in the laboratory. This is example of the scaffolder format
+being both computationally tractable while being simple to edit manually. \pb
 
@@
 
-Scaffolder is software aimed at both bioinformaticians and biologists familiar
-with the command line who wish to build a genome scaffold from a set of
-contigs. The Scaffolder file format maintains the genome scaffold as a concise
-and readable text representation that allows third-parties to see how the
-genome sequence was scaffolded. This file format also allows a broad overview
-of which sequences were included and how they are ordered in the genome
-scaffold, something not possible to deduce from a megabase-length string of
-nucleotide characters. Scaffolder furthers increases the ease of
-reproducibility in genome projects by allowing the scaffold super-sequence to
-be reliably reproduced from the same scaffold file. The YAML syntax for writing
-the scaffold file is also standardised and simple to manipulate
-programmatically. This thereby means the scaffolding process follows the Unix
-tenet of ``If your data structures are good enough, the algorithm to manipulate
-them should be trivial.''
+Scaffolder is software, written in Ruby, aimed at both bioinformaticians and
+biologists familiar with the command line who wish to build a genome scaffold
+from a set of nucleotide sequences. The Scaffolder file format maintains the
+genome scaffold as a concise and readable text representation that allows
+third-parties to see how the genome sequence was scaffolded. This file format
+also allows a broad overview of which sequences were included and how they are
+ordered in the genome scaffold, something not possible to deduce from a
+megabase-length string of nucleotide characters. Scaffolder furthers increases
+the ease of reproducibility in genome projects by allowing the scaffold
+super-sequence to be reliably reproduced from the same scaffold file. The YAML
+syntax for writing the scaffold file is also standardised and simple to
+manipulate programmatically. This thereby means the scaffolding process follows
+the Unix tenet of ``If your data structures are good enough, the algorithm to
+manipulate them should be trivial.''
 
